### 2011-04-22 paper/main.tex
Initial commit.  Added the draft of the paper and associated files.

\documentclass{article}
\usepackage{nips10submit_e,times}
%\documentstyle[nips07submit_09,times]{article}
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage{lscape}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
%
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\DeclareMathOperator*{\argmax}{arg\,max}

\title{Supervised Topic Modeling in Clinical Text}

\author{
Perotte A\hspace{1cm} Li Y \hspace{1cm} Pivovarov R \hspace{1cm} Weiskopf N\hspace{1cm} Wood F\\
Columbia University, New York, NY 10027, USA \\
\texttt{\{\{ajp9009, yil7003, rip7002, ngw7001\}@dbmi., fwood@stat.\}columbia.edu}
%\texttt{pfau@neurotheory.columbia.edu} 
%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\X}{\mathcal{X}}

\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}
Current medical record keeping technology relies heavily upon human capacity to effectively summarize and infer information from free-text physician notes.  We propose a novel method to suggest diagnostic code assignment for patient visits, based upon narrative medical notes.  We applied a supervised latent Dirichlet allocation model to a corpora of free-text medical notes from NewYork - Presbyterian Hospital to infer a set of specific ICD-9 codes for each patient note.  Evaluation of the predictions were conducted by comparison to a gold-standard set of ICD-9s assigned to a set of patient notes.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
Despite the growing emphasis on meaningful use of technology in medicine, many aspects of medical record-keeping remain a manual process.  Diagnostic coding for billing and insurance purposes is often handled by professional medical coders who must explore a patient's extensive clinical record before assigning the proper codes.  So while electronic health records (EHRs) should be adopted by most medical institutions within the next several years, largely due to the provisions of HITECH under the American Recovery and Reinvestment Act \citep{Blumenthal2009}, there has been little movement forward in automating medical coding.

In this paper we describe the use of a topic model based on supervised latent Dirichlet allocation (sLDA) to identify topics within narrative discharge summaries and to automate the assignment of diagnostic codes, specifically International Classification of Disease 9th Revision, Clinical Modification (ICD-9-CM) codes.  There are a number of advantages to this approach.  First, manually coding diagnoses is a time-consuming and notoriously unreliable process.  Many diagnoses are omitted in the final record, and a high error rate is found even in the principal diagnoses \citep{Surjan1999}.

While there have been some attempts to automatically classify groups of patients as a potential preliminary step to ICD-9 code assignment \citep{Ruch2008, FreitasJunior2006, RibeiroNeto2001, Brown2006}, fully automatic assignment of ICD-9 codes to medical text became a more prevalent research topic only in the last few years. A subset of earlier work proposed various methods on small corpora, based on a few specific diseases \citep{Rao2003} but the most recent and promising work on the subject was inspired by the 2007 Medical NLP Challenge: “International Challenge: Classifying Clinical Free Text Using Natural Language Processing” (website). Most of the classification strategies included word matching and rule-based algorithms. \citep{Goldstein2007, Crammer2007, Farkas2008}.  The data set given to the participants consisted only of documents that were 1-2 lines each and all of the documents were radiology reports - clearly limiting the scope of potential ICD-9 codes which could be assigned. The only paper which has attempted to work with a document scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.  Lita proposed support vector machine and bayesian ridge regression methods to assign appropriate labels to the documents but did not utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

An automated process would ideally produce a more complete and accurate diagnosis lists.  Also, this model will reveal information about the medical records themselves.  For example, we may gain an understanding of what a specific code actually means in terms of clinical narratives.  Similarly, viewing the distribution of topics over discharge summaries may reveal information about the latent structure of clinician documentation.  Lastly, the sLDA model would provide a novel approach to dealing with the problem of high dimensionality when representing narrative text in a vector space specifically by reducing dimensions from an entire vocabulary of potentially tens of thousands of words to a set of several dozen topics.

\section{Methods}

\subsection{Data}
Our data set was gathered from the clinical data warehouse of NewYork - Presbyterian Hospital. The data consisted of free-text discharge summaries and their respective ICD-9 codes. A discharge summary is a clinical report prepared by a physician or other health professional at the conclusion of a hospital stay or series of treatments. The note outlines the patient’s chief complaint, diagnostic findings, therapy administered, patient’s response to the chosen therapy, the treatment plan and the recommendations upon discharge. The ICD-9 codes used to structure the discharge summary data are part of a controlled terminology which is the international standard diagnostic classification for epidemiological, health management, and clinical purposes (http://www.who.int/classifications/icd/en/). The codes are classified in a rooted-tree structure, with each edge representing an is-a relationship between parent and child, such that the parent diagnosis subsumes the child diagnosis.  In the hospital, ICD-9 codes are generated manually by trained medical coders, who review all the information in the discharge summary.  For the purposes of sLDA, ICD-9 codes will be used as labels for discharge summaries.

We worked within the guidelines of the Health Insurance Portability and Accountability Act (HIPAA), which protects patient privacy and the security of potentially identifying medical material, known as personal health information (PHI).  HIPAA covers any information within a medical record that was created, used, or disclosed during the course of providing a health care service and that can be used to identify an individual.  Before beginning data processing, we generated a PHI-free dataset (see Data Pre-processing below).

\subsection{Pre-Processing}
Patient discharge summaries and their associated ICD-9 diagnoses are stored in two different places in the NewYork - Presbyterian data warehouse, and so had to be linked together before being fed into the sLDA algorithm.  Each discharge note and set of diagnoses were assigned a patient unique identifier (PUID) and a visit unique identifier (VUID), allowing the two types of data to be linked.

Natural Language Processsing (NLP) techniques were used to process the free-text discharge summaries.  First, the Natural Language Toolkit (http://www.nltk.org/) was used to tokenize the text.  Next, feature selection was performed using a term frequency - inverse document frequency (TF-IDF) algorithm on the entire document set and sorting the words by their TF-IDF values.  The top 10,000 words were manually evaluated to eliminate all potentially identifying information.  Finally, each discharge summary was converted to a bag-of-words, listing the frequencies of the remaining, free of protected health information, top 10,000 words.

Preparation of the diagnostic codes involved inference over the ICD-9 hierarchy.  The is-a relationships of the hierarchy allowed us to make two important assumptions.  First, if a diagnosis was observed to be present, all of its ancestors could be assumed to be present as well (e.g., if a patient had malignant hypertension, it could be assumed that they also had essential hypertension.  Second, if a diagnosis was observed to be absent, it could be assumed that all of its descendants were also absent (e.g. if a patient did not have essential hypertension, it could be assumed that they did not have malignant hypertension).  Unfortunately, ICD-9 code observations never include observations of disease absence.  ICD-9 codes are only documented when the condition is observed to be present.  Additionally, ICD-9 codes are known to have relatively low sensitivity; conditions that are present are often not documented in a set of ICD-9 codes (Surjan, 1999). Given these facts, we made the following assumptions regarding each visit: recorded diagnoses and their ancestors were labeled as true; diagnoses that were observed at some time for a patient but not at the current visit were labeled as unobserved; and diagnoses that had never been listed for a patient were labeled as false for all of that patient’s visits. This last assumption captures the belief that parts of the ICD-9 hierarchy that are never observed for a particular patient are almost certain to be false.  Additionally, for computational purposes, we decided not to include an ICD-9 code at all if neither it nor one of its descendants had been assigned to a patient in any of the records in our dataset.

\subsection{Supervised Topic Models}
Latent dirichlet allocation (LDA) is a generative probabilistic model of corpora that represents documents as a mixed membership bag-of-words. Also known as topic models, these models infer the latent structure, or topics, of documents in a corpus. Each document is represented as a collection of words, generated from a set of topic assignments (one for each word), where each topic assignment is drawn from a distribution over topics \citep{Blei2003}.

This model, supervised latent dirichlet allocation (sLDA), builds on LDA by incorporating an exponential family response variable. Although there are many models for making predictions based on free text, sLDA is unique in that it is a generative model, it represents documents as a mixed-membership, and constrains the inference of the latent structure of the documents by its predictability of the response variable.  In other words, sLDA infers topics such that the model is capable of a high predictive likelihood for words in a document and the response variable associated with a document.  This approach has been shown to outperform both LASSO (L1 regularized least squares regression) and LDA followed by least squares regression \citep{BleiMcAuliffe2008}.

Here, we augment the sLDA model such that the supervised signal is distribution over the ICD-9 code tree, which is an is-a hierarchy \citep{ICD9NCBO}. An is-a hierarchy is represented by the tree data structure where each node has only a single parent and nodes cannot be parents of ancestors (ie. there are no loops). In this particular case, the ICD-9 code hierarchy is also partially a prefix trie where the labels for certain nodes are prefixes for child nodes. Given that this rule does not apply to all nodes in the hierarchy, we did not use this feature to determine the structure of the hierarchy.  Instead we acquired a dataset that explicitly defined the relationships between the nodes of the hierarchy \citep{ICD9NCBO}. In documentation of ICD-9 codes for billing purposes, only a subset of the nodes can be used, however the nodes higher in the hierarchy contain semantic information about the categories of codes that are their descendants. For this reason, we included these nodes in our model.

\subsection{Generative Model}

\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=4in]{model_1}
   \includegraphics[width=3in]{model_2}  
   \caption{adapted sLDA model}
   \label{fig:example}
\end{figure}

Given a fixed set of parameters: $\beta_{1:K}$, the K topics, $\alpha$,
the parameter for the document specific dirichlet distribution, and
$\eta$, the response variable parameter, the generative process for
documents and responses is as follows: 
\begin{enumerate}
\item Draw topic proportions $\theta\mid\alpha\sim Dir\left(\alpha\right)$ 
\item For each word:

\begin{enumerate}
\item Draw topic assignment $z_{n}\mid\theta\sim Mult\left(\theta\right)$ 
\item Draw word $w_{n}\mid z_{n},\beta_{1:K}\sim Mult\left(\beta_{z_{n}}\right)$ 
\end{enumerate}
\item For each ICD-9 code from the root of the hierarchy and recursively
descending the tree:

\begin{enumerate}
\item Draw regression coefficient $\eta_{i}\mid\mu\sim\mathcal{N}\left(\mu,1\right)$
\item Draw a response variable $y_{i}\mid\bar{z},\eta_{i},Y,\xi\sim\frac{\Phi\left(\eta_{i}^{T}\bar{z}\right)\psi\left(Y\right)}{\mathcal{Z}}$
where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\psi\left(Y\right)=\exp\left\{ \sum_{\left\{ p,c\right\} }f(p,c)\right\} $
and $\begin{cases}
\begin{array}{c}
f(0,0)=\xi_{0}\\
f(0,1)=-\infty\\
f(1,0)=\xi_{1}\\
f(1,1)=\xi_{2}\end{array}\end{cases}$
\end{enumerate}
\end{enumerate}
We will employ a data augmentation scheme with auxiliary variables
$a_{i}$ in the probit model such that:

\begin{equation}
y_{i}\sim\begin{cases}
\begin{array}{c}
1,\\
0,\end{array} & \begin{array}{c}
a_{i}\geq0\\
a_{i}<0\end{array}\end{cases}\end{equation}

\begin{equation}
a_{i}\sim\mathcal{N}\left(\eta_{i}^{T}\bar{z},1\right)\end{equation}

\begin{equation}
\eta_{i}\sim\mathcal{N}\left(\mu=0,1\right)\end{equation}

\subsection{Posterior Inference}

Given an observation of a set of ICD-9 codes and a document, the posterior
distribution for the latent variables is given by Equation 4.

The denominator for this distribution is the marginal probability
of the data and cannot be solved in closed form. This is often the
case in evaluating posterior distributions of non-trivial probabilistic
models. We will appeal to one of the common methods for approximating
posterior distributions in the face of intractable normalization factors:
markov chain monte carlo (MCMC). Since in this model it is possible
to sample from the conditional distributions for all variables we
will use the gibbs sampling algorithm to obtain an approximation to
this posterior. In particular, we will derive a Rao-Blackwellized
gibbs sampler for the supervised topic model for general exponential
family response variables and then for our model with a probit response
variable.

\subsection{Rao-Blackwellization}

To derive the Rao-Blackwellized gibbs sampler in general, we integrate
over the parameters $\theta$ and $\phi_{1:K}$ resulting in the
collapsed joint distribution shown in Equations 5-8.

\begin{landscape}
\centering     % optional, probably makes it look better to have it centered on the page

\begin{small}

\begin{equation}
p\left(\theta,z_{1:N}\mid w_{1:N},y_{1:I},\phi_{1:K},\eta_{1:I},\alpha,\beta,\mu,\xi\right)=\frac{p\left(\theta\mid\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)}{\int_{\theta}p\left(\theta\mid\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)d\theta}\end{equation}

\begin{equation}
\intop_{\theta}\intop_{\phi_{1:K}}p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\theta},\mathbf{\phi},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)d\theta d\phi_{1:K}=\intop_{\theta}\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\left\{ p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}\left[p\left(z_{m,n}\mid\theta_{m}\right)p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)\right]\prod_{i=1}^{I}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right\} \prod_{i=1}^{I}p\left(\eta_{i};\mu\right)d\theta d\phi_{1:K}\end{equation}

\begin{equation}
p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\prod_{n=1}^{N}p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)d\phi_{1:K}\intop_{\theta}\prod_{m=1}^{M}p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}p\left(z_{m,n}\mid\theta_{m}\right)d\theta\end{equation}

\begin{equation}
=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}

\begin{equation}
=\prod_{i=1}^{I}\left[\mathcal{N}\left(\eta_{i}\mid\mu,1\right)\prod_{m=1}^{M}h\left(y\right)\exp\left\{ \left(\eta_{i}^{T}\bar{z}\right)y-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}

\end{small}
\end{landscape}

\subsection{Gibbs Sampling}

To derive the gibbs sampler for the general exponential family form
as well as the probit regression model, we derive the individual conditional
probability distributions for all latent variables and parameters.

\subsubsection{$p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)$}

For the purposes of sampling, we will be able to derive a representation
of the joint distribution isolating a particular latent variable,
$z$, for a word instance, $n$, in a document instance, $m$. The
conditional probability with respect to this latent variable is proportional
to the joint distribution up to a constant.

\begin{equation}
p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)\propto p\left(z_{m,n},\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)\end{equation}

Due to the factorization of this model we can rewrite the joint distribution
as the following:

\begin{equation}
\propto\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)p\left(y_{m,i}\mid z_{m,1:N},\eta_{i}\xi\right)\right]p\left(z_{m,n},\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\end{equation}

We isolate only terms that depend on $z_{m,n}$ and absorb all other
constant terms into the normalization constant \citep{Griffiths04}.

\begin{equation}
\propto\prod_{i=1}^{I}\left[p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

Here, $n_{m,v}^{k,-\left(m,n\right)}$ represents the count of word
$v$ in document $m$ assigned to topic $k$ omitting the $(m,n)^{th}$
word count. For exponential family distributions, the normalization
contant, $h(y_{m,i})$, does not depend on $z_{m,n}$.

\begin{equation}
\propto\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\label{eq:z_likelihood}\end{equation}

Given this expression, $p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)$
can be sampled through enumeration as seen in Equation 13.  In the case of probit regression, the expression for \ref{eq:z_likelihood} evaluates
to Equation 14.  Equivalently we can parameterize the model with an auxiliary variable
$a_{i}$, resulting in Equation 15.

%\begin{equation}
%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)=\frac{\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}{\sum_{k=1}^{K}\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}\end{equation}

%\begin{equation}
%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\Phi\left(\eta_{i}^{T}\bar{z}\right)^{y_{m,i}}\left(1-\Phi\left(\eta_{i}^{T}\bar{z}\right)\right)^{1-y_{m,i}}\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

%\begin{equation}
%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\mathbf{a},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\delta\left(sign\left(a_{m,i}\right)=y_{m,i}\right)\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

\begin{landscape}

\centering     % optional, probably makes it look better to have it centered on the page
\begin{equation}
p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)=\frac{\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}{\sum_{k=1}^{K}\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}\end{equation}

\begin{equation}
p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\Phi\left(\eta_{i}^{T}\bar{z}\right)^{y_{m,i}}\left(1-\Phi\left(\eta_{i}^{T}\bar{z}\right)\right)^{1-y_{m,i}}\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

\begin{equation}
p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\mathbf{a},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\delta\left(sign\left(a_{m,i}\right)=2y_{m,i}-1\right)\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}

\end{landscape}

\subsubsection{$p\left(\eta_{i}\mid\mathbf{z},\mathbf{Y},\mu\right)$ or $p\left(\eta_{i}\mid\mathbf{z},\mathbf{a},\mu\right)$in
the augmented probit regression model}

Given that $\eta_{i}$ and $a_{m,i}$ are distributed normally, this
posterior distribution is also normal. In the case for general exponential
family distributions, $\eta_{i}$ can remain a parameter without a
prior, fit with maximimum likelihood in the usual fashion.

\begin{equation}
p\left(\eta_{i}\mid\mathbf{z},\mathbf{a},\mu\right)=\mathcal{N}\left(\eta_{i}\mid\hat{\mathbf{\mu}_{i}},\hat{\mathbf{S}_{i}}\right)\end{equation}

\[
\hat{\mathbf{\mu}_{i}}=\hat{\mathbf{S}_{i}}\bar{\mathbf{Z}}^{T}\mathbf{a}_{\left(\bullet\right),i}\]

\[
\hat{\mathbf{S}_{i}}^{-1}=\mathbf{I}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\]

\subsubsection{$p\left(a_{m,i}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$ in
the augmented probit regression model}

In the augmented probit regression model, the posterior distribution
of $a_{i}$ is distributed according to a truncated normal distribution.

\[
p\left(a_{m,i}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)=trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)\]

\subsubsection{\textmd{$p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)$}}

In our model, response variables are not always observed and are treated
as latent and sampled where appropriate. There are two factors influencing
predictions of the response variable, $y_{m,i}$. There is an undirected
model enforcing the aforementioned constraints and providing a prior
and there is the probit regression.

\begin{equation}
p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)\propto\psi\left(\mathbf{Y}\right)\delta\left(sign\left(a_{m,i}\right)=y_{m,i}\right)\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1\right)\end{equation}

\begin{equation}
p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)\propto\psi\left(\mathbf{Y}\right)trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)\end{equation}

Again, this conditional distribution can be evaluated through enumerations
and normalization.

\begin{equation}
p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)=\frac{\psi\left(\mathbf{Y}\right)trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)}{\sum_{y_{m,i}}\psi\left(\mathbf{Y}\right)trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)}\end{equation}

\section{Results}
%We ran the Gibbs sampler for two days on 500 documents with 20 topics to learn the sLDA model which we then followed by a prediction of ICD-9 codes for 100 documents. The prediction resulted in 38 documents with at least 1 ICD-9 code assignment while some had over 200 assignments.

%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=4in]{number_of_codes_histogram} 
%   \caption{distribution of ICD-9 code assignments over all of the 100 documents predicted}
%   \label{fig:example}
%\end{figure}

%The accuracy of the prediction was tested against the actual ICD-9 codes in the EHR record for each discharge summary.  Different statistical measures were applied to asses the predictions.

%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
%   \centering
%   \includegraphics[width=4in]{signal_detection} 
%   \caption{sensitivity, specificity, precision, recall and F-measure.}
%   \label{fig:example}
%\end{figure}

%Although the sensitivty is not as high as we would have hoped, it is important to note that the sLDA was run on a very limited sample of documents (500) due to time constraints.  Additionally, the “gold standard” data used are actual hospital results which are very noisy given the incompleteness and introduction of human error.

\section{Conclusion}
%We have proposed a new method for prediction of ICD-9 codes given a medical discharge summary note.  Our method explicitly takes advantage of the ICD-9 hierarchical IS-A relationships as well as the word distributions in each document.  We used Gibbs sampling for a supervised latent dirichlet allocation model to learn a distribution of words over twenty topics and their appropriate ICD-9 codes (response variable). The model was then tested on 100 discharge summaries for which the ICD-9 diagnoses were known.  The amount of ICD-9 codes produced for some discharge summaries was quite high because of the way the model worked with the IS-A hierarchy, highlighting the completeness of ICD-9 coding when using this sLDA model.  Overall, the model had extremely good specificity while the sensitivity was lower than anticipated.  Given our limited set of documents, and very noisy gold standard it would not be accurate to discount the use of sLDA in medical document classification and automatic coding.
%In future work, we plan to significantly expand the set of documents used to train the model (up to 15,000), as well as run several different iterations to learn the optimal topic number.

%A more automated approach for coding patient visits with the ICD-9 hierarchy has great promise to  reduce health care costs and increase completeness of medical coded data.  We hope that this implementation of supervised LDA will serve as preliminary work for a much larger study with more documents in the future.

\begin{small}
\bibliographystyle{plainnat}
\bibliography{refs} 
\end{small}
\end{document}

### 2011-04-25 paper/main.tex
Updated generative model and minor fixes

@@
 
-\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{model_1}
-   \includegraphics[width=3in]{model_2}  
+%
+\begin{figure}[h]
+%tbp] %  figure placement: here, top, bottom, or page
+ \centering \includegraphics[width=4in]{model_1} \includegraphics[width=3in]{model_2}
 \caption{adapted sLDA model}
+
 \label{fig:example} 
@@
 
-Given a fixed set of parameters: $\beta_{1:K}$, the K topics, $\alpha$,
-the parameter for the document specific dirichlet distribution, and
-$\eta$, the response variable parameter, the generative process for
-documents and responses is as follows: 
+Given the number of topics, K, the global prior over topic proportions,
+$\alpha'$, and the prior over topics, $\gamma$, the generative process
+for documents and responses is as follows: 
 \begin{enumerate}
-\item Draw topic proportions $\theta\mid\alpha\sim Dir\left(\alpha\right)$ 
+\item For each topic:
+
+\begin{enumerate}
+\item Draw a distribution over words $\beta_{k}\sim Dir\left(u,\gamma\right)$
+\end{enumerate}
+\item For each ICD9 Code:
+
+\begin{enumerate}
+\item Draw regression coefficient $\eta_{i}\mid\mu,\sigma\sim\mathcal{N}\left(\mu,\sigma\right)$ 
+\end{enumerate}
+\item Draw a prior over topic proportions $m\mid\alpha'\sim Dir\left(u,\alpha'\right)$ 
+\item For each document:
+
+\begin{enumerate}
+\item Draw topic proportions $\theta_{d}\mid\alpha\sim Dir\left(m,\alpha\right)$ 
 \item For each word:
@@
 \begin{enumerate}
-\item Draw topic assignment $z_{n}\mid\theta\sim Mult\left(\theta\right)$ 
-\item Draw word $w_{n}\mid z_{n},\beta_{1:K}\sim Mult\left(\beta_{z_{n}}\right)$ 
+\item Draw topic assignment $z_{n,d}\mid\theta_{d}\sim Mult\left(\theta_{d}\right)$ 
+\item Draw word $w_{n,d}\mid z_{n,d},\beta_{1:K}\sim Mult\left(\beta_{z_{n}}\right)$ 
 \end{enumerate}
@@
 \begin{enumerate}
-\item Draw regression coefficient $\eta_{i}\mid\mu\sim\mathcal{N}\left(\mu,1\right)$
-\item Draw a response variable $y_{i}\mid\bar{z},\eta_{i},Y,\xi\sim\frac{\Phi\left(\eta_{i}^{T}\bar{z}\right)\psi\left(Y\right)}{\mathcal{Z}}$
-where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\psi\left(Y\right)=\exp\left\{ \sum_{\left\{ p,c\right\} }f(p,c)\right\} $
-and $\begin{cases}
-\begin{array}{c}
-f(0,0)=\xi_{0}\\
-f(0,1)=-\infty\\
-f(1,0)=\xi_{1}\\
-f(1,1)=\xi_{2}\end{array}\end{cases}$
+\item Draw a response variable $y_{i}\mid\bar{z},\eta_{i},y_{parent}\sim\Phi(\eta_{i}^{T}\bar{z},y_{parent})$
+where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\Phi$ refers to a
+conditional probit model.
+\end{enumerate}
 \end{enumerate}
@@
 We will employ a data augmentation scheme with auxiliary variables
-$a_{i}$ in the probit model such that:
+$a_{i}$ in the probit model where:
 
@@
 0,\end{array} & \begin{array}{c}
-a_{i}\geq0\\
-a_{i}<0\end{array}\end{cases}\end{equation}
+a_{i}>0\; and\; y_{parent}=1\\
+otherwise\end{array}\end{cases}\end{equation}
 
@@
 a_{i}\sim\mathcal{N}\left(\eta_{i}^{T}\bar{z},1\right)\end{equation}
-
-\begin{equation}
-\eta_{i}\sim\mathcal{N}\left(\mu=0,1\right)\end{equation}
 
@@
 posterior distributions in the face of intractable normalization factors:
-markov chain monte carlo (MCMC). Since in this model it is possible
-to sample from the conditional distributions for all variables we
-will use the gibbs sampling algorithm to obtain an approximation to
-this posterior. In particular, we will derive a Rao-Blackwellized
-gibbs sampler for the supervised topic model for general exponential
-family response variables and then for our model with a probit response
-variable.
+Markov chain Monte Carlo (MCMC) sampling. Since in this model it is
+possible to sample from the conditional distributions for all variables
+we will use the Gibbs sampling algorithm to obtain an approximation
+to this posterior. In particular, we will derive a collapsed Gibbs
+sampler for the supervised topic model.
 
@@
 
-To derive the Rao-Blackwellized gibbs sampler in general, we integrate
-over the parameters $\theta$ and $\phi_{1:K}$ resulting in the
-collapsed joint distribution shown in Equations 5-8.
+To derive the Gibbs sampler in general, we integrate over the parameters
+$\theta$ and $\phi_{1:K}$ resulting in the collapsed joint distribution
+shown in Equations 5-8.
 
-\begin{landscape}
-\centering     % optional, probably makes it look better to have it centered on the page
+\begin{landscape} \centering % optional, probably makes it look better to have it centered on the page
 
@@
 
-\end{small}
-\end{landscape}
+\end{small} \end{landscape}
 
@@
 
-To derive the gibbs sampler for the general exponential family form
-as well as the probit regression model, we derive the individual conditional
-probability distributions for all latent variables and parameters.
+To derive the Gibbs sampler we evaluate the individual conditional
+probability distributions for all unobserved variables.
 
@@
 \begin{equation}
-\propto\prod_{i=1}^{I}\left[p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}
+\propto\prod_{i=1}^{I}\left[p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\cdot\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}
 
@@
 Given this expression, $p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)$
-can be sampled through enumeration as seen in Equation 13.  In the case of probit regression, the expression for \ref{eq:z_likelihood} evaluates
-to Equation 14.  Equivalently we can parameterize the model with an auxiliary variable
-$a_{i}$, resulting in Equation 15.
+can be sampled through enumeration as seen in Equation 13. In the
+case of probit regression, the expression for \ref{eq:z_likelihood}
+evaluates to Equation 14. Equivalently we can parameterize the model
+with an auxiliary variable $a_{i}$, resulting in Equation 15.
 
@@
 \section{Results}
+
 %We ran the Gibbs sampler for two days on 500 documents with 20 topics to learn the sLDA model which we then followed by a prediction of ICD-9 codes for 100 documents.
@@
 \section{Conclusion}
+
 %We have proposed a new method for prediction of ICD-9 codes given a medical discharge summary note.
@@
 
-\begin{small}
-\bibliographystyle{plainnat}
+\begin{small} \bibliographystyle{plainnat} \bibliographystyle{plainnat}
 \bibliography{refs}

### 2011-05-10 paper/main.tex
Shuffled things around, updated model.

@@
 \begin{abstract}
-Current medical record keeping technology relies heavily upon human capacity to effectively summarize and infer information from free-text physician notes.
-We propose a novel method to suggest diagnostic code assignment for patient visits, based upon narrative medical notes.
-We applied a supervised latent Dirichlet allocation model to a corpora of free-text medical notes from NewYork - Presbyterian Hospital to infer a set of specific ICD-9 codes for each patient note.
-Evaluation of the predictions were conducted by comparison to a gold-standard set of ICD-9s assigned to a set of patient notes.
+Current medical record keeping technology relies heavily upon human
+capacity to effectively summarize and infer information from free-text
+physician notes. We propose a novel method to suggest diagnostic code
+assignment for patient visits, based upon narrative medical notes.
+We applied a supervised latent Dirichlet allocation model to a corpora
+of free-text medical notes from NewYork - Presbyterian Hospital to
+infer a set of specific ICD-9 codes for each patient note. Evaluation
+of the predictions were conducted by comparison to a gold-standard
+set of ICD-9s assigned to a set of patient notes. 
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:introduction}
-Despite the growing emphasis on meaningful use of technology in medicine, many aspects of medical record-keeping remain a manual process.
-Diagnostic coding for billing and insurance purposes is often handled by professional medical coders who must explore a patient's extensive clinical record before assigning the proper codes.
-So while electronic health records (EHRs) should be adopted by most medical institutions within the next several years, largely due to the provisions of HITECH under the American Recovery and Reinvestment Act \citep{Blumenthal2009}, there has been little movement forward in automating medical coding.
 
-In this paper we describe the use of a topic model based on supervised latent Dirichlet allocation (sLDA) to identify topics within narrative discharge summaries and to automate the assignment of diagnostic codes, specifically International Classification of Disease 9th Revision, Clinical Modification (ICD-9-CM) codes.
-There are a number of advantages to this approach.
-First, manually coding diagnoses is a time-consuming and notoriously unreliable process.
-Many diagnoses are omitted in the final record, and a high error rate is found even in the principal diagnoses \citep{Surjan1999}.
+\label{sec:introduction} Despite the growing emphasis on meaningful
+use of technology in medicine, many aspects of medical record-keeping
+remain a manual process. Diagnostic coding for billing and insurance
+purposes is often handled by professional medical coders who must
+explore a patient's extensive clinical record before assigning the
+proper codes. So while electronic health records (EHRs) should be
+adopted by most medical institutions within the next several years,
+largely due to the provisions of HITECH under the American Recovery
+and Reinvestment Act \citep{Blumenthal2009}, there has been little
+movement forward in automating medical coding.
 
-While there have been some attempts to automatically classify groups of patients as a potential preliminary step to ICD-9 code assignment \citep{Ruch2008, FreitasJunior2006, RibeiroNeto2001, Brown2006}, fully automatic assignment of ICD-9 codes to medical text became a more prevalent research topic only in the last few years.
-A subset of earlier work proposed various methods on small corpora, based on a few specific diseases \citep{Rao2003} but the most recent and promising work on the subject was inspired by the 2007 Medical NLP Challenge: “International Challenge: Classifying Clinical Free Text Using Natural Language Processing” (website).
-Most of the classification strategies included word matching and rule-based algorithms.
-\citep{Goldstein2007, Crammer2007, Farkas2008}.
-The data set given to the participants consisted only of documents that were 1-2 lines each and all of the documents were radiology reports - clearly limiting the scope of potential ICD-9 codes which could be assigned.
-The only paper which has attempted to work with a document scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
-Lita proposed support vector machine and bayesian ridge regression methods to assign appropriate labels to the documents but did not utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
+In this paper we describe the use of a topic model based on supervised
+latent Dirichlet allocation (sLDA) to identify topics within narrative
+discharge summaries and to automate the assignment of diagnostic codes,
+specifically International Classification of Disease 9th Revision,
+Clinical Modification (ICD-9-CM) codes. There are a number of advantages
+to this approach. First, manually coding diagnoses is a time-consuming
+and notoriously unreliable process. Many diagnoses are omitted in
+the final record, and a high error rate is found even in the principal
+diagnoses \citep{Surjan1999}.
 
-An automated process would ideally produce a more complete and accurate diagnosis lists.
-Also, this model will reveal information about the medical records themselves.
-For example, we may gain an understanding of what a specific code actually means in terms of clinical narratives.
-Similarly, viewing the distribution of topics over discharge summaries may reveal information about the latent structure of clinician documentation.
-Lastly, the sLDA model would provide a novel approach to dealing with the problem of high dimensionality when representing narrative text in a vector space specifically by reducing dimensions from an entire vocabulary of potentially tens of thousands of words to a set of several dozen topics.
+While there have been some attempts to automatically classify groups
+of patients as a potential preliminary step to ICD-9 code assignment
+\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
+automatic assignment of ICD-9 codes to medical text became a more
+prevalent research topic only in the last few years. A subset of earlier
+work proposed various methods on small corpora, based on a few specific
+diseases \citep{Rao2003} but the most recent and promising work on
+the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
+Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
+(website). Most of the classification strategies included word matching
+and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
+The data set given to the participants consisted only of documents
+that were 1-2 lines each and all of the documents were radiology reports
+- clearly limiting the scope of potential ICD-9 codes which could
+be assigned. The only paper which has attempted to work with a document
+scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
+Lita proposed support vector machine and bayesian ridge regression
+methods to assign appropriate labels to the documents but did not
+utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
+
+An automated process would ideally produce a more complete and accurate
+diagnosis lists. Also, this model will reveal information about the
+medical records themselves. For example, we may gain an understanding
+of what a specific code actually means in terms of clinical narratives.
+Similarly, viewing the distribution of topics over discharge summaries
+may reveal information about the latent structure of clinician documentation.
+Lastly, the sLDA model would provide a novel approach to dealing with
+the problem of high dimensionality when representing narrative text
+in a vector space specifically by reducing dimensions from an entire
+vocabulary of potentially tens of thousands of words to a set of several
+dozen topics.
+
+\section{Background}
+
+\subsection{Data}
+
+Our data set was gathered from the clinical data warehouse of NewYork
+- Presbyterian Hospital. The data consisted of free-text discharge
+summaries and their respective ICD-9 codes. A discharge summary is
+a clinical report prepared by a physician or other health professional
+at the conclusion of a hospital stay or series of treatments. The
+note outlines the patient\textquoteright{}s chief complaint, diagnostic
+findings, therapy administered, patient\textquoteright{}s response
+to the chosen therapy, the treatment plan and the recommendations
+upon discharge. The ICD-9 codes used to structure the discharge summary
+data are part of a controlled terminology which is the international
+standard diagnostic classification for epidemiological, health management,
+and clinical purposes (http://www.who.int/classifications/icd/en/).
+The codes are classified in a rooted-tree structure, with each edge
+representing an is-a relationship between parent and child, such that
+the parent diagnosis subsumes the child diagnosis. In the hospital,
+ICD-9 codes are generated manually by trained medical coders, who
+review all the information in the discharge summary. For the purposes
+of sLDA, ICD-9 codes will be used as labels for discharge summaries.
+
+We worked within the guidelines of the Health Insurance Portability
+and Accountability Act (HIPAA), which protects patient privacy and
+the security of potentially identifying medical material, known as
+personal health information (PHI). HIPAA covers any information within
+a medical record that was created, used, or disclosed during the course
+of providing a health care service and that can be used to identify
+an individual. Before beginning data processing, we generated a PHI-free
+dataset (see Data Pre-processing below).
+
+\subsection{Pre-Processing}
+
+Patient discharge summaries and their associated ICD-9 diagnoses are
+stored in two different places in the NewYork - Presbyterian data
+warehouse, and so had to be linked together before being fed into
+the sLDA algorithm. Each discharge note and set of diagnoses were
+assigned a patient unique identifier (PUID) and a visit unique identifier
+(VUID), allowing the two types of data to be linked.
+
+Natural Language Processsing (NLP) techniques were used to process
+the free-text discharge summaries. First, the Natural Language Toolkit
+(http://www.nltk.org/) was used to tokenize the text. Next, feature
+selection was performed using a term frequency - inverse document
+frequency (TF-IDF) algorithm on the entire document set and sorting
+the words by their TF-IDF values. The top 10,000 words were manually
+evaluated to eliminate all potentially identifying information. Finally,
+each discharge summary was converted to a bag-of-words, listing the
+frequencies of the remaining, free of protected health information,
+top 10,000 words.
+
+Preparation of the diagnostic codes involved inference over the ICD-9
+hierarchy. The is-a relationships of the hierarchy allowed us to make
+two important assumptions. First, if a diagnosis was observed to be
+present, all of its ancestors could be assumed to be present as well
+(e.g., if a patient had malignant hypertension, it could be assumed
+that they also had essential hypertension. Second, if a diagnosis
+was observed to be absent, it could be assumed that all of its descendants
+were also absent (e.g. if a patient did not have essential hypertension,
+it could be assumed that they did not have malignant hypertension).
+Unfortunately, ICD-9 code observations never include observations
+of disease absence. ICD-9 codes are only documented when the condition
+is observed to be present. Additionally, ICD-9 codes are known to
+have relatively low sensitivity; conditions that are present are often
+not documented in a set of ICD-9 codes (Surjan, 1999). Given these
+facts, we made the following assumptions regarding each visit: recorded
+diagnoses and their ancestors were labeled as true; diagnoses that
+were observed at some time for a patient but not at the current visit
+were labeled as unobserved; and diagnoses that had never been listed
+for a patient were labeled as false for all of that patient\textquoteright{}s
+visits. This last assumption captures the belief that parts of the
+ICD-9 hierarchy that are never observed for a particular patient are
+almost certain to be false. Additionally, for computational purposes,
+we decided not to include an ICD-9 code at all if neither it nor one
+of its descendants had been assigned to a patient in any of the records
+in our dataset.
+
+\subsection{ICD-9 Code Hierarchy}
+
+Here, we augment the sLDA model such that the supervised signal is
+distribution over the ICD-9 code tree, which is an is-a hierarchy
+\citep{ICD9NCBO}. An is-a hierarchy is represented by the tree data
+structure where each node has only a single parent and nodes cannot
+be parents of ancestors (ie. there are no loops). In this particular
+case, the ICD-9 code hierarchy is also partially a prefix trie where
+the labels for certain nodes are prefixes for child nodes. Given that
+this rule does not apply to all nodes in the hierarchy, we did not
+use this feature to determine the structure of the hierarchy. Instead
+we acquired a dataset that explicitly defined the relationships between
+the nodes of the hierarchy \citep{ICD9NCBO}. In documentation of
+ICD-9 codes for billing purposes, only a subset of the nodes can be
+used, however the nodes higher in the hierarchy contain semantic information
+about the categories of codes that are their descendants. For this
+reason, we included these nodes in our model.
 
 \section{Methods}
-
-\subsection{Data}
-Our data set was gathered from the clinical data warehouse of NewYork - Presbyterian Hospital.
-The data consisted of free-text discharge summaries and their respective ICD-9 codes.
-A discharge summary is a clinical report prepared by a physician or other health professional at the conclusion of a hospital stay or series of treatments.
-The note outlines the patient’s chief complaint, diagnostic findings, therapy administered, patient’s response to the chosen therapy, the treatment plan and the recommendations upon discharge.
-The ICD-9 codes used to structure the discharge summary data are part of a controlled terminology which is the international standard diagnostic classification for epidemiological, health management, and clinical purposes (http://www.who.int/classifications/icd/en/).
-The codes are classified in a rooted-tree structure, with each edge representing an is-a relationship between parent and child, such that the parent diagnosis subsumes the child diagnosis.
-In the hospital, ICD-9 codes are generated manually by trained medical coders, who review all the information in the discharge summary.
-For the purposes of sLDA, ICD-9 codes will be used as labels for discharge summaries.
-
-We worked within the guidelines of the Health Insurance Portability and Accountability Act (HIPAA), which protects patient privacy and the security of potentially identifying medical material, known as personal health information (PHI).
-HIPAA covers any information within a medical record that was created, used, or disclosed during the course of providing a health care service and that can be used to identify an individual.
-Before beginning data processing, we generated a PHI-free dataset (see Data Pre-processing below).
-
-\subsection{Pre-Processing}
-Patient discharge summaries and their associated ICD-9 diagnoses are stored in two different places in the NewYork - Presbyterian data warehouse, and so had to be linked together before being fed into the sLDA algorithm.
-Each discharge note and set of diagnoses were assigned a patient unique identifier (PUID) and a visit unique identifier (VUID), allowing the two types of data to be linked.
-
-Natural Language Processsing (NLP) techniques were used to process the free-text discharge summaries.
-First, the Natural Language Toolkit (http://www.nltk.org/) was used to tokenize the text.
-Next, feature selection was performed using a term frequency - inverse document frequency (TF-IDF) algorithm on the entire document set and sorting the words by their TF-IDF values.
-The top 10,000 words were manually evaluated to eliminate all potentially identifying information.
-Finally, each discharge summary was converted to a bag-of-words, listing the frequencies of the remaining, free of protected health information, top 10,000 words.
-
-Preparation of the diagnostic codes involved inference over the ICD-9 hierarchy.
-The is-a relationships of the hierarchy allowed us to make two important assumptions.
-First, if a diagnosis was observed to be present, all of its ancestors could be assumed to be present as well (e.g., if a patient had malignant hypertension, it could be assumed that they also had essential hypertension.
-Second, if a diagnosis was observed to be absent, it could be assumed that all of its descendants were also absent (e.g.
-if a patient did not have essential hypertension, it could be assumed that they did not have malignant hypertension).
-Unfortunately, ICD-9 code observations never include observations of disease absence.
-ICD-9 codes are only documented when the condition is observed to be present.
-Additionally, ICD-9 codes are known to have relatively low sensitivity; conditions that are present are often not documented in a set of ICD-9 codes (Surjan, 1999).
-Given these facts, we made the following assumptions regarding each visit: recorded diagnoses and their ancestors were labeled as true; diagnoses that were observed at some time for a patient but not at the current visit were labeled as unobserved; and diagnoses that had never been listed for a patient were labeled as false for all of that patient’s visits.
-This last assumption captures the belief that parts of the ICD-9 hierarchy that are never observed for a particular patient are almost certain to be false.
-Additionally, for computational purposes, we decided not to include an ICD-9 code at all if neither it nor one of its descendants had been assigned to a patient in any of the records in our dataset.
-
-\subsection{Supervised Topic Models}
-Latent dirichlet allocation (LDA) is a generative probabilistic model of corpora that represents documents as a mixed membership bag-of-words.
-Also known as topic models, these models infer the latent structure, or topics, of documents in a corpus.
-Each document is represented as a collection of words, generated from a set of topic assignments (one for each word), where each topic assignment is drawn from a distribution over topics \citep{Blei2003}.
-
-This model, supervised latent dirichlet allocation (sLDA), builds on LDA by incorporating an exponential family response variable.
-Although there are many models for making predictions based on free text, sLDA is unique in that it is a generative model, it represents documents as a mixed-membership, and constrains the inference of the latent structure of the documents by its predictability of the response variable.
-In other words, sLDA infers topics such that the model is capable of a high predictive likelihood for words in a document and the response variable associated with a document.
-This approach has been shown to outperform both LASSO (L1 regularized least squares regression) and LDA followed by least squares regression \citep{BleiMcAuliffe2008}.
-
-Here, we augment the sLDA model such that the supervised signal is distribution over the ICD-9 code tree, which is an is-a hierarchy \citep{ICD9NCBO}.
-An is-a hierarchy is represented by the tree data structure where each node has only a single parent and nodes cannot be parents of ancestors (ie.
-there are no loops).
-In this particular case, the ICD-9 code hierarchy is also partially a prefix trie where the labels for certain nodes are prefixes for child nodes.
-Given that this rule does not apply to all nodes in the hierarchy, we did not use this feature to determine the structure of the hierarchy.
-Instead we acquired a dataset that explicitly defined the relationships between the nodes of the hierarchy \citep{ICD9NCBO}.
-In documentation of ICD-9 codes for billing purposes, only a subset of the nodes can be used, however the nodes higher in the hierarchy contain semantic information about the categories of codes that are their descendants.
-For this reason, we included these nodes in our model.
-
-\subsection{Generative Model}
 
@@
 %tbp] %  figure placement: here, top, bottom, or page
- \centering \includegraphics[width=4in]{model_1} \includegraphics[width=3in]{model_2}
-\caption{adapted sLDA model}
+ \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}
 
@@
 
-Given the number of topics, K, the global prior over topic proportions,
-$\alpha'$, and the prior over topics, $\gamma$, the generative process
-for documents and responses is as follows: 
+Given the number of topics, K, and broad gamma priors on hyperparameters,
+the generative process is as follows: 
 \begin{enumerate}
-\item For each topic:
+\item For each topic, $k$:
 
 \begin{enumerate}
-\item Draw a distribution over words $\beta_{k}\sim Dir\left(u,\gamma\right)$
+\item Draw a distribution over words $\phi_{k}\sim Dir_{V}\left(\mathbf{1},\gamma\right)$ 
 \end{enumerate}
-\item For each ICD9 Code:
+\item For each ICD9 code, $c$, at all levels in the tree, $l$:
 
 \begin{enumerate}
-\item Draw regression coefficient $\eta_{i}\mid\mu,\sigma\sim\mathcal{N}\left(\mu,\sigma\right)$ 
+\item Draw regression coefficient $\eta_{i_{l,c}}\mid\sigma\sim\mathcal{N}_{K}\left(-\mathbf{1},\sigma\right)$ 
 \end{enumerate}
-\item Draw a prior over topic proportions $m\mid\alpha'\sim Dir\left(u,\alpha'\right)$ 
-\item For each document:
+\item Draw a prior over topic proportions $\beta\mid\alpha'\sim Dir_{K}\left(\mathbf{1},\alpha'\right)$ 
+\item For each document, $d$:
 
 \begin{enumerate}
-\item Draw topic proportions $\theta_{d}\mid\alpha\sim Dir\left(m,\alpha\right)$ 
-\item For each word:
+\item Draw topic proportions $\theta_{d}\mid\beta,\alpha\sim Dir_{K}\left(\beta,\alpha\right)$ 
+\item For each word, $n$:
 
 \begin{enumerate}
-\item Draw topic assignment $z_{n,d}\mid\theta_{d}\sim Mult\left(\theta_{d}\right)$ 
-\item Draw word $w_{n,d}\mid z_{n,d},\beta_{1:K}\sim Mult\left(\beta_{z_{n}}\right)$ 
+\item Draw topic assignment $z_{n,d}\mid\theta_{d}\sim Mult_{K}\left(\theta_{d}\right)$ 
+\item Draw word $w_{n,d}\mid z_{n,d},\beta_{1:K}\sim Mult_{V}\left(\beta_{z_{n}}\right)$ 
 \end{enumerate}
-\item For each ICD-9 code from the root of the hierarchy and recursively
-descending the tree:
+\item For each level of the ICD-9 code tree, $l$:
 
 \begin{enumerate}
-\item Draw a response variable $y_{i}\mid\bar{z},\eta_{i},y_{parent}\sim\Phi(\eta_{i}^{T}\bar{z},y_{parent})$
-where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\Phi$ refers to a
-conditional probit model.
+\item For each ICD-9 code at this level, $c$:
+
+\begin{enumerate}
+\item Draw a latent variable \[
+a_{d,i_{l,c}}\sim\begin{cases}
+\mathcal{N}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=1\\
+trunc\mathcal{N}^{-}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=-1\end{cases}\]
+ where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\mathcal{I}=\left\{ i_{0},i_{1},...,i_{\mathcal{L}}\right\} $
+and $i_{l}=\left\{ i_{l,0},i_{l,1},...,i_{l,C_{l}}\right\} $
+\item Draw a response variable $y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\sim\begin{cases}
+\begin{array}{c}
+1,\\
+-1,\end{array} & \begin{array}{c}
+a_{d,i_{l,c}}>0\\
+otherwise\end{array}\end{cases}$
 \end{enumerate}
@@
 \end{enumerate}
-We will employ a data augmentation scheme with auxiliary variables
-$a_{i}$ in the probit model where:
-
-\begin{equation}
-y_{i}\sim\begin{cases}
-\begin{array}{c}
-1,\\
-0,\end{array} & \begin{array}{c}
-a_{i}>0\; and\; y_{parent}=1\\
-otherwise\end{array}\end{cases}\end{equation}
-
-\begin{equation}
-a_{i}\sim\mathcal{N}\left(\eta_{i}^{T}\bar{z},1\right)\end{equation}
+\end{enumerate}
+The generative model for the ICD-9 codes is equivalent to a probit
+regression model. In our case, the regression is conditional on the
+parents according to the constraints of the ICD-9 code tree. The latent
+variable utilized here is also known as an auxiliary variable.
 
@@
 Given an observation of a set of ICD-9 codes and a document, the posterior
-distribution for the latent variables is given by Equation 4.
+distribution for the latent variables is given by Equation .
+
+\begin{equation}
+p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
+
+\[
+=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]
 
@@
 to this posterior. In particular, we will derive a collapsed Gibbs
-sampler for the supervised topic model.
-
-\subsection{Rao-Blackwellization}
-
-To derive the Gibbs sampler in general, we integrate over the parameters
-$\theta$ and $\phi_{1:K}$ resulting in the collapsed joint distribution
-shown in Equations 5-8.
-
-\begin{landscape} \centering % optional, probably makes it look better to have it centered on the page
-
-\begin{small}
-
-\begin{equation}
-p\left(\theta,z_{1:N}\mid w_{1:N},y_{1:I},\phi_{1:K},\eta_{1:I},\alpha,\beta,\mu,\xi\right)=\frac{p\left(\theta\mid\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)}{\int_{\theta}p\left(\theta\mid\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)d\theta}\end{equation}
-
-\begin{equation}
-\intop_{\theta}\intop_{\phi_{1:K}}p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\theta},\mathbf{\phi},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)d\theta d\phi_{1:K}=\intop_{\theta}\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\left\{ p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}\left[p\left(z_{m,n}\mid\theta_{m}\right)p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)\right]\prod_{i=1}^{I}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right\} \prod_{i=1}^{I}p\left(\eta_{i};\mu\right)d\theta d\phi_{1:K}\end{equation}
-
-\begin{equation}
-p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\prod_{n=1}^{N}p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)d\phi_{1:K}\intop_{\theta}\prod_{m=1}^{M}p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}p\left(z_{m,n}\mid\theta_{m}\right)d\theta\end{equation}
-
-\begin{equation}
-=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}
-
-\begin{equation}
-=\prod_{i=1}^{I}\left[\mathcal{N}\left(\eta_{i}\mid\mu,1\right)\prod_{m=1}^{M}h\left(y\right)\exp\left\{ \left(\eta_{i}^{T}\bar{z}\right)y-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}
-
-\end{small} \end{landscape}
+sampler.
 
@@
 
-To derive the Gibbs sampler we evaluate the individual conditional
-probability distributions for all unobserved variables.
+We derive a collapsed Gibbs sampler for this model, marginalizing
+$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
+in LDA models see \citet{Griffiths04}. To derive the Gibbs sampler
+we evaluate the individual conditional probability distributions for
+all unobserved variables.
 
-\subsubsection{$p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)$}
+\subsubsection{$p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}
 
@@
 of the joint distribution isolating a particular latent variable,
-$z$, for a word instance, $n$, in a document instance, $m$. The
+$z$, for a word instance, $n$, in a document instance, $d$. The
 conditional probability with respect to this latent variable is proportional
-to the joint distribution up to a constant.
+to the joint distribution of its markov blanket up to a constant.
 
 \begin{equation}
-p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)\propto p\left(z_{m,n},\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)\end{equation}
+p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
 
-Due to the factorization of this model we can rewrite the joint distribution
+Due to the factorization of this model, we can rewrite the joint distribution
 as the following:
@@
 \begin{equation}
-\propto\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)p\left(y_{m,i}\mid z_{m,1:N},\eta_{i}\xi\right)\right]p\left(z_{m,n},\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\end{equation}
+\propto\prod_{i_{l,c}\in\mathcal{I}}p\left(a_{i_{l,c}}\mid\mathbf{z},\eta_{i_{l,c}}\right)p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right)\end{equation}
 
@@
 \begin{equation}
-\propto\prod_{i=1}^{I}\left[p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\cdot\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}
+\propto\prod_{i_{l,c}\in\mathcal{I}}\exp\left\{ -\frac{\left(\bar{z}^{T}\eta_{i_{l,c}}-a_{i_{l,c}}\right)^{2}}{2}\right\} p\left(a_{i_{l,c}}\mid\eta_{i_{l,c}}\right)\left(n_{d,\left(\cdot\right)}^{k,-\left(d,n\right)}+\alpha\beta_{k}\right)\frac{n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma\right)}\label{eq:z-likelihood}\end{equation}
 
-Here, $n_{m,v}^{k,-\left(m,n\right)}$ represents the count of word
-$v$ in document $m$ assigned to topic $k$ omitting the $(m,n)^{th}$
-word count. For exponential family distributions, the normalization
-contant, $h(y_{m,i})$, does not depend on $z_{m,n}$.
+Here, $n_{d,v}^{k,-\left(d,n\right)}$ represents the count of word
+$v$ in document $d$ assigned to topic $k$ omitting the $(d,n)^{th}$
+word count.
+
+Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$
+can be sampled through enumeration.
+
+\subsubsection{$p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)$}
+
+Given that $\eta_{i_{l,c}}$ and $a_{d,i_{l,c}}$ are distributed
+normally, this posterior distribution is also normal. We evaluated
+the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
 
 \begin{equation}
-\propto\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\label{eq:z_likelihood}\end{equation}
+p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
 
-Given this expression, $p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)$
-can be sampled through enumeration as seen in Equation 13. In the
-case of probit regression, the expression for \ref{eq:z_likelihood}
-evaluates to Equation 14. Equivalently we can parameterize the model
-with an auxiliary variable $a_{i}$, resulting in Equation 15.
+\[
+\hat{\mu}_{i}=\hat{\mathbf{\Sigma}}_{i}\left(-\mathbf{1}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{\left(\cdot\right),i_{l,c}}\right)\]
 
-%\begin{equation}
-%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)=\frac{\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}{\sum_{k=1}^{K}\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}\end{equation}
+\[
+\hat{\Sigma}_{i}^{-1}=\mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\]
 
-%\begin{equation}
-%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\Phi\left(\eta_{i}^{T}\bar{z}\right)^{y_{m,i}}\left(1-\Phi\left(\eta_{i}^{T}\bar{z}\right)\right)^{1-y_{m,i}}\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}
+\subsubsection{$p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$
+and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
 
-%\begin{equation}
-%p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\mathbf{a},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\delta\left(sign\left(a_{m,i}\right)=y_{m,i}\right)\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}
-
-\begin{landscape}
-
-\centering % optional, probably makes it look better to have it centered on the page
-\begin{equation}
-p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{Y},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)=\frac{\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}{\sum_{k=1}^{K}\exp\left\{ \sum_{i=1}^{I}\left(\eta_{i}^{T}\bar{z}\right)y_{m,i}-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}}\end{equation}
+In the augmented probit regression model, the posterior distribution
+of $a_{i_{l,c}}$ is distributed according to a truncated normal distribution
+where the response variable is observed.
 
 \begin{equation}
-p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\Phi\left(\eta_{i}^{T}\bar{z}\right)^{y_{m,i}}\left(1-\Phi\left(\eta_{i}^{T}\bar{z}\right)\right)^{1-y_{m,i}}\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}
+p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)=\begin{cases}
+trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=1\\
+trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=-1\end{cases}\end{equation}
+
+However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
+be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
+chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
+a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
+-1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
+$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
+since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
+$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
+to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
+sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
 
 \begin{equation}
-p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{w},\mathbf{\eta},\mathbf{a},\alpha,\beta,\mu\right)\propto\prod_{i=1}^{I}\left[\delta\left(sign\left(a_{m,i}\right)=2y_{m,i}-1\right)\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1\right)\right]\left(n_{m,\left(\bullet\right)}^{k,-\left(m,n\right)}+\alpha_{k}\right)\frac{n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,w_{m,n}}}{\sum_{v=1}^{V}\left(n_{\left(\bullet\right),w_{m,n}}^{k,-\left(m,n\right)}+\beta_{k,v}\right)}\end{equation}
-
-\end{landscape}
-
-\subsubsection{$p\left(\eta_{i}\mid\mathbf{z},\mathbf{Y},\mu\right)$ or $p\left(\eta_{i}\mid\mathbf{z},\mathbf{a},\mu\right)$in
-the augmented probit regression model}
-
-Given that $\eta_{i}$ and $a_{m,i}$ are distributed normally, this
-posterior distribution is also normal. In the case for general exponential
-family distributions, $\eta_{i}$ can remain a parameter without a
-prior, fit with maximimum likelihood in the usual fashion.
+p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
 
 \begin{equation}
-p\left(\eta_{i}\mid\mathbf{z},\mathbf{a},\mu\right)=\mathcal{N}\left(\eta_{i}\mid\hat{\mathbf{\mu}_{i}},\hat{\mathbf{S}_{i}}\right)\end{equation}
+p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
+
+\begin{equation}
+p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
+1, & y_{parent_{l,c}}=-1\\
+0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
 
 \[
-\hat{\mathbf{\mu}_{i}}=\hat{\mathbf{S}_{i}}\bar{\mathbf{Z}}^{T}\mathbf{a}_{\left(\bullet\right),i}\]
-
-\[
-\hat{\mathbf{S}_{i}}^{-1}=\mathbf{I}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\]
-
-\subsubsection{$p\left(a_{m,i}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$ in
-the augmented probit regression model}
-
-In the augmented probit regression model, the posterior distribution
-of $a_{i}$ is distributed according to a truncated normal distribution.
-
-\[
-p\left(a_{m,i}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)=trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)\]
-
-\subsubsection{\textmd{$p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)$}}
-
-In our model, response variables are not always observed and are treated
-as latent and sampled where appropriate. There are two factors influencing
-predictions of the response variable, $y_{m,i}$. There is an undirected
-model enforcing the aforementioned constraints and providing a prior
-and there is the probit regression.
+p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
 
 \begin{equation}
-p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)\propto\psi\left(\mathbf{Y}\right)\delta\left(sign\left(a_{m,i}\right)=y_{m,i}\right)\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1\right)\end{equation}
+=\begin{cases}
+\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
+trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right),\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
+trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
+0 & otherwise\end{cases}\end{equation}
+
+where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
+variables excluding the response variable being sampled.
+
+\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha',\alpha\right)$}
+
+In our model, we place a hierarchical Dirichlet prior over topic assignments.
+This prior shares many features with the hierarchical Dirichlet process
+and inference over this distribution proceeds in a very similar fashion.
+This flexible distribution allows for an asymmetric prior over document
+level distributions over topics \citet{WallachMiMc2009}.
+
+Posterior inference was performed using the {}``direct assignment''
+method of \citet{TehJorBea2006}.
 
 \begin{equation}
-p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)\propto\psi\left(\mathbf{Y}\right)trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)\end{equation}
-
-Again, this conditional distribution can be evaluated through enumerations
-and normalization.
+\beta\sim Dir\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)\end{equation}
 
 \begin{equation}
-p\left(y_{m,i}\mid\mathbf{\eta},\mathbf{a},\xi\right)=\frac{\psi\left(\mathbf{Y}\right)trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)}{\sum_{y_{m,i}}\psi\left(\mathbf{Y}\right)trunc\mathcal{N}\left(a_{m,i}\mid\eta_{i}^{T}\bar{z,}1,y_{m,i}\right)}\end{equation}
+p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}
+
+where $s\left(n,m\right)$ represents stirling numbers of the first
+kind.
+
+\subsubsection{$p\left(\alpha;\lambda\right)$, $p\left(\alpha';\lambda\right)$,
+$p\left(\beta;\lambda\right)$}
+
+All hyperparameters were given broad gamma priors $\left(\lambda=\left\{ shape=2,scale=1000\right\} \right)$
+and sampled via the Metropolis-Hastings algorithm.
+
+\subsection{Prediction}
+
+\subsection{Related Work}
+
+Latent dirichlet allocation (LDA) is a generative probabilistic model
+of corpora that represents documents as a mixed membership bag-of-words.
+Also known as topic models, these models infer the latent structure,
+or topics, of documents in a corpus. Each document is represented
+as a collection of words, generated from a set of topic assignments
+(one for each word), where each topic assignment is drawn from a distribution
+over topics \citep{Blei2003}.
+
+This model, supervised latent dirichlet allocation (sLDA), builds
+on LDA by incorporating an exponential family response variable. Although
+there are many models for making predictions based on free text, sLDA
+is unique in that it is a generative model, it represents documents
+as a mixed-membership, and constrains the inference of the latent
+structure of the documents by its predictability of the response variable.
+In other words, sLDA infers topics such that the model is capable
+of a high predictive likelihood for words in a document and the response
+variable associated with a document. This approach has been shown
+to outperform both LASSO (L1 regularized least squares regression)
+and LDA followed by least squares regression \citep{BleiMcAuliffe2008}.
+
+Other models - predicting document links, other supervised latent
+variable models
 
@@
 \begin{small} \bibliographystyle{plainnat} \bibliographystyle{plainnat}
+\bibliographystyle{plainnat}
 \bibliography{refs}
+
 \end{small}
+
+\newpage{}
+
+\subsection{Rao-Blackwellization}
+
+To derive the Gibbs sampler in general, we integrate over the parameters
+$\theta$ and $\phi_{1:K}$ resulting in the collapsed joint distribution
+shown in Equations 5-8.
+
+\begin{landscape} \centering % optional, probably makes it look better to have it centered on the page
+
+\begin{small}
+
+\begin{equation}
+p\left(\theta,z_{1:N}\mid w_{1:N},y_{1:I},\phi_{1:K},\eta_{1:I},\alpha,\beta,\mu,\xi\right)=\frac{p\left(\theta\mid\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)}{\int_{\theta}p\left(\theta\mid\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)d\theta}\end{equation}
+
+\begin{equation}
+\intop_{\theta}\intop_{\phi_{1:K}}p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\theta},\mathbf{\phi},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)d\theta d\phi_{1:K}=\intop_{\theta}\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\left\{ p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}\left[p\left(z_{m,n}\mid\theta_{m}\right)p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)\right]\prod_{i=1}^{I}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right\} \prod_{i=1}^{I}p\left(\eta_{i};\mu\right)d\theta d\phi_{1:K}\end{equation}
+
+\begin{equation}
+p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\prod_{n=1}^{N}p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)d\phi_{1:K}\intop_{\theta}\prod_{m=1}^{M}p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}p\left(z_{m,n}\mid\theta_{m}\right)d\theta\end{equation}
+
+\begin{equation}
+=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}\end{equation}
+
+\begin{equation}
+=\prod_{i=1}^{I}\left[\mathcal{N}\left(\eta_{i}\mid\mu,1\right)\prod_{m=1}^{M}h\left(y\right)\exp\left\{ \left(\eta_{i}^{T}\bar{z}\right)y-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}
+
+\end{small} \end{landscape}
 \end{document}
+
+%\begin{align}
+%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right) & =\\
+% & \frac{p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)p\left(\theta\mid\beta,\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}{\int_{\alpha,\alpha',\gamma}p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)\int_{\theta}p\left(\theta\mid\beta,\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}\end{align}

### 2011-05-12 paper/main.tex
Rearranged some stuff and started to NIPS-ify

@@
 
-\title{Supervised Topic Modeling in Clinical Text}
+\title{Hierarchically Supervised Latent Dirichlet Allocation}
+%Supervised Topic Modeling in Clinical Text}
 
 \author{
-Perotte A\hspace{1cm} Li Y \hspace{1cm} Pivovarov R \hspace{1cm} Weiskopf N\hspace{1cm} Wood F\\
+Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhaddad \hspace{1cm} Frank Wood\\
 Columbia University, New York, NY 10027, USA \\
-\texttt{\{\{ajp9009, yil7003, rip7002, ngw7001\}@dbmi., fwood@stat.\}columbia.edu}
+\texttt{\{ajp9009@dbmi,bartlett@stat,noemie@dbmi,fwood@stat\}.columbia.edu}
 %\texttt{pfau@neurotheory.columbia.edu} 
@@
 \begin{abstract}
-Current medical record keeping technology relies heavily upon human
-capacity to effectively summarize and infer information from free-text
-physician notes. We propose a novel method to suggest diagnostic code
-assignment for patient visits, based upon narrative medical notes.
-We applied a supervised latent Dirichlet allocation model to a corpora
-of free-text medical notes from NewYork - Presbyterian Hospital to
-infer a set of specific ICD-9 codes for each patient note. Evaluation
-of the predictions were conducted by comparison to a gold-standard
-set of ICD-9s assigned to a set of patient notes. 
+The benefits of supervision in topic modeling 
+%Current medical record keeping technology relies heavily upon human
+%capacity to effectively summarize and infer information from free-text
+%physician notes. We propose a novel method to suggest diagnostic code
+%assignment for patient visits, based upon narrative medical notes.
+%We applied a supervised latent Dirichlet allocation model to a corpora
+%of free-text medical notes from NewYork - Presbyterian Hospital to
+%infer a set of specific ICD-9 codes for each patient note. Evaluation
+%of the predictions were conducted by comparison to a gold-standard
+%set of ICD-9s assigned to a set of patient notes. 
 \end{abstract}
@@
 
-\label{sec:introduction} Despite the growing emphasis on meaningful
-use of technology in medicine, many aspects of medical record-keeping
-remain a manual process. Diagnostic coding for billing and insurance
-purposes is often handled by professional medical coders who must
-explore a patient's extensive clinical record before assigning the
-proper codes. So while electronic health records (EHRs) should be
-adopted by most medical institutions within the next several years,
-largely due to the provisions of HITECH under the American Recovery
-and Reinvestment Act \citep{Blumenthal2009}, there has been little
-movement forward in automating medical coding.
+\label{sec:introduction}
+
+\begin{itemize}
+\item Benefits of combining human categorization information into ``topic models''
+\item 
+\end{itemize}
+
+% Informatics journal paper
+%
+% Despite the growing emphasis on meaningful
+%use of technology in medicine, many aspects of medical record-keeping
+%remain a manual process. Diagnostic coding for billing and insurance
+%purposes is often handled by professional medical coders who must
+%explore a patient's extensive clinical record before assigning the
+%proper codes. So while electronic health records (EHRs) should be
+%adopted by most medical institutions within the next several years,
+%largely due to the provisions of HITECH under the American Recovery
+%and Reinvestment Act \citep{Blumenthal2009}, there has been little
+%movement forward in automating medical coding.
 
@@
 diagnoses \citep{Surjan1999}.
-
-While there have been some attempts to automatically classify groups
-of patients as a potential preliminary step to ICD-9 code assignment
-\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
-automatic assignment of ICD-9 codes to medical text became a more
-prevalent research topic only in the last few years. A subset of earlier
-work proposed various methods on small corpora, based on a few specific
-diseases \citep{Rao2003} but the most recent and promising work on
-the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
-Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
-(website). Most of the classification strategies included word matching
-and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
-The data set given to the participants consisted only of documents
-that were 1-2 lines each and all of the documents were radiology reports
-- clearly limiting the scope of potential ICD-9 codes which could
-be assigned. The only paper which has attempted to work with a document
-scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
-Lita proposed support vector machine and bayesian ridge regression
-methods to assign appropriate labels to the documents but did not
-utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
 
@@
 
-\subsection{Data}
+\section{Methods}
 
-Our data set was gathered from the clinical data warehouse of NewYork
-- Presbyterian Hospital. The data consisted of free-text discharge
-summaries and their respective ICD-9 codes. A discharge summary is
-a clinical report prepared by a physician or other health professional
-at the conclusion of a hospital stay or series of treatments. The
-note outlines the patient\textquoteright{}s chief complaint, diagnostic
-findings, therapy administered, patient\textquoteright{}s response
-to the chosen therapy, the treatment plan and the recommendations
-upon discharge. The ICD-9 codes used to structure the discharge summary
-data are part of a controlled terminology which is the international
-standard diagnostic classification for epidemiological, health management,
-and clinical purposes (http://www.who.int/classifications/icd/en/).
-The codes are classified in a rooted-tree structure, with each edge
-representing an is-a relationship between parent and child, such that
-the parent diagnosis subsumes the child diagnosis. In the hospital,
-ICD-9 codes are generated manually by trained medical coders, who
-review all the information in the discharge summary. For the purposes
-of sLDA, ICD-9 codes will be used as labels for discharge summaries.
-
-We worked within the guidelines of the Health Insurance Portability
-and Accountability Act (HIPAA), which protects patient privacy and
-the security of potentially identifying medical material, known as
-personal health information (PHI). HIPAA covers any information within
-a medical record that was created, used, or disclosed during the course
-of providing a health care service and that can be used to identify
-an individual. Before beginning data processing, we generated a PHI-free
-dataset (see Data Pre-processing below).
-
-\subsection{Pre-Processing}
-
-Patient discharge summaries and their associated ICD-9 diagnoses are
-stored in two different places in the NewYork - Presbyterian data
-warehouse, and so had to be linked together before being fed into
-the sLDA algorithm. Each discharge note and set of diagnoses were
-assigned a patient unique identifier (PUID) and a visit unique identifier
-(VUID), allowing the two types of data to be linked.
-
-Natural Language Processsing (NLP) techniques were used to process
-the free-text discharge summaries. First, the Natural Language Toolkit
-(http://www.nltk.org/) was used to tokenize the text. Next, feature
-selection was performed using a term frequency - inverse document
-frequency (TF-IDF) algorithm on the entire document set and sorting
-the words by their TF-IDF values. The top 10,000 words were manually
-evaluated to eliminate all potentially identifying information. Finally,
-each discharge summary was converted to a bag-of-words, listing the
-frequencies of the remaining, free of protected health information,
-top 10,000 words.
-
-Preparation of the diagnostic codes involved inference over the ICD-9
-hierarchy. The is-a relationships of the hierarchy allowed us to make
-two important assumptions. First, if a diagnosis was observed to be
-present, all of its ancestors could be assumed to be present as well
-(e.g., if a patient had malignant hypertension, it could be assumed
-that they also had essential hypertension. Second, if a diagnosis
-was observed to be absent, it could be assumed that all of its descendants
-were also absent (e.g. if a patient did not have essential hypertension,
-it could be assumed that they did not have malignant hypertension).
-Unfortunately, ICD-9 code observations never include observations
-of disease absence. ICD-9 codes are only documented when the condition
-is observed to be present. Additionally, ICD-9 codes are known to
-have relatively low sensitivity; conditions that are present are often
-not documented in a set of ICD-9 codes (Surjan, 1999). Given these
-facts, we made the following assumptions regarding each visit: recorded
-diagnoses and their ancestors were labeled as true; diagnoses that
-were observed at some time for a patient but not at the current visit
-were labeled as unobserved; and diagnoses that had never been listed
-for a patient were labeled as false for all of that patient\textquoteright{}s
-visits. This last assumption captures the belief that parts of the
-ICD-9 hierarchy that are never observed for a particular patient are
-almost certain to be false. Additionally, for computational purposes,
-we decided not to include an ICD-9 code at all if neither it nor one
-of its descendants had been assigned to a patient in any of the records
-in our dataset.
-
-\subsection{ICD-9 Code Hierarchy}
-
-Here, we augment the sLDA model such that the supervised signal is
-distribution over the ICD-9 code tree, which is an is-a hierarchy
-\citep{ICD9NCBO}. An is-a hierarchy is represented by the tree data
-structure where each node has only a single parent and nodes cannot
-be parents of ancestors (ie. there are no loops). In this particular
-case, the ICD-9 code hierarchy is also partially a prefix trie where
-the labels for certain nodes are prefixes for child nodes. Given that
-this rule does not apply to all nodes in the hierarchy, we did not
-use this feature to determine the structure of the hierarchy. Instead
-we acquired a dataset that explicitly defined the relationships between
-the nodes of the hierarchy \citep{ICD9NCBO}. In documentation of
-ICD-9 codes for billing purposes, only a subset of the nodes can be
-used, however the nodes higher in the hierarchy contain semantic information
-about the categories of codes that are their descendants. For this
-reason, we included these nodes in our model.
-
-\section{Methods}
+% generalize to arbitrary codes (not just ICD-9's
 
@@
 
+\section{Experiments}
+
+\subsection{Data}
+
+Our data set was gathered from the clinical data warehouse of NewYork
+- Presbyterian Hospital. The data consisted of free-text discharge
+summaries and their respective ICD-9 codes. A discharge summary is
+a clinical report prepared by a physician or other health professional
+at the conclusion of a hospital stay or series of treatments. The
+note outlines the patient\textquoteright{}s chief complaint, diagnostic
+findings, therapy administered, patient\textquoteright{}s response
+to the chosen therapy, the treatment plan and the recommendations
+upon discharge. The ICD-9 codes used to structure the discharge summary
+data are part of a controlled terminology which is the international
+standard diagnostic classification for epidemiological, health management,
+and clinical purposes (http://www.who.int/classifications/icd/en/).
+The codes are classified in a rooted-tree structure, with each edge
+representing an is-a relationship between parent and child, such that
+the parent diagnosis subsumes the child diagnosis. In the hospital,
+ICD-9 codes are generated manually by trained medical coders, who
+review all the information in the discharge summary. For the purposes
+of sLDA, ICD-9 codes will be used as labels for discharge summaries.
+
+We worked within the guidelines of the Health Insurance Portability
+and Accountability Act (HIPAA), which protects patient privacy and
+the security of potentially identifying medical material, known as
+personal health information (PHI). HIPAA covers any information within
+a medical record that was created, used, or disclosed during the course
+of providing a health care service and that can be used to identify
+an individual. Before beginning data processing, we generated a PHI-free
+dataset (see Data Pre-processing below).
+
+\subsection{Pre-Processing}
+
+Patient discharge summaries and their associated ICD-9 diagnoses are
+stored in two different places in the NewYork - Presbyterian data
+warehouse, and so had to be linked together before being fed into
+the sLDA algorithm. Each discharge note and set of diagnoses were
+assigned a patient unique identifier (PUID) and a visit unique identifier
+(VUID), allowing the two types of data to be linked.
+
+Natural Language Processsing (NLP) techniques were used to process
+the free-text discharge summaries. First, the Natural Language Toolkit
+(http://www.nltk.org/) was used to tokenize the text. Next, feature
+selection was performed using a term frequency - inverse document
+frequency (TF-IDF) algorithm on the entire document set and sorting
+the words by their TF-IDF values. The top 10,000 words were manually
+evaluated to eliminate all potentially identifying information. Finally,
+each discharge summary was converted to a bag-of-words, listing the
+frequencies of the remaining, free of protected health information,
+top 10,000 words.
+
+Preparation of the diagnostic codes involved inference over the ICD-9
+hierarchy. The is-a relationships of the hierarchy allowed us to make
+two important assumptions. First, if a diagnosis was observed to be
+present, all of its ancestors could be assumed to be present as well
+(e.g., if a patient had malignant hypertension, it could be assumed
+that they also had essential hypertension. Second, if a diagnosis
+was observed to be absent, it could be assumed that all of its descendants
+were also absent (e.g. if a patient did not have essential hypertension,
+it could be assumed that they did not have malignant hypertension).
+Unfortunately, ICD-9 code observations never include observations
+of disease absence. ICD-9 codes are only documented when the condition
+is observed to be present. Additionally, ICD-9 codes are known to
+have relatively low sensitivity; conditions that are present are often
+not documented in a set of ICD-9 codes (Surjan, 1999). Given these
+facts, we made the following assumptions regarding each visit: recorded
+diagnoses and their ancestors were labeled as true; diagnoses that
+were observed at some time for a patient but not at the current visit
+were labeled as unobserved; and diagnoses that had never been listed
+for a patient were labeled as false for all of that patient\textquoteright{}s
+visits. This last assumption captures the belief that parts of the
+ICD-9 hierarchy that are never observed for a particular patient are
+almost certain to be false. Additionally, for computational purposes,
+we decided not to include an ICD-9 code at all if neither it nor one
+of its descendants had been assigned to a patient in any of the records
+in our dataset.
+
+\subsection{ICD-9 Code Hierarchy}
+
+Here, we augment the sLDA model such that the supervised signal is
+distribution over the ICD-9 code tree, which is an is-a hierarchy
+\citep{ICD9NCBO}. An is-a hierarchy is represented by the tree data
+structure where each node has only a single parent and nodes cannot
+be parents of ancestors (ie. there are no loops). In this particular
+case, the ICD-9 code hierarchy is also partially a prefix trie where
+the labels for certain nodes are prefixes for child nodes. Given that
+this rule does not apply to all nodes in the hierarchy, we did not
+use this feature to determine the structure of the hierarchy. Instead
+we acquired a dataset that explicitly defined the relationships between
+the nodes of the hierarchy \citep{ICD9NCBO}. In documentation of
+ICD-9 codes for billing purposes, only a subset of the nodes can be
+used, however the nodes higher in the hierarchy contain semantic information
+about the categories of codes that are their descendants. For this
+reason, we included these nodes in our model.
+
 \subsection{Related Work}
@@
 variable models
+
+% move to related work?
+While there have been some attempts to automatically classify groups
+of patients as a potential preliminary step to ICD-9 code assignment
+\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
+automatic assignment of ICD-9 codes to medical text became a more
+prevalent research topic only in the last few years. A subset of earlier
+work proposed various methods on small corpora, based on a few specific
+diseases \citep{Rao2003} but the most recent and promising work on
+the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
+Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
+(website). Most of the classification strategies included word matching
+and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
+The data set given to the participants consisted only of documents
+that were 1-2 lines each and all of the documents were radiology reports
+- clearly limiting the scope of potential ICD-9 codes which could
+be assigned. The only paper which has attempted to work with a document
+scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
+Lita proposed support vector machine and bayesian ridge regression
+methods to assign appropriate labels to the documents but did not
+utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
 

### 2011-05-13 paper/main.tex
Minor changes

@@
 \author{
-Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhaddad \hspace{1cm} Frank Wood\\
+Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhadad \hspace{1cm} Frank Wood\\
 Columbia University, New York, NY 10027, USA \\
@@
 \mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
-trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right),\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
+trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
 trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\

### 2011-05-19 paper/main.tex
Minor equation fixes
updated main.tex with comments
added a few lines to the intro

@@
 
+There exist surprisingly many sources of unstructured text data that have been partially or completely categorized by human editors.
+Examples include hierarchical directories of webpages \citep{DMOZ}, large hierarchically annotated product catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), manually annotated patient medical records, and many more.
+In this work we show how to combine these two sources of information in a single model that allows us to, amongst other things, automatically annotate and/or categorize new text documents, effectively inserting them into the category
+
 \begin{itemize}
 \item Benefits of combining human categorization information into ``topic models''
-\item 
+\item LDA solved free text
+\item supervised LDA improves LDA (extra info) and allows new inference (predict links, etc.)
+\item amazon, freshdirect, netflix, dmoz, pandora (music genome)
 \end{itemize}
@@
 \begin{equation}
-\propto\prod_{i_{l,c}\in\mathcal{I}}\exp\left\{ -\frac{\left(\bar{z}^{T}\eta_{i_{l,c}}-a_{i_{l,c}}\right)^{2}}{2}\right\} p\left(a_{i_{l,c}}\mid\eta_{i_{l,c}}\right)\left(n_{d,\left(\cdot\right)}^{k,-\left(d,n\right)}+\alpha\beta_{k}\right)\frac{n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma\right)}\label{eq:z-likelihood}\end{equation}
+\propto\prod_{i_{l,c}\in\mathcal{I}}\exp\left\{ -\frac{\left(\bar{z}^{T}\eta_{i_{l,c}}-a_{i_{l,c}}\right)^{2}}{2}\right\} \left(n_{d,\left(\cdot\right)}^{k,-\left(d,n\right)}+\alpha\beta_{k}\right)\frac{n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma\right)}\label{eq:z-likelihood}\end{equation}
 

### 2011-05-22 paper/introduction.tex
moved two images around (renamed) and created a separate introduction file

--- 2011-05-19 paper/main.tex
@@
-\documentclass{article}
-\usepackage{nips10submit_e,times}
-%\documentstyle[nips07submit_09,times]{article}
-\usepackage[square,numbers]{natbib}
-\usepackage{amsmath, epsfig}
-\usepackage{amsfonts}
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage{amsfonts}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{easybmat}
-\usepackage{footmisc}
-\usepackage{lscape}
-\renewcommand\algorithmiccomment[1]{// \textit{#1}}
-%
-\newcommand{\ignore}[1]{}
-\newcommand{\comment}[1]{}
-\DeclareMathOperator*{\argmax}{arg\,max}
-
-\title{Hierarchically Supervised Latent Dirichlet Allocation}
-%Supervised Topic Modeling in Clinical Text}
-
-\author{
-Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhadad \hspace{1cm} Frank Wood\\
-Columbia University, New York, NY 10027, USA \\
-\texttt{\{ajp9009@dbmi,bartlett@stat,noemie@dbmi,fwood@stat\}.columbia.edu}
-%\texttt{pfau@neurotheory.columbia.edu} 
-%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
-}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to \LaTeX{} to determine where to break
-% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
-% puts 3 of 4 authors names on the first line, and the last on the second
-% line, try using \AND instead of \And before the third author name.
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-\newcommand{\X}{\mathcal{X}}
-
-\nipsfinalcopy
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-The benefits of supervision in topic modeling 
-%Current medical record keeping technology relies heavily upon human
-%capacity to effectively summarize and infer information from free-text
-%physician notes. We propose a novel method to suggest diagnostic code
-%assignment for patient visits, based upon narrative medical notes.
-%We applied a supervised latent Dirichlet allocation model to a corpora
-%of free-text medical notes from NewYork - Presbyterian Hospital to
-%infer a set of specific ICD-9 codes for each patient note. Evaluation
-%of the predictions were conducted by comparison to a gold-standard
-%set of ICD-9s assigned to a set of patient notes. 
-\end{abstract}
-
-\section{Introduction}
-
-\label{sec:introduction}
-
 There exist surprisingly many sources of unstructured text data that have been partially or completely categorized by human editors.
@@
 dozen topics.
-
-\section{Background}
-
-\section{Methods}
-
-% generalize to arbitrary codes (not just ICD-9's
-
-%
-\begin{figure}[h]
-%tbp] %  figure placement: here, top, bottom, or page
- \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}
-
-\label{fig:example} 
-\end{figure}
-
-Given the number of topics, K, and broad gamma priors on hyperparameters,
-the generative process is as follows: 
-\begin{enumerate}
-\item For each topic, $k$:
-
-\begin{enumerate}
-\item Draw a distribution over words $\phi_{k}\sim Dir_{V}\left(\mathbf{1},\gamma\right)$ 
-\end{enumerate}
-\item For each ICD9 code, $c$, at all levels in the tree, $l$:
-
-\begin{enumerate}
-\item Draw regression coefficient $\eta_{i_{l,c}}\mid\sigma\sim\mathcal{N}_{K}\left(-\mathbf{1},\sigma\right)$ 
-\end{enumerate}
-\item Draw a prior over topic proportions $\beta\mid\alpha'\sim Dir_{K}\left(\mathbf{1},\alpha'\right)$ 
-\item For each document, $d$:
-
-\begin{enumerate}
-\item Draw topic proportions $\theta_{d}\mid\beta,\alpha\sim Dir_{K}\left(\beta,\alpha\right)$ 
-\item For each word, $n$:
-
-\begin{enumerate}
-\item Draw topic assignment $z_{n,d}\mid\theta_{d}\sim Mult_{K}\left(\theta_{d}\right)$ 
-\item Draw word $w_{n,d}\mid z_{n,d},\beta_{1:K}\sim Mult_{V}\left(\beta_{z_{n}}\right)$ 
-\end{enumerate}
-\item For each level of the ICD-9 code tree, $l$:
-
-\begin{enumerate}
-\item For each ICD-9 code at this level, $c$:
-
-\begin{enumerate}
-\item Draw a latent variable \[
-a_{d,i_{l,c}}\sim\begin{cases}
-\mathcal{N}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=1\\
-trunc\mathcal{N}^{-}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=-1\end{cases}\]
- where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\mathcal{I}=\left\{ i_{0},i_{1},...,i_{\mathcal{L}}\right\} $
-and $i_{l}=\left\{ i_{l,0},i_{l,1},...,i_{l,C_{l}}\right\} $
-\item Draw a response variable $y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\sim\begin{cases}
-\begin{array}{c}
-1,\\
--1,\end{array} & \begin{array}{c}
-a_{d,i_{l,c}}>0\\
-otherwise\end{array}\end{cases}$
-\end{enumerate}
-\end{enumerate}
-\end{enumerate}
-\end{enumerate}
-The generative model for the ICD-9 codes is equivalent to a probit
-regression model. In our case, the regression is conditional on the
-parents according to the constraints of the ICD-9 code tree. The latent
-variable utilized here is also known as an auxiliary variable.
-
-\subsection{Posterior Inference}
-
-Given an observation of a set of ICD-9 codes and a document, the posterior
-distribution for the latent variables is given by Equation .
-
-\begin{equation}
-p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
-
-\[
-=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]
-
-The denominator for this distribution is the marginal probability
-of the data and cannot be solved in closed form. This is often the
-case in evaluating posterior distributions of non-trivial probabilistic
-models. We will appeal to one of the common methods for approximating
-posterior distributions in the face of intractable normalization factors:
-Markov chain Monte Carlo (MCMC) sampling. Since in this model it is
-possible to sample from the conditional distributions for all variables
-we will use the Gibbs sampling algorithm to obtain an approximation
-to this posterior. In particular, we will derive a collapsed Gibbs
-sampler.
-
-\subsection{Gibbs Sampling}
-
-We derive a collapsed Gibbs sampler for this model, marginalizing
-$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
-in LDA models see \citet{Griffiths04}. To derive the Gibbs sampler
-we evaluate the individual conditional probability distributions for
-all unobserved variables.
-
-\subsubsection{$p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}
-
-For the purposes of sampling, we will be able to derive a representation
-of the joint distribution isolating a particular latent variable,
-$z$, for a word instance, $n$, in a document instance, $d$. The
-conditional probability with respect to this latent variable is proportional
-to the joint distribution of its markov blanket up to a constant.
-
-\begin{equation}
-p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
-
-Due to the factorization of this model, we can rewrite the joint distribution
-as the following:
-
-\begin{equation}
-\propto\prod_{i_{l,c}\in\mathcal{I}}p\left(a_{i_{l,c}}\mid\mathbf{z},\eta_{i_{l,c}}\right)p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right)\end{equation}
-
-We isolate only terms that depend on $z_{m,n}$ and absorb all other
-constant terms into the normalization constant \citep{Griffiths04}.
-
-\begin{equation}
-\propto\prod_{i_{l,c}\in\mathcal{I}}\exp\left\{ -\frac{\left(\bar{z}^{T}\eta_{i_{l,c}}-a_{i_{l,c}}\right)^{2}}{2}\right\} \left(n_{d,\left(\cdot\right)}^{k,-\left(d,n\right)}+\alpha\beta_{k}\right)\frac{n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma\right)}\label{eq:z-likelihood}\end{equation}
-
-Here, $n_{d,v}^{k,-\left(d,n\right)}$ represents the count of word
-$v$ in document $d$ assigned to topic $k$ omitting the $(d,n)^{th}$
-word count.
-
-Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$
-can be sampled through enumeration.
-
-\subsubsection{$p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)$}
-
-Given that $\eta_{i_{l,c}}$ and $a_{d,i_{l,c}}$ are distributed
-normally, this posterior distribution is also normal. We evaluated
-the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
-
-\begin{equation}
-p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
-
-\[
-\hat{\mu}_{i}=\hat{\mathbf{\Sigma}}_{i}\left(-\mathbf{1}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{\left(\cdot\right),i_{l,c}}\right)\]
-
-\[
-\hat{\Sigma}_{i}^{-1}=\mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\]
-
-\subsubsection{$p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$
-and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
-
-In the augmented probit regression model, the posterior distribution
-of $a_{i_{l,c}}$ is distributed according to a truncated normal distribution
-where the response variable is observed.
-
-\begin{equation}
-p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)=\begin{cases}
-trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=1\\
-trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=-1\end{cases}\end{equation}
-
-However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
-be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
-chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
-a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
--1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
-$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
-since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
-$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
-to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
-sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
-
-\begin{equation}
-p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
-
-\begin{equation}
-p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
-
-\begin{equation}
-p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
-1, & y_{parent_{l,c}}=-1\\
-0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
-
-\[
-p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
-
-\begin{equation}
-=\begin{cases}
-\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
-trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
-trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
-0 & otherwise\end{cases}\end{equation}
-
-where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
-variables excluding the response variable being sampled.
-
-\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha',\alpha\right)$}
-
-In our model, we place a hierarchical Dirichlet prior over topic assignments.
-This prior shares many features with the hierarchical Dirichlet process
-and inference over this distribution proceeds in a very similar fashion.
-This flexible distribution allows for an asymmetric prior over document
-level distributions over topics \citet{WallachMiMc2009}.
-
-Posterior inference was performed using the {}``direct assignment''
-method of \citet{TehJorBea2006}.
-
-\begin{equation}
-\beta\sim Dir\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)\end{equation}
-
-\begin{equation}
-p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}
-
-where $s\left(n,m\right)$ represents stirling numbers of the first
-kind.
-
-\subsubsection{$p\left(\alpha;\lambda\right)$, $p\left(\alpha';\lambda\right)$,
-$p\left(\beta;\lambda\right)$}
-
-All hyperparameters were given broad gamma priors $\left(\lambda=\left\{ shape=2,scale=1000\right\} \right)$
-and sampled via the Metropolis-Hastings algorithm.
-
-\subsection{Prediction}
-
-\section{Experiments}
-
-\subsection{Data}
-
-Our data set was gathered from the clinical data warehouse of NewYork
-- Presbyterian Hospital. The data consisted of free-text discharge
-summaries and their respective ICD-9 codes. A discharge summary is
-a clinical report prepared by a physician or other health professional
-at the conclusion of a hospital stay or series of treatments. The
-note outlines the patient\textquoteright{}s chief complaint, diagnostic
-findings, therapy administered, patient\textquoteright{}s response
-to the chosen therapy, the treatment plan and the recommendations
-upon discharge. The ICD-9 codes used to structure the discharge summary
-data are part of a controlled terminology which is the international
-standard diagnostic classification for epidemiological, health management,
-and clinical purposes (http://www.who.int/classifications/icd/en/).
-The codes are classified in a rooted-tree structure, with each edge
-representing an is-a relationship between parent and child, such that
-the parent diagnosis subsumes the child diagnosis. In the hospital,
-ICD-9 codes are generated manually by trained medical coders, who
-review all the information in the discharge summary. For the purposes
-of sLDA, ICD-9 codes will be used as labels for discharge summaries.
-
-We worked within the guidelines of the Health Insurance Portability
-and Accountability Act (HIPAA), which protects patient privacy and
-the security of potentially identifying medical material, known as
-personal health information (PHI). HIPAA covers any information within
-a medical record that was created, used, or disclosed during the course
-of providing a health care service and that can be used to identify
-an individual. Before beginning data processing, we generated a PHI-free
-dataset (see Data Pre-processing below).
-
-\subsection{Pre-Processing}
-
-Patient discharge summaries and their associated ICD-9 diagnoses are
-stored in two different places in the NewYork - Presbyterian data
-warehouse, and so had to be linked together before being fed into
-the sLDA algorithm. Each discharge note and set of diagnoses were
-assigned a patient unique identifier (PUID) and a visit unique identifier
-(VUID), allowing the two types of data to be linked.
-
-Natural Language Processsing (NLP) techniques were used to process
-the free-text discharge summaries. First, the Natural Language Toolkit
-(http://www.nltk.org/) was used to tokenize the text. Next, feature
-selection was performed using a term frequency - inverse document
-frequency (TF-IDF) algorithm on the entire document set and sorting
-the words by their TF-IDF values. The top 10,000 words were manually
-evaluated to eliminate all potentially identifying information. Finally,
-each discharge summary was converted to a bag-of-words, listing the
-frequencies of the remaining, free of protected health information,
-top 10,000 words.
-
-Preparation of the diagnostic codes involved inference over the ICD-9
-hierarchy. The is-a relationships of the hierarchy allowed us to make
-two important assumptions. First, if a diagnosis was observed to be
-present, all of its ancestors could be assumed to be present as well
-(e.g., if a patient had malignant hypertension, it could be assumed
-that they also had essential hypertension. Second, if a diagnosis
-was observed to be absent, it could be assumed that all of its descendants
-were also absent (e.g. if a patient did not have essential hypertension,
-it could be assumed that they did not have malignant hypertension).
-Unfortunately, ICD-9 code observations never include observations
-of disease absence. ICD-9 codes are only documented when the condition
-is observed to be present. Additionally, ICD-9 codes are known to
-have relatively low sensitivity; conditions that are present are often
-not documented in a set of ICD-9 codes (Surjan, 1999). Given these
-facts, we made the following assumptions regarding each visit: recorded
-diagnoses and their ancestors were labeled as true; diagnoses that
-were observed at some time for a patient but not at the current visit
-were labeled as unobserved; and diagnoses that had never been listed
-for a patient were labeled as false for all of that patient\textquoteright{}s
-visits. This last assumption captures the belief that parts of the
-ICD-9 hierarchy that are never observed for a particular patient are
-almost certain to be false. Additionally, for computational purposes,
-we decided not to include an ICD-9 code at all if neither it nor one
-of its descendants had been assigned to a patient in any of the records
-in our dataset.
-
-\subsection{ICD-9 Code Hierarchy}
-
-Here, we augment the sLDA model such that the supervised signal is
-distribution over the ICD-9 code tree, which is an is-a hierarchy
-\citep{ICD9NCBO}. An is-a hierarchy is represented by the tree data
-structure where each node has only a single parent and nodes cannot
-be parents of ancestors (ie. there are no loops). In this particular
-case, the ICD-9 code hierarchy is also partially a prefix trie where
-the labels for certain nodes are prefixes for child nodes. Given that
-this rule does not apply to all nodes in the hierarchy, we did not
-use this feature to determine the structure of the hierarchy. Instead
-we acquired a dataset that explicitly defined the relationships between
-the nodes of the hierarchy \citep{ICD9NCBO}. In documentation of
-ICD-9 codes for billing purposes, only a subset of the nodes can be
-used, however the nodes higher in the hierarchy contain semantic information
-about the categories of codes that are their descendants. For this
-reason, we included these nodes in our model.
-
-\subsection{Related Work}
-
-Latent dirichlet allocation (LDA) is a generative probabilistic model
-of corpora that represents documents as a mixed membership bag-of-words.
-Also known as topic models, these models infer the latent structure,
-or topics, of documents in a corpus. Each document is represented
-as a collection of words, generated from a set of topic assignments
-(one for each word), where each topic assignment is drawn from a distribution
-over topics \citep{Blei2003}.
-
-This model, supervised latent dirichlet allocation (sLDA), builds
-on LDA by incorporating an exponential family response variable. Although
-there are many models for making predictions based on free text, sLDA
-is unique in that it is a generative model, it represents documents
-as a mixed-membership, and constrains the inference of the latent
-structure of the documents by its predictability of the response variable.
-In other words, sLDA infers topics such that the model is capable
-of a high predictive likelihood for words in a document and the response
-variable associated with a document. This approach has been shown
-to outperform both LASSO (L1 regularized least squares regression)
-and LDA followed by least squares regression \citep{BleiMcAuliffe2008}.
-
-Other models - predicting document links, other supervised latent
-variable models
-
-% move to related work?
-While there have been some attempts to automatically classify groups
-of patients as a potential preliminary step to ICD-9 code assignment
-\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
-automatic assignment of ICD-9 codes to medical text became a more
-prevalent research topic only in the last few years. A subset of earlier
-work proposed various methods on small corpora, based on a few specific
-diseases \citep{Rao2003} but the most recent and promising work on
-the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
-Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
-(website). Most of the classification strategies included word matching
-and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
-The data set given to the participants consisted only of documents
-that were 1-2 lines each and all of the documents were radiology reports
-- clearly limiting the scope of potential ICD-9 codes which could
-be assigned. The only paper which has attempted to work with a document
-scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
-Lita proposed support vector machine and bayesian ridge regression
-methods to assign appropriate labels to the documents but did not
-utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
-
-\section{Results}
-
-%We ran the Gibbs sampler for two days on 500 documents with 20 topics to learn the sLDA model which we then followed by a prediction of ICD-9 codes for 100 documents.
-The prediction resulted in 38 documents with at least 1 ICD-9 code assignment while some had over 200 assignments.
-
-%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
-%   \centering
-%   \includegraphics[width=4in]{number_of_codes_histogram} 
-%   \caption{distribution of ICD-9 code assignments over all of the 100 documents predicted}
-%   \label{fig:example}
-%\end{figure}
-
-%The accuracy of the prediction was tested against the actual ICD-9 codes in the EHR record for each discharge summary.  Different statistical measures were applied to asses the predictions.
-
-%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
-%   \centering
-%   \includegraphics[width=4in]{signal_detection} 
-%   \caption{sensitivity, specificity, precision, recall and F-measure.}
-%   \label{fig:example}
-%\end{figure}
-
-%Although the sensitivty is not as high as we would have hoped, it is important to note that the sLDA was run on a very limited sample of documents (500) due to time constraints.
-Additionally, the “gold standard” data used are actual hospital results which are very noisy given the incompleteness and introduction of human error.
-
-\section{Conclusion}
-
-%We have proposed a new method for prediction of ICD-9 codes given a medical discharge summary note.
-Our method explicitly takes advantage of the ICD-9 hierarchical IS-A relationships as well as the word distributions in each document.
-We used Gibbs sampling for a supervised latent dirichlet allocation model to learn a distribution of words over twenty topics and their appropriate ICD-9 codes (response variable).
-The model was then tested on 100 discharge summaries for which the ICD-9 diagnoses were known.
-The amount of ICD-9 codes produced for some discharge summaries was quite high because of the way the model worked with the IS-A hierarchy, highlighting the completeness of ICD-9 coding when using this sLDA model.
-Overall, the model had extremely good specificity while the sensitivity was lower than anticipated.
-Given our limited set of documents, and very noisy gold standard it would not be accurate to discount the use of sLDA in medical document classification and automatic coding.
-%In future work, we plan to significantly expand the set of documents used to train the model (up to 15,000), as well as run several different iterations to learn the optimal topic number.
-
-%A more automated approach for coding patient visits with the ICD-9 hierarchy has great promise to  reduce health care costs and increase completeness of medical coded data.
-We hope that this implementation of supervised LDA will serve as preliminary work for a much larger study with more documents in the future.
-
-\begin{small} \bibliographystyle{plainnat} \bibliographystyle{plainnat}
-\bibliographystyle{plainnat}
-\bibliography{refs}
-
-\end{small}
-
-\newpage{}
-
-\subsection{Rao-Blackwellization}
-
-To derive the Gibbs sampler in general, we integrate over the parameters
-$\theta$ and $\phi_{1:K}$ resulting in the collapsed joint distribution
-shown in Equations 5-8.
-
-\begin{landscape} \centering % optional, probably makes it look better to have it centered on the page
-
-\begin{small}
-
-\begin{equation}
-p\left(\theta,z_{1:N}\mid w_{1:N},y_{1:I},\phi_{1:K},\eta_{1:I},\alpha,\beta,\mu,\xi\right)=\frac{p\left(\theta\mid\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)}{\int_{\theta}p\left(\theta\mid\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)d\theta}\end{equation}
-
-\begin{equation}
-\intop_{\theta}\intop_{\phi_{1:K}}p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\theta},\mathbf{\phi},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)d\theta d\phi_{1:K}=\intop_{\theta}\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\left\{ p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}\left[p\left(z_{m,n}\mid\theta_{m}\right)p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)\right]\prod_{i=1}^{I}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right\} \prod_{i=1}^{I}p\left(\eta_{i};\mu\right)d\theta d\phi_{1:K}\end{equation}
-
-\begin{equation}
-p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\prod_{n=1}^{N}p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)d\phi_{1:K}\intop_{\theta}\prod_{m=1}^{M}p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}p\left(z_{m,n}\mid\theta_{m}\right)d\theta\end{equation}
-
-\begin{equation}
-=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}\end{equation}
-
-\begin{equation}
-=\prod_{i=1}^{I}\left[\mathcal{N}\left(\eta_{i}\mid\mu,1\right)\prod_{m=1}^{M}h\left(y\right)\exp\left\{ \left(\eta_{i}^{T}\bar{z}\right)y-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}
-
-\end{small} \end{landscape}
-\end{document}
-
-%\begin{align}
-%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right) & =\\
-% & \frac{p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)p\left(\theta\mid\beta,\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}{\int_{\alpha,\alpha',\gamma}p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)\int_{\theta}p\left(\theta\mid\beta,\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}\end{align}

### 2011-05-22 paper/introduction.tex
Added to the introduction

@@
-There exist surprisingly many sources of unstructured text data that have been partially or completely categorized by human editors.
-Examples include hierarchical directories of webpages \citep{DMOZ}, large hierarchically annotated product catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), manually annotated patient medical records, and many more.
-In this work we show how to combine these two sources of information in a single model that allows us to, amongst other things, automatically annotate and/or categorize new text documents, effectively inserting them into the category
+There exist many sources of unstructured data that have been partially or completely categorized by human editors.
+In this paper we focus on unstructured textual data that has been, at least in part, manually categorized.
+Examples include but are not limited to webpages and curated hierarchical directories of the same \citep{DMOZ}, product descriptions and catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), and patient hospital treatment transcripts and  codes applied to them for bookkeeping and insurance purposes (e.g.~hospital discharge records and the International Classification of Disease 9th Revision,
+Clinical Modification (ICD-9-CM) codes assigned to them\cite{}).
+In this work we show how to combine these two sources of information using a single model that allows one, among other things, to automatically categorize new text documents, suggest labels that might be inaccurate, and compute improved similarities between documents for information retrieval purposes.
+The models and techniques that we develop in this paper are applicable in other domains as well, for instance, unstructured representations of data that have been hierarchically classified (e.g.
+image catalogs with bags of features image representations).
+
+In this work we extent supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
+sLDA is latent Dirichlet allocation (LDA) augmented with per document ``supervision''  often taking the form of a single numerical or categorical ``label.''  More generally this ``supervision'' can be seen as extra data generated about a document;  for instance its quality or relevance (e.g.~online reviews), marks given to written work (e.g.~graded essays), or the number of times a web document is linked.
+These labels are usually modeled as having been generated conditioned on the mix of topics found in a document.
+It has been demonstrated that the signal provided by this supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{}.
+
+Our main contribution is to show how to utilize supervision in the form of  hierarchical and (often) multiple labelings in a similar manner.
+As one concrete example consider web retailers.
+They often have both a browse-able hierarchy and free-text descriptions of all products they sell.
+The situation of each product in the product hierarchy (often multiply situated) can be seen as a form of multiple labeling and as a
+similar products based on the similarity of their textual description is an u.
+An equivalent challenge, particularly for larger retailers, is to situate the merchandise in as many categories as possible.
+
+In this paper we describe the use of a topic model based on supervised
+latent Dirichlet allocation (sLDA) to identify topics within narrative
+discharge summaries and to automate the assignment of diagnostic codes,
+specifically International Classification of Disease 9th Revision,
+Clinical Modification (ICD-9-CM) codes. There are a number of advantages
+to this approach. First, manually coding diagnoses is a time-consuming
+and notoriously unreliable process. Many diagnoses are omitted in
+the final record, and a high error rate is found even in the principal
+diagnoses \citep{Surjan1999}.
 
@@
 
-In this paper we describe the use of a topic model based on supervised
-latent Dirichlet allocation (sLDA) to identify topics within narrative
-discharge summaries and to automate the assignment of diagnostic codes,
-specifically International Classification of Disease 9th Revision,
-Clinical Modification (ICD-9-CM) codes. There are a number of advantages
-to this approach. First, manually coding diagnoses is a time-consuming
-and notoriously unreliable process. Many diagnoses are omitted in
-the final record, and a high error rate is found even in the principal
-diagnoses \citep{Surjan1999}.
-
-An automated process would ideally produce a more complete and accurate
-diagnosis lists. Also, this model will reveal information about the
-medical records themselves. For example, we may gain an understanding
-of what a specific code actually means in terms of clinical narratives.
-Similarly, viewing the distribution of topics over discharge summaries
-may reveal information about the latent structure of clinician documentation.
-Lastly, the sLDA model would provide a novel approach to dealing with
-the problem of high dimensionality when representing narrative text
-in a vector space specifically by reducing dimensions from an entire
-vocabulary of potentially tens of thousands of words to a set of several
-dozen topics.
+%An automated process would ideally produce a more complete and accurate
+%diagnosis lists. Also, this model will reveal information about the
+%medical records themselves. For example, we may gain an understanding
+%of what a specific code actually means in terms of clinical narratives.
+%Similarly, viewing the distribution of topics over discharge summaries
+%may reveal information about the latent structure of clinician documentation.
+%Lastly, the sLDA model would provide a novel approach to dealing with
+%the problem of high dimensionality when representing narrative text
+%in a vector space specifically by reducing dimensions from an entire
+%vocabulary of potentially tens of thousands of words to a set of several
+%dozen topics.

### 2011-05-22 paper/main.tex
Added to the introduction

@@
 
-There exist surprisingly many sources of unstructured text data that have been partially or completely categorized by human editors.
-Examples include hierarchical directories of webpages \citep{DMOZ}, large hierarchically annotated product catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), manually annotated patient medical records, and many more.
-In this work we show how to combine these two sources of information in a single model that allows us to, amongst other things, automatically annotate and/or categorize new text documents, effectively inserting them into the category
-
-\begin{itemize}
-\item Benefits of combining human categorization information into ``topic models''
-\item LDA solved free text
-\item supervised LDA improves LDA (extra info) and allows new inference (predict links, etc.)
-\item amazon, freshdirect, netflix, dmoz, pandora (music genome)
-\end{itemize}
-
-% Informatics journal paper
-%
-% Despite the growing emphasis on meaningful
-%use of technology in medicine, many aspects of medical record-keeping
-%remain a manual process. Diagnostic coding for billing and insurance
-%purposes is often handled by professional medical coders who must
-%explore a patient's extensive clinical record before assigning the
-%proper codes. So while electronic health records (EHRs) should be
-%adopted by most medical institutions within the next several years,
-%largely due to the provisions of HITECH under the American Recovery
-%and Reinvestment Act \citep{Blumenthal2009}, there has been little
-%movement forward in automating medical coding.
-
-In this paper we describe the use of a topic model based on supervised
-latent Dirichlet allocation (sLDA) to identify topics within narrative
-discharge summaries and to automate the assignment of diagnostic codes,
-specifically International Classification of Disease 9th Revision,
-Clinical Modification (ICD-9-CM) codes. There are a number of advantages
-to this approach. First, manually coding diagnoses is a time-consuming
-and notoriously unreliable process. Many diagnoses are omitted in
-the final record, and a high error rate is found even in the principal
-diagnoses \citep{Surjan1999}.
-
-An automated process would ideally produce a more complete and accurate
-diagnosis lists. Also, this model will reveal information about the
-medical records themselves. For example, we may gain an understanding
-of what a specific code actually means in terms of clinical narratives.
-Similarly, viewing the distribution of topics over discharge summaries
-may reveal information about the latent structure of clinician documentation.
-Lastly, the sLDA model would provide a novel approach to dealing with
-the problem of high dimensionality when representing narrative text
-in a vector space specifically by reducing dimensions from an entire
-vocabulary of potentially tens of thousands of words to a set of several
-dozen topics.
+\input{introduction}
 
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{Combined_vs_Separate_Learning.png}
+\caption{default}
+\label{default}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{TP_Rate.png}
+\caption{default}
+\label{default}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{FP_Rate.png}
+\caption{default}
+\label{default}
+\end{center}
+\end{figure}
+
 %We ran the Gibbs sampler for two days on 500 documents with 20 topics to learn the sLDA model which we then followed by a prediction of ICD-9 codes for 100 documents.

### 2011-05-22 paper/abstract.tex
working on abstract and stuff

We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply labeled bag-of-words data.  Out-of-sample label prediction is the primary goal of this work; however, improved dimensionality reduction is also of interest.  We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA) in the same manner as was done in supervised LDA (SLDA) prior art.  We find that the additional supervision signal that comes from multiple, hierarchically constrained labels substantially improves out-of-sample label prediction in medical document labeling and product categorization tasks.  Additionally, held-out likelihood

%Topic models are unsupervised models well suited to data exploration and information retrieval.  Supervised topic models use side-information, labels and the like, to improve the learned representations.

%The benefits of supervision in topic modeling 
%Current medical record keeping technology relies heavily upon human
%capacity to effectively summarize and infer information from free-text
%physician notes. We propose a novel method to suggest diagnostic code
%assignment for patient visits, based upon narrative medical notes.
%We applied a supervised latent Dirichlet allocation model to a corpora
%of free-text medical notes from NewYork - Presbyterian Hospital to
%infer a set of specific ICD-9 codes for each patient note. Evaluation
%of the predictions were conducted by comparison to a gold-standard
%set of ICD-9s assigned to a set of patient notes.

### 2011-05-22 paper/introduction.tex
working on abstract and stuff

@@
 In this paper we focus on unstructured textual data that has been, at least in part, manually categorized.
-Examples include but are not limited to webpages and curated hierarchical directories of the same \citep{DMOZ}, product descriptions and catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), and patient hospital treatment transcripts and  codes applied to them for bookkeeping and insurance purposes (e.g.~hospital discharge records and the International Classification of Disease 9th Revision,
-Clinical Modification (ICD-9-CM) codes assigned to them\cite{}).
-In this work we show how to combine these two sources of information using a single model that allows one, among other things, to automatically categorize new text documents, suggest labels that might be inaccurate, and compute improved similarities between documents for information retrieval purposes.
-The models and techniques that we develop in this paper are applicable in other domains as well, for instance, unstructured representations of data that have been hierarchically classified (e.g.
-image catalogs with bags of features image representations).
+Examples include but are not limited to webpages and curated hierarchical directories of the same \citep{DMOZ}, product descriptions and catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), and patient hospital treatment transcripts and  codes applied to them for bookkeeping and insurance purposes (e.g.~hospital discharge records with International Classification of Disease 9th Revision,
+Clinical Modification (ICD-9-CM) codes assigned \cite{}).
+In this work we show how to combine these two sources of information using a single model that allows one to automatically categorize new text documents, suggest labels that might be inaccurate, compute improved similarities between documents for information retrieval purposes, and more.
+The models and techniques that we develop in this paper are applicable in other domains as well, namely, any unstructured representations of data that have been hierarchically classified (e.g.~image catalogs with bag-of-feature image representations).
 
-In this work we extent supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
-sLDA is latent Dirichlet allocation (LDA) augmented with per document ``supervision''  often taking the form of a single numerical or categorical ``label.''  More generally this ``supervision'' can be seen as extra data generated about a document;  for instance its quality or relevance (e.g.~online reviews), marks given to written work (e.g.~graded essays), or the number of times a web document is linked.
-These labels are usually modeled as having been generated conditioned on the mix of topics found in a document.
+In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
+sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision;''  often taking the form of a single numerical or categorical ``label.''  More generally this ``supervision'' can be seen as extra data  about a document;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
+These labels are usually generatively modeled as having been conditional draw given an inferred document-specific topic mixture.
 It has been demonstrated that the signal provided by this supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{}.
@@
 Our main contribution is to show how to utilize supervision in the form of  hierarchical and (often) multiple labelings in a similar manner.
-As one concrete example consider web retailers.
-They often have both a browse-able hierarchy and free-text descriptions of all products they sell.
-The situation of each product in the product hierarchy (often multiply situated) can be seen as a form of multiple labeling and as a
-similar products based on the similarity of their textual description is an u.
-An equivalent challenge, particularly for larger retailers, is to situate the merchandise in as many categories as possible.
+Consider web retail data.
+Web retailers often have both a browse-able product hierarchy and free-text descriptions for all products they sell.
+The situation of each product in a product hierarchy (often multiply situated) constitutes a multiple, hierarchical labeling of the free-text product descriptions.
+We hypothesis that hierarchical labels should, at least in theory, provide better signal than the simpler unstructured supervision previously considered.
+Results from applying our model to medical record and web retail data suggests that this is likely to be the case.
+In particular, we observed big gains in our primary goal of out-of-sample label prediction when using hierarchical supervision.
 
-In this paper we describe the use of a topic model based on supervised
-latent Dirichlet allocation (sLDA) to identify topics within narrative
-discharge summaries and to automate the assignment of diagnostic codes,
-specifically International Classification of Disease 9th Revision,
-Clinical Modification (ICD-9-CM) codes. There are a number of advantages
-to this approach. First, manually coding diagnoses is a time-consuming
-and notoriously unreliable process. Many diagnoses are omitted in
-the final record, and a high error rate is found even in the principal
-diagnoses \citep{Surjan1999}.
+The remainder of this paper is structured as follows.
+In Section~\ref{sec:methods} we introduce hierarchically supervised LDA (HSLDA), in Section~\ref{sec:related_work} we review related work, and in Section~\ref{sec:experiments} we apply HSLDA to health care and web retail data, showing predictive performance and improved topic generation.
 
-\begin{itemize}
-\item Benefits of combining human categorization information into ``topic models''
-\item LDA solved free text
-\item supervised LDA improves LDA (extra info) and allows new inference (predict links, etc.)
-\item amazon, freshdirect, netflix, dmoz, pandora (music genome)
-\end{itemize}
+%In this paper we describe the use of a topic model based on supervised
+%latent Dirichlet allocation (sLDA) to identify topics within narrative
+%discharge summaries and to automate the assignment of diagnostic codes,
+%specifically International Classification of Disease 9th Revision,
+%Clinical Modification (ICD-9-CM) codes. There are a number of advantages
+%to this approach. First, manually coding diagnoses is a time-consuming
+%and notoriously unreliable process. Many diagnoses are omitted in
+%the final record, and a high error rate is found even in the principal
+%diagnoses \citep{Surjan1999}.
+
+%\begin{itemize}
+%\item Benefits of combining human categorization information into ``topic models''
+%\item LDA solved free text
+%\item supervised LDA improves LDA (extra info) and allows new inference (predict links, etc.)
+%\item amazon, freshdirect, netflix, dmoz, pandora (music genome)
+%\end{itemize}
 

### 2011-05-22 paper/main.tex
I was an idiot and edited this simultaneously to what my awesome collaborator Frank was doing.
working on abstract and stuff

@@
 \begin{abstract}
-The benefits of supervision in topic modeling 
-%Current medical record keeping technology relies heavily upon human
-%capacity to effectively summarize and infer information from free-text
-%physician notes. We propose a novel method to suggest diagnostic code
-%assignment for patient visits, based upon narrative medical notes.
-%We applied a supervised latent Dirichlet allocation model to a corpora
-%of free-text medical notes from NewYork - Presbyterian Hospital to
-%infer a set of specific ICD-9 codes for each patient note. Evaluation
-%of the predictions were conducted by comparison to a gold-standard
-%set of ICD-9s assigned to a set of patient notes. 
+\input{abstract}
 \end{abstract}
@@
 \section{Introduction}
-
 \label{sec:introduction}
-
 \input{introduction}
 
-\section{Background}
+%\section{Background}
 
 \section{Methods}
+\label{sec:methods}
 
@@
 \subsection{Related Work}
-
+\label{sec:related_work}
 Latent dirichlet allocation (LDA) is a generative probabilistic model
@@
 \section{Results}
+\label{sec:results}
 
@@
 
-\section{Conclusion}
-
+\section{Discussion}
+\label{sec:discussion}
+\input{discussion}
 %We have proposed a new method for prediction of ICD-9 codes given a medical discharge summary note.

### 2011-05-22 paper/discussion.tex
added figures to tex.  win!

\begin{itemize}
\item what about the nonparametric version of this?
\end{itemize}

### 2011-05-22 paper/abstract.tex
a first draft (sucky) abstract and introduction

@@
-We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply labeled bag-of-words data.
+We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply-labeled bag-of-words data.
 Out-of-sample label prediction is the primary goal of this work; however, improved dimensionality reduction is also of interest.
-We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA) in the same manner as was done in supervised LDA (SLDA) prior art.
-We find that the additional supervision signal that comes from multiple, hierarchically constrained labels substantially improves out-of-sample label prediction in medical document labeling and product categorization tasks.
-Additionally, held-out likelihood
+We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA).
+We find that the additional signal that comes from multiple, hierarchically constrained labels substantially improves out-of-sample label prediction in comparison to supervised LDA approaches that don't utilize information derived from the structure of the label space.
+We demonstrate HSLDA on large-scale data from medical document labeling and retail product categorization tasks and show both improved label prediction performance and show evidence that the learned topic model improves as a result of using this signal too.
 

### 2011-05-22 paper/main.tex
added figures to tex.  win!
a first draft (sucky) abstract and introduction

@@
 
-\section{Methods}
-\label{sec:methods}
+\section{Model}
+\label{sec:model}
 
@@
 
-\subsection{Posterior Inference}
+\section{Inference}
+\label{sec:inference}
 
@@
 
-\subsection{Gibbs Sampling}
+\subsection{Gibbs Sampler}
 
@@
 
+\subsection{Evaluation}
+
 \section{Results}
@@
 \begin{center}
-\includegraphics[width=1\textwidth]{Combined_vs_Separate_Learning.png}
-\caption{default}
-\label{default}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=1\textwidth]{TP_Rate.png}
-\caption{default}
-\label{default}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=1\textwidth]{FP_Rate.png}
-\caption{default}
-\label{default}
+\subfigure[]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
+\subfigure[]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
+\subfigure[]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+\subfigure[]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
+\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  In all figures solid is HSLDA, dashed is ADLER, dotted is ADLER, and dot-dashed is ADLER.}
+\label{fig:icd9}
 \end{center}

### 2011-05-22 paper/main.tex
added additional amazon figures

@@
 \begin{center}
-\subfigure[]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
-\subfigure[]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
-\subfigure[]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
-\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  In all figures solid is HSLDA, dashed is ADLER, dotted is ADLER, and dot-dashed is ADLER.}
+\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
+\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
+\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
+\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.
+In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.
+Top row:  \subref{1a} includes ancestor prediction performance,  \subref{1b} results are for given (leaf) labels alone.
+Bottom row: \subref{1c} are the sensitivity curves from \subref{1b} aligned on threshold value,  \subref{1d} are the 1-specificity curves from \subref{1b} aligned on threshold value.}
+\label{fig:icd9}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[][\label{2a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors_amazon}}
+\subfigure[][\label{2b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs_amazon}}
+\subfigure[][\label{2c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs_amazon}}
+\subfigure[][\label{2d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
+\caption{Out-of-sample Amazon product code predictions from product free-text descriptions.
+In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.
+Top row:  \subref{2a} includes ancestor prediction performance,  \subref{2b} results are for given (leaf) labels alone.
+Bottom row: \subref{2c} are the sensitivity curves from \subref{2b} aligned on threshold value,  \subref{2d} are the 1-specificity curves from \subref{2b} aligned on threshold value.}
 \label{fig:icd9}

### 2011-05-24 paper/main.tex
adding changes made to modeling section

@@
 \label{sec:model}
+We define here a hierarchically supervised LDA model.
+We assume a pre-specified set of labels $\mathcal{L}$.
+Each document is assigned a response of either -1 or 1 for at least one, but potentially many, label(s) in $\mathcal{L}$.
+A response of 1 or -1 to label $l$ indicates if the document respectively is or is not $l$.
+The label $l$ for a document $d$ will be used interchangeably to refer the observed response of document $d$ to label $l$.
+The label set is assumed to be an ``is a" hieararchy.
+This means that if a label $l_1$ is a parent of label $l_2$ in the hierarchy and document $d$ has a positive response to the label $l_2$ then document $d$ also has a positive response to the label $l_1$.
+Seen in a negative light, if document $d$ has a negative response to the $l_1$ label then document $d$ also has a negative response to the $l_2$ label.
+To capture this hierarchical structure we model the labeling of documents using a generative cascade of conditional probit regression models.
+Approximate inference is performed using Gibbs sampling.
 
 % generalize to arbitrary codes (not just ICD-9's
-
+%
 %
@@
  \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}
-
 \label{fig:example} 
 \end{figure}
+%
+The fixed parameters of the model are the number of topics $K$, the number of unique words in the vocabulary $V$, the number of documents $D$, and the standard deviation $\sigma$ used in a normal prior distribution.
+The hyper-parameters $\alpha^\prime$, $\alpha$, and $\gamma$ are weight parameters for Dirichlet prior distributions.
+We will denote the $K$-dimensional Dirichlet distribution as Dir$_K(\cdot)$ and the $K$ dimensional normal distribution as $\mathcal{N}_k(\cdot)$.
 
-Given the number of topics, K, and broad gamma priors on hyperparameters,
-the generative process is as follows: 
+We will now describe the stochastic generative process which defines our model.  The graphical model is show in Figure~\ref{fig:example}. 
+%
+%Given the number of topics, K, and broad gamma priors on hyperparameters,
+%the generative process is as follows: 
 \begin{enumerate}
-\item For each topic, $k$:
+	\item For each topic $k = 1,\ldots, K$:
+	\begin{itemize}
+		\item Draw a distribution over words $\mathbf{\phi_{k}}\sim {\rm Dir}_{V}\left(\gamma \mathbf{1}\right)$, where $\mathbf{1}$ is a vector of ones of length $V$
+	\end{itemize}
 
-\begin{enumerate}
-\item Draw a distribution over words $\phi_{k}\sim Dir_{V}\left(\mathbf{1},\gamma\right)$ 
+	\item For each label $l \in \mathcal{L}$:
+	\begin{itemize}
+		\item Draw the regression coefficients $\mathbf{\eta_l} \ | \ \sigma \sim \mathcal{N}_{K} \left( -\mathbf{1}, \sigma I_K \right)$, where $I_K$ is the $K$ dimensional identity matrix
+	\end{itemize}
+
+	\item Draw a prior over topic proportions $\mathbf{\beta}\mid\alpha'\sim {\rm Dir}_{K} \left(\alpha^{\prime} \mathbf{1} \right)$
+
+	\item For each document $d = 1, \ldots, D$:
+	\begin{itemize} 
+		\item Draw topic proportions $\mathbf{\theta_{d}} \ | \ \mathbf{\beta},\alpha \sim {\rm Dir}_{K}\left(\alpha \mathbf{\beta} \right)$ 
+		\item For $n = 1, \ldots, N_d$:
+		\begin{itemize}
+			\item Draw topic assignment $z_{n,d} \ | \ \mathbf{\theta_{d}} \sim {\rm Multinomial} \left(\mathbf{\theta_{d}} \right)$ 
+			\item Draw word $w_{n,d} \ | \ z_{n,d},\mathbf{\beta_{1:K}} \sim {\rm Multinomial} \left(\mathbf{\beta_{z_{n,d}}} \right)$ 
+		\end{itemize}
+
+		\item For each label $l \in \mathcal{L}$:
+		\begin{itemize}
+			\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$ 
+			\item Set the response variable \begin{equation*} y_{l, d} \ | \ a_{l, d} = \begin{cases}
+																										 \ \ \ 1 & \text{if $a_{l,d} > 0$ and $y_{{\rm parent}(l),d} = 1$} \\
+																										-1 & \text{otherwise}
+																								\end{cases}  
+															 \end{equation*}															 
+		\end{itemize}
+	\end{itemize}
 \end{enumerate}
-\item For each ICD9 code, $c$, at all levels in the tree, $l$:
-
-\begin{enumerate}
-\item Draw regression coefficient $\eta_{i_{l,c}}\mid\sigma\sim\mathcal{N}_{K}\left(-\mathbf{1},\sigma\right)$ 
-\end{enumerate}
-\item Draw a prior over topic proportions $\beta\mid\alpha'\sim Dir_{K}\left(\mathbf{1},\alpha'\right)$ 
-\item For each document, $d$:
-
-\begin{enumerate}
-\item Draw topic proportions $\theta_{d}\mid\beta,\alpha\sim Dir_{K}\left(\beta,\alpha\right)$ 
-\item For each word, $n$:
-
-\begin{enumerate}
-\item Draw topic assignment $z_{n,d}\mid\theta_{d}\sim Mult_{K}\left(\theta_{d}\right)$ 
-\item Draw word $w_{n,d}\mid z_{n,d},\beta_{1:K}\sim Mult_{V}\left(\beta_{z_{n}}\right)$ 
-\end{enumerate}
-\item For each level of the ICD-9 code tree, $l$:
-
-\begin{enumerate}
-\item For each ICD-9 code at this level, $c$:
-
-\begin{enumerate}
-\item Draw a latent variable \[
-a_{d,i_{l,c}}\sim\begin{cases}
-\mathcal{N}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=1\\
-trunc\mathcal{N}^{-}\left(\bar{z}^{T}\eta_{i_{l,c}},1\right), & y_{parent_{l,c}}=-1\end{cases}\]
- where $\bar{z}=N^{-1}\sum_{n=1}^{N}z_{n}$ and $\mathcal{I}=\left\{ i_{0},i_{1},...,i_{\mathcal{L}}\right\} $
-and $i_{l}=\left\{ i_{l,0},i_{l,1},...,i_{l,C_{l}}\right\} $
-\item Draw a response variable $y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\sim\begin{cases}
-\begin{array}{c}
-1,\\
--1,\end{array} & \begin{array}{c}
-a_{d,i_{l,c}}>0\\
-otherwise\end{array}\end{cases}$
-\end{enumerate}
-\end{enumerate}
-\end{enumerate}
-\end{enumerate}
-The generative model for the ICD-9 codes is equivalent to a probit
-regression model. In our case, the regression is conditional on the
-parents according to the constraints of the ICD-9 code tree. The latent
-variable utilized here is also known as an auxiliary variable.
-
+%
+This type of generative model is known as a probit regression model.
+Probit regression is like logistic regression except instead of modeling the logit of $P(y = 1)$ using a linear form we model $\Phi^{-1}(P(y = 1))$ using a linear form, where $\Phi(\cdot)$ is the CDF for a standard normal distribution.
+In this case, the regression is conditional on the parents according to the constraints of the labeling hierarchy.
+The latent variables $a_{l,d}$ utilized here are also known as an auxiliary variables because the are introduced to make exact Gibbs sampling possible and are not of primary interest.
+%
 \section{Inference}
 \label{sec:inference}
+%
+In the Bayesian approach to statistical modeling the primary task of inference is to find the posterior distribution over the unobserved parameters of the model.
+However, it will often be the case that the set of labels $\mathcal{L}$ is not fully observed for every document.
+We will define $\mathcal{L}_d$ to be the subset of labels which have been observed for document $d$.
+It is straightforward to integrate out the variables $a_{l^\prime,d}$ and $y_{l^\prime,d}$ for $l^\prime \in \mathcal{L} \backslash \mathcal{L}_d$ from the full generative model.
+We can also integrate out the parameters $\mathbf{\phi_{1:K}}$ and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}.
+Therefore, in our model the latent variables are $\mathbf{z} = \{ z_{1:N_d,d} \}_{d = 1, \ldots, D},\mathbf{\eta} = \{ \mathbf{\eta_l} \}_{l \in \mathcal{L}},\mathbf{a} = \{ a_{l,d}\}_{l \in \mathcal{L}, d = 1, \ldots, D}, \mathbf{\beta}, \alpha, \alpha^{\prime}$ and $\gamma$.
 
-Given an observation of a set of ICD-9 codes and a document, the posterior
-distribution for the latent variables is given by Equation .
-
-\begin{equation}
-p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
-
-\[
-=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]
-
-The denominator for this distribution is the marginal probability
-of the data and cannot be solved in closed form. This is often the
-case in evaluating posterior distributions of non-trivial probabilistic
-models. We will appeal to one of the common methods for approximating
-posterior distributions in the face of intractable normalization factors:
-Markov chain Monte Carlo (MCMC) sampling. Since in this model it is
-possible to sample from the conditional distributions for all variables
-we will use the Gibbs sampling algorithm to obtain an approximation
-to this posterior. In particular, we will derive a collapsed Gibbs
-sampler.
-
+The posterior distribution we seek cannot be solved in closed form.
+This is often the case in evaluating posterior distributions of non-trivial probabilistic models.
+We will appeal to one of the common methods for approximating posterior distributions in the face of intractable normalization factors: Markov chain Monte Carlo (MCMC) sampling.
+Since in this model it is possible to sample from the conditional distributions for all variables we will use the Gibbs sampling algorithm to obtain an approximation to this posterior.
+In particular, we will derive a collapsed Gibbs sampler.
+%
+%
+%Given an observation of a set of observed labels and a document, the posterior distribution for the latent variables is
+%
+%\begin{equation}
+%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
+%\[
+%=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]
+%
 \subsection{Gibbs Sampler}
 
-We derive a collapsed Gibbs sampler for this model, marginalizing
-$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
-in LDA models see \citet{Griffiths04}. To derive the Gibbs sampler
-we evaluate the individual conditional probability distributions for
-all unobserved variables.
+We derive a collapsed Gibbs sampler for this model by considering the individual conditional probability distributions for
+each of the unobserved variables.  We use the notation $\mathbf{z}_{-(n,d)}$ to denote $\mathbf{z} \backslash z_{n,d}$.
+%, marginalizing
+%$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
+%in LDA models see \citet{Griffiths04}. 
+\subsubsection{$p\left(z_{n,d} \mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}
+%
+First we consider the conditional distribution of the assignment variable for each word $n = 1, \ldots, N_d$ in documents $d = 1, \ldots, D$.
+The conditional distribution does not include $\theta_{1:D}$ and $\phi_{1:K}$ because they have been integrated out as in the collapsed Gibbs sampler \citep{Griffiths04}.
+The conditional distribution of $z_{n,d}$ is proportional to the joint distribution of its markov blanket, explicitly this means
+%\begin{equation}
+%p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
+%
+%
+%Due to the factorization of this model, we can rewrite the joint distribution
+%as the following:
+%
+\begin{equation}
+p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto \prod_{l \in \mathcal{L}_d}p\left(a_{l,d}\mid\mathbf{z},\eta_l \right)p\left(z_{n,d} \ | \ \mathbf{z_{-\left(n, d \right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right).
+\end{equation}
+%
+The product is only over the subset of labels $\mathcal{L}_d$ which have been observed for document $d$.
+By isolating terms that depend on $z_{n,d}$ and absorbing all other terms into a normalizing constant \citep{Griffiths04} we find
+%
+\begin{eqnarray}
+\lefteqn{p\left(z_{n,d} = k \mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w}, \mathbf{\eta},\alpha,\beta,\gamma\right)\propto} \nonumber\\ 
+& \hspace{2cm}  \left(n_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\beta_{k}\right)\frac{n_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(n_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)} \prod_{l \in \mathcal{L}_d} \exp\left\{ -\frac{\left(\bar z_d^{T}\eta_{l}-a_{l,d} \right)^{2}}{2}\right\}\label{eq:z-likelihood}.
+\end{eqnarray}
+%
+Here, $n_{v,d}^{k,-\left(n, d^{\prime} \right)}$ represents the number of words of type $v$ in document $d$ assigned to topic $k$ omitting the $n^{\rm th}$ word of document $d^{\prime}$.
+The notation $(\cdot)$ in the subscript means the count resulting from summing over the omitted subscript variable.
+Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\mathbf{\beta},\gamma\right)$ can be sampled through enumeration.
+%
+\subsubsection{$p\left(\mathbf{\eta_l}\mid\mathbf{z},\mathbf{a},\sigma\right)$}
+%
+We now consider the conditional distribution of the regression coefficients $\mathbf{eta_l}$ for $l \in \mathcal{L}$.
+Given that $\mathbf{\eta_l}$ and $a_{l,d}$ are distributed normally, the posterior distribution of $\mathbf{\eta_l}$ is normally distributed with mean $\hat \mu$ and covariance $\hat{\mathbf{\Sigma}}$ such that
+% (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
+\begin{eqnarray}
+\hat{\mathbf{\Sigma}}^{-1} &=&\mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}} \\
+\hat \mu &=&\hat{\mathbf{\Sigma}}_{i}\left(-\mathbf{1}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\mathbf{a}_l\right). 
+\end{eqnarray}
+This is a standard result from normal Bayesian linear regression \citep{BishopPRML}.
+Here, $\bar{\mathbf{Z}}$ is a $D \times K$ matrix such that row $d$ of $\mathbf{\bar Z}$ is $\bar z_d$, and $\mathbf{a}_l = [a_{l,1}, a_{l,2}, \ldots, a_{l,D}]^T$.
+%
+%
+%
+%p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
+%
+%\[
+%
+%
+%
+%\[
+\subsubsection{$p\left(a_{l,d}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$}
+%and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
+%
+The auxiliary variables $a_{l,d}$ must be sampled for documents $d = 1, \ldots, D$ and $l \in \mathcal{L}_d$.  The conditional posterior distribution of $a_{l,d}$ is the truncated normal distribution
+%
+\begin{equation}
+p\left(a_{l,d,} \ | \ \mathbf{z},\mathbf{Y},\mathbf{\eta}\right) \propto \frac{1}{\sqrt{2\pi}} exp \left\{ \frac{-1}{2}\left(a_{l,d} - \mathbf{\eta_l}^T\mathbf{\bar z}_d\right) \right\} I \left(a_{l,d} y_{l,d} > 0\right).
+\end{equation}
+This conditional distribution can be sampled using an inverse CDF method.
+%
+%However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
+%be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
+%chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
+%a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
+%-1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
+%$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
+%since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
+%$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
+%to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
+%sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
+%
+%\begin{equation}
+%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
+%
+%
+%\begin{equation}
+%p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
+%
+%
+%\begin{equation}
+%p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
+%1, & y_{parent_{l,c}}=-1\\
+%0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
+%
+%
+%\[
+%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
+%
+%
+%\begin{equation}
+%=\begin{cases}
+%\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
+%trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
+%trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
+%0 & otherwise\end{cases}\end{equation}
+%
+%
+%where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
+%variables excluding the response variable being sampled.
+%
+%
+\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha',\alpha\right)$}
+%
+In our model, we place a hierarchical Dirichlet prior over topic assignments.
+This prior shares many features with the hierarchical Dirichlet process and inference over this distribution proceeds in a very similar fashion.
+This flexible distribution allows for an asymmetric prior over document level distributions over topics \citep{WallachMiMc2009}.
 
-\subsubsection{$p\left(z_{m,n}\mid\mathbf{z_{-\left(m,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}
-
-For the purposes of sampling, we will be able to derive a representation
-of the joint distribution isolating a particular latent variable,
-$z$, for a word instance, $n$, in a document instance, $d$. The
-conditional probability with respect to this latent variable is proportional
-to the joint distribution of its markov blanket up to a constant.
-
+Posterior inference is performed using the {}``direct assignment'' method of \citet{TehJorBea2006}.
+%
 \begin{equation}
-p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
-
-Due to the factorization of this model, we can rewrite the joint distribution
-as the following:
-
+\beta\sim {\rm Dir}\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)
+\end{equation}
+%
 \begin{equation}
-\propto\prod_{i_{l,c}\in\mathcal{I}}p\left(a_{i_{l,c}}\mid\mathbf{z},\eta_{i_{l,c}}\right)p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right)\end{equation}
-
-We isolate only terms that depend on $z_{m,n}$ and absorb all other
-constant terms into the normalization constant \citep{Griffiths04}.
-
-\begin{equation}
-\propto\prod_{i_{l,c}\in\mathcal{I}}\exp\left\{ -\frac{\left(\bar{z}^{T}\eta_{i_{l,c}}-a_{i_{l,c}}\right)^{2}}{2}\right\} \left(n_{d,\left(\cdot\right)}^{k,-\left(d,n\right)}+\alpha\beta_{k}\right)\frac{n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma}{\sum_{v=1}^{V}\left(n_{\left(\cdot\right),w_{d,n}}^{k,-\left(d,n\right)}+\gamma\right)}\label{eq:z-likelihood}\end{equation}
-
-Here, $n_{d,v}^{k,-\left(d,n\right)}$ represents the count of word
-$v$ in document $d$ assigned to topic $k$ omitting the $(d,n)^{th}$
-word count.
-
-Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$
-can be sampled through enumeration.
-
-\subsubsection{$p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)$}
-
-Given that $\eta_{i_{l,c}}$ and $a_{d,i_{l,c}}$ are distributed
-normally, this posterior distribution is also normal. We evaluated
-the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
-
-\begin{equation}
-p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
-
-\[
-\hat{\mu}_{i}=\hat{\mathbf{\Sigma}}_{i}\left(-\mathbf{1}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{\left(\cdot\right),i_{l,c}}\right)\]
-
-\[
-\hat{\Sigma}_{i}^{-1}=\mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\]
-
-\subsubsection{$p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$
-and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
-
-In the augmented probit regression model, the posterior distribution
-of $a_{i_{l,c}}$ is distributed according to a truncated normal distribution
-where the response variable is observed.
-
-\begin{equation}
-p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)=\begin{cases}
-trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=1\\
-trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\eta_{i}^{T}\bar{z,}\mathbf{1},y_{d,i_{l,c}}\right) & if\quad y_{d,i_{l,c}}=-1\end{cases}\end{equation}
-
-However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
-be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
-chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
-a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
--1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
-$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
-since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
-$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
-to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
-sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
-
-\begin{equation}
-p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
-
-\begin{equation}
-p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
-
-\begin{equation}
-p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
-1, & y_{parent_{l,c}}=-1\\
-0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
-
-\[
-p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
-
-\begin{equation}
-=\begin{cases}
-\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
-trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
-trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
-0 & otherwise\end{cases}\end{equation}
-
-where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
-variables excluding the response variable being sampled.
-
-\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha',\alpha\right)$}
-
-In our model, we place a hierarchical Dirichlet prior over topic assignments.
-This prior shares many features with the hierarchical Dirichlet process
-and inference over this distribution proceeds in a very similar fashion.
-This flexible distribution allows for an asymmetric prior over document
-level distributions over topics \citet{WallachMiMc2009}.
-
-Posterior inference was performed using the {}``direct assignment''
-method of \citet{TehJorBea2006}.
-
-\begin{equation}
-\beta\sim Dir\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)\end{equation}
-
-\begin{equation}
-p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}
-
-where $s\left(n,m\right)$ represents stirling numbers of the first
-kind.
-
-\subsubsection{$p\left(\alpha;\lambda\right)$, $p\left(\alpha';\lambda\right)$,
-$p\left(\beta;\lambda\right)$}
-
-All hyperparameters were given broad gamma priors $\left(\lambda=\left\{ shape=2,scale=1000\right\} \right)$
-and sampled via the Metropolis-Hastings algorithm.
-
+p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}
+\end{equation}
+%
+%
+where $s\left(n,m\right)$ represents stirling numbers of the first kind.
+%
+\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha' \right)$, $p\left(\gamma \right)$}
+%
+The hyperparameters $\alpha$, $\alpha^\prime$, and $\gamma$ are given broad ${\rm Gamma}(2, 1000)$ prior distributions and sampled via the Metropolis-Hastings algorithm.
+%
 \subsection{Prediction}

### 2011-05-25 paper/main.tex
added a few words
Updated Experiments

@@
 
-\subsection{Data}
+In the section we describe the application of HSLDA for prediction
+in two hierarchically structured domains. Firstly, we describe using
+discharge summaries to predict diagnoses, encoded as ICD-9 codes.
+Discharge summaries are documents that are authored by clinicians
+to summarize the course of someone who has been hospitalized. ICD-9
+codes are used mainly for billing purposes to indicate the conditions
+for which a patient was treated. Secondly, we describe using Amazon.com
+product descriptions to predict product categories.
 
-Our data set was gathered from the clinical data warehouse of NewYork
-- Presbyterian Hospital. The data consisted of free-text discharge
-summaries and their respective ICD-9 codes. A discharge summary is
-a clinical report prepared by a physician or other health professional
-at the conclusion of a hospital stay or series of treatments. The
-note outlines the patient\textquoteright{}s chief complaint, diagnostic
-findings, therapy administered, patient\textquoteright{}s response
-to the chosen therapy, the treatment plan and the recommendations
-upon discharge. The ICD-9 codes used to structure the discharge summary
-data are part of a controlled terminology which is the international
-standard diagnostic classification for epidemiological, health management,
-and clinical purposes (http://www.who.int/classifications/icd/en/).
-The codes are classified in a rooted-tree structure, with each edge
-representing an is-a relationship between parent and child, such that
-the parent diagnosis subsumes the child diagnosis. In the hospital,
+\subsection{Diagnosis Prediction}
+
+Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
+Hospital. The data consisted of free-text discharge summaries and
+their respective ICD-9 codes. A discharge summary is a clinical report
+prepared by a physician or other health professional at the conclusion
+of a hospital stay or series of treatments. The note outlines the
+patient's chief complaint, diagnostic findings, therapy administered,
+patient's response to the chosen therapy, the treatment plan and
+the recommendations upon discharge. The ICD-9 codes used to structure
+the discharge summary data are part of a controlled terminology which
+is the international standard diagnostic classification for epidemiological,
+health management, and clinical purposes. The ICD-9 codes are organized
+in a rooted-tree structure, with each edge representing an is-a relationship
+between parent and child, such that the parent diagnosis subsumes
+the child diagnosis. For example, the code representing {}``Pneumonia
+due to adenovirus'' is a child of the code representing {}``Viral
+pneumonia'' where the former is a type of the latter. In the hospital,
 ICD-9 codes are generated manually by trained medical coders, who
-review all the information in the discharge summary. For the purposes
-of sLDA, ICD-9 codes will be used as labels for discharge summaries.
+review all the information in the discharge summary.
+
+The text of the discharge summaries were pre-processed such that each
+document would be represented as counts over a 10,000 word vocabulary.
+The Natural Language Toolkit was used to tokenize the text. A vocabulary
+was identified by first sorting terms based on a global term frequency-inverse
+document frequency measure. The top 10,000 words which were not identifying
+in some way (a name, place, or identifying number) were selected for
+inclusion in the vocabulary.
+
+For each hospitalization there are usually several ICD-9 codes assigned
+for billing purposes. These codes are known to be quite specific but
+not very sensitive {[}citation{]}. Regardless of that fact, this is
+one of the only sources for information on patient diagnoses aside
+from the free text. Aside from prediction, one of the goals is to
+compare the sensitivity of predictions from the HSLDA model in comparison
+to the codes in a case where a test closer to ground truth is available.
+For this we will compare whether predictions for the ICD-9 code associated
+with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia
+was chosen because hemoglobin values are readily available and the
+definition of anemia according the World Health Organization is approximately
+12.5, with a threshold of 12 for women and 13 for men {[}citation{]}.
 
@@
 of providing a health care service and that can be used to identify
-an individual. Before beginning data processing, we generated a PHI-free
-dataset (see Data Pre-processing below).
+an individual. This study was approved by the Institutional Review
+Board.
 
-\subsection{Pre-Processing}
+\subsection{Product Category Prediction}
 
-Patient discharge summaries and their associated ICD-9 diagnoses are
-stored in two different places in the NewYork - Presbyterian data
-warehouse, and so had to be linked together before being fed into
-the sLDA algorithm. Each discharge note and set of diagnoses were
-assigned a patient unique identifier (PUID) and a visit unique identifier
-(VUID), allowing the two types of data to be linked.
+Data for these experiments were obtained partially from the Stanford
+Network Analysis Platform (SNAP) Amazon product metadata dataset and
+partially directly from the the Amazon.com website. The product ID's
+and categorizations were obtained from the SNAP dataset and the product
+descriptions were obtained directly from the website.
 
-Natural Language Processsing (NLP) techniques were used to process
-the free-text discharge summaries. First, the Natural Language Toolkit
-(http://www.nltk.org/) was used to tokenize the text. Next, feature
-selection was performed using a term frequency - inverse document
-frequency (TF-IDF) algorithm on the entire document set and sorting
-the words by their TF-IDF values. The top 10,000 words were manually
-evaluated to eliminate all potentially identifying information. Finally,
-each discharge summary was converted to a bag-of-words, listing the
-frequencies of the remaining, free of protected health information,
-top 10,000 words.
+We were able to deduce the structure of the hierarchy for the Amazon.com
+products directly since all ancestors in the hierarchy were included
+with each category label. For example, {}``DVD / Genres / Science
+Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
+for the DVD, {}``The Time Machine''. Each product was labeled with
+multiple categories.
 
-Preparation of the diagnostic codes involved inference over the ICD-9
-hierarchy. The is-a relationships of the hierarchy allowed us to make
-two important assumptions. First, if a diagnosis was observed to be
-present, all of its ancestors could be assumed to be present as well
-(e.g., if a patient had malignant hypertension, it could be assumed
-that they also had essential hypertension. Second, if a diagnosis
-was observed to be absent, it could be assumed that all of its descendants
-were also absent (e.g. if a patient did not have essential hypertension,
-it could be assumed that they did not have malignant hypertension).
-Unfortunately, ICD-9 code observations never include observations
-of disease absence. ICD-9 codes are only documented when the condition
-is observed to be present. Additionally, ICD-9 codes are known to
-have relatively low sensitivity; conditions that are present are often
-not documented in a set of ICD-9 codes (Surjan, 1999). Given these
-facts, we made the following assumptions regarding each visit: recorded
-diagnoses and their ancestors were labeled as true; diagnoses that
-were observed at some time for a patient but not at the current visit
-were labeled as unobserved; and diagnoses that had never been listed
-for a patient were labeled as false for all of that patient\textquoteright{}s
-visits. This last assumption captures the belief that parts of the
-ICD-9 hierarchy that are never observed for a particular patient are
-almost certain to be false. Additionally, for computational purposes,
-we decided not to include an ICD-9 code at all if neither it nor one
-of its descendants had been assigned to a patient in any of the records
-in our dataset.
+The vocabulary for this experiment was created by including the most
+frequent 30K words omitting stopwords.
 
-\subsection{ICD-9 Code Hierarchy}
+\subsection{Evaluation}
 
-Here, we augment the sLDA model such that the supervised signal is
-distribution over the ICD-9 code tree, which is an is-a hierarchy
-\citep{ICD9NCBO}. An is-a hierarchy is represented by the tree data
-structure where each node has only a single parent and nodes cannot
-be parents of ancestors (ie. there are no loops). In this particular
-case, the ICD-9 code hierarchy is also partially a prefix trie where
-the labels for certain nodes are prefixes for child nodes. Given that
-this rule does not apply to all nodes in the hierarchy, we did not
-use this feature to determine the structure of the hierarchy. Instead
-we acquired a dataset that explicitly defined the relationships between
-the nodes of the hierarchy \citep{ICD9NCBO}. In documentation of
-ICD-9 codes for billing purposes, only a subset of the nodes can be
-used, however the nodes higher in the hierarchy contain semantic information
-about the categories of codes that are their descendants. For this
-reason, we included these nodes in our model.
+The two main methods of evaluation for the model are prediction and
+topic quality. We compare model performance against 3 similar models
+to demonstrate that each component of the model is important for performance.
+Specifically, we evaluate models including independent regressors
++ sLDA (hierarchical constraints on labels ignored), HSLDA fit by
+running LDA first then running tree-conditional regressions, and HSLDA
+fit with fixed random regression parameters.
+
+\subsubsection{Prediction}
+
+Given the expectation o
+
+\subsubsection{Topic Quality and Character}
 

### 2011-05-26 paper/discussion.tex
a first draft (sucky) abstract and introduction
Updated Experiment section and other minor things

@@
+We have described a mixed membership model with hierarchical supervision.
+We have demonstrated this model in the context of document modeling with hierarchical multi-label supervision.
+Such a model is appropriate in domains where there are hierarchical constraints among the labels such as is the case in an IS-A hierarchy.
+
+...
+
 \begin{itemize}
 \item what about the nonparametric version of this?
+\item discuss the broader goal, from the beginning of search engine time, to combine categorization and free text.  this, to our knowledge, is the first principled approach to doing so
 \end{itemize}

### 2011-05-26 paper/main.tex
Updated Experiment section and other minor things

@@
 		\begin{itemize}
-			\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$ 
+			\item Draw $a_{l,d}\mid z_{1:N_{d},d},\eta_{l},y_{parent(l),d}\sim\begin{cases}
+				\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right), & y_{parent(l)}=1\\
+				\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ 
+			%\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$ 
 			\item Set the response variable \begin{equation*} y_{l, d} \ | \ a_{l, d} = \begin{cases}
@@
 %
-\subsection{Prediction}
 
@@
 patient's chief complaint, diagnostic findings, therapy administered,
-patient's response to the chosen therapy, the treatment plan and
-the recommendations upon discharge. The ICD-9 codes used to structure
+patient's response to the chosen therapy, the treatment plan and the
+recommendations upon discharge. The ICD-9 codes used to structure
 the discharge summary data are part of a controlled terminology which
@@
 
-Given the expectation o
+The two measures for predictive performance used here include the
+true positive rate and the false positive rate. We evaluate model
+performance on held out data. A more ideal evaluation of performance
+would include a manually labeled hierarchy since it is well known
+that ICD-9 codes have a relatively low sensitivity.
+
+Performance was evaluated against $p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$.
 
@@
 
+TBD
+
 \subsection{Related Work}
-\label{sec:related_work}
-Latent dirichlet allocation (LDA) is a generative probabilistic model
-of corpora that represents documents as a mixed membership bag-of-words.
-Also known as topic models, these models infer the latent structure,
-or topics, of documents in a corpus. Each document is represented
-as a collection of words, generated from a set of topic assignments
-(one for each word), where each topic assignment is drawn from a distribution
-over topics \citep{Blei2003}.
 
-This model, supervised latent dirichlet allocation (sLDA), builds
-on LDA by incorporating an exponential family response variable. Although
-there are many models for making predictions based on free text, sLDA
-is unique in that it is a generative model, it represents documents
-as a mixed-membership, and constrains the inference of the latent
-structure of the documents by its predictability of the response variable.
-In other words, sLDA infers topics such that the model is capable
-of a high predictive likelihood for words in a document and the response
-variable associated with a document. This approach has been shown
-to outperform both LASSO (L1 regularized least squares regression)
-and LDA followed by least squares regression \citep{BleiMcAuliffe2008}.
+\label{sec:related_work} Latent dirichlet allocation (LDA) is a generative
+probabilistic model of corpora that represents documents as a mixed
+membership bag-of-words. Also known as topic models, these models
+infer the latent structure, or topics, of documents in a corpus. Each
+document is represented as a collection of words, generated from a
+set of topic assignments (one for each word), where each topic assignment
+is drawn from a distribution over topics \citep{Blei2003}.
 
-Other models - predicting document links, other supervised latent
-variable models
+Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
+an exponential family response variable. Although there are many models
+for making predictions based on free text, sLDA is unique in that
+it is a generative model, it represents documents as a mixed-membership,
+and constrains the inference of the latent structure of the documents
+by its predictability of the response variable. In other words, sLDA
+infers topics such that the model is capable of a high predictive
+likelihood for words in a document and the response variable associated
+with a document. This approach has been shown to outperform both LASSO
+(L1 regularized least squares regression) and LDA followed by least
+squares regression \citep{BleiMcAuliffe2008}.
+
+There have been many models that incorporate both latent models of
+text and some form of supervision {[}citations{]}. One set of models
+that are particularly relevant to HSLDA are Chang and Blei's {[}citation{]}
+hierarchical models for document networks. In that family of models,
+they encountered a similar scenario where the lack of a link did not
+truly indicate absence. In our model, the lack of a code in the hierarchy
+being assigned does not necessarily indicate absence. Therefore, as
+in the work of Chang and Blei, we employ regularization in the form
+of a negative prior on the regression parameters to provide a prior
+that indicates a bias towards being truly negative in the absence
+of a code.
 
@@
 diseases \citep{Rao2003} but the most recent and promising work on
-the subject was inspired by the 2007 Medical NLP Challenge: \textquotedblleft{}International
+the subject was inspired by the 2007 Medical NLP Challenge: {}``International
 Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
@@
 utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
-
-\subsection{Evaluation}
 
@@
 
-\newpage{}
-
-\subsection{Rao-Blackwellization}
-
-To derive the Gibbs sampler in general, we integrate over the parameters
-$\theta$ and $\phi_{1:K}$ resulting in the collapsed joint distribution
-shown in Equations 5-8.
-
-\begin{landscape} \centering % optional, probably makes it look better to have it centered on the page
-
-\begin{small}
-
-\begin{equation}
-p\left(\theta,z_{1:N}\mid w_{1:N},y_{1:I},\phi_{1:K},\eta_{1:I},\alpha,\beta,\mu,\xi\right)=\frac{p\left(\theta\mid\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)}{\int_{\theta}p\left(\theta\mid\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\beta\right)\right)\left(\prod_{i=1}^{I}p\left(y_{i}\mid z_{1:N},\eta_{i},\xi\right)p\left(\eta_{i}\mid\mu\right)\right)d\theta}\end{equation}
-
-\begin{equation}
-\intop_{\theta}\intop_{\phi_{1:K}}p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\theta},\mathbf{\phi},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)d\theta d\phi_{1:K}=\intop_{\theta}\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\left\{ p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}\left[p\left(z_{m,n}\mid\theta_{m}\right)p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)\right]\prod_{i=1}^{I}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right\} \prod_{i=1}^{I}p\left(\eta_{i};\mu\right)d\theta d\phi_{1:K}\end{equation}
-
-\begin{equation}
-p\left(\mathbf{Y},\mathbf{w},\mathbf{z},\mathbf{\eta},\alpha,\beta,\mu,\xi\right)=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\intop_{\phi_{1:K}}\prod_{k=1}^{K}p\left(\phi_{k};\beta\right)\prod_{m=1}^{M}\prod_{n=1}^{N}p\left(w_{m,n}\mid\phi_{z_{m,n}}\right)d\phi_{1:K}\intop_{\theta}\prod_{m=1}^{M}p\left(\theta_{m};\alpha\right)\prod_{n=1}^{N}p\left(z_{m,n}\mid\theta_{m}\right)d\theta\end{equation}
-
-\begin{equation}
-=\prod_{i=1}^{I}\left[p\left(\eta_{i};\mu\right)\prod_{m=1}^{M}p\left(y_{m,i}\mid z_{m,1:N},\eta_{i},\xi\right)\right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\cdot\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\cdot\right)}^{k}+\alpha_{k}\right)}\end{equation}
-
-\begin{equation}
-=\prod_{i=1}^{I}\left[\mathcal{N}\left(\eta_{i}\mid\mu,1\right)\prod_{m=1}^{M}h\left(y\right)\exp\left\{ \left(\eta_{i}^{T}\bar{z}\right)y-A\left(\eta_{i}^{T}\bar{z}\right)\right\} \right]\prod_{k=1}^{K}\frac{\Gamma\left(\sum_{v=1}^{V}\beta_{v}\right)}{\prod_{v=1}^{V}\Gamma\left(\beta_{v}\right)}\frac{\prod_{v=1}^{V}\Gamma\left(n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}{\Gamma\left(\sum_{v=1}^{V}n_{\left(\bullet\right),v}^{k}+\beta_{v}\right)}\prod_{m=1}^{M}\frac{\Gamma\left(\sum_{k=1}^{K}\alpha_{k}\right)}{\prod_{k=1}^{K}\Gamma\left(\alpha_{k}\right)}\frac{\prod_{k=1}^{K}\Gamma\left(n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}{\Gamma\left(\sum_{k=1}^{K}n_{m,\left(\bullet\right)}^{k}+\alpha_{k}\right)}\end{equation}
-
-\end{small} \end{landscape}
 \end{document}
-
-%\begin{align}
-%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right) & =\\
-% & \frac{p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)p\left(\theta\mid\beta,\alpha\right)\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}{\int_{\alpha,\alpha',\gamma}p\left(\alpha,\alpha',\gamma;\lambda\right)p\left(\beta\mid\alpha'\right)\int_{\theta}p\left(\theta\mid\beta,\alpha\right)\sum_{k=1}^{K}\left(\prod_{n=1}^{N}p\left(z_{n}\mid\theta\right)p\left(w_{n}\mid z_{n},\phi_{1:K}\right)\right)\left(\prod_{k=1}^{K}p\left(\phi_{k}\mid\gamma\right)\right)\left(\prod_{i_{l,c}\in\mathcal{I}}p\left(y_{i_{l,c}}\mid z_{1:N},\eta_{i_{l,c}}\right)p\left(\eta_{i_{l,c}};\sigma\right)\right)}\end{align}

### 2011-05-29 paper/abstract.tex
Updated and rearranged

@@
 We find that the additional signal that comes from multiple, hierarchically constrained labels substantially improves out-of-sample label prediction in comparison to supervised LDA approaches that don't utilize information derived from the structure of the label space.
-We demonstrate HSLDA on large-scale data from medical document labeling and retail product categorization tasks and show both improved label prediction performance and show evidence that the learned topic model improves as a result of using this signal too.
+We demonstrate HSLDA on large-scale data from medical document labeling and retail product categorization tasks.
+We show improved label prediction performance and evidence that the learned topic model also improves as a result of using this signal.
 

### 2011-05-29 paper/introduction.tex
Updated and rearranged

@@
 Examples include but are not limited to webpages and curated hierarchical directories of the same \citep{DMOZ}, product descriptions and catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), and patient hospital treatment transcripts and  codes applied to them for bookkeeping and insurance purposes (e.g.~hospital discharge records with International Classification of Disease 9th Revision,
-Clinical Modification (ICD-9-CM) codes assigned \cite{}).
+Clinical Modification (ICD-9-CM) codes assigned \cite{ICD9}).
 In this work we show how to combine these two sources of information using a single model that allows one to automatically categorize new text documents, suggest labels that might be inaccurate, compute improved similarities between documents for information retrieval purposes, and more.
@@
 In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
-sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision;''  often taking the form of a single numerical or categorical ``label.''  More generally this ``supervision'' can be seen as extra data  about a document;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
-These labels are usually generatively modeled as having been conditional draw given an inferred document-specific topic mixture.
-It has been demonstrated that the signal provided by this supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{}.
+sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this ``supervision'' can be seen as extra data  about a document;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
+These labels are usually generatively modeled as having been conditionally draw given an inferred document-specific topic mixture.
+It has been demonstrated that the signal provided by this supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
 
@@
 The remainder of this paper is structured as follows.
-In Section~\ref{sec:methods} we introduce hierarchically supervised LDA (HSLDA), in Section~\ref{sec:related_work} we review related work, and in Section~\ref{sec:experiments} we apply HSLDA to health care and web retail data, showing predictive performance and improved topic generation.
+In Section~\ref{sec:related_work} we review related work, in Section~\ref{sec:model} we introduce hierarchically supervised LDA (HSLDA), and in Section~\ref{sec:experiments} we apply HSLDA to health care and web retail data, showing predictive performance and improved topic generation.
 

### 2011-05-29 paper/main.tex
Updated and rearranged

@@
 
+\subsection{Related Work}
+
+\label{sec:related_work} Latent dirichlet allocation (LDA) is a generative
+probabilistic model of corpora that represents documents as a mixed
+membership bag-of-words. Also known as topic models, these models
+infer the latent structure, or topics, of documents in a corpus. Each
+document is represented as a collection of words, generated from a
+set of topic assignments (one for each word), where each topic assignment
+is drawn from a distribution over topics \citep{Blei2003}.
+
+Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
+an exponential family response variable. Although there are many models
+for making predictions based on free text, sLDA is unique in that
+it is a generative model, it represents documents as a mixed-membership,
+and constrains the inference of the latent structure of the documents
+by its predictability of the response variable. In other words, sLDA
+infers topics such that the model is capable of a high predictive
+likelihood for words in a document and the response variable associated
+with a document. This approach has been shown to outperform both LASSO
+(L1 regularized least squares regression) and LDA followed by least
+squares regression \citep{BleiMcAuliffe2008}.
+
+There have been many models that incorporate both latent models of
+text and some form of supervision \citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}.
+One set of models that are particularly relevant to HSLDA are Chang
+and Blei's hierarchical models for document networks (Relational Topic
+Models). In that family of models, they encountered a similar scenario
+where the lack of a link did not truly indicate absence. In hierarchically
+labeled data, negative labels are uncommon and the lack of a label
+in the hierarchy is not equivalent to a negative label. Therefore,
+as in the work of Chang and Blei, we employ regularization to account
+for the lack of negative labels. This will be discussed further in
+\ref{sec:model}.
+
+% move to related work?
+While there have been some attempts to automatically classify groups
+of patients as a potential preliminary step to ICD-9 code assignment
+\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
+automatic assignment of ICD-9 codes to medical text became a more
+prevalent research topic only in the last few years. A subset of earlier
+work proposed various methods on small corpora, based on a few specific
+diseases \citep{Rao2003} but the most recent and promising work on
+the subject was inspired by the 2007 Medical NLP Challenge: {}``International
+Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
+(website). Most of the classification strategies included word matching
+and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
+The data set given to the participants consisted only of documents
+that were 1-2 lines each and all of the documents were radiology reports
+- clearly limiting the scope of potential ICD-9 codes which could
+be assigned. The only paper which has attempted to work with a document
+scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
+Lita proposed support vector machine and bayesian ridge regression
+methods to assign appropriate labels to the documents but did not
+utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
+
 \section{Model}
-\label{sec:model}
-We define here a hierarchically supervised LDA model.
-We assume a pre-specified set of labels $\mathcal{L}$.
-Each document is assigned a response of either -1 or 1 for at least one, but potentially many, label(s) in $\mathcal{L}$.
-A response of 1 or -1 to label $l$ indicates if the document respectively is or is not $l$.
-The label $l$ for a document $d$ will be used interchangeably to refer the observed response of document $d$ to label $l$.
-The label set is assumed to be an ``is a" hieararchy.
-This means that if a label $l_1$ is a parent of label $l_2$ in the hierarchy and document $d$ has a positive response to the label $l_2$ then document $d$ also has a positive response to the label $l_1$.
-Seen in a negative light, if document $d$ has a negative response to the $l_1$ label then document $d$ also has a negative response to the $l_2$ label.
-To capture this hierarchical structure we model the labeling of documents using a generative cascade of conditional probit regression models.
-Approximate inference is performed using Gibbs sampling.
+
+\label{sec:model} We define here a hierarchically supervised LDA
+model. Although we will focus on document modeling in our description
+and experiments, this model applies equally well to other collections
+of discrete data with hierarchically constrained labels.
+
+We assume a pre-specified set of labels $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
+Each document is assigned a response of either -1 or 1 for at least
+one, but potentially many labels in $\mathcal{L}$. The label, $l$,
+for a document, $d$, will be used interchangeably to refer to the
+observed response of document $d$ to label $l$. The label set is
+assumed to be structured as an {}``is-a'' hieararchy.
+To understand this, consider a hierarchy where label $l_{1}$ is a
+parent of label $l_{2}$. If document $d$ has a positive response
+to label $l_{2}$ then it will also have a positive response to label
+$l_{1}$. Conversely, if document $d$ has a negative response to
+label $l_{1}$ then it will also has a negative response to $l_{2}$.
+To capture this hierarchical structure we model the labeling of documents
+using a generative cascade of conditional probit regression models.
 
 % generalize to arbitrary codes (not just ICD-9's
-%
 %
@@
  \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}
+
 \label{fig:example} 
 \end{figure}
-%
-The fixed parameters of the model are the number of topics $K$, the number of unique words in the vocabulary $V$, the number of documents $D$, and the standard deviation $\sigma$ used in a normal prior distribution.
-The hyper-parameters $\alpha^\prime$, $\alpha$, and $\gamma$ are weight parameters for Dirichlet prior distributions.
-We will denote the $K$-dimensional Dirichlet distribution as Dir$_K(\cdot)$ and the $K$ dimensional normal distribution as $\mathcal{N}_k(\cdot)$.
 
-We will now describe the stochastic generative process which defines our model.  The graphical model is show in Figure~\ref{fig:example}. 
-%
+The fixed parameters of the model are the number of topics $K$, the
+number of unique words in the vocabulary $V$, the number of documents
+$D$, as well as the mean, $\mu$, and the standard deviation, $\sigma$,
+used in a normal prior distribution. The hyper-parameters $\alpha^{\prime}$,
+$\alpha$, and $\gamma$ are weight parameters for Dirichlet prior
+distributions. We will denote the $K$-dimensional Dirichlet distribution
+as Dir$_{K}(\cdot)$ and the $K$ dimensional normal distribution
+as $\mathcal{N}_{K}(\cdot)$.
+
+We will now describe the stochastic generative process which defines
+our model. The graphical model is show in Figure~\ref{fig:example}.
 %Given the number of topics, K, and broad gamma priors on hyperparameters,
 %the generative process is as follows:
+
 \begin{enumerate}
-	\item For each topic $k = 1,\ldots, K$:
+\item For each topic $k=1,\ldots,K$:
+
 \begin{itemize}
-		\item Draw a distribution over words $\mathbf{\phi_{k}}\sim {\rm Dir}_{V}\left(\gamma \mathbf{1}\right)$, where $\mathbf{1}$ is a vector of ones of length $V$
+\item Draw a distribution over words $\mathbf{\phi_{k}}\sim{\rm Dir}_{V}\left(\gamma\mathbf{1}\right)$,
+where $\mathbf{1}$ is a vector of ones of length $V$ 
 \end{itemize}
+\item For each label $l\in\mathcal{L}$:
 
-	\item For each label $l \in \mathcal{L}$:
 \begin{itemize}
-		\item Draw the regression coefficients $\mathbf{\eta_l} \ | \ \sigma \sim \mathcal{N}_{K} \left( -\mathbf{1}, \sigma I_K \right)$, where $I_K$ is the $K$ dimensional identity matrix
+\item Draw the regression coefficients $\mathbf{\eta_{l}}\mid\sigma\sim\mathcal{N}_{K}\left(\mu I_{K},\sigma I_{K}\right)$,
+where $I_{K}$ is the $K$ dimensional identity matrix 
 \end{itemize}
+\item Draw a prior over topic proportions $\mathbf{\beta}\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}\right)$
+\item For each document $d=1,\ldots,D$:
 
-	\item Draw a prior over topic proportions $\mathbf{\beta}\mid\alpha'\sim {\rm Dir}_{K} \left(\alpha^{\prime} \mathbf{1} \right)$
+\begin{itemize}
+\item Draw topic proportions $\mathbf{\theta_{d}}\mid\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\mathbf{\beta}\right)$ 
+\item For $n=1,\ldots,N_{d}$:
 
-	\item For each document $d = 1, \ldots, D$:
 \begin{itemize}
-		\item Draw topic proportions $\mathbf{\theta_{d}} \ | \ \mathbf{\beta},\alpha \sim {\rm Dir}_{K}\left(\alpha \mathbf{\beta} \right)$ 
-		\item For $n = 1, \ldots, N_d$:
-		\begin{itemize}
-			\item Draw topic assignment $z_{n,d} \ | \ \mathbf{\theta_{d}} \sim {\rm Multinomial} \left(\mathbf{\theta_{d}} \right)$ 
-			\item Draw word $w_{n,d} \ | \ z_{n,d},\mathbf{\beta_{1:K}} \sim {\rm Multinomial} \left(\mathbf{\beta_{z_{n,d}}} \right)$ 
+\item Draw topic assignment $z_{n,d}\mid\mathbf{\theta_{d}}\sim{\rm Multinomial}\left(\mathbf{\theta_{d}}\right)$ 
+\item Draw word $w_{n,d}\mid z_{n,d},\mathbf{\beta_{1:K}}\sim{\rm Multinomial}\left(\mathbf{\beta_{z_{n,d}}}\right)$ 
 \end{itemize}
+\item For each label $l\in\mathcal{L}$:
 
-		\item For each label $l \in \mathcal{L}$:
 \begin{itemize}
@@
 \mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right), & y_{parent(l)}=1\\
-				\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ 
-			%\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$ 
-			\item Set the response variable \begin{equation*} y_{l, d} \ | \ a_{l, d} = \begin{cases}
-																										 \ \ \ 1 & \text{if $a_{l,d} > 0$ and $y_{{\rm parent}(l),d} = 1$} \\
-																										-1 & \text{otherwise}
-																								\end{cases}  
-															 \end{equation*}															 
+\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
+
+\item Set the response variable \[
+y_{l,d}\mid a_{l,d}=\begin{cases}
+\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}and \ensuremath{y_{{\rm parent}(l),d}=1}}\\
+-1 & \text{otherwise}\end{cases}\]
+
 \end{itemize}
@@
 \end{enumerate}
-%
 This type of generative model is known as a probit regression model.
-Probit regression is like logistic regression except instead of modeling the logit of $P(y = 1)$ using a linear form we model $\Phi^{-1}(P(y = 1))$ using a linear form, where $\Phi(\cdot)$ is the CDF for a standard normal distribution.
-In this case, the regression is conditional on the parents according to the constraints of the labeling hierarchy.
-The latent variables $a_{l,d}$ utilized here are also known as an auxiliary variables because the are introduced to make exact Gibbs sampling possible and are not of primary interest.
-%
+Probit regression models are a type of discriminative probabilistic
+model similar to logistic regression. However, instead of using the
+logistic sigmoid as the link function, the probit regression model
+uses the CDF for a standard normal distribution - the inverse of which
+is known as the probit. In this case, the regression is conditional
+on the parents according to the constraints of the labeling hierarchy.
+The latent variables $a_{l,d}$ utilized here are also known as an
+auxiliary variables because the are introduced to make exact Gibbs
+sampling possible and are not of primary interest.
+
+Given that negative labels are uncommon and that the absence of a
+label is not equivalent to a negative label, we apply an informative
+prior to the regression parameters, $\mathbf{\eta}_{\mathcal{L}}$,
+in the form of a negative prior that encodes a bias towards being
+truly negative in the absence of a label.
+
 \section{Inference}
-\label{sec:inference}
-%
-In the Bayesian approach to statistical modeling the primary task of inference is to find the posterior distribution over the unobserved parameters of the model.
-However, it will often be the case that the set of labels $\mathcal{L}$ is not fully observed for every document.
-We will define $\mathcal{L}_d$ to be the subset of labels which have been observed for document $d$.
-It is straightforward to integrate out the variables $a_{l^\prime,d}$ and $y_{l^\prime,d}$ for $l^\prime \in \mathcal{L} \backslash \mathcal{L}_d$ from the full generative model.
-We can also integrate out the parameters $\mathbf{\phi_{1:K}}$ and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}.
-Therefore, in our model the latent variables are $\mathbf{z} = \{ z_{1:N_d,d} \}_{d = 1, \ldots, D},\mathbf{\eta} = \{ \mathbf{\eta_l} \}_{l \in \mathcal{L}},\mathbf{a} = \{ a_{l,d}\}_{l \in \mathcal{L}, d = 1, \ldots, D}, \mathbf{\beta}, \alpha, \alpha^{\prime}$ and $\gamma$.
+
+\label{sec:inference} In the Bayesian approach to statistical modeling,
+the primary task of inference is to find the posterior distribution
+over the unobserved parameters of the model. However, it is often
+possible and desirable to integrate over certain variables in the
+model, also known as collapsing. In our model, it will often be the
+case that the set of labels $\mathcal{L}$ is not fully observed for
+every document. We will define $\mathcal{L}_{d}$ to be the subset
+of labels which have been observed for document $d$. It is straightforward
+to integrate out the variables $a_{l^{\prime},d}$ and $y_{l^{\prime},d}$
+for $l^{\prime}\in\mathcal{L}\backslash\mathcal{L}_{d}$ from the
+full generative model. We can also integrate out the parameters $\mathbf{\phi_{1:K}}$
+and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}. Therefore,
+in our model the latent variables are $\mathbf{z}=\{z_{1:N_{d},d}\}_{d=1,\ldots,D},\mathbf{\eta}=\{\mathbf{\eta_{l}}\}_{l\in\mathcal{L}},\mathbf{a}=\{a_{l^{\prime},d}\}_{l^{\prime}\in\mathcal{L}_{d},d=1,\ldots,D},\mathbf{\beta},\alpha,\alpha^{\prime}$
+and $\gamma$.
 
 The posterior distribution we seek cannot be solved in closed form.
-This is often the case in evaluating posterior distributions of non-trivial probabilistic models.
-We will appeal to one of the common methods for approximating posterior distributions in the face of intractable normalization factors: Markov chain Monte Carlo (MCMC) sampling.
-Since in this model it is possible to sample from the conditional distributions for all variables we will use the Gibbs sampling algorithm to obtain an approximation to this posterior.
-In particular, we will derive a collapsed Gibbs sampler.
-%
-%
+This is often the case in evaluating posterior distributions of non-trivial
+probabilistic models. We will appeal to one of the common methods
+for approximating posterior distributions in the face of intractable
+normalization factors: Markov chain Monte Carlo (MCMC) sampling. Since
+in this model it is possible to sample from the conditional distributions
+for all variables we will use the Gibbs sampling algorithm to obtain
+an approximation to this posterior. %In particular, we will derive a collapsed Gibbs sampler.
 %Given an observation of a set of observed labels and a document, the posterior distribution for the latent variables is
-%
 %\begin{equation}
@@
 %=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]
-%
+
 \subsection{Gibbs Sampler}
 
-We derive a collapsed Gibbs sampler for this model by considering the individual conditional probability distributions for
-each of the unobserved variables.  We use the notation $\mathbf{z}_{-(n,d)}$ to denote $\mathbf{z} \backslash z_{n,d}$.
-%, marginalizing
+We derive a collapsed Gibbs sampler for this model by considering
+the individual conditional probability distributions for each of the
+unobserved variables. We use the notation $\mathbf{z}_{-(n,d)}$ to
+denote $\mathbf{z}\backslash z_{n,d}$. %, marginalizing
 %$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
 %in LDA models see \citet{Griffiths04}.
-\subsubsection{$p\left(z_{n,d} \mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}
-%
-First we consider the conditional distribution of the assignment variable for each word $n = 1, \ldots, N_d$ in documents $d = 1, \ldots, D$.
-The conditional distribution does not include $\theta_{1:D}$ and $\phi_{1:K}$ because they have been integrated out as in the collapsed Gibbs sampler \citep{Griffiths04}.
-The conditional distribution of $z_{n,d}$ is proportional to the joint distribution of its markov blanket, explicitly this means
-%\begin{equation}
+
+\subsubsection{$p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}
+
+First we consider the conditional distribution of the assignment variable
+for each word $n=1,\ldots,N_{d}$ in documents $d=1,\ldots,D$. The
+conditional distribution does not include $\theta_{1:D}$ and $\phi_{1:K}$
+because they have been integrated out as in the collapsed Gibbs sampler
+\citep{Griffiths04}. The conditional distribution of $z_{n,d}$ is
+proportional to the joint distribution of its markov blanket. %\begin{equation}
 %p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
-%
-%
 %Due to the factorization of this model, we can rewrite the joint distribution
 %as the following:
-%
 \begin{equation}
-p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto \prod_{l \in \mathcal{L}_d}p\left(a_{l,d}\mid\mathbf{z},\eta_l \right)p\left(z_{n,d} \ | \ \mathbf{z_{-\left(n, d \right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right).
-\end{equation}
-%
-The product is only over the subset of labels $\mathcal{L}_d$ which have been observed for document $d$.
-By isolating terms that depend on $z_{n,d}$ and absorbing all other terms into a normalizing constant \citep{Griffiths04} we find
-%
+p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto\prod_{l\in\mathcal{L}_{d}}p\left(a_{l,d}\mid\mathbf{z},\eta_{l}\right)p\left(z_{n,d}\ |\ \mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right).\end{equation}
+ The product is only over the subset of labels $\mathcal{L}_{d}$
+which have been observed for document $d$. By isolating terms that
+depend on $z_{n,d}$ and absorbing all other terms into a normalizing
+constant as in \citep{Griffiths04} we find \begin{eqnarray}
+\lefteqn{p\left(z_{n,d}=k\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto}\nonumber \\
+ & \hspace{2cm}\left(n_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\beta_{k}\right)\frac{n_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(n_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)}\prod_{l\in\mathcal{L}_{d}}\exp\left\{ -\frac{\left(\bar{z}_{d}^{T}\eta_{l}-a_{l,d}\right)^{2}}{2}\right\} .\label{eq:z-likelihood}\end{eqnarray}
+ Here, $n_{v,d}^{k,-\left(n,d^{\prime}\right)}$ represents the number
+of words of type $v$ in document $d$ assigned to topic $k$ omitting
+the $n^{{\rm th}}$ word of document $d^{\prime}$. The notation $(\cdot)$
+in the subscript means the count resulting from summing over the omitted
+subscript variable. Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\mathbf{\beta},\gamma\right)$
+can be sampled through enumeration.
+
+\subsubsection{$p\left(\mathbf{\eta_{l}}\mid\mathbf{z},\mathbf{a},\sigma\right)$}
+
+We now consider the conditional distribution of the regression coefficients
+$\mathbf{\eta_{l}}$ for $l\in\mathcal{L}$. Given that $\mathbf{\eta_{l}}$
+and $a_{l,d}$ are distributed normally, the posterior distribution
+of $\mathbf{\eta_{l}}$ is normally distributed with mean $\hat{\mu}$
+and covariance $\hat{\mathbf{\Sigma}}$ such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
 \begin{eqnarray}
-\lefteqn{p\left(z_{n,d} = k \mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w}, \mathbf{\eta},\alpha,\beta,\gamma\right)\propto} \nonumber\\ 
-& \hspace{2cm}  \left(n_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\beta_{k}\right)\frac{n_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(n_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)} \prod_{l \in \mathcal{L}_d} \exp\left\{ -\frac{\left(\bar z_d^{T}\eta_{l}-a_{l,d} \right)^{2}}{2}\right\}\label{eq:z-likelihood}.
-\end{eqnarray}
-%
-Here, $n_{v,d}^{k,-\left(n, d^{\prime} \right)}$ represents the number of words of type $v$ in document $d$ assigned to topic $k$ omitting the $n^{\rm th}$ word of document $d^{\prime}$.
-The notation $(\cdot)$ in the subscript means the count resulting from summing over the omitted subscript variable.
-Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\mathbf{\beta},\gamma\right)$ can be sampled through enumeration.
-%
-\subsubsection{$p\left(\mathbf{\eta_l}\mid\mathbf{z},\mathbf{a},\sigma\right)$}
-%
-We now consider the conditional distribution of the regression coefficients $\mathbf{eta_l}$ for $l \in \mathcal{L}$.
-Given that $\mathbf{\eta_l}$ and $a_{l,d}$ are distributed normally, the posterior distribution of $\mathbf{\eta_l}$ is normally distributed with mean $\hat \mu$ and covariance $\hat{\mathbf{\Sigma}}$ such that
-% (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
-\begin{eqnarray}
-\hat{\mathbf{\Sigma}}^{-1} &=&\mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}} \\
-\hat \mu &=&\hat{\mathbf{\Sigma}}_{i}\left(-\mathbf{1}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\mathbf{a}_l\right). 
-\end{eqnarray}
-This is a standard result from normal Bayesian linear regression \citep{BishopPRML}.
-Here, $\bar{\mathbf{Z}}$ is a $D \times K$ matrix such that row $d$ of $\mathbf{\bar Z}$ is $\bar z_d$, and $\mathbf{a}_l = [a_{l,1}, a_{l,2}, \ldots, a_{l,D}]^T$.
-%
-%
-%
+\hat{\mathbf{\Sigma}}^{-1} & = & \mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\\
+\hat{\mu} & = & \hat{\mathbf{\Sigma}}_{i}\left(\mathbf{1}\frac{\mu}{\sigma}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{l}\right).\end{eqnarray}
+ This is a standard result from normal Bayesian linear regression
+\citep{BishopPRML}. Here, $\bar{\mathbf{Z}}$ is a $D\times K$ matrix
+such that row $d$ of $\mathbf{\bar{Z}}$ is $\bar{z}_{d}$, and $\mathbf{a}_{l}=[a_{l,1},a_{l,2},\ldots,a_{l,D}]^{T}$.
 %p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
-%
 %\[
-%
-%
-%
 %\[
+
 \subsubsection{$p\left(a_{l,d}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$}
+
 %and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
-%
-The auxiliary variables $a_{l,d}$ must be sampled for documents $d = 1, \ldots, D$ and $l \in \mathcal{L}_d$.  The conditional posterior distribution of $a_{l,d}$ is the truncated normal distribution
-%
-\begin{equation}
-p\left(a_{l,d,} \ | \ \mathbf{z},\mathbf{Y},\mathbf{\eta}\right) \propto \frac{1}{\sqrt{2\pi}} exp \left\{ \frac{-1}{2}\left(a_{l,d} - \mathbf{\eta_l}^T\mathbf{\bar z}_d\right) \right\} I \left(a_{l,d} y_{l,d} > 0\right).
-\end{equation}
-This conditional distribution can be sampled using an inverse CDF method.
-%
-%However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
+The auxiliary variables $a_{l,d}$ must be sampled for documents $d=1,\ldots,D$
+and $l\in\mathcal{L}_{d}$. The conditional posterior distribution
+of $a_{l,d}$ is the truncated normal distribution \begin{equation}
+p\left(a_{l,d,}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\propto\frac{1}{\sqrt{2\pi}}exp\left\{ -\frac{1}{2}\left(a_{l,d}-\mathbf{\eta_{l}}^{T}\mathbf{\bar{z}}_{d}\right)\right\} I\left(a_{l,d}y_{l,d}>0\right).\end{equation}
+ This conditional distribution can be sampled using an inverse CDF
+method. %However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
 %be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
@@
 %sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
-%
 %\begin{equation}
 %p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
-%
-%
 %\begin{equation}
 %p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
-%
-%
 %\begin{equation}
@@
 %0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
-%
-%
 %\[
 %p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
-%
-%
 %\begin{equation}
@@
 %0 & otherwise\end{cases}\end{equation}
-%
-%
 %where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
 %variables excluding the response variable being sampled.
-%
-%
-\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha',\alpha\right)$}
-%
+
+\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha^{\prime},\alpha\right)$}
+
 In our model, we place a hierarchical Dirichlet prior over topic assignments.
-This prior shares many features with the hierarchical Dirichlet process and inference over this distribution proceeds in a very similar fashion.
-This flexible distribution allows for an asymmetric prior over document level distributions over topics \citep{WallachMiMc2009}.
+This flexible distribution allows for an asymmetric prior over document
+level distributions over topics \citep{WallachMiMc2009}. This prior
+shares many features with the hierarchical Dirichlet process and inference
+over this distribution proceeds in a very similar fashion.
 
-Posterior inference is performed using the {}``direct assignment'' method of \citet{TehJorBea2006}.
-%
+Posterior inference is performed using the {}``direct assignment''
+method of \citet{TehJorBea2006}. \begin{equation}
+\beta\sim{\rm Dir}\left(m_{\left(\cdot\right),1}+\alpha^{\prime},m_{\left(\cdot\right),2}+\alpha^{\prime},\ldots,m_{\left(\cdot\right),K}+\alpha^{\prime}\right)\end{equation}
  \begin{equation}
-\beta\sim {\rm Dir}\left(m_{\left(\cdot\right),1},m_{\left(\cdot\right),2},\ldots,m_{\left(\cdot\right),K}\right)
-\end{equation}
-%
-\begin{equation}
-p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}
-\end{equation}
-%
-%
-where $s\left(n,m\right)$ represents stirling numbers of the first kind.
-%
-\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha' \right)$, $p\left(\gamma \right)$}
-%
-The hyperparameters $\alpha$, $\alpha^\prime$, and $\gamma$ are given broad ${\rm Gamma}(2, 1000)$ prior distributions and sampled via the Metropolis-Hastings algorithm.
-%
+p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}
+ where $s\left(n,m\right)$ represents stirling numbers of the first
+kind.
+
+\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha'\right)$, $p\left(\gamma\right)$}
+
+The hyperparameters $\alpha$, $\alpha^{\prime}$, and $\gamma$ are
+given broad ${\rm Gamma}(1,1000)$ prior distributions and sampled
+via the Metropolis-Hastings algorithm.
 
@@
 
+\label{sec:experiments}
+
+%\begin{itemize}
+%\item Describe Data for both Experiments
+%\item Describe the Four different conditions
+%\item Describe how they are all evaluated, with and without hierarchy
+%\end{itemize}
 In the section we describe the application of HSLDA for prediction
@@
 Discharge summaries are documents that are authored by clinicians
-to summarize the course of someone who has been hospitalized. ICD-9
-codes are used mainly for billing purposes to indicate the conditions
-for which a patient was treated. Secondly, we describe using Amazon.com
-product descriptions to predict product categories.
+to summarize the course of a hospitalization. ICD-9 codes are used
+mainly for billing purposes to indicate the conditions for which a
+patient was treated. Secondly, we describe using Amazon.com product
+descriptions to predict product categories.
 
-\subsection{Diagnosis Prediction}
+\subsection{Data and Pre-Processing}
+
+\subsubsection{Diagnosis Prediction}
 
@@
 for billing purposes. These codes are known to be quite specific but
-not very sensitive {[}citation{]}. Regardless of that fact, this is
-one of the only sources for information on patient diagnoses aside
-from the free text. Aside from prediction, one of the goals is to
-compare the sensitivity of predictions from the HSLDA model in comparison
-to the codes in a case where a test closer to ground truth is available.
-For this we will compare whether predictions for the ICD-9 code associated
-with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia
-was chosen because hemoglobin values are readily available and the
-definition of anemia according the World Health Organization is approximately
-12.5, with a threshold of 12 for women and 13 for men {[}citation{]}.
+not very sensitive \citep{Birmetal2005}. Regardless of that fact,
+this is one of the only sources for information on patient diagnoses
+aside from the free text.
+%Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available.
+For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes.
+Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].
 
@@
 
-\subsection{Product Category Prediction}
+\subsubsection{Product Category Prediction}
 
 Data for these experiments were obtained partially from the Stanford
-Network Analysis Platform (SNAP) Amazon product metadata dataset and
-partially directly from the the Amazon.com website. The product ID's
-and categorizations were obtained from the SNAP dataset and the product
-descriptions were obtained directly from the website.
+Network Analysis Platform (SNAP) Amazon product metadata dataset \citep{SNAP}
+and partially directly from the the Amazon.com website \citep{AMAZON}.
+The product ID's and categorizations were obtained from the SNAP dataset
+and the product descriptions were obtained directly from the website.
 
@@
 
+\subsection{Comparison Models}
+
+We applied HSLDA, along with three other closely related models, to
+the clinical data and the retail product data. Specifically, we evaluate
+models including sLDA with independent regressors (hierarchical constraints
+on labels ignored), HSLDA fit by performing LDA followed by tree-conditional
+regressions, and HSLDA fit with fixed random regression parameters.
+We will refer to the three comparison models as the sLDA model, the
+separate HSLDA model, and the random HSLDA model, respectively. These
+models were chosen as to highlight performance in absence of hierarchical
+constraints, the effect of the combined inference procedure, and regression
+performance attributable solely to the hierarchical constraints.
+
+We evaluated model performance for all three models with a range of
+values for $\mu$ ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\} )$,
+the mean prior parameter for regression parameters.
+
 \subsection{Evaluation}
-
-The two main methods of evaluation for the model are prediction and
-topic quality. We compare model performance against 3 similar models
-to demonstrate that each component of the model is important for performance.
-Specifically, we evaluate models including independent regressors
-+ sLDA (hierarchical constraints on labels ignored), HSLDA fit by
-running LDA first then running tree-conditional regressions, and HSLDA
-fit with fixed random regression parameters.
+%\begin{itemize}
+%\item Talk about hierarchy and no hierarchy
+%\item Make explicit calculation of sensitivity and 1-specificity
+%\end{itemize}
+For each dataset, a held out set of 1,000 documents and accompanying
+labels were used for evaluation. The two main methods of evaluation
+for the model are prediction and topic quality. To evaluate predictive
+performance for all comparison models equivalently, each model was
+evaluated with two methods. The first method augmented the observed
+labels in the held out set as well as their ancestors in the hierarchy
+as
 
@@
 TBD
-
-\subsection{Related Work}
-
-\label{sec:related_work} Latent dirichlet allocation (LDA) is a generative
-probabilistic model of corpora that represents documents as a mixed
-membership bag-of-words. Also known as topic models, these models
-infer the latent structure, or topics, of documents in a corpus. Each
-document is represented as a collection of words, generated from a
-set of topic assignments (one for each word), where each topic assignment
-is drawn from a distribution over topics \citep{Blei2003}.
-
-Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
-an exponential family response variable. Although there are many models
-for making predictions based on free text, sLDA is unique in that
-it is a generative model, it represents documents as a mixed-membership,
-and constrains the inference of the latent structure of the documents
-by its predictability of the response variable. In other words, sLDA
-infers topics such that the model is capable of a high predictive
-likelihood for words in a document and the response variable associated
-with a document. This approach has been shown to outperform both LASSO
-(L1 regularized least squares regression) and LDA followed by least
-squares regression \citep{BleiMcAuliffe2008}.
-
-There have been many models that incorporate both latent models of
-text and some form of supervision {[}citations{]}. One set of models
-that are particularly relevant to HSLDA are Chang and Blei's {[}citation{]}
-hierarchical models for document networks. In that family of models,
-they encountered a similar scenario where the lack of a link did not
-truly indicate absence. In our model, the lack of a code in the hierarchy
-being assigned does not necessarily indicate absence. Therefore, as
-in the work of Chang and Blei, we employ regularization in the form
-of a negative prior on the regression parameters to provide a prior
-that indicates a bias towards being truly negative in the absence
-of a code.
-
-% move to related work?
-While there have been some attempts to automatically classify groups
-of patients as a potential preliminary step to ICD-9 code assignment
-\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
-automatic assignment of ICD-9 codes to medical text became a more
-prevalent research topic only in the last few years. A subset of earlier
-work proposed various methods on small corpora, based on a few specific
-diseases \citep{Rao2003} but the most recent and promising work on
-the subject was inspired by the 2007 Medical NLP Challenge: {}``International
-Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
-(website). Most of the classification strategies included word matching
-and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
-The data set given to the participants consisted only of documents
-that were 1-2 lines each and all of the documents were radiology reports
-- clearly limiting the scope of potential ICD-9 codes which could
-be assigned. The only paper which has attempted to work with a document
-scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
-Lita proposed support vector machine and bayesian ridge regression
-methods to assign appropriate labels to the documents but did not
-utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
 

### 2011-05-30 paper/main.tex
Updated figure and minor changes to evaluation

@@
 \subsection{Evaluation}
+
 %\begin{itemize}
@@
 performance for all comparison models equivalently, each model was
-evaluated with two methods. The first method augmented the observed
-labels in the held out set as well as their ancestors in the hierarchy
-as
-
-\subsubsection{Prediction}
+evaluated with two methods. The first evaluation method augments the
+observed labels in the held out set with their ancestors and considers
+all other non-existant labels to be negative. The second method ignores
+the ancestors of the observed labels in the held out set and considers
+all other non-existant labels to be negative. This uniform treatment
+of ancestors allows for a fair comparison of the models.
 
 The two measures for predictive performance used here include the
-true positive rate and the false positive rate. We evaluate model
-performance on held out data. A more ideal evaluation of performance
-would include a manually labeled hierarchy since it is well known
-that ICD-9 codes have a relatively low sensitivity.
-
-Performance was evaluated against $p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$.
-
-\subsubsection{Topic Quality and Character}
-
-TBD
+true positive rate and the false positive rate evaluated based on
+$p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$ for each label in each
+model.
 

### 2011-05-31 paper/main.tex
Split off model section

@@
 \section{Model}
-
-\label{sec:model} We define here a hierarchically supervised LDA
-model. Although we will focus on document modeling in our description
-and experiments, this model applies equally well to other collections
-of discrete data with hierarchically constrained labels.
-
-We assume a pre-specified set of labels $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
-Each document is assigned a response of either -1 or 1 for at least
-one, but potentially many labels in $\mathcal{L}$. The label, $l$,
-for a document, $d$, will be used interchangeably to refer to the
-observed response of document $d$ to label $l$. The label set is
-assumed to be structured as an {}``is-a'' hieararchy.
-To understand this, consider a hierarchy where label $l_{1}$ is a
-parent of label $l_{2}$. If document $d$ has a positive response
-to label $l_{2}$ then it will also have a positive response to label
-$l_{1}$. Conversely, if document $d$ has a negative response to
-label $l_{1}$ then it will also has a negative response to $l_{2}$.
-To capture this hierarchical structure we model the labeling of documents
-using a generative cascade of conditional probit regression models.
-
-% generalize to arbitrary codes (not just ICD-9's
-%
-\begin{figure}[h]
-%tbp] %  figure placement: here, top, bottom, or page
- \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}
-
-\label{fig:example} 
-\end{figure}
-
-The fixed parameters of the model are the number of topics $K$, the
-number of unique words in the vocabulary $V$, the number of documents
-$D$, as well as the mean, $\mu$, and the standard deviation, $\sigma$,
-used in a normal prior distribution. The hyper-parameters $\alpha^{\prime}$,
-$\alpha$, and $\gamma$ are weight parameters for Dirichlet prior
-distributions. We will denote the $K$-dimensional Dirichlet distribution
-as Dir$_{K}(\cdot)$ and the $K$ dimensional normal distribution
-as $\mathcal{N}_{K}(\cdot)$.
-
-We will now describe the stochastic generative process which defines
-our model. The graphical model is show in Figure~\ref{fig:example}.
-%Given the number of topics, K, and broad gamma priors on hyperparameters,
-%the generative process is as follows:
-
-\begin{enumerate}
-\item For each topic $k=1,\ldots,K$:
-
-\begin{itemize}
-\item Draw a distribution over words $\mathbf{\phi_{k}}\sim{\rm Dir}_{V}\left(\gamma\mathbf{1}\right)$,
-where $\mathbf{1}$ is a vector of ones of length $V$ 
-\end{itemize}
-\item For each label $l\in\mathcal{L}$:
-
-\begin{itemize}
-\item Draw the regression coefficients $\mathbf{\eta_{l}}\mid\sigma\sim\mathcal{N}_{K}\left(\mu I_{K},\sigma I_{K}\right)$,
-where $I_{K}$ is the $K$ dimensional identity matrix 
-\end{itemize}
-\item Draw a prior over topic proportions $\mathbf{\beta}\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}\right)$
-\item For each document $d=1,\ldots,D$:
-
-\begin{itemize}
-\item Draw topic proportions $\mathbf{\theta_{d}}\mid\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\mathbf{\beta}\right)$ 
-\item For $n=1,\ldots,N_{d}$:
-
-\begin{itemize}
-\item Draw topic assignment $z_{n,d}\mid\mathbf{\theta_{d}}\sim{\rm Multinomial}\left(\mathbf{\theta_{d}}\right)$ 
-\item Draw word $w_{n,d}\mid z_{n,d},\mathbf{\beta_{1:K}}\sim{\rm Multinomial}\left(\mathbf{\beta_{z_{n,d}}}\right)$ 
-\end{itemize}
-\item For each label $l\in\mathcal{L}$:
-
-\begin{itemize}
-\item Draw $a_{l,d}\mid z_{1:N_{d},d},\eta_{l},y_{parent(l),d}\sim\begin{cases}
-\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right), & y_{parent(l)}=1\\
-\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
-
-\item Set the response variable \[
-y_{l,d}\mid a_{l,d}=\begin{cases}
-\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}and \ensuremath{y_{{\rm parent}(l),d}=1}}\\
--1 & \text{otherwise}\end{cases}\]
-
-\end{itemize}
-\end{itemize}
-\end{enumerate}
-This type of generative model is known as a probit regression model.
-Probit regression models are a type of discriminative probabilistic
-model similar to logistic regression. However, instead of using the
-logistic sigmoid as the link function, the probit regression model
-uses the CDF for a standard normal distribution - the inverse of which
-is known as the probit. In this case, the regression is conditional
-on the parents according to the constraints of the labeling hierarchy.
-The latent variables $a_{l,d}$ utilized here are also known as an
-auxiliary variables because the are introduced to make exact Gibbs
-sampling possible and are not of primary interest.
-
-Given that negative labels are uncommon and that the absence of a
-label is not equivalent to a negative label, we apply an informative
-prior to the regression parameters, $\mathbf{\eta}_{\mathcal{L}}$,
-in the form of a negative prior that encodes a bias towards being
-truly negative in the absence of a label.
+\label{sec:model}
+\input{model}
 

### 2011-05-31 paper/model.tex
Split off model section

\label{sec:model} We define here a hierarchically supervised LDA
model. Although we will focus on document modeling in our description
and experiments, this model applies equally well to other collections
of discrete data with hierarchically constrained labels.

We assume a pre-specified set of labels $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
Each document is assigned a response of either -1 or 1 for at least
one, but potentially many labels in $\mathcal{L}$. The label, $l$,
for a document, $d$, will be used interchangeably to refer to the
observed response of document $d$ to label $l$. The label set is
assumed to be structured as an {}``is-a'' hieararchy.
To understand this, consider a hierarchy where label $l_{1}$ is a
parent of label $l_{2}$. If document $d$ has a positive response
to label $l_{2}$ then it will also have a positive response to label
$l_{1}$. Conversely, if document $d$ has a negative response to
label $l_{1}$ then it will also has a negative response to $l_{2}$.
To capture this hierarchical structure we model the labeling of documents
using a generative cascade of conditional probit regression models.

% generalize to arbitrary codes (not just ICD-9's
%
\begin{figure}[h]
%tbp] %  figure placement: here, top, bottom, or page
 \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}

\label{fig:example} 
\end{figure}

The fixed parameters of the model are the number of topics $K$, the
number of unique words in the vocabulary $V$, the number of documents
$D$, as well as the mean, $\mu$, and the standard deviation, $\sigma$,
used in a normal prior distribution. The hyper-parameters $\alpha^{\prime}$,
$\alpha$, and $\gamma$ are weight parameters for Dirichlet prior
distributions. We will denote the $K$-dimensional Dirichlet distribution
as Dir$_{K}(\cdot)$ and the $K$ dimensional normal distribution
as $\mathcal{N}_{K}(\cdot)$.

We will now describe the stochastic generative process which defines
our model. The graphical model is show in Figure~\ref{fig:example}.
%Given the number of topics, K, and broad gamma priors on hyperparameters,
%the generative process is as follows:

\begin{enumerate}
\item For each topic $k=1,\ldots,K$:

\begin{itemize}
\item Draw a distribution over words $\mathbf{\phi_{k}}\sim{\rm Dir}_{V}\left(\gamma\mathbf{1}\right)$,
where $\mathbf{1}$ is a vector of ones of length $V$ 
\end{itemize}
\item For each label $l\in\mathcal{L}$:

\begin{itemize}
\item Draw the regression coefficients $\mathbf{\eta_{l}}\mid\sigma\sim\mathcal{N}_{K}\left(\mu I_{K},\sigma I_{K}\right)$,
where $I_{K}$ is the $K$ dimensional identity matrix 
\end{itemize}
\item Draw a prior over topic proportions $\mathbf{\beta}\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}\right)$
\item For each document $d=1,\ldots,D$:

\begin{itemize}
\item Draw topic proportions $\mathbf{\theta_{d}}\mid\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\mathbf{\beta}\right)$ 
\item For $n=1,\ldots,N_{d}$:

\begin{itemize}
\item Draw topic assignment $z_{n,d}\mid\mathbf{\theta_{d}}\sim{\rm Multinomial}\left(\mathbf{\theta_{d}}\right)$ 
\item Draw word $w_{n,d}\mid z_{n,d},\mathbf{\beta_{1:K}}\sim{\rm Multinomial}\left(\mathbf{\beta_{z_{n,d}}}\right)$ 
\end{itemize}
\item For each label $l\in\mathcal{L}$:

\begin{itemize}
\item Draw $a_{l,d}\mid z_{1:N_{d},d},\eta_{l},y_{parent(l),d}\sim\begin{cases}
\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right), & y_{parent(l)}=1\\
\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$

\item Set the response variable \[
y_{l,d}\mid a_{l,d}=\begin{cases}
\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}and \ensuremath{y_{{\rm parent}(l),d}=1}}\\
-1 & \text{otherwise}\end{cases}\]

\end{itemize}
\end{itemize}
\end{enumerate}
This type of generative model is known as a probit regression model.
Probit regression models are a type of discriminative probabilistic
model similar to logistic regression. However, instead of using the
logistic sigmoid as the link function, the probit regression model
uses the CDF for a standard normal distribution - the inverse of which
is known as the probit. In this case, the regression is conditional
on the parents according to the constraints of the labeling hierarchy.
The latent variables $a_{l,d}$ utilized here are also known as an
auxiliary variables because the are introduced to make exact Gibbs
sampling possible and are not of primary interest.

Given that negative labels are uncommon and that the absence of a
label is not equivalent to a negative label, we apply an informative
prior to the regression parameters, $\mathbf{\eta}_{\mathcal{L}}$,
in the form of a negative prior that encodes a bias towards being
truly negative in the absence of a label.

### 2011-05-31 paper/model.tex
fixed most of the bolding issues in model.tex

@@
 number of unique words in the vocabulary $V$, the number of documents
-$D$, as well as the mean, $\mu$, and the standard deviation, $\sigma$,
+$D$, as well as the mean, $\boldsymbol\mu$, and the standard deviation, $\sigma$,
 used in a normal prior distribution. The hyper-parameters $\alpha^{\prime}$,
@@
 \begin{enumerate}
-\item For each topic $k=1,\ldots,K$:
+\item For each topic $k=1,\ldots,K$
 
 \begin{itemize}
-\item Draw a distribution over words $\mathbf{\phi_{k}}\sim{\rm Dir}_{V}\left(\gamma\mathbf{1}\right)$,
-where $\mathbf{1}$ is a vector of ones of length $V$ 
+\item Draw a distribution over words $\boldsymbol\phi_{k}\sim{\rm Dir}_{V}(\gamma\mathbf{1}_V)$%,
+%where $\mathbf{1}$ is a vector of ones of length $V$ 
+\end{itemize}
+\item For each label $l\in\mathcal{L}$
+
+\begin{itemize}
+\item Draw a regression coefficient $\boldsymbol\eta_{l}\mid\mu,\sigma\sim\mathcal{N}_{K}(\mu \mathbf{1}_K,\sigma \mathbf{I}_{K})$,
+where $\mathbf{I}_{K}$ is the $K$ dimensional identity matrix 
+\end{itemize}
+\item Draw the global topic proportions $\boldsymbol\beta\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}_K\right)$
+\item For each document $d=1,\ldots,D$
+
+\begin{itemize}
+\item Draw topic proportions $\boldsymbol\theta_d\mid\boldsymbol\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\boldsymbol\beta\right)$ 
+\item For $n=1,\ldots,N_{d}$
+
+\begin{itemize}
+\item Draw topic assignment $z_{n,d}\mid\boldsymbol\theta_d\sim{\rm Multinomial}(\boldsymbol\theta_d)$ 
+\item Draw word $w_{n,d}\mid z_{n,d},\boldsymbol\beta_{1:K}\sim{\rm Multinomial}(\boldsymbol\beta_{z_{n,d}})$ 
 \end{itemize}
@@
 \begin{itemize}
-\item Draw the regression coefficients $\mathbf{\eta_{l}}\mid\sigma\sim\mathcal{N}_{K}\left(\mu I_{K},\sigma I_{K}\right)$,
-where $I_{K}$ is the $K$ dimensional identity matrix 
-\end{itemize}
-\item Draw a prior over topic proportions $\mathbf{\beta}\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}\right)$
-\item For each document $d=1,\ldots,D$:
-
-\begin{itemize}
-\item Draw topic proportions $\mathbf{\theta_{d}}\mid\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\mathbf{\beta}\right)$ 
-\item For $n=1,\ldots,N_{d}$:
-
-\begin{itemize}
-\item Draw topic assignment $z_{n,d}\mid\mathbf{\theta_{d}}\sim{\rm Multinomial}\left(\mathbf{\theta_{d}}\right)$ 
-\item Draw word $w_{n,d}\mid z_{n,d},\mathbf{\beta_{1:K}}\sim{\rm Multinomial}\left(\mathbf{\beta_{z_{n,d}}}\right)$ 
-\end{itemize}
-\item For each label $l\in\mathcal{L}$:
-
-\begin{itemize}
-\item Draw $a_{l,d}\mid z_{1:N_{d},d},\eta_{l},y_{parent(l),d}\sim\begin{cases}
-\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right), & y_{parent(l)}=1\\
-\mathcal{N}\left(\bar{z}^{T}\eta_{l},1\right)I\left(a_{l,d}<0\right), & y_{parent(l)}=-1\end{cases}$ where $\bar{z_{d}}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \eta_l \sim \mathcal{N} \left(\bar z_d^{T} \eta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
+\item Draw $a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\beta_{l},y_{\mathrm{pa}(l),d}\sim\begin{cases}
+\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\beta_{l},1), & y_{\mathrm{pa}(l)}=1\\
+\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\beta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l)}=-1\end{cases}$ \\where $\bar{z}_{d}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
 
@@
 y_{l,d}\mid a_{l,d}=\begin{cases}
-\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}and \ensuremath{y_{{\rm parent}(l),d}=1}}\\
+\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0} and \ensuremath{y_{{\rm \mathrm{pa}}(l),d}=1}}\\
 -1 & \text{otherwise}\end{cases}\]
@@
 label is not equivalent to a negative label, we apply an informative
-prior to the regression parameters, $\mathbf{\eta}_{\mathcal{L}}$,
+prior to the regression parameters, $\mathbf{\boldsymbol\beta}_{\mathcal{L}}$,
 in the form of a negative prior that encodes a bias towards being

### 2011-05-31 paper/main.tex
Split inference section out.

@@
 \section{Inference}
-
-\label{sec:inference} In the Bayesian approach to statistical modeling,
-the primary task of inference is to find the posterior distribution
-over the unobserved parameters of the model. However, it is often
-possible and desirable to integrate over certain variables in the
-model, also known as collapsing. In our model, it will often be the
-case that the set of labels $\mathcal{L}$ is not fully observed for
-every document. We will define $\mathcal{L}_{d}$ to be the subset
-of labels which have been observed for document $d$. It is straightforward
-to integrate out the variables $a_{l^{\prime},d}$ and $y_{l^{\prime},d}$
-for $l^{\prime}\in\mathcal{L}\backslash\mathcal{L}_{d}$ from the
-full generative model. We can also integrate out the parameters $\mathbf{\phi_{1:K}}$
-and $\mathbf{\theta_{1:D}}$ as in \citet{Griffiths04}. Therefore,
-in our model the latent variables are $\mathbf{z}=\{z_{1:N_{d},d}\}_{d=1,\ldots,D},\mathbf{\eta}=\{\mathbf{\eta_{l}}\}_{l\in\mathcal{L}},\mathbf{a}=\{a_{l^{\prime},d}\}_{l^{\prime}\in\mathcal{L}_{d},d=1,\ldots,D},\mathbf{\beta},\alpha,\alpha^{\prime}$
-and $\gamma$.
-
-The posterior distribution we seek cannot be solved in closed form.
-This is often the case in evaluating posterior distributions of non-trivial
-probabilistic models. We will appeal to one of the common methods
-for approximating posterior distributions in the face of intractable
-normalization factors: Markov chain Monte Carlo (MCMC) sampling. Since
-in this model it is possible to sample from the conditional distributions
-for all variables we will use the Gibbs sampling algorithm to obtain
-an approximation to this posterior. %In particular, we will derive a collapsed Gibbs sampler.
-%Given an observation of a set of observed labels and a document, the posterior distribution for the latent variables is
-%\begin{equation}
-%p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
-%\[
-%=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\eta,\alpha,\alpha',\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]
-
-\subsection{Gibbs Sampler}
-
-We derive a collapsed Gibbs sampler for this model by considering
-the individual conditional probability distributions for each of the
-unobserved variables. We use the notation $\mathbf{z}_{-(n,d)}$ to
-denote $\mathbf{z}\backslash z_{n,d}$. %, marginalizing
-%$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
-%in LDA models see \citet{Griffiths04}.
-
-\subsubsection{$p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)$}
-
-First we consider the conditional distribution of the assignment variable
-for each word $n=1,\ldots,N_{d}$ in documents $d=1,\ldots,D$. The
-conditional distribution does not include $\theta_{1:D}$ and $\phi_{1:K}$
-because they have been integrated out as in the collapsed Gibbs sampler
-\citep{Griffiths04}. The conditional distribution of $z_{n,d}$ is
-proportional to the joint distribution of its markov blanket. %\begin{equation}
-%p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\end{equation}
-%Due to the factorization of this model, we can rewrite the joint distribution
-%as the following:
-\begin{equation}
-p\left(z_{n,d}\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto\prod_{l\in\mathcal{L}_{d}}p\left(a_{l,d}\mid\mathbf{z},\eta_{l}\right)p\left(z_{n,d}\ |\ \mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\alpha,\beta,\gamma\right).\end{equation}
- The product is only over the subset of labels $\mathcal{L}_{d}$
-which have been observed for document $d$. By isolating terms that
-depend on $z_{n,d}$ and absorbing all other terms into a normalizing
-constant as in \citep{Griffiths04} we find \begin{eqnarray}
-\lefteqn{p\left(z_{n,d}=k\mid\mathbf{z_{-\left(n,d\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\beta,\gamma\right)\propto}\nonumber \\
- & \hspace{2cm}\left(n_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\beta_{k}\right)\frac{n_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(n_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)}\prod_{l\in\mathcal{L}_{d}}\exp\left\{ -\frac{\left(\bar{z}_{d}^{T}\eta_{l}-a_{l,d}\right)^{2}}{2}\right\} .\label{eq:z-likelihood}\end{eqnarray}
- Here, $n_{v,d}^{k,-\left(n,d^{\prime}\right)}$ represents the number
-of words of type $v$ in document $d$ assigned to topic $k$ omitting
-the $n^{{\rm th}}$ word of document $d^{\prime}$. The notation $(\cdot)$
-in the subscript means the count resulting from summing over the omitted
-subscript variable. Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\eta},\alpha,\mathbf{\beta},\gamma\right)$
-can be sampled through enumeration.
-
-\subsubsection{$p\left(\mathbf{\eta_{l}}\mid\mathbf{z},\mathbf{a},\sigma\right)$}
-
-We now consider the conditional distribution of the regression coefficients
-$\mathbf{\eta_{l}}$ for $l\in\mathcal{L}$. Given that $\mathbf{\eta_{l}}$
-and $a_{l,d}$ are distributed normally, the posterior distribution
-of $\mathbf{\eta_{l}}$ is normally distributed with mean $\hat{\mu}$
-and covariance $\hat{\mathbf{\Sigma}}$ such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
-\begin{eqnarray}
-\hat{\mathbf{\Sigma}}^{-1} & = & \mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}\\
-\hat{\mu} & = & \hat{\mathbf{\Sigma}}_{i}\left(\mathbf{1}\frac{\mu}{\sigma}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{l}\right).\end{eqnarray}
- This is a standard result from normal Bayesian linear regression
-\citep{BishopPRML}. Here, $\bar{\mathbf{Z}}$ is a $D\times K$ matrix
-such that row $d$ of $\mathbf{\bar{Z}}$ is $\bar{z}_{d}$, and $\mathbf{a}_{l}=[a_{l,1},a_{l,2},\ldots,a_{l,D}]^{T}$.
-%p\left(\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
-%\[
-%\[
-
-\subsubsection{$p\left(a_{l,d}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)$}
-
-%and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
-The auxiliary variables $a_{l,d}$ must be sampled for documents $d=1,\ldots,D$
-and $l\in\mathcal{L}_{d}$. The conditional posterior distribution
-of $a_{l,d}$ is the truncated normal distribution \begin{equation}
-p\left(a_{l,d,}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\propto\frac{1}{\sqrt{2\pi}}exp\left\{ -\frac{1}{2}\left(a_{l,d}-\mathbf{\eta_{l}}^{T}\mathbf{\bar{z}}_{d}\right)\right\} I\left(a_{l,d}y_{l,d}>0\right).\end{equation}
- This conditional distribution can be sampled using an inverse CDF
-method. %However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
-%be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
-%chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
-%a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
-%-1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
-%$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
-%since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
-%$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
-%to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
-%sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
-%\begin{equation}
-%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\eta}\right)\end{equation}
-%\begin{equation}
-%p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
-%\begin{equation}
-%p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
-%1, & y_{parent_{l,c}}=-1\\
-%0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
-%\[
-%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\eta}\right)\]
-%\begin{equation}
-%=\begin{cases}
-%\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
-%trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
-%trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
-%0 & otherwise\end{cases}\end{equation}
-%where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
-%variables excluding the response variable being sampled.
-
-\subsubsection{$p\left(\beta\mid\mathbf{z},\alpha^{\prime},\alpha\right)$}
-
-In our model, we place a hierarchical Dirichlet prior over topic assignments.
-This flexible distribution allows for an asymmetric prior over document
-level distributions over topics \citep{WallachMiMc2009}. This prior
-shares many features with the hierarchical Dirichlet process and inference
-over this distribution proceeds in a very similar fashion.
-
-Posterior inference is performed using the {}``direct assignment''
-method of \citet{TehJorBea2006}. \begin{equation}
-\beta\sim{\rm Dir}\left(m_{\left(\cdot\right),1}+\alpha^{\prime},m_{\left(\cdot\right),2}+\alpha^{\prime},\ldots,m_{\left(\cdot\right),K}+\alpha^{\prime}\right)\end{equation}
- \begin{equation}
-p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\beta\right)=\frac{\Gamma\left(\alpha\beta_{k}\right)}{\Gamma\left(\alpha\beta_{k}+n_{d,k}\right)}s\left(n_{d,k},m\right)\left(\alpha\beta_{k}\right)^{m}\end{equation}
- where $s\left(n,m\right)$ represents stirling numbers of the first
-kind.
-
-\subsubsection{$p\left(\alpha\right)$, $p\left(\alpha'\right)$, $p\left(\gamma\right)$}
-
-The hyperparameters $\alpha$, $\alpha^{\prime}$, and $\gamma$ are
-given broad ${\rm Gamma}(1,1000)$ prior distributions and sampled
-via the Metropolis-Hastings algorithm.
-
+\label{sec:inference}
+\input{inference}
 \section{Experiments}

### 2011-05-31 paper/related_work.tex
moved related work around

--- 2011-05-31 paper/main.tex
@@
-\documentclass{article}
-\usepackage{nips10submit_e,times}
-%\documentstyle[nips07submit_09,times]{article}
-\usepackage[square,numbers]{natbib}
-\usepackage{amsmath, epsfig}
-\usepackage{amsfonts}
-\usepackage{subfigure}
-\usepackage{graphicx}
-\usepackage{amsfonts}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{easybmat}
-\usepackage{footmisc}
-\usepackage{lscape}
-\renewcommand\algorithmiccomment[1]{// \textit{#1}}
-%
-\newcommand{\ignore}[1]{}
-\newcommand{\comment}[1]{}
-\DeclareMathOperator*{\argmax}{arg\,max}
-
-\title{Hierarchically Supervised Latent Dirichlet Allocation}
-%Supervised Topic Modeling in Clinical Text}
-
-\author{
-Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhadad \hspace{1cm} Frank Wood\\
-Columbia University, New York, NY 10027, USA \\
-\texttt{\{ajp9009@dbmi,bartlett@stat,noemie@dbmi,fwood@stat\}.columbia.edu}
-%\texttt{pfau@neurotheory.columbia.edu} 
-%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
-}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to \LaTeX{} to determine where to break
-% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
-% puts 3 of 4 authors names on the first line, and the last on the second
-% line, try using \AND instead of \And before the third author name.
-
-\newcommand{\fix}{\marginpar{FIX}}
-\newcommand{\new}{\marginpar{NEW}}
-\newcommand{\X}{\mathcal{X}}
-
-\nipsfinalcopy
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-\section{Introduction}
-\label{sec:introduction}
-\input{introduction}
-
-%\section{Background}
-
-\subsection{Related Work}
-
-\label{sec:related_work} Latent dirichlet allocation (LDA) is a generative
+Latent dirichlet allocation (LDA) is a generative
 probabilistic model of corpora that represents documents as a mixed
@@
 % move to related work?
-While there have been some attempts to automatically classify groups
-of patients as a potential preliminary step to ICD-9 code assignment
-\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
-automatic assignment of ICD-9 codes to medical text became a more
-prevalent research topic only in the last few years. A subset of earlier
-work proposed various methods on small corpora, based on a few specific
-diseases \citep{Rao2003} but the most recent and promising work on
-the subject was inspired by the 2007 Medical NLP Challenge: {}``International
-Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
-(website). Most of the classification strategies included word matching
-and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
-The data set given to the participants consisted only of documents
-that were 1-2 lines each and all of the documents were radiology reports
-- clearly limiting the scope of potential ICD-9 codes which could
-be assigned. The only paper which has attempted to work with a document
-scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
-Lita proposed support vector machine and bayesian ridge regression
-methods to assign appropriate labels to the documents but did not
-utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
-
-\section{Model}
-\label{sec:model}
-\input{model}
-
-\section{Inference}
-\label{sec:inference}
-\input{inference}
-\section{Experiments}
-
-\label{sec:experiments}
-
-%\begin{itemize}
-%\item Describe Data for both Experiments
-%\item Describe the Four different conditions
-%\item Describe how they are all evaluated, with and without hierarchy
-%\end{itemize}
-In the section we describe the application of HSLDA for prediction
-in two hierarchically structured domains. Firstly, we describe using
-discharge summaries to predict diagnoses, encoded as ICD-9 codes.
-Discharge summaries are documents that are authored by clinicians
-to summarize the course of a hospitalization. ICD-9 codes are used
-mainly for billing purposes to indicate the conditions for which a
-patient was treated. Secondly, we describe using Amazon.com product
-descriptions to predict product categories.
-
-\subsection{Data and Pre-Processing}
-
-\subsubsection{Diagnosis Prediction}
-
-Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
-Hospital. The data consisted of free-text discharge summaries and
-their respective ICD-9 codes. A discharge summary is a clinical report
-prepared by a physician or other health professional at the conclusion
-of a hospital stay or series of treatments. The note outlines the
-patient's chief complaint, diagnostic findings, therapy administered,
-patient's response to the chosen therapy, the treatment plan and the
-recommendations upon discharge. The ICD-9 codes used to structure
-the discharge summary data are part of a controlled terminology which
-is the international standard diagnostic classification for epidemiological,
-health management, and clinical purposes. The ICD-9 codes are organized
-in a rooted-tree structure, with each edge representing an is-a relationship
-between parent and child, such that the parent diagnosis subsumes
-the child diagnosis. For example, the code representing {}``Pneumonia
-due to adenovirus'' is a child of the code representing {}``Viral
-pneumonia'' where the former is a type of the latter. In the hospital,
-ICD-9 codes are generated manually by trained medical coders, who
-review all the information in the discharge summary.
-
-The text of the discharge summaries were pre-processed such that each
-document would be represented as counts over a 10,000 word vocabulary.
-The Natural Language Toolkit was used to tokenize the text. A vocabulary
-was identified by first sorting terms based on a global term frequency-inverse
-document frequency measure. The top 10,000 words which were not identifying
-in some way (a name, place, or identifying number) were selected for
-inclusion in the vocabulary.
-
-For each hospitalization there are usually several ICD-9 codes assigned
-for billing purposes. These codes are known to be quite specific but
-not very sensitive \citep{Birmetal2005}. Regardless of that fact,
-this is one of the only sources for information on patient diagnoses
-aside from the free text.
-%Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available.
-For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes.
-Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].
-
-We worked within the guidelines of the Health Insurance Portability
-and Accountability Act (HIPAA), which protects patient privacy and
-the security of potentially identifying medical material, known as
-personal health information (PHI). HIPAA covers any information within
-a medical record that was created, used, or disclosed during the course
-of providing a health care service and that can be used to identify
-an individual. This study was approved by the Institutional Review
-Board.
-
-\subsubsection{Product Category Prediction}
-
-Data for these experiments were obtained partially from the Stanford
-Network Analysis Platform (SNAP) Amazon product metadata dataset \citep{SNAP}
-and partially directly from the the Amazon.com website \citep{AMAZON}.
-The product ID's and categorizations were obtained from the SNAP dataset
-and the product descriptions were obtained directly from the website.
-
-We were able to deduce the structure of the hierarchy for the Amazon.com
-products directly since all ancestors in the hierarchy were included
-with each category label. For example, {}``DVD / Genres / Science
-Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
-for the DVD, {}``The Time Machine''. Each product was labeled with
-multiple categories.
-
-The vocabulary for this experiment was created by including the most
-frequent 30K words omitting stopwords.
-
-\subsection{Comparison Models}
-
-We applied HSLDA, along with three other closely related models, to
-the clinical data and the retail product data. Specifically, we evaluate
-models including sLDA with independent regressors (hierarchical constraints
-on labels ignored), HSLDA fit by performing LDA followed by tree-conditional
-regressions, and HSLDA fit with fixed random regression parameters.
-We will refer to the three comparison models as the sLDA model, the
-separate HSLDA model, and the random HSLDA model, respectively. These
-models were chosen as to highlight performance in absence of hierarchical
-constraints, the effect of the combined inference procedure, and regression
-performance attributable solely to the hierarchical constraints.
-
-We evaluated model performance for all three models with a range of
-values for $\mu$ ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\} )$,
-the mean prior parameter for regression parameters.
-
-\subsection{Evaluation}
-
-%\begin{itemize}
-%\item Talk about hierarchy and no hierarchy
-%\item Make explicit calculation of sensitivity and 1-specificity
-%\end{itemize}
-For each dataset, a held out set of 1,000 documents and accompanying
-labels were used for evaluation. The two main methods of evaluation
-for the model are prediction and topic quality. To evaluate predictive
-performance for all comparison models equivalently, each model was
-evaluated with two methods. The first evaluation method augments the
-observed labels in the held out set with their ancestors and considers
-all other non-existant labels to be negative. The second method ignores
-the ancestors of the observed labels in the held out set and considers
-all other non-existant labels to be negative. This uniform treatment
-of ancestors allows for a fair comparison of the models.
-
-The two measures for predictive performance used here include the
-true positive rate and the false positive rate evaluated based on
-$p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$ for each label in each
-model.
-
-\section{Results}
-\label{sec:results}
-
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
-\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
-\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
-\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.
-In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.
-Top row:  \subref{1a} includes ancestor prediction performance,  \subref{1b} results are for given (leaf) labels alone.
-Bottom row: \subref{1c} are the sensitivity curves from \subref{1b} aligned on threshold value,  \subref{1d} are the 1-specificity curves from \subref{1b} aligned on threshold value.}
-\label{fig:icd9}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[][\label{2a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors_amazon}}
-\subfigure[][\label{2b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs_amazon}}
-\subfigure[][\label{2c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs_amazon}}
-\subfigure[][\label{2d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
-\caption{Out-of-sample Amazon product code predictions from product free-text descriptions.
-In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.
-Top row:  \subref{2a} includes ancestor prediction performance,  \subref{2b} results are for given (leaf) labels alone.
-Bottom row: \subref{2c} are the sensitivity curves from \subref{2b} aligned on threshold value,  \subref{2d} are the 1-specificity curves from \subref{2b} aligned on threshold value.}
-\label{fig:icd9}
-\end{center}
-\end{figure}
-
-%We ran the Gibbs sampler for two days on 500 documents with 20 topics to learn the sLDA model which we then followed by a prediction of ICD-9 codes for 100 documents.
-The prediction resulted in 38 documents with at least 1 ICD-9 code assignment while some had over 200 assignments.
-
-%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
-%   \centering
-%   \includegraphics[width=4in]{number_of_codes_histogram} 
-%   \caption{distribution of ICD-9 code assignments over all of the 100 documents predicted}
-%   \label{fig:example}
-%\end{figure}
-
-%The accuracy of the prediction was tested against the actual ICD-9 codes in the EHR record for each discharge summary.  Different statistical measures were applied to asses the predictions.
-
-%\begin{figure}[h]%tbp] %  figure placement: here, top, bottom, or page
-%   \centering
-%   \includegraphics[width=4in]{signal_detection} 
-%   \caption{sensitivity, specificity, precision, recall and F-measure.}
-%   \label{fig:example}
-%\end{figure}
-
-%Although the sensitivty is not as high as we would have hoped, it is important to note that the sLDA was run on a very limited sample of documents (500) due to time constraints.
-Additionally, the “gold standard” data used are actual hospital results which are very noisy given the incompleteness and introduction of human error.
-
-\section{Discussion}
-\label{sec:discussion}
-\input{discussion}
-%We have proposed a new method for prediction of ICD-9 codes given a medical discharge summary note.
-Our method explicitly takes advantage of the ICD-9 hierarchical IS-A relationships as well as the word distributions in each document.
-We used Gibbs sampling for a supervised latent dirichlet allocation model to learn a distribution of words over twenty topics and their appropriate ICD-9 codes (response variable).
-The model was then tested on 100 discharge summaries for which the ICD-9 diagnoses were known.
-The amount of ICD-9 codes produced for some discharge summaries was quite high because of the way the model worked with the IS-A hierarchy, highlighting the completeness of ICD-9 coding when using this sLDA model.
-Overall, the model had extremely good specificity while the sensitivity was lower than anticipated.
-Given our limited set of documents, and very noisy gold standard it would not be accurate to discount the use of sLDA in medical document classification and automatic coding.
-%In future work, we plan to significantly expand the set of documents used to train the model (up to 15,000), as well as run several different iterations to learn the optimal topic number.
-
-%A more automated approach for coding patient visits with the ICD-9 hierarchy has great promise to  reduce health care costs and increase completeness of medical coded data.
-We hope that this implementation of supervised LDA will serve as preliminary work for a much larger study with more documents in the future.
-
-\begin{small} \bibliographystyle{plainnat} \bibliographystyle{plainnat}
-\bibliographystyle{plainnat}
-\bibliography{refs}
-
-\end{small}
-
-\end{document}
+%While there have been some attempts to automatically classify groups
+%of patients as a potential preliminary step to ICD-9 code assignment
+%\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
+%automatic assignment of ICD-9 codes to medical text became a more
+%prevalent research topic only in the last few years. A subset of earlier
+%work proposed various methods on small corpora, based on a few specific
+%diseases \citep{Rao2003} but the most recent and promising work on
+%the subject was inspired by the 2007 Medical NLP Challenge: {}``International
+%Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
+%(website). Most of the classification strategies included word matching
+%and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
+%The data set given to the participants consisted only of documents
+%that were 1-2 lines each and all of the documents were radiology reports
+%- clearly limiting the scope of potential ICD-9 codes which could
+%be assigned. The only paper which has attempted to work with a document
+%scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
+%Lita proposed support vector machine and bayesian ridge regression
+%methods to assign appropriate labels to the documents but did not
+%utilize the ICD-9 hierarchy to leverage more comprehensive predictions.

### 2011-05-31 paper/main.tex

@@
 
-\subsection{Related Work}
-
-\label{sec:related_work} Latent dirichlet allocation (LDA) is a generative
-probabilistic model of corpora that represents documents as a mixed
-membership bag-of-words. Also known as topic models, these models
-infer the latent structure, or topics, of documents in a corpus. Each
-document is represented as a collection of words, generated from a
-set of topic assignments (one for each word), where each topic assignment
-is drawn from a distribution over topics \citep{Blei2003}.
-
-Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
-an exponential family response variable. Although there are many models
-for making predictions based on free text, sLDA is unique in that
-it is a generative model, it represents documents as a mixed-membership,
-and constrains the inference of the latent structure of the documents
-by its predictability of the response variable. In other words, sLDA
-infers topics such that the model is capable of a high predictive
-likelihood for words in a document and the response variable associated
-with a document. This approach has been shown to outperform both LASSO
-(L1 regularized least squares regression) and LDA followed by least
-squares regression \citep{BleiMcAuliffe2008}.
-
-There have been many models that incorporate both latent models of
-text and some form of supervision \citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}.
-One set of models that are particularly relevant to HSLDA are Chang
-and Blei's hierarchical models for document networks (Relational Topic
-Models). In that family of models, they encountered a similar scenario
-where the lack of a link did not truly indicate absence. In hierarchically
-labeled data, negative labels are uncommon and the lack of a label
-in the hierarchy is not equivalent to a negative label. Therefore,
-as in the work of Chang and Blei, we employ regularization to account
-for the lack of negative labels. This will be discussed further in
-\ref{sec:model}.
-
-% move to related work?
-While there have been some attempts to automatically classify groups
-of patients as a potential preliminary step to ICD-9 code assignment
-\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
-automatic assignment of ICD-9 codes to medical text became a more
-prevalent research topic only in the last few years. A subset of earlier
-work proposed various methods on small corpora, based on a few specific
-diseases \citep{Rao2003} but the most recent and promising work on
-the subject was inspired by the 2007 Medical NLP Challenge: {}``International
-Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
-(website). Most of the classification strategies included word matching
-and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
-The data set given to the participants consisted only of documents
-that were 1-2 lines each and all of the documents were radiology reports
-- clearly limiting the scope of potential ICD-9 codes which could
-be assigned. The only paper which has attempted to work with a document
-scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
-Lita proposed support vector machine and bayesian ridge regression
-methods to assign appropriate labels to the documents but did not
-utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
-
 \section{Model}
@@
 \input{inference}
+
+\section{Related Work}
+
+\label{sec:related_work} 
+\input{related_work}
+
 \section{Experiments}

### 2011-05-31 paper/abstract.tex
updated the abstract -- is it better?

@@
-We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply-labeled bag-of-words data.
-Out-of-sample label prediction is the primary goal of this work; however, improved dimensionality reduction is also of interest.
-We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA).
-We find that the additional signal that comes from multiple, hierarchically constrained labels substantially improves out-of-sample label prediction in comparison to supervised LDA approaches that don't utilize information derived from the structure of the label space.
+We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply-labeled bag-of-word data.
+Examples of such data include web pages and their placement in link directories, product descriptions and placement(s) in product hierarchies, and free-text clinical records and diagnosis codes assigned to them.
+Out-of-sample label prediction is the primary goal of this work, but
+improved lower-dimensional representations of the bag-of-words data is also of interest.  
+%We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA).  
+We find that using the signal from hierarchical labels substantially improves out-of-sample label prediction in comparison to other models that don't utilize the structure of the labels.
 We demonstrate HSLDA on large-scale data from medical document labeling and retail product categorization tasks.
-We show improved label prediction performance and evidence that the learned topic model also improves as a result of using this signal.
+We show improved label prediction performance and evidence that the learned topics also improve.
 

### 2011-05-31 paper/model.tex
Updated graphical model
updated the abstract -- is it better?
making a sweep through the paper, partial commit

@@
-\label{sec:model} We define here a hierarchically supervised LDA
-model. Although we will focus on document modeling in our description
-and experiments, this model applies equally well to other collections
-of discrete data with hierarchically constrained labels.
+\label{sec:model} 
+%We define here a hierarchically supervised LDA
+%model. Although we will focus on document modeling in our description
+%and experiments, this model applies equally well to other collections
+%of discrete data with hierarchically constrained labels.
+
+HSLDA is a model for hierarchically, multiply labeled, bag-of-words data.
+We will refer to the grouped bag-of-word data as a document.
+Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.
+Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the ordered set of $N_d$ observations in document $d$.
+Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.
 
@@
 %
-\begin{figure}[h]
+\begin{figure}[t]
 %tbp] %  figure placement: here, top, bottom, or page
- \centering \includegraphics[scale=0.3]{Graphical_Model2} \caption{adapted sLDA model}
+ \centering \includegraphics[scale=0.4]{Graphical_Model-final} \caption{HSLDA graphical model}
 
@@
 \item Draw $a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\beta_{l},y_{\mathrm{pa}(l),d}\sim\begin{cases}
-\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\beta_{l},1), & y_{\mathrm{pa}(l)}=1\\
-\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\beta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l)}=-1\end{cases}$ \\where $\bar{z}_{d}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
+\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\eta_{l},1), & y_{\mathrm{pa}(l)}=1\\
+\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\eta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l)}=-1\end{cases}$ \\where $\bar{z}_{d}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
 

### 2011-05-31 paper/introduction.tex

@@
 In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
-sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this ``supervision'' can be seen as extra data  about a document;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
-These labels are usually generatively modeled as having been conditionally draw given an inferred document-specific topic mixture.
-It has been demonstrated that the signal provided by this supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
+sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this supervision is just extra per document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
+These labels are usually generatively modeled as having been conditionally drawn from some distribution that depends on the document-specific topic mixture.
+It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
 
@@
 The situation of each product in a product hierarchy (often multiply situated) constitutes a multiple, hierarchical labeling of the free-text product descriptions.
-We hypothesis that hierarchical labels should, at least in theory, provide better signal than the simpler unstructured supervision previously considered.
-Results from applying our model to medical record and web retail data suggests that this is likely to be the case.
-In particular, we observed big gains in our primary goal of out-of-sample label prediction when using hierarchical supervision.
+We hypothesize that such hierarchical labels should, at least in theory, provide better supervision than the simpler unstructured labels previously considered.
+Results from applying our model to both medical record and web retail data suggests that this is likely the case.
+In particular, we observe gains in our primary goal of out-of-sample label prediction that result specifically from leveraging hierarchical supervision.
 
 The remainder of this paper is structured as follows.
-In Section~\ref{sec:related_work} we review related work, in Section~\ref{sec:model} we introduce hierarchically supervised LDA (HSLDA), and in Section~\ref{sec:experiments} we apply HSLDA to health care and web retail data, showing predictive performance and improved topic generation.
+Section~\ref{sec:model}  introduces hierarchically supervised LDA (HSLDA), Section~\ref{sec:inference} details a sampling approach to inference in HSLDA,  Section~\ref{sec:related_work} reviews related work, and Section~\ref{sec:experiments} shows results from applying HSLDA to health care and web retail data.
 

### 2011-05-31 paper/model.tex
slight changes to model
cleaning up Nick's disastrous "help"

@@
 
-HSLDA is a model for hierarchically, multiply labeled, bag-of-words data.
-We will refer to the grouped bag-of-word data as a document.
+HSLDA is a model for hierarchically, multiply-labeled, bag-of-words data.
+We will refer to individual groups of bag-of-word data as documents.
 Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.
-Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the ordered set of $N_d$ observations in document $d$.
+Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document $d$.
 Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.
+Let the set of labels be $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
+Each label $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
+ We will for exposition purposes assume that this label set has hard ``is-a'' parent-child constraints (explained later), although this assumption can be relaxed at the cost of more complicated inference.
+Such a label hierarchy forms a multiply rooted tree.
+Without loss of generality we will consider a trees with a single root $r\in\mathcal{L}$.
+Each document has a response $y_{l,d} \in \{-1,1\}$ to every label which indicates whether the label applies to document $d$ or not.
 
-We assume a pre-specified set of labels $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
-Each document is assigned a response of either -1 or 1 for at least
-one, but potentially many labels in $\mathcal{L}$. The label, $l$,
-for a document, $d$, will be used interchangeably to refer to the
-observed response of document $d$ to label $l$. The label set is
-assumed to be structured as an {}``is-a'' hieararchy.
-To understand this, consider a hierarchy where label $l_{1}$ is a
-parent of label $l_{2}$. If document $d$ has a positive response
+The is-a hierarchical constraint is a hard constraint that 
+document $d$ has a positive response
 to label $l_{2}$ then it will also have a positive response to label
@@
 using a generative cascade of conditional probit regression models.
+
+Each document is assigned a response of either -1 or 1 for at least
+one, but potentially many labels in $\mathcal{L}$. 
+The label, $l$,
+for a document, $d$, will be used interchangeably to refer to the
+observed response of document $d$ to label $l$.
 
@@
 \item Draw topic assignment $z_{n,d}\mid\boldsymbol\theta_d\sim{\rm Multinomial}(\boldsymbol\theta_d)$ 
-\item Draw word $w_{n,d}\mid z_{n,d},\boldsymbol\beta_{1:K}\sim{\rm Multinomial}(\boldsymbol\beta_{z_{n,d}})$ 
+\item Draw word $w_{n,d}\mid z_{n,d},\boldsymbol\phi_{1:K}\sim{\rm Multinomial}(\boldsymbol\phi_{z_{n,d}})$ 
 \end{itemize}
@@
 \begin{itemize}
-\item Draw $a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\beta_{l},y_{\mathrm{pa}(l),d}\sim\begin{cases}
+\item Draw $a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\eta_{l},y_{\mathrm{pa}(l),d}\sim\begin{cases}
 \mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\eta_{l},1), & y_{\mathrm{pa}(l)}=1\\

### 2011-05-31 paper/main.tex
Updated graphical model - moved experiment to new file

@@
 \label{sec:experiments}
-
-%\begin{itemize}
-%\item Describe Data for both Experiments
-%\item Describe the Four different conditions
-%\item Describe how they are all evaluated, with and without hierarchy
-%\end{itemize}
-In the section we describe the application of HSLDA for prediction
-in two hierarchically structured domains. Firstly, we describe using
-discharge summaries to predict diagnoses, encoded as ICD-9 codes.
-Discharge summaries are documents that are authored by clinicians
-to summarize the course of a hospitalization. ICD-9 codes are used
-mainly for billing purposes to indicate the conditions for which a
-patient was treated. Secondly, we describe using Amazon.com product
-descriptions to predict product categories.
-
-\subsection{Data and Pre-Processing}
-
-\subsubsection{Diagnosis Prediction}
-
-Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
-Hospital. The data consisted of free-text discharge summaries and
-their respective ICD-9 codes. A discharge summary is a clinical report
-prepared by a physician or other health professional at the conclusion
-of a hospital stay or series of treatments. The note outlines the
-patient's chief complaint, diagnostic findings, therapy administered,
-patient's response to the chosen therapy, the treatment plan and the
-recommendations upon discharge. The ICD-9 codes used to structure
-the discharge summary data are part of a controlled terminology which
-is the international standard diagnostic classification for epidemiological,
-health management, and clinical purposes. The ICD-9 codes are organized
-in a rooted-tree structure, with each edge representing an is-a relationship
-between parent and child, such that the parent diagnosis subsumes
-the child diagnosis. For example, the code representing {}``Pneumonia
-due to adenovirus'' is a child of the code representing {}``Viral
-pneumonia'' where the former is a type of the latter. In the hospital,
-ICD-9 codes are generated manually by trained medical coders, who
-review all the information in the discharge summary.
-
-The text of the discharge summaries were pre-processed such that each
-document would be represented as counts over a 10,000 word vocabulary.
-The Natural Language Toolkit was used to tokenize the text. A vocabulary
-was identified by first sorting terms based on a global term frequency-inverse
-document frequency measure. The top 10,000 words which were not identifying
-in some way (a name, place, or identifying number) were selected for
-inclusion in the vocabulary.
-
-For each hospitalization there are usually several ICD-9 codes assigned
-for billing purposes. These codes are known to be quite specific but
-not very sensitive \citep{Birmetal2005}. Regardless of that fact,
-this is one of the only sources for information on patient diagnoses
-aside from the free text.
-%Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available.
-For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes.
-Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].
-
-We worked within the guidelines of the Health Insurance Portability
-and Accountability Act (HIPAA), which protects patient privacy and
-the security of potentially identifying medical material, known as
-personal health information (PHI). HIPAA covers any information within
-a medical record that was created, used, or disclosed during the course
-of providing a health care service and that can be used to identify
-an individual. This study was approved by the Institutional Review
-Board.
-
-\subsubsection{Product Category Prediction}
-
-Data for these experiments were obtained partially from the Stanford
-Network Analysis Platform (SNAP) Amazon product metadata dataset \citep{SNAP}
-and partially directly from the the Amazon.com website \citep{AMAZON}.
-The product ID's and categorizations were obtained from the SNAP dataset
-and the product descriptions were obtained directly from the website.
-
-We were able to deduce the structure of the hierarchy for the Amazon.com
-products directly since all ancestors in the hierarchy were included
-with each category label. For example, {}``DVD / Genres / Science
-Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
-for the DVD, {}``The Time Machine''. Each product was labeled with
-multiple categories.
-
-The vocabulary for this experiment was created by including the most
-frequent 30K words omitting stopwords.
-
-\subsection{Comparison Models}
-
-We applied HSLDA, along with three other closely related models, to
-the clinical data and the retail product data. Specifically, we evaluate
-models including sLDA with independent regressors (hierarchical constraints
-on labels ignored), HSLDA fit by performing LDA followed by tree-conditional
-regressions, and HSLDA fit with fixed random regression parameters.
-We will refer to the three comparison models as the sLDA model, the
-separate HSLDA model, and the random HSLDA model, respectively. These
-models were chosen as to highlight performance in absence of hierarchical
-constraints, the effect of the combined inference procedure, and regression
-performance attributable solely to the hierarchical constraints.
-
-We evaluated model performance for all three models with a range of
-values for $\mu$ ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\} )$,
-the mean prior parameter for regression parameters.
-
-\subsection{Evaluation}
-
-%\begin{itemize}
-%\item Talk about hierarchy and no hierarchy
-%\item Make explicit calculation of sensitivity and 1-specificity
-%\end{itemize}
-For each dataset, a held out set of 1,000 documents and accompanying
-labels were used for evaluation. The two main methods of evaluation
-for the model are prediction and topic quality. To evaluate predictive
-performance for all comparison models equivalently, each model was
-evaluated with two methods. The first evaluation method augments the
-observed labels in the held out set with their ancestors and considers
-all other non-existant labels to be negative. The second method ignores
-the ancestors of the observed labels in the held out set and considers
-all other non-existant labels to be negative. This uniform treatment
-of ancestors allows for a fair comparison of the models.
-
-The two measures for predictive performance used here include the
-true positive rate and the false positive rate evaluated based on
-$p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$ for each label in each
-model.
+\input{experiments}
 

### 2011-05-31 paper/experiments.tex
added experiments.tex

%\begin{itemize}
%\item Describe Data for both Experiments
%\item Describe the Four different conditions
%\item Describe how they are all evaluated, with and without hierarchy
%\end{itemize}
In the section we describe the application of HSLDA for prediction
in two hierarchically structured domains. Firstly, we describe using
discharge summaries to predict diagnoses, encoded as ICD-9 codes.
Discharge summaries are documents that are authored by clinicians
to summarize the course of a hospitalization. ICD-9 codes are used
mainly for billing purposes to indicate the conditions for which a
patient was treated. Secondly, we describe using Amazon.com product
descriptions to predict product categories.

\subsection{Data and Pre-Processing}

\subsubsection{Diagnosis Prediction}

Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
Hospital. The data consisted of free-text discharge summaries and
their respective ICD-9 codes. A discharge summary is a clinical report
prepared by a physician or other health professional at the conclusion
of a hospital stay or series of treatments. The note outlines the
patient's chief complaint, diagnostic findings, therapy administered,
patient's response to the chosen therapy, the treatment plan and the
recommendations upon discharge. The ICD-9 codes used to structure
the discharge summary data are part of a controlled terminology which
is the international standard diagnostic classification for epidemiological,
health management, and clinical purposes. The ICD-9 codes are organized
in a rooted-tree structure, with each edge representing an is-a relationship
between parent and child, such that the parent diagnosis subsumes
the child diagnosis. For example, the code representing {}``Pneumonia
due to adenovirus'' is a child of the code representing {}``Viral
pneumonia'' where the former is a type of the latter. In the hospital,
ICD-9 codes are generated manually by trained medical coders, who
review all the information in the discharge summary.

The text of the discharge summaries were pre-processed such that each
document would be represented as counts over a 10,000 word vocabulary.
The Natural Language Toolkit was used to tokenize the text. A vocabulary
was identified by first sorting terms based on a global term frequency-inverse
document frequency measure. The top 10,000 words which were not identifying
in some way (a name, place, or identifying number) were selected for
inclusion in the vocabulary.

For each hospitalization there are usually several ICD-9 codes assigned
for billing purposes. These codes are known to be quite specific but
not very sensitive \citep{Birmetal2005}. Regardless of that fact,
this is one of the only sources for information on patient diagnoses
aside from the free text. %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available. For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].

We worked within the guidelines of the Health Insurance Portability
and Accountability Act (HIPAA), which protects patient privacy and
the security of potentially identifying medical material, known as
personal health information (PHI). HIPAA covers any information within
a medical record that was created, used, or disclosed during the course
of providing a health care service and that can be used to identify
an individual. This study was approved by the Institutional Review
Board.

\subsubsection{Product Category Prediction}

Data for these experiments were obtained partially from the Stanford
Network Analysis Platform (SNAP) Amazon product metadata dataset \citep{SNAP}
and partially directly from the the Amazon.com website \citep{AMAZON}.
The product ID's and categorizations were obtained from the SNAP dataset
and the product descriptions were obtained directly from the website.

We were able to deduce the structure of the hierarchy for the Amazon.com
products directly since all ancestors in the hierarchy were included
with each category label. For example, {}``DVD / Genres / Science
Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
for the DVD, {}``The Time Machine''. Each product was labeled with
multiple categories.

The vocabulary for this experiment was created by including the most
frequent 30K words omitting stopwords.

\subsection{Comparison Models}

We applied HSLDA, along with three other closely related models, to
the clinical data and the retail product data. Specifically, we evaluate
models including sLDA with independent regressors (hierarchical constraints
on labels ignored), HSLDA fit by performing LDA followed by tree-conditional
regressions, and HSLDA fit with fixed random regression parameters.
We will refer to the three comparison models as the sLDA model, the
separate HSLDA model, and the random HSLDA model, respectively. These
models were chosen as to highlight performance in absence of hierarchical
constraints, the effect of the combined inference procedure, and regression
performance attributable solely to the hierarchical constraints.

We evaluated model performance for all three models with a range of
values for $\mu$ ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\} )$,
the mean prior parameter for regression parameters.

\subsection{Evaluation}

%\begin{itemize}
%\item Talk about hierarchy and no hierarchy
%\item Make explicit calculation of sensitivity and 1-specificity
%\end{itemize}
For each dataset, a held out set of 1,000 documents and accompanying
labels were used for evaluation. The two main methods of evaluation
for the model are prediction and topic quality. To evaluate predictive
performance for all comparison models equivalently, each model was
evaluated with two methods. The first evaluation method augments the
observed labels in the held out set with their ancestors and considers
all other non-existant labels to be negative. The second method ignores
the ancestors of the observed labels in the held out set and considers
all other non-existant labels to be negative. This uniform treatment
of ancestors allows for a fair comparison of the models.

The two measures for predictive performance used here include the
true positive rate and the false positive rate evaluated based on
$p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$ for each label in each
model.

### 2011-06-01 paper/experiments.tex
Modified the experimental section (data and preprocessing) Tightened intro Added an accent on my name

@@
 %\end{itemize}
-In the section we describe the application of HSLDA for prediction
-in two hierarchically structured domains. Firstly, we describe using
-discharge summaries to predict diagnoses, encoded as ICD-9 codes.
-Discharge summaries are documents that are authored by clinicians
-to summarize the course of a hospitalization. ICD-9 codes are used
-mainly for billing purposes to indicate the conditions for which a
-patient was treated. Secondly, we describe using Amazon.com product
-descriptions to predict product categories.
+
+We experimented with HSLDA for prediction in two domains: predicting medical
+diagnosis codes from hospital discharge summaries and predicting product
+categories from Amazon.com product descriptions.
+
+%In the section we describe the application of HSLDA for prediction
+%in two hierarchically structured domains. Firstly, we describe using
+%discharge summaries to predict diagnoses, encoded as ICD-9 codes.
+%Discharge summaries are documents that are authored by clinicians
+%to summarize the course of a hospitalization. ICD-9 codes are used
+%mainly for billing purposes to indicate the conditions for which a
+%patient was treated. Secondly, we describe using Amazon.com product
+%descriptions to predict product categories.
 
@@
 
-Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
-Hospital. The data consisted of free-text discharge summaries and
-their respective ICD-9 codes. A discharge summary is a clinical report
-prepared by a physician or other health professional at the conclusion
-of a hospital stay or series of treatments. The note outlines the
-patient's chief complaint, diagnostic findings, therapy administered,
-patient's response to the chosen therapy, the treatment plan and the
-recommendations upon discharge. The ICD-9 codes used to structure
-the discharge summary data are part of a controlled terminology which
-is the international standard diagnostic classification for epidemiological,
-health management, and clinical purposes. The ICD-9 codes are organized
-in a rooted-tree structure, with each edge representing an is-a relationship
-between parent and child, such that the parent diagnosis subsumes
-the child diagnosis. For example, the code representing {}``Pneumonia
-due to adenovirus'' is a child of the code representing {}``Viral
-pneumonia'' where the former is a type of the latter. In the hospital,
-ICD-9 codes are generated manually by trained medical coders, who
-review all the information in the discharge summary.
+Discharge summaries are authored by clinicians to summarize the course of a
+hospitalized patient. The summaries typically contain a record of the patient
+complains, findings and diagnoses, along with treatment and hospital course.
+For each admission trained medical coders review the information in the
+discharge summary and assign a series of diagnoses codes. Coding follows the
+ICD-9-CM controlled terminology, an international diagnostic classification for
+epidemiological, health management, and clinical
+purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm} As such, the ICD-9
+codes constitute a labeling of a patient's diagnoses based on a discharge
+summary. The ICD-9 codes are organized in a rooted-tree structure, with each
+edge representing an is-a relationship between parent and child, such that the
+parent diagnosis subsumes the child diagnosis. For example, the code for
+{}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
+pneumonia,'' where the former is a type of the latter.  It is worth noting that
+the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
+tend to be more specific than sensitive in their
+assignments~\cite{Birmetal2005}, and sometimes make
+mistakes~\cite{FarzandipourEtAl10}.
 
-The text of the discharge summaries were pre-processed such that each
-document would be represented as counts over a 10,000 word vocabulary.
-The Natural Language Toolkit was used to tokenize the text. A vocabulary
-was identified by first sorting terms based on a global term frequency-inverse
-document frequency measure. The top 10,000 words which were not identifying
-in some way (a name, place, or identifying number) were selected for
-inclusion in the vocabulary.
+The task of automatic ICD-9 coding has been investigated in the clinical
+domain. Much of the work was triggered by the 2007 medical NLP community
+challenge~\cite{Challenge07}. The data in the challenge, however, differs from 
+ours in its scope. The datasets were smaller (1,000 training and 1,000 testing
+documents) and focused on radiology reports with a restricted number of ICD-9
+codes (45 of them, compared to 7K+ in our dataset). Methods ranged from manual
+rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
+Other work had leveraged larger datasets and experimented with K-nearest
+neighbor, Naive Bayes, support vector machines, Bayesian Ridge Regression, as
+well as simple keyword mappings, all with promising
+results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Ruch2008,Lita2008}.
 
-For each hospitalization there are usually several ICD-9 codes assigned
-for billing purposes. These codes are known to be quite specific but
-not very sensitive \citep{Birmetal2005}. Regardless of that fact,
-this is one of the only sources for information on patient diagnoses
-aside from the free text.
+Our dataset was gathered from the clinical data warehouse of
+NewYork-Presbyterian Hospital. It consists of 6,000 discharge summaries and
+their associated ICD-9 codes (7,298 distinct codes overall), representing all
+the discharges from the hospital in 2009. Summaries have 8.39 associated ICD-9
+codes on average (std dev=5.01) and contain an average of 536.57 terms after
+preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
+summaries for training and 1,000 for testing.
+
+The text of the discharge summaries was tokenized with
+NLTK.\footnote{http://www.nltk.org} Vocabulary was determined as the top 10,000
+tokens with highest document frequency (exclusive of names,
+places and other identifying numbers). Each discharge summary is thus
+represented as counts over the 10,000-word vocabulary. The study was approved
+by the Institutional Review Board and follows HIPAA (Health
+Insurance Portability and Accountability Act) privacy guidelines.
+
+%For each hospitalization there are usually several ICD-9 codes assigned
+%for billing purposes. These codes are known to be quite specific but
+%not very sensitive \citep{}. Regardless of that fact,
+%this is one of the only sources for information on patient diagnoses
+%aside from the free text.
 %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available.
@@
 
-We worked within the guidelines of the Health Insurance Portability
-and Accountability Act (HIPAA), which protects patient privacy and
-the security of potentially identifying medical material, known as
-personal health information (PHI). HIPAA covers any information within
-a medical record that was created, used, or disclosed during the course
-of providing a health care service and that can be used to identify
-an individual. This study was approved by the Institutional Review
-Board.
-
 \subsubsection{Product Category Prediction}
 
-Data for these experiments were obtained partially from the Stanford
-Network Analysis Platform (SNAP) Amazon product metadata dataset \citep{SNAP}
-and partially directly from the the Amazon.com website \citep{AMAZON}.
-The product ID's and categorizations were obtained from the SNAP dataset
-and the product descriptions were obtained directly from the website.
+In this experiment, we look at product descriptions and their
+categorizations according to a product hierarchy. Product ID's
+and categorizations were obtained from the Stanford Network Analysis Platform
+(SNAP) dataset~\citep{SNAP}. Product descriptions were crawled from the
+amazon.com website directly. We were able to deduce the structure of the
+hierarchy for the Amazon.com products directly since all ancestors in the
+hierarchy were included with each category label. For example, {}``DVD / Genres
+/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
+for the DVD, {}``The Time Machine.''
 
-We were able to deduce the structure of the hierarchy for the Amazon.com
-products directly since all ancestors in the hierarchy were included
-with each category label. For example, {}``DVD / Genres / Science
-Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
-for the DVD, {}``The Time Machine''. Each product was labeled with
-multiple categories.
-
-The vocabulary for this experiment was created by including the most
-frequent 30K words omitting stopwords.
+Our dataset contains 15,130 product descriptions for training and 1,000
+for testing. The product descriptions are shorter than the discharge summaries
+(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique codes.
+Products are assigned on average 9.01 codes (std dev=4.91). The vocabulary
+consists of the most frequent 30,000 words omitting stopwords.
 

### 2011-06-01 paper/introduction.tex
Modified the experimental section (data and preprocessing) Tightened intro Added an accent on my name

@@
-There exist many sources of unstructured data that have been partially or completely categorized by human editors.
-In this paper we focus on unstructured textual data that has been, at least in part, manually categorized.
-Examples include but are not limited to webpages and curated hierarchical directories of the same \citep{DMOZ}, product descriptions and catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), and patient hospital treatment transcripts and  codes applied to them for bookkeeping and insurance purposes (e.g.~hospital discharge records with International Classification of Disease 9th Revision,
-Clinical Modification (ICD-9-CM) codes assigned \cite{ICD9}).
-In this work we show how to combine these two sources of information using a single model that allows one to automatically categorize new text documents, suggest labels that might be inaccurate, compute improved similarities between documents for information retrieval purposes, and more.
-The models and techniques that we develop in this paper are applicable in other domains as well, namely, any unstructured representations of data that have been hierarchically classified (e.g.~image catalogs with bag-of-feature image representations).
+There exist many sources of unstructured data that have been partially or
+completely categorized by human editors.  In this paper we focus on
+unstructured textual data that has been, at least in part, manually
+categorized.  Examples include but are not limited to webpages and curated
+hierarchical directories of the same \citep{DMOZ}, product descriptions and
+catalogs~\citep{SNAP}, and patient records and their associated billing codes.
+In this work we show how to combine these two sources of information using a
+single model that allows one to automatically categorize new text documents,
+suggest labels that might be inaccurate, compute improved similarities between
+documents for information retrieval purposes, and more. The models and
+techniques that we develop in this paper are applicable in other domains as
+well, namely, any unstructured representations of data that have been
+hierarchically classified (e.g.~image catalogs with bag-of-feature image
+representations).
 

### 2011-06-01 paper/main.tex

@@
 \author{
-Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} Noemie Elhadad \hspace{1cm} Frank Wood\\
+Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} No\'emie Elhadad \hspace{1cm} Frank Wood\\
 Columbia University, New York, NY 10027, USA \\

### 2011-06-01 paper/model.tex
took a pass through the model section

@@
 Such a label hierarchy forms a multiply rooted tree.
-Without loss of generality we will consider a trees with a single root $r\in\mathcal{L}$.
-Each document has a response $y_{l,d} \in \{-1,1\}$ to every label which indicates whether the label applies to document $d$ or not.
+Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.
+Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.
+In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.
+In the applications we consider, only positive label applications are observed.
 
-The is-a hierarchical constraint is a hard constraint that 
-document $d$ has a positive response
-to label $l_{2}$ then it will also have a positive response to label
-$l_{1}$. Conversely, if document $d$ has a negative response to
-label $l_{1}$ then it will also has a negative response to $l_{2}$.
-To capture this hierarchical structure we model the labeling of documents
-using a generative cascade of conditional probit regression models.
-
-Each document is assigned a response of either -1 or 1 for at least
-one, but potentially many labels in $\mathcal{L}$. 
-The label, $l$,
-for a document, $d$, will be used interchangeably to refer to the
-observed response of document $d$ to label $l$.
+The constraints imposed by an is-a label hierarchy are that if the $l$th label is applied to document $d$, i.e.~$y_{l,d} = 1$, then all labels in the label hierarchy up to the root are also applied to document $d$, i.e.~$y_{\mathrm{pa}(l),d} = 1, y_{\mathrm{pa}(\mathrm{pa}(l)),d} = 1, \ldots, y_{r,d}=1.$  Conversely, if a label $l'$ is marked as not applying to a document then no descendant of that label may be applied to the same.
+We assume that at least one label is applied to every document.
+This is illustrated in Figure~\ref{fig:graphical_model} where the root label is always applied but only some of the descendant labelings are observed as having been applied (diagonal hashing indicates that potentially some of the plated variables are observed).
+%To reiterate, these is-a label hierarchy constraints can be relaxed.
+%To capture this hierarchical structure we model the labeling of documents
+%using a generative cascade of conditional probit regression models.
 
@@
 
-\label{fig:example} 
+\label{fig:graphical_model} 
 \end{figure}
 
-The fixed parameters of the model are the number of topics $K$, the
-number of unique words in the vocabulary $V$, the number of documents
-$D$, as well as the mean, $\boldsymbol\mu$, and the standard deviation, $\sigma$,
-used in a normal prior distribution. The hyper-parameters $\alpha^{\prime}$,
-$\alpha$, and $\gamma$ are weight parameters for Dirichlet prior
-distributions. We will denote the $K$-dimensional Dirichlet distribution
-as Dir$_{K}(\cdot)$ and the $K$ dimensional normal distribution
-as $\mathcal{N}_{K}(\cdot)$.
+In HSLDA, the bag-of-words document data is modeled using the LDA mixed-membership mixture model with global topic estimation.
+Label responses are generated using a conditional hierarchy of probit regressors.
+The HSLDA graphical model is given in Figure~\ref{fig:graphical_model}.
+In it and the following $K$ is the number of LDA "topics" (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$ is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific distribution over topics, $\boldsymbol\beta$ is a global distribution over topics, Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution, $\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution, $\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the $d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator function that takes the value $1$ if its argument is true and $0$ otherwise.
+The following procedure describes how to generate from the HSLDA generative model.
 
-We will now describe the stochastic generative process which defines
-our model. The graphical model is show in Figure~\ref{fig:example}.
+ %as well as the mean, $\boldsymbol\mu$, and the standard deviation, $\sigma$,
+%used in a normal prior distribution. The hyper-parameters $\alpha^{\prime}$,
+%$\alpha$, and $\gamma$ are weight parameters for Dirichlet prior
+%distributions.
+
+%We will now describe the stochastic generative process which defines
+%our model. The 
 %Given the number of topics, K, and broad gamma priors on hyperparameters,
@@
 \begin{itemize}
-\item Draw a regression coefficient $\boldsymbol\eta_{l}\mid\mu,\sigma\sim\mathcal{N}_{K}(\mu \mathbf{1}_K,\sigma \mathbf{I}_{K})$,
-where $\mathbf{I}_{K}$ is the $K$ dimensional identity matrix 
+\item Draw a label application coefficient $\boldsymbol\eta_{l}\mid\mu,\sigma\sim\mathcal{N}_{K}(\mu \mathbf{1}_K,\sigma \mathbf{I}_{K})$  
 \end{itemize}
@@
 \end{itemize}
-\item For each label $l\in\mathcal{L}$:
+\item Set $y_{r,d} = 1$
+\item For each label $l$ in a breadth first traversal of $\mathcal{L}$ starting at the children of  root $r$
 
@@
 \item Draw $a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\eta_{l},y_{\mathrm{pa}(l),d}\sim\begin{cases}
-\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\eta_{l},1), & y_{\mathrm{pa}(l)}=1\\
-\mathcal{N}(\bar{\mathbf{z}}^{T}\boldsymbol\eta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l)}=-1\end{cases}$ \\where $\bar{z}_{d}=N_{d}^{-1}\sum_{n=1}^{N}z_{n,d}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
+\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1), & y_{\mathrm{pa}(l),d}=1\\
+\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l),d}=-1\end{cases}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
 
-\item Set the response variable \[
+\item Apply label $l$ to document $d$ according to $a_{l,d}$ \[\hspace{-2cm}
 y_{l,d}\mid a_{l,d}=\begin{cases}
-\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0} and \ensuremath{y_{{\rm \mathrm{pa}}(l),d}=1}}\\
+\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}} \\% and \ensuremath{y_{{\rm \mathrm{pa}}(l),d}=1}}\\
 -1 & \text{otherwise}\end{cases}\]
@@
 \end{enumerate}
-This type of generative model is known as a probit regression model.
-Probit regression models are a type of discriminative probabilistic
-model similar to logistic regression. However, instead of using the
-logistic sigmoid as the link function, the probit regression model
-uses the CDF for a standard normal distribution - the inverse of which
-is known as the probit. In this case, the regression is conditional
-on the parents according to the constraints of the labeling hierarchy.
-The latent variables $a_{l,d}$ utilized here are also known as an
-auxiliary variables because the are introduced to make exact Gibbs
-sampling possible and are not of primary interest.
 
-Given that negative labels are uncommon and that the absence of a
-label is not equivalent to a negative label, we apply an informative
-prior to the regression parameters, $\mathbf{\boldsymbol\beta}_{\mathcal{L}}$,
-in the form of a negative prior that encodes a bias towards being
-truly negative in the absence of a label.
+Here $\bar{\mathbf{z}}_d^T = [\bar{z}_{1}, \ldots, \bar{z}_k, \ldots, \bar{z}_K]$ is the empirical topic distribution for document $d$, in which each entry is the percentage of the words in that document that come from topic $k$, $\bar{z}_{k}=N_{d}^{-1}\sum_{n=1}^{N_d}\mathbb{I}(z_{n,d}=k).$
+
+The second half of step 4 is a substantial part of our contribution to the general class of supervised LDA models.
+Here each document is generatively labeled using a hierarchy of conditionally dependent probit regressors \cite{gelman?,probit_regression?}.
+For every label $l \in \mathcal{L}$ both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to decide whether or not  label $l$ is to be applied to  document $d$ as well.
+Note that label $y_{l,d}$ can only be applied to document $d$ if its parent label $\mathrm{pa}(l)$ is also applied (these expressions are specific to is-a constraints but can be modified to accommodate different constraints).
+The regression coefficients $\boldsymbol\eta_l$ are independent a priori, however, the hierarchical coupling in this model induces a posteriori dependence.
+The net effect of this is that label predictors deeper in the label hierarchy are able to focus on finding specific, conditional labeling features.
+We believe this to be a significant source of the empirical label prediction improvement we observe experimentally.
+We test this hypothesis in Section~\ref{sec:experiments}.
+
+Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations.
+In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:eta_l_cov}\eqref{eqn:eta_l_mean} which are analytic.
+This simplifies posterior inference substantially.
+
+%Probit regression is similar to logistic regression except that it uses the standard normal distribution cumulative distribution function (CDF) rather than the logistic sigmoid link function.
+The inverse of the normal CDF is is known as the probit function.
+%The latent variables $a_{l,d}$ are
+%auxiliary variables because the are introduced to make exact Gibbs
+%sampling possible and are not of primary interest.
+
+%This type of generative model is known as a probit regression model.
+%Probit regression models are a type of discriminative probabilistic
+%model similar to logistic regression. However, instead of using the
+%logistic sigmoid as the link function, the probit regression model
+%uses the CDF for a standard normal distribution - the inverse of which
+%is known as the probit. In this case, the regression is conditional
+%on the parents according to the constraints of the labeling hierarchy.
+
+In the common case where no negative labels are observed (like the example applications we consider in Section~\ref{sec:experiments}), the model must be explicitly biased towards generating data that has negative labels in order to keep it from learning to assign all labels to all documents.
+This is a common problem in modeling unbalanced data.
+To see how this model can be biased in this way we draw the readers attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.
+Because $\bar{\mathbf{z}}_d$ is always positive, setting $\mu$ to a negative value results in a bias towards negative labelings, i.e.~for large negative values of $\mu$, all labels become a priori more likely to be negative ($y_{l,d}=-1$).
+We explore the ability of $\mu$ to bias out-of-sample label prediction performance in Section~\ref{sec:experiments}.
+
+%we apply an informative
+%prior to the regression parameters, $\mathbf{\boldsymbol\beta}_{\mathcal{L}}$,
+%in the form of a negative prior that encodes a bias towards being
+%truly negative in the absence of a label.

### 2011-06-01 paper/model.tex
took a pass at inference.

@@
 Let the set of labels be $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
-Each label $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
+Each label except root labels $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
  We will for exposition purposes assume that this label set has hard ``is-a'' parent-child constraints (explained later), although this assumption can be relaxed at the cost of more complicated inference.
@@
 Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations.
-In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:eta_l_cov}\eqref{eqn:eta_l_mean} which are analytic.
+In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:regression_param_conditional} which are analytic.
 This simplifies posterior inference substantially.

### 2011-06-01 paper/related_work.tex
related work

@@
 Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
-an exponential family response variable. Although there are many models
-for making predictions based on free text, sLDA is unique in that
-it is a generative model, it represents documents as a mixed-membership,
-and constrains the inference of the latent structure of the documents
-by its predictability of the response variable. In other words, sLDA
-infers topics such that the model is capable of a high predictive
-likelihood for words in a document and the response variable associated
-with a document. This approach has been shown to outperform both LASSO
+supervision in the form of an observed exponential family response 
+variable per document. As a result, sLDA
+infers topics such that the model predicts the response variable while
+improing word likelihood.  This approach has been shown to outperform both LASSO
 (L1 regularized least squares regression) and LDA followed by least
@@
 
-There have been many models that incorporate both latent models of
-text and some form of supervision \citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}.
-One set of models that are particularly relevant to HSLDA are Chang
-and Blei's hierarchical models for document networks (Relational Topic
-Models). In that family of models, they encountered a similar scenario
-where the lack of a link did not truly indicate absence. In hierarchically
-labeled data, negative labels are uncommon and the lack of a label
-in the hierarchy is not equivalent to a negative label. Therefore,
-as in the work of Chang and Blei, we employ regularization to account
-for the lack of negative labels. This will be discussed further in
+Other models that incorporate both and some form of supervision include
+LabeledLDA\citep{Ramage2009}, DiscLDA\citep{DiscLDA}, models applied to 
+computer vision and document networks\citep{wangbleifeifei08,RelationalLDA}.
+
+%One set of models that are particularly relevant to HSLDA are Chang
+%and Blei's hierarchical models for document networks (Relational Topic
+%Models). In that family of models, they encountered a similar scenario
+%where the lack of a link did not truly indicate absence. In hierarchically
+%labeled data, negative labels are uncommon and the lack of a label
+%in the hierarchy is not equivalent to a negative label. Therefore,
+%as in the work of Chang and Blei, we employ regularization to account
+%for the lack of negative labels. This will be discussed further in
 \ref{sec:model}.

### 2011-06-01 paper/related_work.tex
related work needs work intro needs better explanation of challenges

--- 2011-05-31 paper/related_work.tex
@@
-Latent dirichlet allocation (LDA) is a generative
-probabilistic model of corpora that represents documents as a mixed
-membership bag-of-words. Also known as topic models, these models
-infer the latent structure, or topics, of documents in a corpus. Each
-document is represented as a collection of words, generated from a
-set of topic assignments (one for each word), where each topic assignment
-is drawn from a distribution over topics \citep{Blei2003}.
+While there has been much work in multi-label classification of text and
+text modeling in general, we focus here on topic modeling approaches.
+% Noemie: do we need that much text on LDA?
+Latent Dirichlet allocation (LDA) is a generative probabilistic model which
+represents documents as a mixed-membership bag of word. Each document is
+represented as a collection of words, generated from a set of topic assignments
+(one for each word), where each topic assignment is drawn from a distribution
+over topics~\citep{Blei2003}. sLDA is latent Dirichlet allocation (LDA)
+\cite{Blei2003} augmented with per-document labeling, often taking the form of
+a single numerical or categorical label. Examples of labels include rating
+associated with an online  eview, grades for an essay, and number of times a
+webpage is linked. This approach has been shown to outperform both LASSO
+(L1 regularized least squares regression) and LDA followed by least
+squares regression~\cite{BleiMcAuliffe2008}.
 
-Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
-an exponential family response variable. Although there are many models
-for making predictions based on free text, sLDA is unique in that
-it is a generative model, it represents documents as a mixed-membership,
-and constrains the inference of the latent structure of the documents
-by its predictability of the response variable. In other words, sLDA
-infers topics such that the model is capable of a high predictive
-likelihood for words in a document and the response variable associated
-with a document. This approach has been shown to outperform both LASSO
-(L1 regularized least squares regression) and LDA followed by least
-squares regression \citep{BleiMcAuliffe2008}.
-
-There have been many models that incorporate both latent models of
-text and some form of supervision \citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}.
-One set of models that are particularly relevant to HSLDA are Chang
-and Blei's hierarchical models for document networks (Relational Topic
-Models). In that family of models, they encountered a similar scenario
-where the lack of a link did not truly indicate absence. In hierarchically
-labeled data, negative labels are uncommon and the lack of a label
-in the hierarchy is not equivalent to a negative label. Therefore,
-as in the work of Chang and Blei, we employ regularization to account
-for the lack of negative labels. This will be discussed further in
-\ref{sec:model}.
-
-% move to related work?
-%While there have been some attempts to automatically classify groups
-%of patients as a potential preliminary step to ICD-9 code assignment
-%\citep{Ruch2008,FreitasJunior2006,RibeiroNeto2001,Brown2006}, fully
-%automatic assignment of ICD-9 codes to medical text became a more
-%prevalent research topic only in the last few years. A subset of earlier
-%work proposed various methods on small corpora, based on a few specific
-%diseases \citep{Rao2003} but the most recent and promising work on
-%the subject was inspired by the 2007 Medical NLP Challenge: {}``International
-%Challenge: Classifying Clinical Free Text Using Natural Language Processing\textquotedblright{}
-%(website). Most of the classification strategies included word matching
-%and rule-based algorithms. \citep{Goldstein2007,Crammer2007,Farkas2008}.
-%The data set given to the participants consisted only of documents
-%that were 1-2 lines each and all of the documents were radiology reports
-%- clearly limiting the scope of potential ICD-9 codes which could
-%be assigned. The only paper which has attempted to work with a document
-%scope as large as ours was the 2008 Lita et al publication \citep{Lita2008}.
-%Lita proposed support vector machine and bayesian ridge regression
-%methods to assign appropriate labels to the documents but did not
-%utilize the ICD-9 hierarchy to leverage more comprehensive predictions.
+% Noemie: check out what I wrote here, because I am not really sure this is
+% correct. 
+There have been several models that incorporate both latent models of text and
+some form of
+supervision~\citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}. One set 
+of models that are particularly relevant to HSLDA are Chang and Blei's
+hierarchical models for document networks (Relational Topic Models). In that
+family of models, they encountered a similar scenario where an unselected label 
+does not always indicate absence. In hierarchical labels, this phenomenon is
+even more pervasive -- there are no explicit negative labels, but it is also
+unclear how to treat the parents of selected labels. Like in the work of Chang
+and Blei, we employ regularization to account for the lack of negative
+labeling. In our experiments, we look at the impact of assigning positive and
+negative instances to the ancestors of selected labels.

### 2011-06-01 paper/main.tex

@@
 
+\begin{tabular}{|c|c|}
+\hline 
+HSLDA & sLDA with Independent Regressors\tabularnewline
+\hline 
+children, animation, animated, kids, songs, story, & children, kids, animated, 
+\tabularnewline
+family, new, fun,video, friends, music, live, classic, & animation, songs, video, friends, fun, animals, family, story, 
+\tabularnewline
+adventure, minute, world, magic,animals, song, musical, & new, little, music, like, minute, along, musical, song,
+\tabularnewline
+voiced, feature, along, baby, ages, magical, & adventure, help, classic, baby, child, ages, favorite, sing, learn,
+\tabularnewline
+action, favorite, special, tale, little, child, holiday, & holiday, magical, tale, world, special, parents, live, way, day, boy,
+\tabularnewline
+help, voice, toys, original, learn, characters, & make, old, magic, adventures, home, show, gang, animal, adults, toys,
+\tabularnewline
+computer, sing, like, adults, parents, version, old, first, & many, action
+\tabularnewline
+including, young, 
+\tabularnewline
+\hline
+comedy, funny, comic, hilarious, series, gags, jokes, slapstick, satire, & comedy, show, play, characters, sketches,
+\tabularnewline
+laugh, stars, films, comedies, first, comedian, funniest, show, involving, & funny, mob, performance,
+\tabularnewline
+characters, cast, perfect, mob, performance, play, writer, television, & people, screwball, cast, best, including, stars, sketch, girlfriend,
+\tabularnewline
+satirical, lines, hilariously, boss, routine, silent, funnier, comedic, & hilariously, charming, ensemble, comic, character, star, writer, hit,
+\tabularnewline
+chemistry, comedians, talk, routines, brothers, girlfriend, shorts, & stage, host, classic, romantic, mutants, boss, new, played, together,
+\tabularnewline
+mouse, whose, cinema, screwball, hit, humor, brilliant, gag, new,  & choice, like, karaoke, daughter, vehicle, news, splitting, actress,
+\tabularnewline
+& get, story, hilarious, wacky, featuring, could, side, comedies, comedic\tabularnewline
+\hline 
+thriller, horror, murder, crime, killer, police, mystery, detective, & horror, film, vampire, blood, one, dead, supernatural, killer, evil,
+\tabularnewline
+cop, one, dead, suspense, plot, case, blood, drug, dark, mysterious, & thriller, gore, monster, young, night, terror, mysterious, director,
+\tabularnewline
+death, revenge, violence, town, gang, turns, prison, noir, violent, & movie, films, cult, creepy, terrifying, original, dark, chilling,
+\tabularnewline
+deadly, stars, young, woman, nbsp, body, vampire, night, murders, & genre, atmosphere, victims, classic, murder, haunted, town, mystery,
+\tabularnewline
+kill, criminal, action, becomes, wife, cops, gangster, sex, victims, & story, body, ghost, tale, family, murders, castle, house, nightmare,
+\tabularnewline
+victim, brutal, terror, evil, murdered & budget, eerie, death, doctor, victim, later, bloody, effects
+\tabularnewline
+\hline 
+workout, body, yoga, video, moves, dance, minute, program, fitness, & workout, body, yoga, video, moves, dance, minute,
+\tabularnewline
+easy, poses, muscles, routine, exercises, learn, strength, movements, & program, fitness
+\tabularnewline
+practice, help, step, w, workouts, minutes, techniques, flexibility, & easy, poses, muscles, exercises, learn, routine, strength, movements,
+\tabularnewline
+get, exercise, one, instructor, first, series, time, instruction, & minutes, get, help, practice, step, series, techniques, workouts,
+\tabularnewline
+use, follow, training, breathing, two, basic, back, ll, muscle, fun, & time, flexibility, exercise, instructor, instruction, one, two, fun,
+\tabularnewline
+warm, steps, technique, balance, beginners, designed, using & training, use, first, back, follow, steps, breathing, muscle, basic,
+\tabularnewline
+& using, ll, beginners, three, warm, balance, technique, designed\tabularnewline
+\hline
+\end{tabular}
+
 \begin{figure}[htbp]

### 2011-06-01 paper/related_work.tex

@@
 text modeling in general, we focus here on topic modeling approaches.
-% Noemie: do we need that much text on LDA?
 Latent Dirichlet allocation (LDA) is a generative probabilistic model which
@@
 
+%Latent dirichlet allocation (LDA) is a generative
+%probabilistic model of corpora that represents documents as a mixed
+%membership bag-of-words. Also known as topic models, these models
+%infer the latent structure, or topics, of documents in a corpus. Each
+%document is represented as a collection of words, generated from a
+%set of topic assignments (one for each word), where each topic assignment
+%is drawn from a distribution over topics \citep{Blei2003}.
+
+%Supervised latent Dirichlet allocation (sLDA) builds on LDA by incorporating
+%supervision in the form of an observed exponential family response 
+%variable per document. As a result, sLDA
+%infers topics such that the model predicts the response variable while
+%improing word likelihood.  This approach has been shown to outperform both LASSO
+%(L1 regularized least squares regression) and LDA followed by least
+%squares regression \citep{BleiMcAuliffe2008}.
+
+Other models that incorporate both and some form of supervision include
+LabeledLDA\citep{Ramage2009}, DiscLDA\citep{DiscLDA}, models applied to 
+computer vision and document networks\citep{wangbleifeifei08,RelationalLDA}.
+
+%One set of models that are particularly relevant to HSLDA are Chang
+%and Blei's hierarchical models for document networks (Relational Topic
+%Models). In that family of models, they encountered a similar scenario
+%where the lack of a link did not truly indicate absence. In hierarchically
+%labeled data, negative labels are uncommon and the lack of a label
+%in the hierarchy is not equivalent to a negative label. Therefore,
+%as in the work of Chang and Blei, we employ regularization to account
+%for the lack of negative labels. This will be discussed further in
+\ref{sec:model}.
+
 % Noemie: check out what I wrote here, because I am not really sure this is
 % correct. 
-There have been several models that incorporate both latent models of text and
-some form of
-supervision~\citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}. One set 
-of models that are particularly relevant to HSLDA are Chang and Blei's
-hierarchical models for document networks (Relational Topic Models). In that
-family of models, they encountered a similar scenario where an unselected label 
-does not always indicate absence. In hierarchical labels, this phenomenon is
-even more pervasive -- there are no explicit negative labels, but it is also
-unclear how to treat the parents of selected labels. Like in the work of Chang
-and Blei, we employ regularization to account for the lack of negative
-labeling. In our experiments, we look at the impact of assigning positive and
-negative instances to the ancestors of selected labels.
+%There have been several models that incorporate both latent models of text and
+%some form of
+%supervision~\citep{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}. One set 
+%of models that are particularly relevant to HSLDA are Chang and Blei's
+%hierarchical models for document networks (Relational Topic Models). In that
+%family of models, they encountered a similar scenario where an unselected label 
+%does not always indicate absence. In hierarchical labels, this phenomenon is
+%even more pervasive -- there are no explicit negative labels, but it is also
+%unclear how to treat the parents of selected labels. Like in the work of Chang
+%and Blei, we employ regularization to account for the lack of negative
+%labeling. In our experiments, we look at the impact of assigning positive and
+%negative instances to the ancestors of selected labels.

### 2011-06-01 paper/experiments.tex
minor changes
updates to multiple sections

@@
-%\begin{itemize}
-%\item Describe Data for both Experiments
-%\item Describe the Four different conditions
-%\item Describe how they are all evaluated, with and without hierarchy
-%\end{itemize}
-
 We experimented with HSLDA for prediction in two domains: predicting medical
@@
 
-\subsubsection{Diagnosis Prediction}
+\subsubsection{Discharge Summaries and ICD-9 Codes}
 
@@
 
-\subsubsection{Product Category Prediction}
+\subsubsection{Product Descriptions and Categorizations}
 
-In this experiment, we look at product descriptions and their
-categorizations according to a product hierarchy. Product ID's
-and categorizations were obtained from the Stanford Network Analysis Platform
-(SNAP) dataset~\citep{SNAP}. Product descriptions were crawled from the
-amazon.com website directly. We were able to deduce the structure of the
-hierarchy for the Amazon.com products directly since all ancestors in the
+Amazon.com, an online retail store, organizes its catalog of 
+products in a hierarchy and provides product descriptions
+for most products in their catolog.  Products can be discovered by users
+through query-based searching or through product category exploration. Top-level
+product categories are displayed on the front page of the website and lower
+level categories can be discovered by choosing one of the top-level categories.
+Products can exist in multiple locations in the hierarchy.
+
+In this experiment, we obtained Amazon.com product categorization data from 
+the Stanford Network Analysis Platform (SNAP) dataset~\citep{SNAP}. 
+Product descriptions were obtained separately from the
+Amazon.com website directly. We limited our dataset to the collection of DVDs
+in the product catalog.
+
+We were able to deduce the structure of the
+hierarchy for the Amazon.com products because all ancestors in the
 hierarchy were included with each category label. For example, {}``DVD / Genres
@@
 on labels ignored), HSLDA fit by performing LDA followed by tree-conditional
-regressions, and HSLDA fit with fixed random regression parameters.
-We will refer to the three comparison models as the sLDA model, the
-separate HSLDA model, and the random HSLDA model, respectively. These
-models were chosen as to highlight performance in absence of hierarchical
-constraints, the effect of the combined inference procedure, and regression
+regressions, and HSLDA fit with fixed random regression parameters. These
+models were chosen to highlight several aspects of the model including performance in the absence 
+of hierarchical constraints, the effect of the combined inference procedure, and regression
 performance attributable solely to the hierarchical constraints.
@@
 %\end{itemize}
-For each dataset, a held out set of 1,000 documents and accompanying
-labels were used for evaluation. The two main methods of evaluation
+
+The two main methods of evaluation
 for the model are prediction and topic quality. To evaluate predictive

### 2011-06-01 paper/related_work.tex
updates to multiple sections

--- 2011-06-01 paper/related_work.tex
@@
 
-Other models that incorporate both and some form of supervision include
-LabeledLDA\citep{Ramage2009}, DiscLDA\citep{DiscLDA}, models applied to 
+Other models that incorporate LDA and supervision include
+LabeledLDA\citep{Ramage2009}, DiscLDA\citep{DiscLDA}, and other models applied to 
 computer vision and document networks\citep{wangbleifeifei08,RelationalLDA}. 
+These models, however, do not implement contraints on the label space.
+
+In other work, researchers have classified documents in a hierarchy with naive Bayes classifiers
+and support vector machines. Most of this work has been demonstrated on relatively
+small datasets, small label spaces and has focused on single label classification without
+a model of documents such as LDA\citep{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.
 
@@
 %for the lack of negative labels. This will be discussed further in
-\ref{sec:model}.
 

### 2011-06-01 paper/abstract.tex

--- 2011-05-29 paper/abstract.tex
@@
-We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a model for hierarchically and multiply-labeled bag-of-words data.
-Out-of-sample label prediction is the primary goal of this work; however, improved dimensionality reduction is also of interest.
-We define a model that uses probit regressors on a conditionally dependent label hierarchy tied to latent Dirichlet allocation (LDA).
-We find that the additional signal that comes from multiple, hierarchically constrained labels substantially improves out-of-sample label prediction in comparison to supervised LDA approaches that don't utilize information derived from the structure of the label space.
-We demonstrate HSLDA on large-scale data from medical document labeling and retail product categorization tasks.
-We show improved label prediction performance and evidence that the learned topic model also improves as a result of using this signal.
+We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
+model for hierarchically and multi-labeled bag-of-word data.  Examples of
+such data include pages and their placement in Web directories, product
+descriptions and associated categories from product hierarchies, and free-text
+clinical records and their assigned diagnosis codes. Out-of-sample label
+prediction is the primary goal of this work, but improved lower-dimensional
+representations of the bag-of-word data is also of interest.
+%We define a model that uses probit regressors on a conditionally dependent
+%label hierarchy tied to latent Dirichlet allocation (LDA). 
+We demonstrate HSLDA on large-scale data from clinical document labeling and
+retail product categorization tasks. We show that leveraging the structure from
+hierarchical labels improves out-of-sample label prediction substantially when
+compared to structure-agnostic models. Furthermore, we show evidence
+that the resulting HSLDA topics are more descriptive of the underlying data
+than sLDA topics, which ignore the label hierarchy.
 

### 2011-06-01 paper/introduction.tex

@@
-There exist many sources of unstructured data that have been partially or
-completely categorized by human editors.  In this paper we focus on
-unstructured textual data that has been, at least in part, manually
-categorized.  Examples include but are not limited to webpages and curated
-hierarchical directories of the same \citep{DMOZ}, product descriptions and
-catalogs~\citep{SNAP}, and patient records and their associated billing codes.
-In this work we show how to combine these two sources of information using a
-single model that allows one to automatically categorize new text documents,
-suggest labels that might be inaccurate, compute improved similarities between
-documents for information retrieval purposes, and more. The models and
-techniques that we develop in this paper are applicable in other domains as
-well, namely, any unstructured representations of data that have been
-hierarchically classified (e.g.~image catalogs with bag-of-feature image
-representations).
+The task of multi-label classification, selecting the k-best labels for a given
+instance, has been a topic of research for several years. One simplistic way to
+carry out the classification is through a series of independent binary
+classifiers, but this ignores the many inherent dependencies among the
+labels. Thus, much work has been devoted on incorporating the co-occurence
+patterns of the labels into the classification task. In this paper, we focus on
+multi-label classification, where the labels are organized in a hierarchical
+structure. Scenarios of use include, but are not limited to, placing webpages
+into manually curated Internet directories~\cite{DMOZ}, categorizing images
+according to a taxonomy, tagging product descriptions with catalogue
+information~\cite{SNAP}, and assigning diagnosis codes to clinical
+records~\cite{Challenge07}.
 
-In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
-sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this supervision is just extra per document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
-These labels are usually generatively modeled as having been conditionally drawn from some distribution that depends on the document-specific topic mixture.
-It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
+There are several challenges entailed in incorporating the hierarchical nature
+of labels into the classification task. One pertains to the labeling itself: in
+the datasets (especially real-world, noisy ones), for a given label, instances
+labeled with it contribute positive instance, but it is unclear how to
+determine the negative instances. In particular, how to treat the parent labels
+of the selected ones?
 
-Our main contribution is to show how to utilize supervision in the form of  hierarchical and (often) multiple labelings in a similar manner.
-Consider web retail data.
-Web retailers often have both a browse-able product hierarchy and free-text descriptions for all products they sell.
-The situation of each product in a product hierarchy (often multiply situated) constitutes a multiple, hierarchical labeling of the free-text product descriptions.
-We hypothesize that such hierarchical labels should, at least in theory, provide better supervision than the simpler unstructured labels previously considered.
-Results from applying our model to both medical record and web retail data suggests that this is likely the case.
-In particular, we observe gains in our primary goal of out-of-sample label prediction that result specifically from leveraging hierarchical supervision.
+Our work operates within the framework of topic modeling. Our approach learns
+topic models of the underlying data and labeling strategies in a joint model,
+while leveraging the hierarchical structure of the labels. In particular, we
+extend supervised latent Dirichlet allocation (sLDA)~\cite{BleiMcAuliffe2008}
+to take advantage of hierarchical supervision. We hypothesize that hierarchical
+label information provides more information about labeling than considering
+labels as a flat list.
 
-The remainder of this paper is structured as follows.
-Section~\ref{sec:model}  introduces hierarchically supervised LDA (HSLDA), Section~\ref{sec:inference} details a sampling approach to inference in HSLDA,  Section~\ref{sec:related_work} reviews related work, and Section~\ref{sec:experiments} shows results from applying HSLDA to health care and web retail data.
+We demonstrate our model on large, real-world datasets in the clinical and web
+retail domains. We observe that hierarchical information is valuable when
+incorporated into the learning and improves our primary goal of multi-label
+classification. Following the trend observed in supervised topic modeling, we
+note that the learned topic models are more representative of the underlying
+data in both of our datasets~\cite{BleiMcAuliffe2008}.
+
+The remainder of this paper is as follows. Section~\ref{sec:model}
+introduces hierarchically supervised LDA (HSLDA), while
+Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
+Section~\ref{sec:related_work} reviews related work, and
+Section~\ref{sec:experiments} shows results from applying HSLDA to health care
+and web retail data.
 
@@
 %diagnoses \citep{Surjan1999}.
+
+%Our main contribution is to show how to utilize supervision in the form of
+%hierarchical and (often) multiple labelings in a similar manner. Consider web
+%retail data. Web retailers often have both a browse-able product hierarchy and
+%free-text descriptions for all products they sell. The situation of each
+%product in a product hierarchy (often multiply situated) constitutes a
+%multiple, hierarchical labeling of the free-text product descriptions. We
+%hypothesize that such hierarchical labels should, at least in theory, provide
+%better supervision than the simpler unstructured labels previously considered.
+%Results from applying our model to both medical record and web retail data
+%suggests that this is likely the case. In particular, we observe gains in our
+%primary goal of out-of-sample label prediction that result specifically from
+%leveraging hierarchical supervision.
 

### 2011-06-02 paper/experiments.tex
updates to experimental section

@@
 
-We applied HSLDA, along with three other closely related models, to
-the clinical data and the retail product data. Specifically, we evaluate
-models including sLDA with independent regressors (hierarchical constraints
-on labels ignored), HSLDA fit by performing LDA followed by tree-conditional
-regressions, and HSLDA fit with fixed random regression parameters. These
-models were chosen to highlight several aspects of the model including performance in the absence 
+We evaluated HSLDA along with three other closely related models against 
+these two datasets. The comparison models included sLDA with independent 
+regressors (hierarchical constraints on labels ignored), HSLDA fit by 
+performing LDA followed by tree-conditional regressions, and HSLDA fit 
+with fixed random regression parameters. These models were chosen to 
+highlight several aspects of the model including performance in the absence 
 of hierarchical constraints, the effect of the combined inference procedure, and regression
@@
 
+sLDA with independent regressors is the most salient comparison model
+for our work. The distinguishing factor between HSLDA and sLDA is the
+addition structure imposed on the label space, a distinction that we
+hypothesize will result in a difference in predictive performance. 
+Aside from this, all other features of sLDA are preserved between the two models.
+
+There are two components to HSLDA, LDA and a hierarchically constrained response.
+The second comparison model is HSLDA fit by performing LDA first followed by
+performing inference over the hierarchically constrained label space. In this
+comparison model, the separate inference processes do not allow the responses
+to influence the low dimensional structure inferred by LDA. Combined inference
+has been shown to improve performance in sLDA \citep{BleiMcAuliffe2008}. This
+comparison model examines not the structuring of the label space, but the benefit
+of combined inference over both the documents and the label space.
+
+The last comparison model is HSLDA with fixed and randomly selected regression
+parameters. There is a baseline benefit that the structure in the label space 
+provides for the prediction of labels. This comparison model is intended to
+quantify the contribution of the structure alone.
+
 We evaluated model performance for all three models with a range of
-values for $\mu$ ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\} )$,
-the mean prior parameter for regression parameters.
+values for $\mu$, the mean prior parameter for regression parameters 
+($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$). The number of topics for all
+models was set to 50, the parameters for the prior distributions
+$\alpha$, $\alpha^prime$, and $\gamma$ were a shape parameter of 1
+and a scale parameters of 1000.
 
-\subsection{Evaluation}
+\subsection{Evaluation and Results}
 
-%\begin{itemize}
-%\item Talk about hierarchy and no hierarchy
-%\item Make explicit calculation of sensitivity and 1-specificity
-%\end{itemize}
+%Outline:
+%1. Model evaluation goal, metrics
+%2. Gold standard(s), properties
+%3. Method for prediction w/ details
+%4. Discussion of results
+%5. Connection of results with contribution and generality etc
 
-The two main methods of evaluation
-for the model are prediction and topic quality. To evaluate predictive
-performance for all comparison models equivalently, each model was
-evaluated with two methods. The first evaluation method augments the
-observed labels in the held out set with their ancestors and considers
-all other non-existant labels to be negative. The second method ignores
-the ancestors of the observed labels in the held out set and considers
-all other non-existant labels to be negative. This uniform treatment
-of ancestors allows for a fair comparison of the models.
+We evaluated our model, HSLDA, against the comparison models with a focus
+on predictive performance. Predictive performance was measured with
+standard metrics of performance - sensitiviy (true positive rate) and 
+1-specificity (false positive rate). We evaluate on the two aforementioned
+datasets to demonstrate that model performance generalizes over two very disparate
+domains.
 
-The two measures for predictive performance used here include the
-true positive rate and the false positive rate evaluated based on
-$p\left(y_{l,\hat{d}}\mid w_{1:N,d}\right)$ for each label in each
-model.
+To evaluate the performance of these models, we establish a gold standard for comparison.
+In our evaluation of each dataset, a held out set of 1000 documents and labels were
+reserved for evaluation and predictive performance was evaluated against the observed
+labeling. Specifically, ancestors of observed nodes were ignored, observed nodes 
+were considered positive and unobserved nodes were considered to be negative. The gold
+standard defined in this way will likely lead to a slight overestimation of the number of
+false positives. It is known that ICD-9 codes lack sensitivity and their use as a
+gold standard could lead to correctly positive predictions being labeled as false positives.
+However, given that the label space is often large (as in our examples) these potentially
+erroneous false positives should not skew results significantly.
+
+Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ where $\hat{d}$ represents the test document.
+For efficiency,
+the expectation of this probability distribution was estimated in the following way. Expectations
+of $\mathbf{\bar{z}}_d$ and $\boldsymbol{\eta}_l$ were estimated with samples from the posterior.
+Using these expectations, we performed Gibbs sampling over the hierarchy to acquire predictive
+samples for the documents in the test set.

### 2011-06-02 paper/experiments.tex
minor updates to many sections

@@
 ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$). The number of topics for all
-models was set to 50, the parameters for the prior distributions
-$\alpha$, $\alpha^prime$, and $\gamma$ were a shape parameter of 1
+models was set to 50, the prior distributions of
+$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and $p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1
 and a scale parameters of 1000.

### 2011-06-02 paper/model.tex
minor updates to many sections

@@
 Here each document is generatively labeled using a hierarchy of conditionally dependent probit regressors \cite{gelman?,probit_regression?}.
-For every label $l \in \mathcal{L}$ both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to decide whether or not  label $l$ is to be applied to  document $d$ as well.
+For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.
 Note that label $y_{l,d}$ can only be applied to document $d$ if its parent label $\mathrm{pa}(l)$ is also applied (these expressions are specific to is-a constraints but can be modified to accommodate different constraints).

### 2011-06-02 paper/related_work.tex
minor updates to many sections

@@
 \cite{Blei2003} augmented with per-document labeling, often taking the form of
-a single numerical or categorical label. Examples of labels include rating
-associated with an online  eview, grades for an essay, and number of times a
+a single numerical or categorical label. Examples of labels include ratings
+associated with online reviews, grades for essays, and the number of times a
 webpage is linked. This approach has been shown to outperform both LASSO
@@
 
-In other work, researchers have classified documents in a hierarchy with naive Bayes classifiers
+In other work, researchers have classified documents into a hierarchy with naive Bayes classifiers
 and support vector machines. Most of this work has been demonstrated on relatively

### 2011-06-02 paper/experiments.tex
Changes to experiment section mostly

@@
 
-We evaluated model performance for all three models with a range of
-values for $\mu$, the mean prior parameter for regression parameters 
-($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$). The number of topics for all
+For all of these models, particular attention was payed to the settings of the
+prior parameters for the regression coefficients. These parameters implement an
+important form of regularization in HSLDA. In the setting where there are no negative
+labels, a Gaussian prior over the regression parameters with a negative mean implements
+a prior belief that missing labels are likely to be negative. We evaluated model performance 
+for all three models with a range of values for $\mu$, the mean prior parameter for regression coefficients
+($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$).
+
+The number of topics for all
 models was set to 50, the prior distributions of

### 2011-06-02 paper/experiments.tex
new figures

@@
 labeling. Specifically, ancestors of observed nodes were ignored, observed nodes 
-were considered positive and unobserved nodes were considered to be negative. The gold
-standard defined in this way will likely lead to a slight overestimation of the number of
+were considered positive and unobserved nodes were considered to be negative. This method of 
+defining positive and negative labels was chosen to be as fair as possible to all models
+being compared.  In particular, since the sLDA model does not enforce the hierarchical
+constraints, the models can be compared on a more equal footing by focusing on the observed labels
+as positives. The gold standard defined in this way will likely lead to a slight overestimation of the number of
 false positives. It is known that ICD-9 codes lack sensitivity and their use as a
 gold standard could lead to correctly positive predictions being labeled as false positives.
-However, given that the label space is often large (as in our examples) these potentially
+However, given that the label space is often large (as in our examples) it is a moderate assumption that
 erroneous false positives should not skew results significantly.

### 2011-06-02 paper/abstract.tex
abstract and intro, changed figures, added a bit to experiments

@@
 model for hierarchically and multi-labeled bag-of-word data.  Examples of
-such data include pages and their placement in Web directories, product
+such data include Web pages and their placement in directories, product
 descriptions and associated categories from product hierarchies, and free-text
@@
 hierarchical labels improves out-of-sample label prediction substantially when
-compared to structure-agnostic models. Furthermore, we show evidence
-that the resulting HSLDA topics are more descriptive of the underlying data
-than sLDA topics, which ignore the label hierarchy.
+compared to models that do not. %Furthermore, we show evidence
+%that the resulting HSLDA topics are more descriptive of the underlying data
+%than sLDA topics, which ignore the label hierarchy.
 

### 2011-06-02 paper/experiments.tex
abstract and intro, changed figures, added a bit to experiments

@@
 samples for the documents in the test set.
+
+\begin{figure}[htbp]
+\begin{center}
+%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
+%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
+\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  
+In all figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical 
+constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
+tree-conditional regressions. \subref{} shows predictive performance as a function of the 
+auxiliary variable threshold and \subref{1b} shows predictive perfmance as a function of the
+prior mean on regression parameters.}
+\label{fig:main_clinical_results}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
+\caption{Out-of-sample Amazon product category predictions from product free-text descriptions.  
+In all figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints 
+on labels ignored), and dotted is HSLDA fit by running LDA first then running tree-conditional 
+regressions. \subref{} shows predictive performance as a function of the 
+auxiliary variable threshold and \subref{1b} shows predictive perfmance as a function of the
+prior mean on regression parameters.}
+\label{fig:main_amazon_results}
+\end{center}
+\end{figure}

### 2011-06-02 paper/introduction.tex
abstract and intro, changed figures, added a bit to experiments

--- 2011-05-31 paper/introduction.tex
@@
 
-In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
+%In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
 sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical ``label.''  More generally this supervision is just extra per document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
@@
 
-The remainder of this paper is structured as follows.
-Section~\ref{sec:model}  introduces hierarchically supervised LDA (HSLDA), Section~\ref{sec:inference} details a sampling approach to inference in HSLDA,  Section~\ref{sec:related_work} reviews related work, and Section~\ref{sec:experiments} shows results from applying HSLDA to health care and web retail data.
+The remainder of this paper is as follows. Section~\ref{sec:model}
+introduces hierarchically supervised LDA (HSLDA), while
+Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
+Section~\ref{sec:related_work} reviews related work, and
+Section~\ref{sec:experiments} shows results from applying HSLDA to health care
+and web retail data.
 

### 2011-06-02 paper/main.tex
updates to experimental section
turned off author names
abstract and intro, changed figures, added a bit to experiments

--- 2011-06-01 paper/main.tex
@@
 
-\nipsfinalcopy
+%\nipsfinalcopy
 
@@
 
-\section{Results}
-\label{sec:results}
-
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
-\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
-\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
-\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.
+%\begin{figure}[htbp]
+%\begin{center}
+%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors}}
+%\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs}}
+%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
+%\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.
 In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.
@@
 Bottom row: \subref{1c} are the sensitivity curves from \subref{1b} aligned on threshold value,  \subref{1d} are the 1-specificity curves from \subref{1b} aligned on threshold value.}
-\label{fig:icd9}
-\end{center}
-\end{figure}
+%\label{fig:icd9}
+%\end{center}
+%\end{figure}
 
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[][\label{2a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors_amazon}}
-\subfigure[][\label{2b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs_amazon}}
-\subfigure[][\label{2c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs_amazon}}
-\subfigure[][\label{2d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
-\caption{Out-of-sample Amazon product code predictions from product free-text descriptions.
+%\begin{figure}[htbp]
+%\begin{center}
+%\subfigure[][\label{2a}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_ancestors_amazon}}
+%\subfigure[][\label{2b}]{\includegraphics[width=.49\textwidth]{figs/ROC_comparison_leafs_amazon}}
+%\subfigure[][\label{2c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs_amazon}}
+%\subfigure[][\label{2d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
+%\caption{Out-of-sample Amazon product code predictions from product free-text descriptions.
 In all figures solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints on labels ignored), dotted is HSLDA fit by running LDA first then running tree-conditional regressions, and dot-dashed is HSLDA fit with fixed random regression parameters.
@@
 Bottom row: \subref{2c} are the sensitivity curves from \subref{2b} aligned on threshold value,  \subref{2d} are the 1-specificity curves from \subref{2b} aligned on threshold value.}
-\label{fig:icd9}
-\end{center}
-\end{figure}
+%\label{fig:icd9}
+%\end{center}
+%\end{figure}
 

### 2011-06-02 paper/related_work.tex
f'ed around with related work

--- 2011-06-01 paper/related_work.tex
@@
-While there has been much work in multi-label classification of text and
-text modeling in general, we focus here on topic modeling approaches.
-Latent Dirichlet allocation (LDA) is a generative probabilistic model which
-represents documents as a mixed-membership bag of word. Each document is
-represented as a collection of words, generated from a set of topic assignments
-(one for each word), where each topic assignment is drawn from a distribution
-over topics~\citep{Blei2003}. sLDA is latent Dirichlet allocation (LDA)
-\cite{Blei2003} augmented with per-document labeling, often taking the form of
-a single numerical or categorical label. Examples of labels include rating
-associated with an online  eview, grades for an essay, and number of times a
-webpage is linked. This approach has been shown to outperform both LASSO
+In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
+sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical label.
+More generally this supervision is just extra per-document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
+These labels are usually generatively modeled as a conditional draw from some distribution that depends on each document-specific topic mixture.
+It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
+It also has been demonstrated that sLDA has been shown to outperform both LASSO
 (L1 regularized least squares regression) and LDA followed by least
 squares regression~\cite{BleiMcAuliffe2008}.
+sLDA can be applied to data of the type we consider in this paper; however, doing so requires ignoring the hierarchical dependencies amongst the labels.
+In Section~\ref{sec:experiments} we constrast HSLDA with sLDA applied in this way.
+
+%While there has been much work in multi-label classification of text and
+%text modeling in general, we focus here on topic modeling approaches.
+%Latent Dirichlet allocation (LDA) is a generative probabilistic model which
+%represents documents as a mixed-membership bag of word. Each document is
+%represented as a collection of words, generated from a set of topic assignments
+%(one for each word), where each topic assignment is drawn from a distribution
+%over topics~\citep{Blei2003}. sLDA is latent Dirichlet allocation (LDA)
+%\cite{Blei2003} augmented with per-document labeling, often taking the form of
+%a single numerical or categorical label. Examples of labels include ratings
+%associated with online reviews, grades for essays, and the number of times a
+%webpage is linked. This approach has been shown to outperform both LASSO
+%(L1 regularized least squares regression) and LDA followed by least
+%squares regression~\cite{BleiMcAuliffe2008}.
 
@@
 Other models that incorporate LDA and supervision include
-LabeledLDA\citep{Ramage2009}, DiscLDA\citep{DiscLDA}, and other models applied to 
-computer vision and document networks\citep{wangbleifeifei08,RelationalLDA}. 
-These models, however, do not implement contraints on the label space.
+LabeledLDA\citep{Ramage2009} and DiscLDA\citep{DiscLDA}.  Various applications of these models to 
+computer vision and document networks have been explored \citep{wangbleifeifei08,RelationalLDA} . 
+None of these models, however, leverage dependency structure in the label space.
 
-In other work, researchers have classified documents in a hierarchy with naive Bayes classifiers
+In other work, researchers have classified documents into a hierarchy (a closely related task) with naive Bayes classifiers
 and support vector machines. Most of this work has been demonstrated on relatively
-small datasets, small label spaces and has focused on single label classification without
+small datasets, small label spaces, and has focused on single label classification without
 a model of documents such as LDA\citep{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.

### 2011-06-02 paper/discussion.tex
Wrote a discussion.

--- 2011-05-22 paper/discussion.tex
@@
-\begin{itemize}
-\item what about the nonparametric version of this?
-\end{itemize}
+%Despite the proliferation of LDA related models in recent years, one could argue that the true killer app for LDA has yet to be discovered.
+LDA makes sense for information retrieval (IR) because  the learned latent topic vectors serve as good low-dimensional document representations, and, a posteriori similarity between such vectors can be used to compute the similarity of documents.
+Unfortunately, computational considerations render this approach to IR impractical in most settings.
+
+The SLDA model family, of which HSLDA is a member, can be understood in two different ways.
+One way is to see it as a family of topic models that improve on the topic modeling performance of LDA via the inclusion of observed supervision.
+Another way to see the family is as a set of models that can predict labels for bag-of-words data.
+A large diversity of problems can be expressed as label prediction problems for bag-of-words data.
+A surprisingly large amount of that kind of data possess structured labels, either hierarchically constrained or otherwise.
+That HSLDA directly addresses this kind of data is a large part of the motivation for this work.
+That it outperforms more straightforward approaches should be of interest to practitioners.
+
+Extensions to this work include unbounded topic cardinality variants and relaxations to different kinds of label structure.
+Unbounded topic cardinality variants pose interesting inference challenges.
+Utilizing different kinds of label structure is possible within this framework, but requires relaxing some of the simplifications we made in this paper for expositional purposes.
+
+%We have described a mixed membership model with hierarchical supervision.
+We have demonstrated this model in the context of document modeling with hierarchical multi-label supervision.
+Such a model is appropriate in domains where there are hierarchical constraints among the labels such as is the case in an IS-A hierarchy.
+
+...
+
+%\begin{itemize}
+%\item what about the nonparametric version of this?
+%\item discuss the broader goal, from the beginning of search engine time, to combine categorization and free text.  this, to our knowledge, is the first principled approach to doing so
+%\end{itemize}

### 2011-06-02 paper/experiments.tex
Added to experiments section

@@
 
+The following are the predictive performance results for the clinical
+data given a prior mean for the regression parameters of -1.6. 
+The full HSLDA model had a true positive rate of 0.57 
+and a false positive rate of 0.13, the sLDA model had a true positive
+rate of 0.42 and a false positive rate of 0.07, and the HSLDA model where
+LDA and the regressions were fit separately had a true positive rate of 0.39
+and a false positive rate of 0.08.
+
+These results indicate that the full HSLDA model predicts more of the the
+correct labels at a cost of an increase in the number of false positives
+relative to the comparison models.
+
+The following are the predictive performance results for the retail product
+data given a prior mean for the regression parameters of -2.2. 
+The full HSLDA model had a true positive rate of 0.85 
+and a false positive rate of 0.30, the sLDA model had a true positive
+rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
+LDA and the regressions were fit separately had a true positive rate of 0.77
+and a false positive rate of 0.16. These results follow a similar pattern to the clinical data.
+
+To further explore this tradeoff between the true positive rate and the
+false positive rate we evaluated predictive performance for a range of values
+for two different parameters - the prior mean for the regression coefficients
+and the threshold for the auxiliary variables.  The goal in this analysis
+was to evaluate the performance of these models subject to more or less
+stringent requirements for predicting positive labels. These two parameters 
+have important related functions in the model. The prior mean in combination 
+with the auxiliary variable threshold together encode the strength of the prior
+belief that unobserved labels are likely to be negative. Effectively, the
+prior mean applies negative pressure to the predictions and the auxiliary
+variable threshold determines the cutoff.
+
+For each model type, separate models were fit for each value of the 
+prior mean of the regression coefficients.  In contrast, to evaluate
+predictive performance as a function of the auxiliary variable threshold,
+a single model was fit for each model type and prediction was evaluated
+based on predictive samples drawn subject to different auxiliary variable
+thresholds. These methods are significantly different since the prior mean
+is varied prior to inference and the auxiliary variable threshold is varied
+after inference.  However, as intended, they both highlight model performance
+under more or less stringent requirements for predicting positive labels.
+
+Figures \ref{fig:main_clinical_results} and \ref{fig:main_amazon_results} show the predictive performance
+of HSLDA relative to the three comparison models.
+
 \begin{figure}[htbp]

### 2011-06-02 paper/experiments.tex
slight changes

@@
 on predictive performance. Predictive performance was measured with
-standard metrics of performance - sensitiviy (true positive rate) and 
+standard metrics - sensitiviy (true positive rate) and 
 1-specificity (false positive rate). We evaluate on the two aforementioned
-datasets to demonstrate that model performance generalizes over two very disparate
+datasets to demonstrate that our model generalizes to two different
 domains.
 
-To evaluate the performance of these models, we establish a gold standard for comparison.
-In our evaluation of each dataset, a held out set of 1000 documents and labels were
-reserved for evaluation and predictive performance was evaluated against the observed
-labeling. Specifically, ancestors of observed nodes were ignored, observed nodes 
+To evaluate the performance of these models, we established a gold standard for comparison.
+For each dataset, a held out set of 1000 documents and labels were
+reserved for evaluation and predictive performance was evaluated against a standard derived
+from the observed labeling. Specifically, ancestors of observed nodes were ignored, observed nodes 
 were considered positive and unobserved nodes were considered to be negative. This method of 
@@
 being compared.  In particular, since the sLDA model does not enforce the hierarchical
-constraints, the models can be compared on a more equal footing by focusing on the observed labels
-as positives. The gold standard defined in this way will likely lead to a slight overestimation of the number of
+constraints, the models can be compared on a more equal footing by considering only on the observed labels
+as being positive despite the fact that ancestors must also be positive. The gold standard 
+defined in this way will likely lead to a slight overestimation of the number of
 false positives. It is known that ICD-9 codes lack sensitivity and their use as a
@@
 
-Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ where $\hat{d}$ represents the test document.
-For efficiency,
+Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
+where $\hat{d}$ represents the test document. For efficiency,
 the expectation of this probability distribution was estimated in the following way. Expectations
@@
 is varied prior to inference and the auxiliary variable threshold is varied
-after inference.  However, as intended, they both highlight model performance
+following inference.  However, as intended, they both highlight model performance
 under more or less stringent requirements for predicting positive labels.
 
-Figures \ref{fig:main_clinical_results} and \ref{fig:main_amazon_results} show the predictive performance
-of HSLDA relative to the three comparison models.
+Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
+relative to the three comparison models. In figure %\ref{fig:1b}.
 
-\begin{figure}[htbp]
-\begin{center}
-%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
-%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
-\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records.  
-In all figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical 
-constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
-tree-conditional regressions. \subref{} shows predictive performance as a function of the 
-auxiliary variable threshold and \subref{1b} shows predictive perfmance as a function of the
-prior mean on regression parameters.}
-\label{fig:main_clinical_results}
-\end{center}
-\end{figure}
+%\begin{figure}%[htbp]
+%\begin{center}
+%%\subfloat[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+%\subfigure[][]{\label{1b}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
+%%\subfloat[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+%\subfigure[][]{\label{1d}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+%\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records 
+%(\subref{} and \subref{1b}). Out-of-sample Amazon product category predictions from 
+%product free-text descriptions (\subref{} and \subref{1d}). In all figures, solid is 
+%HSLDA, dashed are independent regressors + sLDA (hierarchical 
+%constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
+%tree-conditional regressions. \subref{} and \subref{} show predictive performance as a function of the 
+%auxiliary variable threshold and \subref{1b} and \subref{1d} show predictive perfmance as a function of the
+%prior mean on regression coefficients.}
+%\label{fig:main_results}
+%\end{center}
+%\end{figure}
 
-\begin{figure}[htbp]
-\begin{center}
-%\subfigure[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][\label{1b}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-%\subfigure[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-%\subfigure[][\label{1d}]{\includegraphics[width=.49\textwidth]{figs/one_m_spec_comparison_leafs}}
-\caption{Out-of-sample Amazon product category predictions from product free-text descriptions.  
-In all figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical constraints 
-on labels ignored), and dotted is HSLDA fit by running LDA first then running tree-conditional 
-regressions. \subref{} shows predictive performance as a function of the 
-auxiliary variable threshold and \subref{1b} shows predictive perfmance as a function of the
-prior mean on regression parameters.}
-\label{fig:main_amazon_results}
-\end{center}
-\end{figure}
+%\begin{figure}[t]
+%\centering
+%\parbox{0.9\columnwidth}{
+% \vspace*{-\subfigtopskip}
+% \subfigure[]{
+%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
+%       \textsf{TEMPLATE Id: 12.01\\
+%         FILE No: ahj\_137\_02\_0424\\
+%         SENTENCE No: S-98\\
+%         RELATION: prediction\\
+%         ANALYSIS TYPE: independent\\
+%         FINDING(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0155626\\
+%         \mbox{~~~~~~}LEX: new acute myocardial infarction\\
+%         PARAMETER(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0520887\\
+%         \mbox{~~~~~~}LEX: ST-segment depression
+%       }}}
+% }\goodgap
+% \subfigure[]{
+%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
+%       \textsf{TEMPLATE Id: 12.02\\
+%         FILE No: ahj\_137\_02\_0424\\
+%         SENTENCE No: S-98\\
+%         RELATION: prediction\\
+%         ANALYSIS TYPE: independent\\
+%         FINDING(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0155626\\
+%         \mbox{~~~~~~}LEX: new acute myocardial infarction\\
+%         PARAMETER(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0002962\\
+%         \mbox{~~~~~~}LEX: prior angina%
+%       }}}
+% }\\     \vspace*{-\subfigtopskip}
+% \subfigure[]{
+%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
+%       \textsf{TEMPLATE Id: 12.03\\
+%         FILE No: ahj\_137\_02\_0424\\
+%         SENTENCE No: S-98\\
+%         RELATION: prediction\\
+%         ANALYSIS TYPE: independent\\
+%         FINDING(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0011065\\
+%         \mbox{~~~~~~}LEX: death\\
+%         PARAMETER(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0520887\\
+%         \mbox{~~~~~~}LEX: ST-segment depression%
+%       }}}
+% }\goodgap
+% \subfigure[]{
+%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
+%       \textsf{TEMPLATE Id: 12.04\\
+%         FILE No: ahj\_137\_02\_0424\\
+%         SENTENCE No: S-98\\
+%         RELATION: prediction\\
+%         ANALYSIS TYPE: independent\\
+%         FINDING(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0011065\\
+%         \mbox{~~~~~~}LEX: death\\
+%         PARAMETER(S):\\
+%         \mbox{~~~}ITEM: \\
+%         \mbox{~~~~~~}CUI: C0002962\\
+%         \mbox{~~~~~~}LEX: prior angina%
+%       }}}
+% }}
+%\caption{A Content Item after atomization.}\label{fig:org-afteratom}
+%\end{figure}

### 2011-06-02 paper/abstract.tex
subfigure intro (not finished)

@@
 We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
-model for hierarchically and multi-labeled bag-of-word data.  Examples of
-such data include Web pages and their placement in directories, product
+model for hierarchically and multiply labeled bag-of-word data.  Examples of
+such data include web pages and their placement in directories, product
 descriptions and associated categories from product hierarchies, and free-text

### 2011-06-02 paper/experiments.tex
subfigure intro (not finished)

@@
 Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
-relative to the three comparison models. In figure %\ref{fig:1b}.
+relative to the three comparison models. In figure \ref{fig:1b}.
 
-%\begin{figure}%[htbp]
-%\begin{center}
-%%\subfloat[][\label{1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-%\subfigure[][]{\label{1b}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
-%%\subfloat[][\label{1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-%\subfigure[][]{\label{1d}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-%\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records 
-%(\subref{} and \subref{1b}). Out-of-sample Amazon product category predictions from 
-%product free-text descriptions (\subref{} and \subref{1d}). In all figures, solid is 
-%HSLDA, dashed are independent regressors + sLDA (hierarchical 
-%constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
-%tree-conditional regressions. \subref{} and \subref{} show predictive performance as a function of the 
-%auxiliary variable threshold and \subref{1b} and \subref{1d} show predictive perfmance as a function of the
-%prior mean on regression coefficients.}
-%\label{fig:main_results}
-%\end{center}
-%\end{figure}
+\begin{figure}%[htbp]
+\begin{center}
+%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
+%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+\subfigure[][]{\label{fig:1d}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records 
+(\subref{1a} and \subref{fig:1b}). Out-of-sample Amazon product category predictions from 
+product free-text descriptions (\subref{1c} and \subref{fig:1d}). In all figures, solid is 
+HSLDA, dashed are independent regressors + sLDA (hierarchical 
+constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
+tree-conditional regressions. \subref{} and \subref{} show predictive performance as a function of the 
+auxiliary variable threshold and \subref{fig:1b} and \subref{fig:1d} show predictive perfmance as a function of the
+prior mean on regression coefficients.}
+\label{fig:main_results}
+\end{center}
+\end{figure}
 

### 2011-06-02 paper/introduction.tex
subfigure intro (not finished)

--- 2011-06-01 paper/introduction.tex
@@
-The task of multi-label classification, selecting the k-best labels for a given
-instance, has been a topic of research for several years. One simplistic way to
-carry out the classification is through a series of independent binary
-classifiers, but this ignores the many inherent dependencies among the
-labels. Thus, much work has been devoted on incorporating the co-occurence
-patterns of the labels into the classification task. In this paper, we focus on
-multi-label classification, where the labels are organized in a hierarchical
-structure. Scenarios of use include, but are not limited to, placing webpages
-into manually curated Internet directories~\cite{DMOZ}, categorizing images
-according to a taxonomy, tagging product descriptions with catalogue
-information~\cite{SNAP}, and assigning diagnosis codes to clinical
-records~\cite{Challenge07}.
+There exist many sources of unstructured data that have been partially or
+completely categorized by human editors.  In this paper we focus on
+unstructured textual data that has been, at least in part, manually
+categorized.  Examples include but are not limited to webpages and curated
+hierarchical directories of the same \citep{DMOZ}, product descriptions and
+catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
+and patient records %hospital treatment transcripts
+and diagnosis codes assigned to them for bookkeeping and insurance purposes
+(e.g.~hospital discharge summaries with 
+International Classification of Disease 9th Revision, Clinical Modification
+(ICD-9-CM) codes assigned \cite{ICD9}).  In this work we show how to combine
+these two sources of information using a single model that allows one to
+categorize new text documents automatically, suggest labels that might be
+inaccurate, compute improved similarities between documents for information
+retrieval purposes, and more. The models and techniques that we develop in
+this paper are applicable in other domains as well, namely, any unstructured
+representations of data that have been hierarchically classified (e.g.,~image
+catalogs with bag-of-feature image representations).
 
-There are several challenges entailed in incorporating the hierarchical nature
-of labels into the classification task. One pertains to the labeling itself: in
-the datasets (especially real-world, noisy ones), for a given label, instances
-labeled with it contribute positive instance, but it is unclear how to
-determine the negative instances. In particular, how to treat the parent labels
-of the selected ones?
+There are several challenges entailed in incorporating a hierarchy of labels.
+Given a large set of potential labels (often thousands), each instance has only
+a small number of labels associated to it. There are no negative labeling
+present in the data naturally, and the absence of a label cannot always be
+interpreted as a negative labeling.
+
+%The task of multi-label classification, selecting the k-best labels for a given
+%instance, has been a topic of research for several years. One simplistic way to
+%carry out the classification is through a series of independent binary
+%classifiers, but this ignores the many inherent dependencies among the
+%labels. Thus, much work has been devoted on incorporating the co-occurence
+%patterns of the labels into the classification task.
 
@@
 label information provides more information about labeling than considering
-labels as a flat list.
+unstructured labels.
 
@@
 incorporated into the learning and improves our primary goal of multi-label
-classification. Following the trend observed in supervised topic modeling, we
-note that the learned topic models are more representative of the underlying
-data in both of our datasets~\cite{BleiMcAuliffe2008}.
+classification. %in two ways
 
@@
 %dozen topics.
+
+%In this work we extend supervised latent Dirichlet allocation (sLDA)
+%\cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA
+%is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per
+%document ``supervision'';  often taking the form of a single numerical or
+%categorical ``label.''  More generally this supervision is just extra per
+%document data;  for instance its quality or relevance (e.g.~online review
+%scores), marks given to written work (e.g.~essay grades), or the number of
+%times a web page is linked.  These labels are usually generatively modeled as
+%having been conditionally drawn from some distribution that depends on the
+%document-specific topic mixture.  It has been demonstrated that the signal
+%provided by such supervision can result in better, task-specific document
+%models and can also lead to good label prediction for out-of-sample data
+%\cite{BleiMcAuliffe2008}.

### 2011-06-02 paper/experiments.tex
figure

@@
 Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
-relative to the three comparison models. In figure \ref{fig:1b}.
+relative to the three comparison models as a function of the prior mean
+on regression coefficient. Figure \ref{fig:1a} demo.
 
@@
 %\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
+\subfigure[][]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
 %\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[][]{\label{fig:1d}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
 \caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records 
-(\subref{1a} and \subref{fig:1b}). Out-of-sample Amazon product category predictions from 
-product free-text descriptions (\subref{1c} and \subref{fig:1d}). In all figures, solid is 
+(\subref{fig:1b}). Out-of-sample Amazon product category predictions from 
+product free-text descriptions (\subref{fig:1b}). In both figures, solid is 
 HSLDA, dashed are independent regressors + sLDA (hierarchical 
 constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
-tree-conditional regressions. \subref{} and \subref{} show predictive performance as a function of the 
-auxiliary variable threshold and \subref{fig:1b} and \subref{fig:1d} show predictive perfmance as a function of the
-prior mean on regression coefficients.}
+tree-conditional regressions.}
 \label{fig:main_results}
@@
 
-%\begin{figure}[t]
-%\centering
-%\parbox{0.9\columnwidth}{
-% \vspace*{-\subfigtopskip}
-% \subfigure[]{
-%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
-%       \textsf{TEMPLATE Id: 12.01\\
-%         FILE No: ahj\_137\_02\_0424\\
-%         SENTENCE No: S-98\\
-%         RELATION: prediction\\
-%         ANALYSIS TYPE: independent\\
-%         FINDING(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0155626\\
-%         \mbox{~~~~~~}LEX: new acute myocardial infarction\\
-%         PARAMETER(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0520887\\
-%         \mbox{~~~~~~}LEX: ST-segment depression
-%       }}}
-% }\goodgap
-% \subfigure[]{
-%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
-%       \textsf{TEMPLATE Id: 12.02\\
-%         FILE No: ahj\_137\_02\_0424\\
-%         SENTENCE No: S-98\\
-%         RELATION: prediction\\
-%         ANALYSIS TYPE: independent\\
-%         FINDING(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0155626\\
-%         \mbox{~~~~~~}LEX: new acute myocardial infarction\\
-%         PARAMETER(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0002962\\
-%         \mbox{~~~~~~}LEX: prior angina%
-%       }}}
-% }\\     \vspace*{-\subfigtopskip}
-% \subfigure[]{
-%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
-%       \textsf{TEMPLATE Id: 12.03\\
-%         FILE No: ahj\_137\_02\_0424\\
-%         SENTENCE No: S-98\\
-%         RELATION: prediction\\
-%         ANALYSIS TYPE: independent\\
-%         FINDING(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0011065\\
-%         \mbox{~~~~~~}LEX: death\\
-%         PARAMETER(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0520887\\
-%         \mbox{~~~~~~}LEX: ST-segment depression%
-%       }}}
-% }\goodgap
-% \subfigure[]{
-%   \fbox{\parbox{0.35\columnwidth}{\scriptsize
-%       \textsf{TEMPLATE Id: 12.04\\
-%         FILE No: ahj\_137\_02\_0424\\
-%         SENTENCE No: S-98\\
-%         RELATION: prediction\\
-%         ANALYSIS TYPE: independent\\
-%         FINDING(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0011065\\
-%         \mbox{~~~~~~}LEX: death\\
-%         PARAMETER(S):\\
-%         \mbox{~~~}ITEM: \\
-%         \mbox{~~~~~~}CUI: C0002962\\
-%         \mbox{~~~~~~}LEX: prior angina%
-%       }}}
-% }}
-%\caption{A Content Item after atomization.}\label{fig:org-afteratom}
-%\end{figure}
+\begin{figure}[t]
+%tbp] %  figure placement: here, top, bottom, or page
+ \centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC Curve for clinical data}

### 2011-06-02 paper/discussion.tex

@@
 That it outperforms more straightforward approaches should be of interest to practitioners.
+
+Variational Bayes has been the predominant estimation approach applied to sLDA models.
+Hierarchical probit regression makes for tractable Markov chain Monte Carlo SLDA inference, a benefit that should extend to other sLDA models should probit regression be used for response variable prediction there too.
+
+The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases it is better to do full joint estimation of HSLDA.
+In alternative interpretation of the same results is that if one is more sensitive to the performance gains that result from exploiting the structure of the labels then one can, in an engineering sense, get nearly as much gain in label prediction performance by first fitting LDA and then fitting a hierarchical probit regression.
+There are applied settings in which this could be advantageous.
 

### 2011-06-02 paper/introduction.tex

@@
 
-There are several challenges entailed in incorporating a hierarchy of labels.
-Given a large set of potential labels (often thousands), each instance has only
-a small number of labels associated to it. There are no negative labeling
-present in the data naturally, and the absence of a label cannot always be
-interpreted as a negative labeling.
+There are several challenges entailed in incorporating a hierarchy of labels
+into the model. Given a large set of potential labels (often thousands), each
+instance has only a small number of labels associated to it. There are no
+naturally occurring negative labeling in the data, and the absence of a label
+cannot always be interpreted as a negative labeling.
 
@@
 topic models of the underlying data and labeling strategies in a joint model,
-while leveraging the hierarchical structure of the labels. In particular, we
-extend supervised latent Dirichlet allocation (sLDA)~\cite{BleiMcAuliffe2008}
-to take advantage of hierarchical supervision. We hypothesize that hierarchical
-label information provides more information about labeling than considering
-unstructured labels.
+while leveraging the hierarchical structure of the labels. For the sake of
+simplicity, we focus on ISA hierarchies, but the model can be
+applied to other hierarchies. We extend supervised latent Dirichlet
+allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
+supervision. We propose an efficient way to incorporate hierarchical
+information into the model. We hypothesize that the context of labels within
+the hierarchy provides valuable information about labeling.
 
@@
 incorporated into the learning and improves our primary goal of multi-label
-classification. %in two ways
+classification. Our results show that a joint, hierarchical model outperforms a
+classification with unstructured labels as well as a disjoint model, where the
+topic modeling and the hierarchical classification are carried out
+independently of each other.
 

### 2011-06-02 paper/abstract.tex
misc

@@
 prediction is the primary goal of this work, but improved lower-dimensional
-representations of the bag-of-word data is also of interest.
+representations of the bag-of-word data are also of interest.
 %We define a model that uses probit regressors on a conditionally dependent

### 2011-06-02 paper/introduction.tex
misc

@@
 completely categorized by human editors.  In this paper we focus on
-unstructured textual data that has been, at least in part, manually
+unstructured text data that has been, at least in part, manually
 categorized.  Examples include but are not limited to webpages and curated
@@
 retrieval purposes, and more. The models and techniques that we develop in
-this paper are applicable in other domains as well, namely, any unstructured
+this paper are applicable in other data as well, namely, any unstructured
 representations of data that have been hierarchically classified (e.g.,~image
@@
 while leveraging the hierarchical structure of the labels. For the sake of
-simplicity, we focus on ISA hierarchies, but the model can be
+simplicity, we focus on is-a hierarchies, but the model can be
 applied to other hierarchies. We extend supervised latent Dirichlet

### 2011-06-02 paper/model.tex
Worked references and cleaned them up significantly.
misc

@@
 
-HSLDA is a model for hierarchically, multiply-labeled, bag-of-words data.
+HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.
 We will refer to individual groups of bag-of-word data as documents.
@@
 Each label except root labels $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
- We will for exposition purposes assume that this label set has hard ``is-a'' parent-child constraints (explained later), although this assumption can be relaxed at the cost of more complicated inference.
+ We will for exposition purposes assume that this label set has hard ``is-a''
+ parent-child constraints (explained later), although this assumption can be
+ relaxed at the cost of more computationally complex inference.
 Such a label hierarchy forms a multiply rooted tree.
@@
 The second half of step 4 is a substantial part of our contribution to the general class of supervised LDA models.
-Here each document is generatively labeled using a hierarchy of conditionally dependent probit regressors \cite{gelman?,probit_regression?}.
+Here each document is generatively labeled using a hierarchy of conditionally dependent probit regressors \cite{gelmanbda04}.
 For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.

### 2011-06-02 paper/model.tex
model, inference tiny edits

@@
 
-HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.
-We will refer to individual groups of bag-of-word data as documents.
-Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.
-Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document $d$.
-Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.
-Let the set of labels be $\mathcal{L}=\left\{ l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
-Each label except root labels $l \in \mathcal{L}$ has a parent $\mathrm{pa}(l) \in \mathcal{L}$ also in the set of labels.
+HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We
+will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d}
+\in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
+= \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document
+$d$.  Let there be $D$ such documents and let the size of the vocabulary be
+$V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
+  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label labels
+$l \in \mathcal{L}$, except root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
+also in the set of labels. 
  We will for exposition purposes assume that this label set has hard ``is-a''
@@
 
-The constraints imposed by an is-a label hierarchy are that if the $l$th label is applied to document $d$, i.e.~$y_{l,d} = 1$, then all labels in the label hierarchy up to the root are also applied to document $d$, i.e.~$y_{\mathrm{pa}(l),d} = 1, y_{\mathrm{pa}(\mathrm{pa}(l)),d} = 1, \ldots, y_{r,d}=1.$  Conversely, if a label $l'$ is marked as not applying to a document then no descendant of that label may be applied to the same.
+The constraints imposed by an is-a label hierarchy are that if the $l$th label is applied to document $d$, i.e.,~$y_{l,d} = 1$, then all labels in the label hierarchy up to the root are also applied to document $d$, i.e.,~$y_{\mathrm{pa}(l),d} = 1, y_{\mathrm{pa}(\mathrm{pa}(l)),d} = 1, \ldots, y_{r,d}=1.$  Conversely, if a label $l'$ is marked as not applying to a document then no descendant of that label may be applied to the same.
 We assume that at least one label is applied to every document.
@@
 
-In HSLDA, the bag-of-words document data is modeled using the LDA mixed-membership mixture model with global topic estimation.
-Label responses are generated using a conditional hierarchy of probit regressors.
-The HSLDA graphical model is given in Figure~\ref{fig:graphical_model}.
-In it and the following $K$ is the number of LDA "topics" (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$ is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific distribution over topics, $\boldsymbol\beta$ is a global distribution over topics, Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution, $\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution, $\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the $d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator function that takes the value $1$ if its argument is true and $0$ otherwise.
-The following procedure describes how to generate from the HSLDA generative model.
+%In HSLDA, the bag-of-word document data is modeled using the LDA
+%mixed-membership mixture model with global topic estimation.
+In HSLDA, documents are modeled using the LDA mixed-membership mixture model
+with global topic estimation. Label responses are generated using a conditional
+hierarchy of probit regressors. The HSLDA graphical model is given in
+Figure~\ref{fig:graphical_model}. In the model, $K$ is the number of LDA
+``topics'' (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$
+is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific
+distribution over topics, $\boldsymbol\beta$ is a global distribution over
+topics, Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution,
+$\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution,
+$\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the
+$d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator
+function that takes the value $1$ if its argument is true and $0$ otherwise.
+The following procedure describes how to generate from the HSLDA generative
+model.
 
@@
 
-The second half of step 4 is a substantial part of our contribution to the general class of supervised LDA models.
-Here each document is generatively labeled using a hierarchy of conditionally dependent probit regressors \cite{gelmanbda04}.
+The second half of step 4 is a substantial part of our contribution to the
+general class of supervised LDA models.
+Here, each document is labeled generatively using a hierarchy of conditionally dependent probit regressors \cite{gelmanbda04}.
 For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.
@@
 This is a common problem in modeling unbalanced data.
-To see how this model can be biased in this way we draw the readers attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.
+To see how this model can be biased in this way we draw the reader's attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.
 Because $\bar{\mathbf{z}}_d$ is always positive, setting $\mu$ to a negative value results in a bias towards negative labelings, i.e.~for large negative values of $\mu$, all labels become a priori more likely to be negative ($y_{l,d}=-1$).

### 2011-06-02 paper/experiments.tex
fixed fig
edits to experiments

@@
-We experimented with HSLDA for prediction in two domains: predicting medical
+We applied HSLDA to data from two domains: predicting medical
 diagnosis codes from hospital discharge summaries and predicting product
@@
 
-Discharge summaries are authored by clinicians to summarize the course of a
-hospitalized patient. The summaries typically contain a record of the patient
-complains, findings and diagnoses, along with treatment and hospital course.
-For each admission trained medical coders review the information in the
+Discharge summaries are authored by clinicians to summarize patient hospitalization course. 
+The summaries typically contain a record of patient
+complaints, findings and diagnoses, along with treatment and hospital course.
+For each hospitalization, trained medical coders review the information in the
 discharge summary and assign a series of diagnoses codes. Coding follows the
@@
 epidemiological, health management, and clinical
-purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm} As such, the ICD-9
-codes constitute a labeling of a patient's diagnoses based on a discharge
-summary. The ICD-9 codes are organized in a rooted-tree structure, with each
+purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm}  The ICD-9 codes are organized in a rooted-tree structure, with each
 edge representing an is-a relationship between parent and child, such that the
@@
 The task of automatic ICD-9 coding has been investigated in the clinical
-domain. Much of the work was triggered by the 2007 medical NLP community
-challenge~\cite{Challenge07}. The data in the challenge, however, differs from 
-ours in its scope. The datasets were smaller (1,000 training and 1,000 testing
-documents) and focused on radiology reports with a restricted number of ICD-9
-codes (45 of them, compared to 7K+ in our dataset). Methods ranged from manual
+domain.  Methods range from manual
 rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
@@
 
+%Much of the work was triggered by the 2007 medical NLP community
+%challenge~\cite{Challenge07}.
+%The data in the challenge, however, differs from 
+%ours in its scope. The datasets were smaller (1,000 training and 1,000 testing
+%documents) and focused on radiology reports with a restricted number of ICD-9
+%codes (45 of them, compared to 7K+ in our dataset).
+
 Our dataset was gathered from the clinical data warehouse of
-NewYork-Presbyterian Hospital. It consists of 6,000 discharge summaries and
+a large metropolitan hospital. It consists of 6,000 discharge summaries and
 their associated ICD-9 codes (7,298 distinct codes overall), representing all
@@
 The text of the discharge summaries was tokenized with
-NLTK.\footnote{http://www.nltk.org} Vocabulary was determined as the top 10,000
-tokens with highest document frequency (exclusive of names,
-places and other identifying numbers). Each discharge summary is thus
-represented as counts over the 10,000-word vocabulary. The study was approved
+NLTK.\footnote{http://www.nltk.org} A fixed vocabulary was formed by taking
+the top 10,000 tokens with highest document frequency (exclusive of names,
+places and other identifying numbers). The study was approved
 by the Institutional Review Board and follows HIPAA (Health
@@
 Amazon.com, an online retail store, organizes its catalog of 
-products in a hierarchy and provides product descriptions
-for most products in their catolog.  Products can be discovered by users
+products in a mulitply-rooted hierarchy and provides textual product descriptions
+for most products.  Products can be discovered by users
 through query-based searching or through product category exploration. Top-level
@@
 
-We were able to deduce the structure of the
-hierarchy for the Amazon.com products because all ancestors in the
-hierarchy were included with each category label. For example, {}``DVD / Genres
-/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
-for the DVD, {}``The Time Machine.''
+%We were able to deduce the structure of the
+%hierarchy for the Amazon.com products because all ancestors in the
+%hierarchy were included with each category label. For example, {}``DVD / Genres
+%/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
+%for the DVD, {}``The Time Machine.''
 
@@
 these two datasets. The comparison models included sLDA with independent 
-regressors (hierarchical constraints on labels ignored), HSLDA fit by 
-performing LDA followed by tree-conditional regressions, and HSLDA fit 
-with fixed random regression parameters. These models were chosen to 
-highlight several aspects of the model including performance in the absence 
-of hierarchical constraints, the effect of the combined inference procedure, and regression
+regressors (hierarchical constraints on labels ignored), HSLDA fit by first
+performing LDA then fitting tree-conditional regressions. These models were chosen to 
+highlight several aspects of HSLDA including performance in the absence 
+of hierarchical constraints, the effect of the combined inference, and regression
 performance attributable solely to the hierarchical constraints.
@@
 addition structure imposed on the label space, a distinction that we
-hypothesize will result in a difference in predictive performance. 
-Aside from this, all other features of sLDA are preserved between the two models.
+hypothesized would result in a difference in predictive performance.
 
@@
 
-The last comparison model is HSLDA with fixed and randomly selected regression
-parameters. There is a baseline benefit that the structure in the label space 
-provides for the prediction of labels. This comparison model is intended to
-quantify the contribution of the structure alone.
+%The last comparison model is HSLDA with fixed and randomly selected regression
+%parameters. There is a baseline benefit that the structure in the label space 
+%provides for the prediction of labels. This comparison model is intended to
+%quantify the contribution of the structure alone.
 
@@
 reserved for evaluation and predictive performance was evaluated against a standard derived
-from the observed labeling. Specifically, ancestors of observed nodes were ignored, observed nodes 
-were considered positive and unobserved nodes were considered to be negative. This method of 
+from the observed labeling. To make the comparison as fair as possible, 
+ancestors of observed nodes were ignored, observed nodes 
+were considered positive and descendents of observed nodes were considered to be negative. This method of 
 defining positive and negative labels was chosen to be as fair as possible to all models
@@
 relative to the three comparison models as a function of the prior mean
-on regression coefficient. Figure \ref{fig:1a} demo.
+on regression coefficients as a receiver operating characteristic (ROC) curve. 
+For low values of the auxiliary variable
+threshold, the models predict labels in a more sensitive and less specific
+manner, creating the points in the upper right corner of the ROC curve. As
+the auxiliary variable threshold is increased, the models predict in a less
+sensitive and more specific manner, creating the points in the lower left
+hand corner of the ROC curve. For all values of the prior mean in both datasets
+HSLDA outperforms sLDA with independent regressors. In the case of HSLDA with
+separately trained regression, HSLDA outperforms in the clinical dataset but
+performs equally well across the board with the retail product dataset.
 
@@
  \centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC Curve for clinical data}
+\label{fig:clinical_roc} 
+\end{figure}

### 2011-06-02 paper/related_work.tex
intro and rel work

@@
-In this work we extend supervised latent Dirichlet allocation (sLDA) \cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.
-sLDA is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per document ``supervision'';  often taking the form of a single numerical or categorical label.
-More generally this supervision is just extra per-document data;  for instance its quality or relevance (e.g.~online review scores), marks given to written work (e.g.~essay grades), or the number of times a web page is linked.
-These labels are usually generatively modeled as a conditional draw from some distribution that depends on each document-specific topic mixture.
-It has been demonstrated that the signal provided by such supervision can result in better, task-specific document models and can also lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
+In this work we extend supervised latent Dirichlet allocation (sLDA)
+\cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision. sLDA
+is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per
+document ``supervision,'' often taking the form of a single numerical or
+categorical label. It has been demonstrated that the signal provided by such
+supervision can result in better, task-specific document models and can also
+lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
 It also has been demonstrated that sLDA has been shown to outperform both LASSO
-(L1 regularized least squares regression) and LDA followed by least
-squares regression~\cite{BleiMcAuliffe2008}.
-sLDA can be applied to data of the type we consider in this paper; however, doing so requires ignoring the hierarchical dependencies amongst the labels.
-In Section~\ref{sec:experiments} we constrast HSLDA with sLDA applied in this way.
+(L1 regularized least squares regression) and LDA followed by least squares
+regression~\cite{BleiMcAuliffe2008}. sLDA can be applied to data of the type
+we consider in this paper; however, doing so requires ignoring the hierarchical
+dependencies amongst the labels. In Section~\ref{sec:experiments} we constrast
+HSLDA with sLDA applied in this way.
 
@@
 LabeledLDA\citep{Ramage2009} and DiscLDA\citep{DiscLDA}.  Various applications of these models to 
-computer vision and document networks have been explored \citep{wangbleifeifei08,RelationalLDA} . 
+computer vision and document networks have been explored
+\citep{wangbleifeifei08,RelationalLDA} .
 None of these models, however, leverage dependency structure in the label space.
@@
 small datasets, small label spaces, and has focused on single label classification without
-a model of documents such as LDA\citep{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.
+a model of documents such as LDA~\citep{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.
 

### 2011-06-02 paper/experiments.tex
Experiments updated

@@
 for most products.  Products can be discovered by users
-through query-based searching or through product category exploration. Top-level
+through free-text search and product category exploration. Top-level
 product categories are displayed on the front page of the website and lower
@@
 We evaluated our model, HSLDA, against the comparison models with a focus
-on predictive performance. Predictive performance was measured with
+on predictive performance on held-out data. Prediction performance was measured with
 standard metrics - sensitiviy (true positive rate) and 
-1-specificity (false positive rate). We evaluate on the two aforementioned
-datasets to demonstrate that our model generalizes to two different
-domains.
+1-specificity (false positive rate).
 
@@
 being compared.  In particular, since the sLDA model does not enforce the hierarchical
-constraints, the models can be compared on a more equal footing by considering only on the observed labels
+constraints, its performance was found to be so poor on joint prediction that it was not included.
+The models can be compared on a more equal footing by considering only on the observed labels
 as being positive despite the fact that ancestors must also be positive. The gold standard 
-defined in this way will likely lead to a slight overestimation of the number of
-false positives. It is known that ICD-9 codes lack sensitivity and their use as a
-gold standard could lead to correctly positive predictions being labeled as false positives.
+defined in this way will likely inflate the number of
+false positives because the labels applied to any particular document are usually not as complete as they could be. 
+ICD-9 codes lack sensitivity and their use as a
+gold standard could lead to correctly positive predictions being labeled as false positives\citep{Birmetal2005}.
 However, given that the label space is often large (as in our examples) it is a moderate assumption that
@@
 Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
-where $\hat{d}$ represents the test document. For efficiency,
+for each test document, $\hat{d}$. For efficiency,
 the expectation of this probability distribution was estimated in the following way. Expectations
@@
 Using these expectations, we performed Gibbs sampling over the hierarchy to acquire predictive
-samples for the documents in the test set.
+samples for the documents in the test set. The true positive rate was calculated as the
+ average expected labeling for gold standard positive labels. The false positive rate was calculated
+as the average expected labeling for gold standard negative labels.
 
-The following are the predictive performance results for the clinical
-data given a prior mean for the regression parameters of -1.6. 
+The following are the performance results for the clinical
+data given a prior mean for the regression parameters of $\mu=-1.6$. 
 The full HSLDA model had a true positive rate of 0.57 
@@
 The following are the predictive performance results for the retail product
-data given a prior mean for the regression parameters of -2.2. 
+data given a prior mean for the regression parameters of $\mu=-2.2$. 
 The full HSLDA model had a true positive rate of 0.85 
@@
 
-To further explore this tradeoff between the true positive rate and the
-false positive rate we evaluated predictive performance for a range of values
-for two different parameters - the prior mean for the regression coefficients
+As sensitivity and specificity can always be traded off we examined sensitivity for a range of values
+for two different parameters - the prior means for the regression coefficients
 and the threshold for the auxiliary variables.  The goal in this analysis
@@
 For each model type, separate models were fit for each value of the 
-prior mean of the regression coefficients.  In contrast, to evaluate
+prior mean of the regression coefficients.  This is a proper Bayesian
+sensitivity analysis. In contrast, to evaluate
 predictive performance as a function of the auxiliary variable threshold,

### 2011-06-02 paper/discussion.tex
Experiments updated
discussion tiny tiny tiny edits

@@
-%Despite the proliferation of LDA related models in recent years, one could argue that the true killer app for LDA has yet to be discovered.
-LDA makes sense for information retrieval (IR) because  the learned latent topic vectors serve as good low-dimensional document representations, and, a posteriori similarity between such vectors can be used to compute the similarity of documents.
-Unfortunately, computational considerations render this approach to IR impractical in most settings.
+%Despite the proliferation of LDA related models in recent years, one could
+%argue that the true killer app for LDA has yet to be discovered.  LDA makes
+%sense for information retrieval (IR) because  the learned latent topic vectors
+%serve as good low-dimensional document representations, and, a posteriori
+%similarity between such vectors can be used to compute the similarity of
+%documents.  Unfortunately, computational considerations render this approach
+%to IR impractical in most settings.
 
-The SLDA model family, of which HSLDA is a member, can be understood in two different ways.
-One way is to see it as a family of topic models that improve on the topic modeling performance of LDA via the inclusion of observed supervision.
-Another way to see the family is as a set of models that can predict labels for bag-of-words data.
-A large diversity of problems can be expressed as label prediction problems for bag-of-words data.
-A surprisingly large amount of that kind of data possess structured labels, either hierarchically constrained or otherwise.
-That HSLDA directly addresses this kind of data is a large part of the motivation for this work.
-That it outperforms more straightforward approaches should be of interest to practitioners.
+The SLDA model family, of which HSLDA is a member, can be understood in two
+different ways. One way is to see it as a family of topic models that improve
+on the topic modeling performance of LDA via the inclusion of observed
+supervision. An alternative, complementary way is to see it as a set of models
+that can predict labels for bag-of-word data. A large diversity of problems
+can be expressed as label prediction problems for bag-of-word data. A
+surprisingly large amount of that kind of data possess structured labels,
+either hierarchically constrained or otherwise. That HSLDA directly addresses
+this kind of data is a large part of the motivation for this work. That it
+outperforms more straightforward approaches should be of interest to
+practitioners.
 
-Variational Bayes has been the predominant estimation approach applied to sLDA models.
-Hierarchical probit regression makes for tractable Markov chain Monte Carlo SLDA inference, a benefit that should extend to other sLDA models should probit regression be used for response variable prediction there too.
+Variational Bayes has been the predominant estimation approach applied to sLDA
+models. Hierarchical probit regression makes for tractable Markov chain Monte
+Carlo SLDA inference, a benefit that should extend to other sLDA models should
+probit regression be used for response variable prediction there too.
 
-The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases it is better to do full joint estimation of HSLDA.
-In alternative interpretation of the same results is that if one is more sensitive to the performance gains that result from exploiting the structure of the labels then one can, in an engineering sense, get nearly as much gain in label prediction performance by first fitting LDA and then fitting a hierarchical probit regression.
-There are applied settings in which this could be advantageous.
+The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
+it is better to do full joint estimation of HSLDA.  An alternative
+interpretation of the same results is that, if one is more sensitive to the
+performance gains that result from exploiting the structure of the labels, then
+one can, in an engineering sense, get nearly as much gain in label prediction
+performance by first fitting LDA and then fitting a hierarchical probit
+regression.  There are applied settings in which this could be advantageous.
 
-Extensions to this work include unbounded topic cardinality variants and relaxations to different kinds of label structure.
-Unbounded topic cardinality variants pose interesting inference challenges.
-Utilizing different kinds of label structure is possible within this framework, but requires relaxing some of the simplifications we made in this paper for expositional purposes.
+Extensions to this work include unbounded topic cardinality variants and
+relaxations to different kinds of label structure.  Unbounded topic cardinality
+variants pose interesting inference challenges.  Utilizing different kinds of
+label structure is possible within this framework, but requires relaxing some
+of the simplifications we made in this paper for expositional purposes.
 
-%We have described a mixed membership model with hierarchical supervision.
-We have demonstrated this model in the context of document modeling with hierarchical multi-label supervision.
-Such a model is appropriate in domains where there are hierarchical constraints among the labels such as is the case in an IS-A hierarchy.
-
-...
+%We have described a mixed membership model with hierarchical supervision. We
+%have demonstrated this model in the context of document modeling with
+%hierarchical multi-label supervision. Such a model is appropriate in domains
+%where there are hierarchical constraints among the labels such as is the case
+%in an IS-A hierarchy.
 

### 2011-06-02 paper/experiments.tex
figure placement and captions

@@
 well as simple keyword mappings, all with promising
-results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Ruch2008,Lita2008}.
+results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Lita2008}.%Ruch2008 omitted
 
@@
 \subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\caption{Out-of-sample ICD-9 code prediction from patient free-text discharge records 
-(\subref{fig:1b}). Out-of-sample Amazon product category predictions from 
-product free-text descriptions (\subref{fig:1b}). In both figures, solid is 
+\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
+\caption{ROC curves for out-of-sample ICD-9 code prediction from patient free-text discharge records 
+(\subref{fig:1a},\subref{fig:clinical_roc}). ROC curve for out-of-sample Amazon product category predictions from 
+product free-text descriptions \subref{fig:1b}. Figures \subref{fig:1a} and \subref{fig:1b} are a function 
+of the prior means of the regression parameters. Figure \subref{fig:clinical_roc} is a function of auxiliary variable threshold. In all figures, solid is 
 HSLDA, dashed are independent regressors + sLDA (hierarchical 
@@
 
-\begin{figure}[t]
-%tbp] %  figure placement: here, top, bottom, or page
- \centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC Curve for clinical data}
-\label{fig:clinical_roc} 
-\end{figure}
+%\begin{figure}[t]
+%%tbp] %  figure placement: here, top, bottom, or page
+% \centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC curve for Out-of-sample ICD-9 code predictions.}
+%\label{fig:clinical_roc} 
+%\end{figure}

### 2011-06-02 paper/introduction.tex
intro and rel work

@@
 and patient records %hospital treatment transcripts
-and diagnosis codes assigned to them for bookkeeping and insurance purposes
-(e.g.~hospital discharge summaries with 
-International Classification of Disease 9th Revision, Clinical Modification
-(ICD-9-CM) codes assigned \cite{ICD9}).  In this work we show how to combine
+and diagnosis codes assigned to them for bookkeeping and insurance purposes.
+In this work we show how to combine 
 these two sources of information using a single model that allows one to
@@
 retrieval purposes, and more. The models and techniques that we develop in
-this paper are applicable in other data as well, namely, any unstructured
+this paper are applicable to other data as well, namely, any unstructured
 representations of data that have been hierarchically classified (e.g.,~image
@@
 cannot always be interpreted as a negative labeling.
-
-%The task of multi-label classification, selecting the k-best labels for a given
-%instance, has been a topic of research for several years. One simplistic way to
-%carry out the classification is through a series of independent binary
-%classifiers, but this ignores the many inherent dependencies among the
-%labels. Thus, much work has been devoted on incorporating the co-occurence
-%patterns of the labels into the classification task.
 

### 2011-06-02 paper/experiments.tex
intermediary commit

--- 2011-06-02 paper/experiments.tex
@@
 
-Discharge summaries are authored by clinicians to summarize patient hospitalization course. 
-The summaries typically contain a record of patient
+Discharge summaries are authored by clinicians to summarize patient
+hospitalization course. The summaries typically contain a record of patient
 complaints, findings and diagnoses, along with treatment and hospital course.
@@
 
-Amazon.com, an online retail store, organizes its catalog of 
-products in a mulitply-rooted hierarchy and provides textual product descriptions
+Amazon.com, an online retail store, organizes its catalog of products in a
+mulitply-rooted hierarchy and provides textual product descriptions 
 for most products. Products can be discovered by users
@@
 
-We evaluated HSLDA along with three other closely related models against 
+We evaluated HSLDA along with three closely related models against 
 these two datasets. The comparison models included sLDA with independent 
 regressors (hierarchical constraints on labels ignored), HSLDA fit by first
-performing LDA then fitting tree-conditional regressions. These models were chosen to 
-highlight several aspects of HSLDA including performance in the absence 
-of hierarchical constraints, the effect of the combined inference, and regression
-performance attributable solely to the hierarchical constraints.
+performing LDA then fitting tree-conditional regressions. These models were
+chosen to  highlight several aspects of HSLDA including performance in the
+absence  of hierarchical constraints, the effect of the combined inference, and
+regression performance attributable solely to the hierarchical constraints.
 
@@
 
-There are two components to HSLDA, LDA and a hierarchically constrained response.
-The second comparison model is HSLDA fit by performing LDA first followed by
-performing inference over the hierarchically constrained label space. In this
-comparison model, the separate inference processes do not allow the responses
-to influence the low dimensional structure inferred by LDA. Combined inference
-has been shown to improve performance in sLDA \citep{BleiMcAuliffe2008}. This
-comparison model examines not the structuring of the label space, but the benefit
-of combined inference over both the documents and the label space.
+There are two components to HSLDA, LDA and a hierarchically constrained
+response. The second comparison model is HSLDA fit by performing LDA first
+followed by performing inference over the hierarchically constrained label
+space. In this comparison model, the separate inference processes do not allow
+the responses to influence the low dimensional structure inferred by LDA.
+Combined inference has been shown to improve performance in sLDA
+\citep{BleiMcAuliffe2008}. This comparison model examines not the structuring
+of the label space, but the benefit of combined inference over both the
+documents and the label space.
 

### 2011-06-02 paper/experiments.tex
edits

@@
 
-We evaluated HSLDA along with three closely related models against 
-these two datasets. The comparison models included sLDA with independent 
+We evaluated HSLDA along with two closely related models against
+the two datasets. The comparison models included sLDA with independent
 regressors (hierarchical constraints on labels ignored), HSLDA fit by first
@@
 
-For all of these models, particular attention was payed to the settings of the
+For all three models, particular attention was given to the settings of the 
 prior parameters for the regression coefficients. These parameters implement an
-important form of regularization in HSLDA. In the setting where there are no negative
-labels, a Gaussian prior over the regression parameters with a negative mean implements
-a prior belief that missing labels are likely to be negative. We evaluated model performance 
-for all three models with a range of values for $\mu$, the mean prior parameter for regression coefficients
+important form of regularization in HSLDA. In the setting where there are no
+negative labels, a Gaussian prior over the regression parameters with a
+negative mean implements a prior belief that missing labels are likely to be
+negative. Thus, we evaluated model performance for all three models with a
+range of values for $\mu$, the mean prior parameter for regression coefficients 
 ($\mu\in\left\{ -3,-2.8,-2.6,\ldots,1\right\}$).
 
-The number of topics for all
-models was set to 50, the prior distributions of
-$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and $p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1
-and a scale parameters of 1000.
+The number of topics for all models was set to 50, the prior distributions of
+$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
+$p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1 and a
+scale parameters of 1000.
+
+\begin{figure}[h]
+\begin{center}
+%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\subfigure[][]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
+%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
+\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
+\caption{ROC curves for out-of-sample ICD-9 code prediction from patient free-text discharge records 
+(\subref{fig:1a},\subref{fig:clinical_roc}). ROC curve for out-of-sample Amazon product category predictions from 
+product free-text descriptions \subref{fig:1b}. Figures \subref{fig:1a} and \subref{fig:1b} are a function 
+of the prior means of the regression parameters. Figure \subref{fig:clinical_roc} is a function of auxiliary variable threshold. In all figures, solid is 
+HSLDA, dashed are independent regressors + sLDA (hierarchical 
+constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
+tree-conditional regressions.}
+\label{fig:main_results}
+\end{center}
+\end{figure}
 
@@
 
-We evaluated our model, HSLDA, against the comparison models with a focus
-on predictive performance on held-out data. Prediction performance was measured with
-standard metrics - sensitiviy (true positive rate) and 
-1-specificity (false positive rate).
+We evaluated our model, HSLDA, against the comparison models with a focus on
+predictive performance on held-out data. Prediction performance was measured
+with standard metrics -- sensitiviy (true positive rate) and 1-specificity
+(false positive rate).
 
-To evaluate the performance of these models, we established a gold standard for comparison.
-For each dataset, a held out set of 1000 documents and labels were
-reserved for evaluation and predictive performance was evaluated against a standard derived
-from the observed labeling. To make the comparison as fair as possible, 
-ancestors of observed nodes were ignored, observed nodes 
-were considered positive and descendents of observed nodes were considered to be negative. This method of 
-defining positive and negative labels was chosen to be as fair as possible to all models
-being compared.  In particular, since the sLDA model does not enforce the hierarchical
-constraints, its performance was found to be so poor on joint prediction that it was not included.
+The gold standard for comparison was derived from the testing set in each
+dataset. To make the comparison as fair as possible among models, ancestors of
+observed 
+nodes in the label hierarchy were ignored, observed nodes were considered
+positive and descendents of observed nodes were considered to be negative. In
+particular, since the sLDA model does not enforce the hierarchical constraints,
+its performance was found to be so poor on joint prediction that it was not
+included. 
 The models can be compared on a more equal footing by considering only on the observed labels
@@
 
-\begin{figure}%[htbp]
-\begin{center}
-%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
-%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
-\caption{ROC curves for out-of-sample ICD-9 code prediction from patient free-text discharge records 
-(\subref{fig:1a},\subref{fig:clinical_roc}). ROC curve for out-of-sample Amazon product category predictions from 
-product free-text descriptions \subref{fig:1b}. Figures \subref{fig:1a} and \subref{fig:1b} are a function 
-of the prior means of the regression parameters. Figure \subref{fig:clinical_roc} is a function of auxiliary variable threshold. In all figures, solid is 
-HSLDA, dashed are independent regressors + sLDA (hierarchical 
-constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
-tree-conditional regressions.}
-\label{fig:main_results}
-\end{center}
-\end{figure}
-
 %\begin{figure}[t]

### 2011-06-02 paper/experiments.tex
more edits!

@@
 dataset. To make the comparison as fair as possible among models, ancestors of
-observed 
-nodes in the label hierarchy were ignored, observed nodes were considered
-positive and descendents of observed nodes were considered to be negative. In
-particular, since the sLDA model does not enforce the hierarchical constraints,
-its performance was found to be so poor on joint prediction that it was not
-included. 
-The models can be compared on a more equal footing by considering only on the observed labels
-as being positive despite the fact that ancestors must also be positive. The gold standard 
-defined in this way will likely inflate the number of
-false positives because the labels applied to any particular document are usually not as complete as they could be. 
-ICD-9 codes lack sensitivity and their use as a
-gold standard could lead to correctly positive predictions being labeled as false positives\citep{Birmetal2005}.
-However, given that the label space is often large (as in our examples) it is a moderate assumption that
-erroneous false positives should not skew results significantly.
+observed nodes in the label hierarchy were ignored, observed nodes were
+considered positive and descendents of observed nodes were considered to be
+negative. In particular, since the sLDA model does not enforce the hierarchical
+constraints, its performance was found to be so poor on joint prediction that
+it was not included. The models can be compared on a more equal footing by
+considering only on the observed labels as being positive, despite the fact
+that, following the hierarchical constraints, ancestors must also be positive.
+Such a gold standard will likely inflate the number of false positives because
+the labels applied to any particular document are usually not as complete as
+they could be.  ICD-9 codes, for instance, lack sensitivity and their use as a
+gold standard could lead to correctly positive predictions being labeled as
+false positives\citep{Birmetal2005}. 
+However, given that the label space is often large (as in our examples) it is a
+moderate assumption that erroneous false positives should not skew results
+significantly.
 
 Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
-for each test document, $\hat{d}$. For efficiency,
-the expectation of this probability distribution was estimated in the following way. Expectations
-of $\mathbf{\bar{z}}_d$ and $\boldsymbol{\eta}_l$ were estimated with samples from the posterior.
-Using these expectations, we performed Gibbs sampling over the hierarchy to acquire predictive
-samples for the documents in the test set. The true positive rate was calculated as the
- average expected labeling for gold standard positive labels. The false positive rate was calculated
+for each test document, $\hat{d}$. For efficiency, the expectation of this
+probability distribution was estimated in the following way. Expectations 
+of $\mathbf{\bar{z}}_d$ and $\boldsymbol{\eta}_l$ were estimated with samples
+from the posterior. Using these expectations, we performed Gibbs sampling over
+the hierarchy to acquire predictive samples for the documents in the test set.
+The true positive rate was calculated as the average expected labeling for
+gold standard positive labels. The false positive rate was calculated 
 as the average expected labeling for gold standard negative labels.
 
-The following are the performance results for the clinical
-data given a prior mean for the regression parameters of $\mu=-1.6$. 
-The full HSLDA model had a true positive rate of 0.57 
+The following are the performance results for the clinical data given a prior
+mean for the regression parameters of $\mu=-1.6$.  The full HSLDA model had a
+true positive rate of 0.57  
 and a false positive rate of 0.13, the sLDA model had a true positive

### 2011-06-02 paper/experiments.tex
done

@@
 
-The following are the performance results for the clinical data given a prior
-mean for the regression parameters of $\mu=-1.6$.  The full HSLDA model had a
-true positive rate of 0.57  
-and a false positive rate of 0.13, the sLDA model had a true positive
-rate of 0.42 and a false positive rate of 0.07, and the HSLDA model where
-LDA and the regressions were fit separately had a true positive rate of 0.39
-and a false positive rate of 0.08.
+Figure~\ref{fig:1a} shows the results for the clinical data. For instance, 
+a prior mean for the regression of $\mu=-1.6$ yields the following performance:
+the full HSLDA model had a 
+true positive rate of 0.57 and a false positive rate of 0.13, the sLDA model
+had a true positive rate of 0.42 and a false positive rate of 0.07, and the
+HSLDA model where LDA and the regressions were fit separately had a true
+positive rate of 0.39 and a false positive rate of 0.08.
 
@@
 
-The following are the predictive performance results for the retail product
-data given a prior mean for the regression parameters of $\mu=-2.2$. 
-The full HSLDA model had a true positive rate of 0.85 
+Figure~\ref{fig:1b} shows the results for the retail product data. For
+instance, a prior mean for the regression of $\mu=-2.2$ yields the following
+performance: the full HSLDA model had a true positive rate of 0.85 
 and a false positive rate of 0.30, the sLDA model had a true positive
@@
 
-As sensitivity and specificity can always be traded off we examined sensitivity for a range of values
-for two different parameters - the prior means for the regression coefficients
-and the threshold for the auxiliary variables.  The goal in this analysis
-was to evaluate the performance of these models subject to more or less
-stringent requirements for predicting positive labels. These two parameters 
-have important related functions in the model. The prior mean in combination 
-with the auxiliary variable threshold together encode the strength of the prior
-belief that unobserved labels are likely to be negative. Effectively, the
-prior mean applies negative pressure to the predictions and the auxiliary
-variable threshold determines the cutoff.
-
+As sensitivity and specificity can always be traded off, we examined
+sensitivity for a range of values for two different parameters -- the prior
+means for the regression coefficients and the threshold for the auxiliary
+variables.  The goal in this analysis was to evaluate the performance of these
+models subject to more or less stringent requirements for predicting positive
+labels. These two parameters have important related functions in the model. The
+prior mean in combination with the auxiliary variable threshold together encode
+the strength of the prior belief that unobserved labels are likely to be
+negative. Effectively, the prior mean applies negative pressure to the
+predictions and the auxiliary variable threshold determines the cutoff. 
 For each model type, separate models were fit for each value of the 
@@
 is varied prior to inference and the auxiliary variable threshold is varied
-following inference.  However, as intended, they both highlight model performance
-under more or less stringent requirements for predicting positive labels.
+following inference.
+%However, as intended, they both highlight model performance 
+%under more or less stringent requirements for predicting positive labels.
 
@@
 on regression coefficients as a receiver operating characteristic (ROC) curve. 
-For low values of the auxiliary variable
-threshold, the models predict labels in a more sensitive and less specific
-manner, creating the points in the upper right corner of the ROC curve. As
-the auxiliary variable threshold is increased, the models predict in a less
-sensitive and more specific manner, creating the points in the lower left
-hand corner of the ROC curve. For all values of the prior mean in both datasets
-HSLDA outperforms sLDA with independent regressors. In the case of HSLDA with
-separately trained regression, HSLDA outperforms in the clinical dataset but
-performs equally well across the board with the retail product dataset.
+For low values of the auxiliary variable threshold, the models predict labels
+in a more sensitive and less specific manner, creating the points in the upper
+right corner of the ROC curve. As the auxiliary variable threshold is
+increased, the models predict in a less sensitive and more specific manner,
+creating the points in the lower left hand corner of the ROC curve. For all
+values of the prior mean in both datasets HSLDA outperforms sLDA with
+independent regressors. In the case of HSLDA with separately trained
+regression, HSLDA outperforms in the clinical dataset but performs equally well
+across the board with the retail product dataset.
 

### 2011-06-03 paper/experiments.tex
near submission fixes

@@
 thresholds. These methods are significantly different since the prior mean
-is varied prior to inference and the auxiliary variable threshold is varied
+is varied prior to inference, and the auxiliary variable threshold is varied
 following inference.
@@
 Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
-relative to the three comparison models as a function of the prior mean
+relative to the two comparison models as a function of the prior mean
 on regression coefficients as a receiver operating characteristic (ROC) curve. 
@@
 creating the points in the lower left hand corner of the ROC curve. For all
-values of the prior mean in both datasets HSLDA outperforms sLDA with
+values of the prior mean in both datasets, HSLDA outperforms sLDA with
 independent regressors. In the case of HSLDA with separately trained

### 2011-09-08 NYAS_abstract/abstract.tex
Added NYAS abstract

We introduce hierarchically supervised latent Dirichlet allocation (HSLDA), a
model for hierarchically and multiply labeled bag-of-word data.  Examples of
such data include web pages and their placement in directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data are also of interest.
%% We demonstrate HSLDA on large-scale data from clinical document labeling and
%% retail product categorization tasks. We show that leveraging the structure from
%% hierarchical labels improves out-of-sample label prediction substantially when
%% compared to models that do not.

Our work operates within the framework of topic modeling. Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. For the sake of
simplicity, we focus on is-a hierarchies, but the model can be
applied to other structured label spaces. We extend supervised latent Dirichlet
allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
supervision. We propose an efficient way to incorporate hierarchical
information into the model. We hypothesize that the context of labels within
the hierarchy provides valuable information about labeling. Other models that incorporate LDA and supervision include
LabeledLDA\citep{Ramage2009} and DiscLDA\citep{DiscLDA}. None of these models, however, leverage dependency structure in the label space.

We demonstrate our model on large, real-world datasets in the clinical and web
retail domains. We observe that hierarchical information is valuable when
incorporated into the learning and improves our primary goal of multi-label
classification. Our results show that a joint, hierarchical model outperforms a
classification with unstructured labels as well as a disjoint model, where the
topic model and the hierarchical classification are inferred
independently of each other.

HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We
will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d}
\in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
= \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document
$d$.  Let there be $D$ such documents and let the size of the vocabulary be
$V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label labels
$l \in \mathcal{L}$, except root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
also in the set of labels. 
 We will for exposition purposes assume that this label set has hard ``is-a''
 parent-child constraints (explained later), although this assumption can be
 relaxed at the cost of more computationally complex inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  In the applications we consider, only positive label applications are observed.

\begin{figure}[t]
%tbp] %  figure placement: here, top, bottom, or page
 \centering \includegraphics[scale=0.25]{Graphical_Model-final} \caption{HSLDA graphical model}

\label{fig:graphical_model} 
\end{figure}

%In HSLDA, the bag-of-word document data is modeled using the LDA
%mixed-membership mixture model with global topic estimation.
In HSLDA, documents are modeled using the LDA mixed-membership mixture model
with global topic estimation. Label responses are generated using a conditional
hierarchy of probit regressors\cite{gelmanbda04}. The HSLDA graphical model is given in
Figure~\ref{fig:graphical_model}. In the model, $K$ is the number of LDA
``topics'' (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$
is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific
distribution over topics, $\boldsymbol\beta$ is a global distribution over
topics, Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution,
$\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution,
$\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the
$d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator
function that takes the value $1$ if its argument is true and $0$ otherwise.

Posterior inference in HSLDA was performed using Gibbs sampling and Markov chain Monte Carlo.  Note that, like in collapsed Gibbs samplers for LDA \cite{Griffiths04}, we have analytically marginalized out the parameters $\boldsymbol{\phi}_{1:K}$
and $\boldsymbol{\theta}_{1:D}$. HSLDA also employs a hierarchical Dirichlet prior over topic assignments (i.e.,~$\boldsymbol\beta$ is estimated from data rather than assumed to be symmetric).  This has been shown to improve the quality and stability of inferred topics \citep{WallachMiMc2009}. The hyperparameters $\alpha$, $\alpha^{\prime}$, and $\gamma$ are
%given broad ${\rm Gamma}(1,1000)$ prior distributions and 
sampled
using Metropolis-Hastings.

We applied HSLDA to data from two domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions. The clinical dataset consists
of 6,000 clinical notes along with associated billing codes that are used
to document conditions that a particular patient was treated for. These billing
codes (7298 distinct codes in our dataset) are organized in an is-a hierarchy. The retail dataset
consists of product descriptions for DVDs from the Amazon.com product catalog. This data
was partially obtained from the Stanford Network Analysis Platform (SNAP) dataset~\citep{SNAP}. The comparison models included sLDA with independent
regressors (hierarchical constraints on labels ignored) and HSLDA fit by first
performing LDA then fitting tree-conditional regressions. The number of topics for all models was set to 50, the prior distributions of
$p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
$p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1 and a
scale parameters of 1000.

\begin{figure}[h]
\begin{center}
%\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][]{\label{fig:1a}\includegraphics[width=.3\textwidth]{figs/clin_pred_varying_mu}}
%\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
\subfigure[][]{\label{fig:1b}\includegraphics[width=.3\textwidth]{figs/amazon_pred_varying_mu}}
\subfigure[][]{\label{fig:clinical_roc}\includegraphics[width=0.3\textwidth]{figs/ROC_comparison_leafs}}
\caption{ROC curves for out-of-sample ICD-9 code prediction from patient free-text discharge records 
(\subref{fig:1a},\subref{fig:clinical_roc}). ROC curve for out-of-sample Amazon product category predictions from 
product free-text descriptions \subref{fig:1b}. Figures \subref{fig:1a} and \subref{fig:1b} are a function 
of the prior means of the regression parameters. Figure \subref{fig:clinical_roc} is a function of auxiliary variable threshold. In all figures, solid is 
HSLDA, dashed are independent regressors + sLDA (hierarchical 
constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
tree-conditional regressions.}
\label{fig:main_results}
\end{center}
\end{figure}

The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
it is better to do full joint estimation of HSLDA.  An alternative
interpretation of the same results is that, if one is more sensitive to the
performance gains that result from exploiting the structure of the labels, then
one can, in an engineering sense, get nearly as much gain in label prediction
performance by first fitting LDA and then fitting a hierarchical probit
regression.  There are applied settings in which this could be advantageous.

### 2011-09-08 NYAS_abstract/introduction.tex
Added NYAS abstract

--- 2011-06-02 paper/introduction.tex
@@
 There exist many sources of unstructured data that have been partially or
-completely categorized by human editors.  In this paper we focus on
+completely categorized by human editors.  In this paper, we focus on
 unstructured text data that has been, at least in part, manually
@@
 There are several challenges entailed in incorporating a hierarchy of labels
-into the model. Given a large set of potential labels (often thousands), each
-instance has only a small number of labels associated to it. There are no
+into the model. Among them, given a large set of potential labels (often thousands), each
+instance has only a small number of labels associated to it. Furthermore, there are no
 naturally occurring negative labeling in the data, and the absence of a label
@@
 simplicity, we focus on is-a hierarchies, but the model can be
-applied to other hierarchies. We extend supervised latent Dirichlet
+applied to other structured label spaces. We extend supervised latent Dirichlet
 allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
@@
 classification with unstructured labels as well as a disjoint model, where the
-topic modeling and the hierarchical classification are carried out
+topic model and the hierarchical classification are inferred
 independently of each other.

### 2011-09-09 NYAS_abstract/abstract.tex
Updated NYAS abstract

@@
 simplicity, we focus on is-a hierarchies, but the model can be
-applied to other structured label spaces. We extend supervised latent Dirichlet
+applied to other structured label spaces. Our work extends supervised latent Dirichlet
 allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
-supervision. We propose an efficient way to incorporate hierarchical
-information into the model. We hypothesize that the context of labels within
-the hierarchy provides valuable information about labeling. Other models that incorporate LDA and supervision include
-LabeledLDA\citep{Ramage2009} and DiscLDA\citep{DiscLDA}. None of these models, however, leverage dependency structure in the label space.
+supervision and proposes an efficient way to incorporate such information into the model. 
+We hypothesize that the context of labels within
+the hierarchy provides valuable information about labeling.
+Other models, such as LabeledLDA\citep{Ramage2009}, incorporate LDA and supervision; However, none of these models leverage dependency structure in the label space.
 
@@
 
-\begin{figure}[t]
-%tbp] %  figure placement: here, top, bottom, or page
- \centering \includegraphics[scale=0.25]{Graphical_Model-final} \caption{HSLDA graphical model}
-
-\label{fig:graphical_model} 
-\end{figure}
-
 %In HSLDA, the bag-of-word document data is modeled using the LDA
@@
 function that takes the value $1$ if its argument is true and $0$ otherwise.
+
+\begin{figure}[t]
+%tbp] %  figure placement: here, top, bottom, or page
+ \centering \includegraphics[scale=0.25]{Graphical_Model-final} \caption{HSLDA graphical model}
+
+\label{fig:graphical_model} 
+\end{figure}
 

### 2011-09-09 NYAS_abstract/abstract.tex
Final NYAS abstract

@@
 the hierarchy provides valuable information about labeling.
-Other models, such as LabeledLDA\citep{Ramage2009}, incorporate LDA and supervision; However, none of these models leverage dependency structure in the label space.
+Other models, such as LabeledLDA~\citep{Ramage2009}, incorporate LDA and supervision; however, none of these models leverage dependency structure in the label space.
 
@@
 $V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
-  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label labels
+  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label
 $l \in \mathcal{L}$, except root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
@@
  We will for exposition purposes assume that this label set has hard ``is-a''
- parent-child constraints (explained later), although this assumption can be
+ parent-child constraints, although this assumption can be
  relaxed at the cost of more computationally complex inference.
@@
 is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific
-distribution over topics, $\boldsymbol\beta$ is a global distribution over
-topics, Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution,
-$\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution,
-$\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the
-$d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator
-function that takes the value $1$ if its argument is true and $0$ otherwise.
+distribution over topics, and $\boldsymbol\beta$ is a global distribution over
+topics.
+%% , Dir$_{K}(\cdot)$ is a $K$-dimensional Dirichlet distribution,
+%% $\mathcal{N}_{K}(\cdot)$ is the $K$-dimensional Normal distribution,
+%% $\mathbf{I}_{K}$ is the $K$ dimensional identity matrix,  $\mathbf{1}_d$ is the
+%% $d$-dimensional vector of all ones, and $\mathbb{I}(\cdot)$ is an indicator
+%% function that takes the value $1$ if its argument is true and $0$ otherwise.
 

### 2011-10-27 NYAS_NIPS_Poster/HSLDA_poster.tex
Added poster and updated paper for camera-ready

\documentclass[portrait]{sciposter}

\usepackage{a0size}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage[english]{babel}
\usepackage{setspace}
\usepackage[pdftex,final]{graphicx}  % graphics (PDFTeX)
\usepackage{epsfig}
\usepackage{dsfont}
\usepackage{bm}
\usepackage{pstricks,pst-grad}
\usepackage{color}
\usepackage[square,numbers]{natbib}

\usepackage{amscd}
%\usepackage{subfigure}

%\def\ManuPath{./}

%this makes the \clines more thick
\newlength{\arrayrulewidthOriginal}
\newcommand{\Cline}[2]{%
  \noalign{\global\setlength{\arrayrulewidthOriginal}{\arrayrulewidth}}%
  \noalign{\global\setlength{\arrayrulewidth}{#1}}\cline{#2}%
  \noalign{\global\setlength{\arrayrulewidth}{\arrayrulewidthOriginal}}}

%\definecolor{BoxCol}{rgb}{0.9,0.9,0.9}
% uncomment for grey background to \section boxes 
% for use with default option boxedsections
\definecolor{BoxCol}{rgb}{0.1,0.1,0.8}
\definecolor{TextColTitle}{rgb}{0.1,0.1,0.8}

\definecolor{SectionCol}{rgb}{1,1,1}
% uncomment for dark blue \section text 
\definecolor{TextColG}{rgb}{0.1,0.0,0.8}
\definecolor{TextColR}{rgb}{0.9,0,0}
\newcommand{\Rcolor}{\textcolor{TextColR}}
\newcommand{\Gcolor}{\textcolor{TextColG}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

%-------------------------------------------------------------------------------------------------------
\title{{\Huge Hierarchically Supervised Latent Dirichlet Allocation}}

\author{
Adler Perotte\hspace{1cm} Nicholas Bartlett \hspace{1cm} No\'emie Elhadad \hspace{1cm} Frank Wood\\
Columbia University, New York, NY 10027, USA \\
%\texttt{\{ajp9009@dbmi,bartlett@stat,noemie@dbmi,fwood@stat\}.columbia.edu}
%\texttt{pfau@neurotheory.columbia.edu} 
%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
}

%\author{D.J. Albers \hspace{2 in} Noemie Elhadad \hspace{2in} George
%  Hripcsak  \hspace{2in} Esteban Tabuk}

%\institute{Columbia University \hspace{2.0 in} Columbia University
%  \hspace{2.0 in} Columbia University \hspace{2.0 in} Courant Institute (NYU)}

%\email{\hspace{1 in}\{pdeleon,vijendra\}@nmsu.edu\hspace{5 in} pucher@ftw.at\hspace{3 in} jyamagis@inf.ed.ac.uk\hspace{0 in}}

% The following commands can be used to alter the default logo settings
%\leftlogo[1.0]{Figures/NMlogo}  % defines logo to left of title (with scale factor)
\rightlogo[1]{crown2}  % same but on right

%-------------------------------------------------------------------------------------------------------
\begin{document}

%define conference poster is presented at (appears as footer)
%\conference{NIPS 2011}

\maketitle
\begin{abstract}
\begin{center}
\parbox{70 cm}{
We introduce hierarchically supervised latent Dirichlet allocation (HSLDA), a
model for hierarchically and multiply labeled bag-of-word data.  Examples of
such data include web pages and their placement in directories, product
descriptions and associated categories from product hierarchies, and free-text
clinical records and their assigned diagnosis codes. Out-of-sample label
prediction is the primary goal of this work, but improved lower-dimensional
representations of the bag-of-word data are also of interest. We demonstrate HSLDA on large-scale data from clinical document labeling and
retail product categorization tasks. We show that leveraging the structure from
hierarchical labels improves out-of-sample label prediction substantially when
compared to models that do not. 
}
\end{center}
\end{abstract}

%%% Begin of Multicols-Enviroment
\begin{multicols}{3}

%-------------------------------------------------------------------------------------------------------
\section{\textsc{Introduction}}

\begin{itemize}
\item Our work operates within the framework of topic modeling. 
\item Our approach learns
topic models of the underlying data and labeling strategies in a joint model,
while leveraging the hierarchical structure of the labels. 
\item For the sake of
simplicity, we focus on is-a hierarchies, but the model can be
applied to other structured label spaces. 
\item Our work extends supervised latent Dirichlet
allocation (sLDA)~\citep{BleiMcAuliffe2008} to take advantage of hierarchical
supervision and proposes an efficient way to incorporate such information into the model. 
\item We demonstrate our model on large, real-world datasets in the clinical and web
retail domains.
\end{itemize}
%We hypothesize that the context of labels within
%the hierarchy provides valuable information about labeling. Other models, such as LabeledLDA~\citep{Ramage2009}, incorporate LDA and supervision; however, none of these models leverage dependency structure in the label space.

%We observe that hierarchical information is valuable when
%incorporated into the learning and improves our primary goal of multi-label
%classification. Our results show that a joint, hierarchical model outperforms a
%classification with unstructured labels as well as a disjoint model, where the
%topic model and the hierarchical classification are inferred
%independently of each other.

%\begin{figure}
%\centering
%\includegraphics[width=18cm]{final_dry.pdf}
%\small{\caption{Dry geostrophic adjustment}}
%\end{figure}

%-------------------------------------------------------------------------------------------------------
\section{\textsc{Data}}

\subsection{Discharge Summaries and ICD-9 Codes}

\begin{figure}
\centering
\includegraphics[width=25cm]{ICD9_crop}
\small{\caption{A Portion of the ICD9 Hierarchy}}
\end{figure}

\begin{itemize}
\item \emph{Text:} Discharge summaries are authored by clinicians to summarize patient hospitalization courses.
\begin{itemize}
\item 10000 unique words, 6000 documents, average of 536.57 terms per document (std dev=300.29)
\end{itemize}
\item \emph{Labels:} Trained medical coders review the information in the discharge summary and assign a series of diagnoses codes from a hierarcharchy called ICD9.
\begin{itemize}
\item 7298 unique codes, 6000 code sets, average of 8.39 codes per set (std dev=5.01)
\end{itemize}
\end{itemize}

\subsection{Product Descriptions and Categorizations}

\begin{figure}
\centering
\includegraphics[width=25cm]{Amazon_crop}
\small{\caption{A Portion of the Amazon.com Hierarchy}}
\end{figure}

\begin{itemize}
\item \emph{Text:} Product descriptions were obtained from the Amazon.com website directly. We limited our dataset to the collection of DVDs in the product catalog.
\begin{itemize}
\item 30000 unique words, 16130 documents, average of 91.89 terms per document (std dev=53.08)
\end{itemize}
\item \emph{Labels:} Amazon.com product categorization data from the Stanford Network Analysis Platform (SNAP) dataset~\citep{SNAP}.
\begin{itemize}
\item 2691 unique categories, 16130 category sets, average of 9.01 codes per set (std dev=4.91) \newline \newline \newline \newline \newline \newline \newline \newline \newline
\end{itemize}
\end{itemize}

%-------------------------------------------------------------------------------------------------------
\section{\textsc{Model}}

\begin{itemize}
\item HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.

\begin{itemize}
\item Let $w_{n,d} \in \Sigma$ be the $n$th observation in the $d$th document.  
\item Let $\mathbf{w}_d = \{w_{1,d},\ldots,w_{1,N_d}\}$ be the  set of $N_d$ observations in document $d$.  
\item Let there be $D$ such documents and let the size of the vocabulary be $V=|\Sigma|$.  
\item Let the set of labels be $\mathcal{L}=\left\{  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. 
%Each label
%$l \in \mathcal{L}$, except root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
%also in the set of labels. 
% We will for exposition purposes assume that this label set has hard ``is-a''
% parent-child constraints, although this assumption can be
% relaxed at the cost of more computationally complex inference.  Such a label hierarchy forms a multiply rooted tree.  Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.  
\begin{itemize}
\item Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.   
\item In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.  
\item In the applications we consider, only positive label applications are observed.  
\item Label responses are generated using a conditional hierarchy of probit regressors\cite{gelmanbda04}.

\end{itemize}
\end{itemize}
\end{itemize}

%In HSLDA, the bag-of-word document data is modeled using the LDA
%mixed-membership mixture model with global topic estimation.
%In HSLDA, documents are modeled using the LDA mixed-membership mixture model
%with global topic estimation.

% The HSLDA graphical model is given in
%Figure~\ref{fig:graphical_model}. In the model, $K$ is the number of LDA
%``topics'' (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$
%is a distribution over ``words,'' $\boldsymbol\theta_d$ is a document-specific
%distribution over topics, and $\boldsymbol\beta$ is a global distribution over
%topics.

\begin{figure}
\centering
\includegraphics[width=25cm]{Graphical_Model-final}
\small{\caption{HSLDA Graphical Model}}
\label{fig:graphical_model} 
\end{figure}

\subsection{Generative Story}

%\begin{small}
\begin{enumerate}
\item For each topic $k=1,\ldots,K$

\begin{itemize}
\item Draw a distribution over words $\boldsymbol\phi_{k}\sim{\rm Dir}_{V}(\gamma\mathbf{1}_V)$%,
%where $\mathbf{1}$ is a vector of ones of length $V$ 
\end{itemize}
\item For each label $l\in\mathcal{L}$

\begin{itemize}
\item Draw a label application coefficient $\boldsymbol\eta_{l}\mid\mu,\sigma\sim\mathcal{N}_{K}(\mu \mathbf{1}_K,\sigma \mathbf{I}_{K})$  
\end{itemize}
\item Draw the global topic proportions $\boldsymbol\beta\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}_K\right)$
\item For each document $d=1,\ldots,D$

\begin{itemize}
\item Draw topic proportions $\boldsymbol\theta_d\mid\boldsymbol\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\boldsymbol\beta\right)$ 
\item For $n=1,\ldots,N_{d}$

\begin{itemize}
\item Draw topic assignment $z_{n,d}\mid\boldsymbol\theta_d\sim{\rm Multinomial}(\boldsymbol\theta_d)$ 
\item Draw word $w_{n,d}\mid z_{n,d},\boldsymbol\phi_{1:K}\sim{\rm Multinomial}(\boldsymbol\phi_{z_{n,d}})$ 
\end{itemize}
\item Set $y_{r,d} = 1$
\item For each label $l$ in a breadth first traversal of $\mathcal{L}$ starting at the children of  root $r$

\begin{itemize}
\item Draw $a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\eta_{l},y_{\mathrm{pa}(l),d}\sim\begin{cases}
\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1), & y_{\mathrm{pa}(l),d}=1\\
\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l),d}=-1\end{cases}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$

\item Apply label $l$ to document $d$ according to $a_{l,d}$ \[\hspace{-2cm}
y_{l,d}\mid a_{l,d}=\begin{cases}
\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}} \\% and \ensuremath{y_{{\rm \mathrm{pa}}(l),d}=1}}\\
-1 & \text{otherwise}\end{cases}\]

\end{itemize}
\end{itemize}
\end{enumerate}
%\end{small}
%--------------------------------------------------------------------------------------------------------

\section{\textsc{Inference}}

We define a collapsed Gibbs sampler for HSLDA where the topic assignments are sampled according to:  
%We have analytically marginalized out the parameters $\boldsymbol{\phi}_{1:K}$
%and $\boldsymbol{\theta}_{1:D}$  in the following expressions.   Let $\mathbf{a}$ be the set of all auxiliary variables, $\mathbf{w}$  the set of all words, $\boldsymbol\eta$  the set of all regression coef%ficients, and  $\mathbf{z}_d\backslash z_{n,d}$  the set $\mathbf{z}_d$ with element $z_{n,d}$ removed.  The conditional posterior distribution of the latent topic indicators is
\begin{eqnarray*}
\lefteqn{p\left(z_{n,d}=k\mid\mathbf{z}_d\backslash z_{n,d},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma\right)\propto}\nonumber \\
 & \hspace{2cm}\left(c_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\boldsymbol\beta_{k}\right)\frac{c_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(c_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)}\prod_{l\in\mathcal{L}_{d}}\exp\left\{ -\frac{\left(\bar{\mathbf{z}}_{d}^{T}\boldsymbol\eta_{l}-a_{l,d}\right)^{2}}{2}\right\} \label{eq:z-likelihood}\end{eqnarray*}
% where $c_{v,d}^{k,-\left(n,d\right)}$ is the number
%of words of type $v$ in document $d$ assigned to topic $k$ omitting
%the $n$th word of document $d$.

%The subscript $(\cdot)$'s
%indicate to sum over the range of the replaced variable, i.e.~$ {c_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}} = \sum_d {c_{w_{n,d},d}^{k,-\left(n,d\right)}}$.  Here $\mathcal{L}_{d}$ is the set of labels which are observed for document $d$.

%Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z}_{-\left(d,n\right)},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma\right)$
%can be sampled through enumeration.

%The conditional posterior distribution of the regression coefficients is given by 
The regression coefficients are sampled according to:
\begin{equation*}
p(\boldsymbol\eta_{l}\mid\mathbf{z},\mathbf{a},\sigma) = \mathcal{N}(\hat{\boldsymbol\mu}_{l},\hat{\mathbf{\Sigma}})\label{eqn:regression_param_conditional}
\end{equation*}
%$\boldsymbol\eta_{l}$ for $l\in\mathcal{L}$. Given that $\boldsymbol\eta_{l}$
%and $a_{l,d}$ are distributed normally, the posterior distribution
%of $\boldsymbol\eta_{l}$ is normally distributed with mean $\hat{\boldsymbol\mu}_{l}$
%and covariance $\hat{\mathbf{\Sigma}}$ such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
where
\begin{equation*}
\hat{\boldsymbol\mu}_{l}  =  \hat{\mathbf{\Sigma}}\left(\mathbf{1}\frac{\mu}{\sigma}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{l}\right) \qquad \hat{\mathbf{\Sigma}}^{-1}  =  \mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}
.\end{equation*}
The auxiliary variables are sampled according to:
\begin{equation*}
p\left(a_{l,d,}\mid\mathbf{z},\mathbf{Y},\mathbf{\boldsymbol\eta}\right)\propto\frac{1}{\sqrt{2\pi}}\exp\left\{ -\frac{1}{2}\left(a_{l,d}-\boldsymbol\eta_{l}^{T}\mathbf{\bar{z}}_{d}\right)\right\} \mathbb{I}\left(a_{l,d}y_{l,d}>0\right).\label{eqn:a_l_d}\end{equation*}

Lastly, the document-level topic distributions are sampled according to:
\begin{equation*}
\boldsymbol\beta\mid\mathbf{z},\alpha^{\prime},\alpha\sim{\rm Dir}\left(m_{\left(\cdot\right),1}+\alpha^{\prime},m_{\left(\cdot\right),2}+\alpha^{\prime},\ldots,m_{\left(\cdot\right),K}+\alpha^{\prime}.\right)\end{equation*}
where the table counts, $m_{d,k}$, are sampled directly.

%-------------------------------------------------------------------------------------------------
\section{\textsc{Experiments}}

\begin{itemize}
\item Comparison models:
\begin{itemize}

\item sLDA with independent regressors (isolates effect of hierarchical constraints)
\item HSLDA fit by first performing LDA then fitting tree-conditional regressions (isolates effect of combined inference)
\end{itemize}
\item Parameters:
\begin{itemize}
\item Number of topics, $K$, set to 50.
\item $p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and $p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1 and a scale parameters of 1000. 
\end{itemize}
\end{itemize}

\subsection{Topic Examples}

%\begin{multicols}{2}
%\begin{center}
%\textbf{Clinical Topics}
%\end{center}
%\begin{center}
%\textbf{Product Topics}
%\end{center}
%\end{multicols}

\begin{small}
\begin{tabular}{|c|c||c|c|}
\hline
\multicolumn{2}{|c||}{\textbf{Clinical Topics}} &
\multicolumn{2}{|c|}{\textbf{Product Topics}} \\ \hline
MASS & WOUND & SERIES & BASEBALL \\
\hline
CANCER & FOOT & EPISODES & TEAM \\
\hline
RIGHT & CELLULITIS & SHOW & GAME \\
\hline
BREAST & ULCER & SEASON & PLAYERS \\
\hline
CHEMOTHERAPY & LEFT & EPISODE & BASKETBALL \\
\hline
METASTATIC & ERYTHEMA & FIRST & SPORT \\
\hline
LEFT & PAIN & TELEVISION & SPORTS \\
\hline
LYMPH & SWELLING & SET & NEW \\
\hline
TUMOR & SKIN & TIME & PLAYER \\
\hline
BIOPSY & RIGHT & TWO & SEASON \\
\hline
CARCINOMA & ABSCESS & SECOND & LEAGUE \\
\hline
LUNG & LEG & ONE & FOOTBALL \\
\hline
CHEMO & OSTEOMYELITIS & CHARACTERS & STARS \\
\hline
ADENOCARCINOMA & TOE & DISC & FANS \\
\hline
NODE & DRAINAGE & GUEST & FIELD \\
\hline
\end{tabular}
\end{small}

\subsection{Prediction}

\begin{figure}
\centering
\includegraphics[width=25cm]{ROC_comparison_leafs}
\small{\caption{ROC curve for held-out ICD9 prediction varying auxiliary variable threshold.}}
\label{ICD9_aux_results}
\end{figure}

\begin{multicols}{2}

\begin{figure}
\centering
\includegraphics[width=11cm]{clin_pred_varying_mu}
\small{\caption{ROC curve for held-out ICD9 prediction varying regression parameter regularization.}}
\label{ICD9_mu_results}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=11cm]{amazon_pred_varying_mu}
\small{\caption{ROC curve for held-out product categorization prediciton varying regression parameter regularization.}}
\label{Amazon_mu_results}
\end{figure}

\end{multicols}

%\begin{figure}[h]
%\begin{center}
%\subfigure[][]{\label{fig:1a}\includegraphics[width=.3\textwidth]{clin_pred_varying_mu}}
%\subfigure[][]{\label{fig:1b}\includegraphics[width=.3\textwidth]{amazon_pred_varying_mu}}
%\subfigure[][]{\label{fig:clinical_roc}\includegraphics[width=0.3\textwidth]{ROC_comparison_leafs}}
%\caption{ROC curves for out-of-sample ICD-9 code prediction from patient free-text discharge records 
%(\subref{fig:1a},\subref{fig:clinical_roc}). ROC curve for out-of-sample Amazon product category predictions from 
%product free-text descriptions \subref{fig:1b}. Figures \subref{fig:1a} and \subref{fig:1b} are a function 
%of the prior means of the regression parameters. Figure \subref{fig:clinical_roc} is a function of auxiliary variable threshold. In all figures, solid is 
%HSLDA, dashed are independent regressors + sLDA (hierarchical 
%constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
%tree-conditional regressions.}
%\label{fig:main_results}
%\end{center}
%\end{figure}

%-------------------------------------------------------------------------------------------------
\section{\textsc{Conclusions}}

\begin{itemize}
\item We present an efficient Gibbs sampler for sLDA based models.
\item The results in Figures~\ref{ICD9_aux_results}, \ref{ICD9_mu_results} and \ref{Amazon_mu_results} suggest that in most cases it is better to do full joint estimation of HSLDA when the label space is strongly structured.
\item One can, in an engineering sense, get nearly as much gain in label prediction performance by first fitting LDA and then fitting a hierarchical probit regression.  There are applied settings in which this could be advantageous.
\end{itemize}

% References

\begin{small} \bibliographystyle{plainnat} \bibliographystyle{plainnat}
\bibliographystyle{plainnat}
\bibliography{refs}

\end{small}

\end{multicols}

\end{document}

### 2011-10-27 paper/experiments.tex
Added poster and updated paper for camera-ready

@@
 Discharge summaries are authored by clinicians to summarize patient
-hospitalization course. The summaries typically contain a record of patient
+hospitalization courses. The summaries typically contain a record of patient
 complaints, findings and diagnoses, along with treatment and hospital course.
@@
 their associated ICD-9 codes (7,298 distinct codes overall), representing all
-the discharges from the hospital in 2009. Summaries have 8.39 associated ICD-9
+the discharges from the hospital in 2009. All included discharge summaries had associated ICD-9 Codes.
+Summaries have 8.39 associated ICD-9
 codes on average (std dev=5.01) and contain an average of 536.57 terms after
@@
 for our work. The distinguishing factor between HSLDA and sLDA is the
-addition structure imposed on the label space, a distinction that we
+additional structure imposed on the label space, a distinction that we
 hypothesized would result in a difference in predictive performance.
@@
 %\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
+\subfigure[][Clinical data performance.]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
 %\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[][]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
-\caption{ROC curves for out-of-sample ICD-9 code prediction from patient free-text discharge records 
-(\subref{fig:1a},\subref{fig:clinical_roc}). ROC curve for out-of-sample Amazon product category predictions from 
-product free-text descriptions \subref{fig:1b}. Figures \subref{fig:1a} and \subref{fig:1b} are a function 
-of the prior means of the regression parameters. Figure \subref{fig:clinical_roc} is a function of auxiliary variable threshold. In all figures, solid is 
-HSLDA, dashed are independent regressors + sLDA (hierarchical 
+\subfigure[][Retail product performance.]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+%\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
+\caption{ROC curves for out-of-sample label prediction varying $\mu$, the prior mean of the regression parameters. 
+In both figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical 
 constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
 tree-conditional regressions.}
-\label{fig:main_results}
+%\label{fig:main_results}
 \end{center}
@@
 predictive performance on held-out data. Prediction performance was measured
-with standard metrics -- sensitiviy (true positive rate) and 1-specificity
+with standard metrics -- sensitivity (true positive rate) and 1-specificity
 (false positive rate).
@@
 negative. In particular, since the sLDA model does not enforce the hierarchical
-constraints, its performance was found to be so poor on joint prediction that
-it was not included. The models can be compared on a more equal footing by
-considering only on the observed labels as being positive, despite the fact
+constraints, we establish a more equal footing by
+considering only the observed labels as being positive, despite the fact
 that, following the hierarchical constraints, ancestors must also be positive.
@@
 gold standard could lead to correctly positive predictions being labeled as
-false positives\citep{Birmetal2005}. 
+false positives~\citep{Birmetal2005}. 
 However, given that the label space is often large (as in our examples) it is a
@@
 probability distribution was estimated in the following way. Expectations 
-of $\mathbf{\bar{z}}_d$ and $\boldsymbol{\eta}_l$ were estimated with samples
+of $\mathbf{\bar{z}}_{\hat{d}}$ and $\boldsymbol{\eta}_l$ were estimated with samples
 from the posterior. Using these expectations, we performed Gibbs sampling over
@@
 as the average expected labeling for gold standard negative labels.
-
-Figure~\ref{fig:1a} shows the results for the clinical data. For instance, 
-a prior mean for the regression of $\mu=-1.6$ yields the following performance:
-the full HSLDA model had a 
-true positive rate of 0.57 and a false positive rate of 0.13, the sLDA model
-had a true positive rate of 0.42 and a false positive rate of 0.07, and the
-HSLDA model where LDA and the regressions were fit separately had a true
-positive rate of 0.39 and a false positive rate of 0.08.
-
-These results indicate that the full HSLDA model predicts more of the the
-correct labels at a cost of an increase in the number of false positives
-relative to the comparison models.
-
-Figure~\ref{fig:1b} shows the results for the retail product data. For
-instance, a prior mean for the regression of $\mu=-2.2$ yields the following
-performance: the full HSLDA model had a true positive rate of 0.85 
-and a false positive rate of 0.30, the sLDA model had a true positive
-rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
-LDA and the regressions were fit separately had a true positive rate of 0.77
-and a false positive rate of 0.16. These results follow a similar pattern to the clinical data.
 
@@
 
-Figure \ref{fig:main_results} shows the predictive performance of HSLDA 
-relative to the two comparison models as a function of the prior mean
-on regression coefficients as a receiver operating characteristic (ROC) curve. 
+\begin{figure}[t]
+%tbp] %  figure placement: here, top, bottom, or page
+\centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC curve for out-of-sample ICD-9 code prediction varying auxiliary variable threshold.
+$\mu = -1.0$ for all three models in this figure.}
+\label{fig:results_auxiliary_var} 
+\end{figure}
+
+Figure~\ref{fig:1a} demonstrates the performance of the model on the clinical data as an ROC curve varying $\mu$.  For instance, 
+a hyperparameter setting of $\mu=-1.6$ yields the following performance:
+the full HSLDA model had a 
+true positive rate of 0.57 and a false positive rate of 0.13, the sLDA model
+had a true positive rate of 0.42 and a false positive rate of 0.07, and the
+HSLDA model where LDA and the regressions were fit separately had a true
+positive rate of 0.39 and a false positive rate of 0.08. These points are highlighted in Figure~\ref{fig:1a}.
+
+These results indicate that the full HSLDA model predicts more of the the
+correct labels at a cost of an increase in the number of false positives
+relative to the comparison models.
+
+Figure~\ref{fig:1b} demonstrates the performance of the model on the retail product data as an ROC curve also varying $\mu$. For
+instance, a hyperparameter setting of $\mu=-2.2$ yields the following
+performance: the full HSLDA model had a true positive rate of 0.85 
+and a false positive rate of 0.30, the sLDA model had a true positive
+rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
+LDA and the regressions were fit separately had a true positive rate of 0.77
+and a false positive rate of 0.16. These results follow a similar pattern to the clinical data. These points are highlighted in Figure~\ref{fig:1b}.
+
+Figure \ref{fig:results_auxiliary_var} shows the predictive performance of HSLDA 
+relative to the two comparison models on the clinical dataset as a function of the auxiliary variable threshold. 
 For low values of the auxiliary variable threshold, the models predict labels
@@
 increased, the models predict in a less sensitive and more specific manner,
-creating the points in the lower left hand corner of the ROC curve. For all
-values of the prior mean in both datasets, HSLDA outperforms sLDA with
-independent regressors. In the case of HSLDA with separately trained
-regression, HSLDA outperforms in the clinical dataset but performs equally well
-across the board with the retail product dataset.
+creating the points in the lower left hand corner of the ROC curve. HSLDA with full joint inference outperforms sLDA with
+independent regressors as well as HSLDA with separately trained
+regression.
 

### 2011-10-27 paper/experiments.tex
Camera ready draft 2

@@
 
-Our dataset was gathered from the clinical data warehouse of
-a large metropolitan hospital. It consists of 6,000 discharge summaries and
+Our dataset was gathered from the NewYork-Presbyterian Hospital clinical data warehouse. 
+It consists of 6,000 discharge summaries and
 their associated ICD-9 codes (7,298 distinct codes overall), representing all
@@
 for testing. The product descriptions are shorter than the discharge summaries
-(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique codes.
-Products are assigned on average 9.01 codes (std dev=4.91). The vocabulary
+(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique categories.
+Products are assigned on average 9.01 categories (std dev=4.91). The vocabulary
 consists of the most frequent 30,000 words omitting stopwords.
@@
 the two datasets. The comparison models included sLDA with independent
-regressors (hierarchical constraints on labels ignored), HSLDA fit by first
+regressors (hierarchical constraints on labels ignored) and HSLDA fit by first
 performing LDA then fitting tree-conditional regressions. These models were
@@
 considered positive and descendents of observed nodes were considered to be
-negative. In particular, since the sLDA model does not enforce the hierarchical
+negative. Note that this is different from our treatment of the observations during inference.
+Since the sLDA model does not enforce the hierarchical
 constraints, we establish a more equal footing by

### 2011-10-27 paper/introduction.tex
Camera ready draft 2

@@
 representations of data that have been hierarchically classified (e.g.,~image
-catalogs with bag-of-feature image representations).
+catalogs with bag-of-feature representations).
 
 There are several challenges entailed in incorporating a hierarchy of labels
-into the model. Given a large set of potential labels (often thousands), each
-instance has only a small number of labels associated to it. There are no
+into the model. Among them, given a large set of potential labels (often thousands), each
+instance has only a small number of labels associated to it. Furthermore, there are no
 naturally occurring negative labeling in the data, and the absence of a label
@@
 while leveraging the hierarchical structure of the labels. For the sake of
-simplicity, we focus on is-a hierarchies, but the model can be
-applied to other hierarchies. We extend supervised latent Dirichlet
+simplicity, we focus on ``is-a'' hierarchies, but the model can be
+applied to other structured label spaces. We extend supervised latent Dirichlet
 allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical

### 2011-10-27 paper/model.tex
Camera ready draft 2

@@
 $V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
-  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label labels
-$l \in \mathcal{L}$, except root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
+  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label
+$l \in \mathcal{L}$, except the root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
 also in the set of labels. 
@@
 In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.
-In the applications we consider, only positive label applications are observed.
+In the applications we consider, only positive labels are observed.
 

### 2011-10-27 paper/related_work.tex
Camera ready draft 2

@@
 Other models that incorporate LDA and supervision include
-LabeledLDA\citep{Ramage2009} and DiscLDA\citep{DiscLDA}.  Various applications of these models to 
-computer vision and document networks have been explored
-\citep{wangbleifeifei08,RelationalLDA} .
+LabeledLDA~\citep{Ramage2009} and DiscLDA~\citep{DiscLDA}.  Various applications of these models to 
+computer vision and document networks have been explored~\citep{wangbleifeifei08,RelationalLDA} .
 None of these models, however, leverage dependency structure in the label space.

### 2012-02-24 Krantz-v3/Chapters/chapter1/discussion.tex
Added booke chapter.

--- 2011-06-02 paper/discussion.tex
@@
 surprisingly large amount of that kind of data possess structured labels,
-either hierarchically constrained or otherwise. That HSLDA directly addresses
-this kind of data is a large part of the motivation for this work. That it
+either hierarchically constrained or otherwise.  HSLDA directly addresses
+this kind of data and works well in practice. That it
 outperforms more straightforward approaches should be of interest to
 practitioners.
-
-Variational Bayes has been the predominant estimation approach applied to sLDA
-models. Hierarchical probit regression makes for tractable Markov chain Monte
-Carlo SLDA inference, a benefit that should extend to other sLDA models should
-probit regression be used for response variable prediction there too.
-
-The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
-it is better to do full joint estimation of HSLDA.  An alternative
-interpretation of the same results is that, if one is more sensitive to the
-performance gains that result from exploiting the structure of the labels, then
-one can, in an engineering sense, get nearly as much gain in label prediction
-performance by first fitting LDA and then fitting a hierarchical probit
-regression.  There are applied settings in which this could be advantageous.
+%
+%Variational Bayes has been the predominant estimation approach applied to sLDA
+%models. Hierarchical probit regression makes for tractable Markov chain Monte
+%Carlo SLDA inference, a benefit that should extend to other sLDA models should
+%probit regression be used for response variable prediction there too.
+%
+%The results in Figures~\ref{fig:1a} and \ref{fig:1b} suggest that in most cases
+%it is better to do full joint estimation of HSLDA.  An alternative
+%interpretation of the same results is that, if one is more sensitive to the
+%performance gains that result from exploiting the structure of the labels, then
+%one can, in an engineering sense, get nearly as much gain in label prediction
+%performance by first fitting LDA and then fitting a hierarchical probit
+%regression.  There are applied settings in which this could be advantageous.
 
@@
 relaxations to different kinds of label structure.  Unbounded topic cardinality
-variants pose interesting inference challenges.  Utilizing different kinds of
-label structure is possible within this framework, but requires relaxing some
-of the simplifications we made in this paper for expositional purposes.
+variants pose interesting inference challenges.  Imposing different kinds of
+label structure constraints is possible within this framework, but requires relaxing some
+of the assumptions we made in deriving the sampling distributions for HSLDA inference.
 

### 2012-02-24 Krantz-v3/Chapters/chapter1/experiments.tex
Added booke chapter.

--- 2011-10-27 paper/experiments.tex
@@
-We applied HSLDA to data from two domains: predicting medical
+Having described HSLDA and given details on how to do inference in it, we turn to demonstrations of using HSLDA to solve problems from  two different domains: predicting medical
 diagnosis codes from hospital discharge summaries and predicting product
@@
 
-\subsection{Data and Pre-Processing}
+\subsection{Hospital Discharge Summaries and ICD-9 Codes}
 
-\subsubsection{Discharge Summaries and ICD-9 Codes}
+ Despite the growing emphasis on meaningful
+use of technology in medicine, many aspects of medical record-keeping
+remain a manual process. In the US, diagnostic coding for billing and insurance
+purposes is often handled by professional medical coders who must
+explore a patient's extensive clinical record before assigning the
+proper codes. %So while electronic health records (EHRs) should be
+%adopted by most medical institutions within the next several years,
+%largely due to the provisions of HITECH under the American Recovery
+%and Reinvestment Act \cite{Blumenthal2009}, there has been little
+%movement forward in automating medical coding.
 
-Discharge summaries are authored by clinicians to summarize patient
-hospitalization courses. The summaries typically contain a record of patient
-complaints, findings and diagnoses, along with treatment and hospital course.
+%Similarly, viewing the distribution of topics over discharge summaries
+%may reveal information about the latent structure of clinician documentation.
+%Lastly, the SLDA model would provide a novel approach to dealing with
+%the problem of high dimensionality when representing narrative text
+%in a vector space specifically by reducing dimensions from an entire
+%vocabulary of potentially tens of thousands of words to a set of several
+%dozen topics.
+
+A specific example of this involves labeling of hospital discharge summaries.  These summaries are authored by clinicians to summarize patient
+hospitalization courses. They typically contain a record of patient
+complaints, findings and diagnoses, along with treatment and hospital course.  The kind of text one might expect to find in such a discharge summary is illustrated by this made-up snippet
+\begin{quote}
+{ADLER Dischage summary text description snippet }
+\end{quote} 
 For each hospitalization, trained medical coders review the information in the
@@
 epidemiological, health management, and clinical
-purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm}  The ICD-9 codes are organized in a rooted-tree structure, with each
+purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm}  These ICD-9 codes are organized in a rooted-tree structure, with each
 edge representing an is-a relationship between parent and child, such that the
@@
 {}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
-pneumonia,'' where the former is a type of the latter.  It is worth noting that
+pneumonia,'' where the former is a type of the latter.  A representative sub-tree of the ICD-9 code tree is shown in Figure ADLER\ref{ADLER}. It is worth noting that
 the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
@@
 
+An automated process would ideally produce a more complete and accurate
+diagnosis lists. %Also, this model will reveal information about the
+%medical records themselves. For example, we may gain an understanding
+%of what a specific code actually means in terms of clinical narratives.
 The task of automatic ICD-9 coding has been investigated in the clinical
-domain.  Methods range from manual
-rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
-Other work had leveraged larger datasets and experimented with K-nearest
-neighbor, Naive Bayes, support vector machines, Bayesian Ridge Regression, as
+domain.  Methods used to solve this problem (besides HSLDA) range from applying manually derived coding rules
+rules to applications of online rule learning approaches \cite{Crammer2007,Goldstein2007,Farkas2008}.
+Many classification schemes have been applied to this problem: K-nearest
+neighbor , Naive Bayes, support vector machines, Bayesian Ridge Regression, as
 well as simple keyword mappings, all with promising
@@
 
-Our dataset was gathered from the NewYork-Presbyterian Hospital clinical data warehouse. 
+The specific dataset we report results for in this chapter was gathered from the NewYork-Presbyterian Hospital clinical data warehouse. 
 It consists of 6,000 discharge summaries and
@@
 
+Here HSLDA is evaluated as a way to understand and model the relationship between a discharge summary and the ICD-9 codes that should be assigned to it.
+We show promising results for automatically assigning ICD-9 codes to hospital discharge records.
+
 %For each hospitalization there are usually several ICD-9 codes assigned
 %for billing purposes. These codes are known to be quite specific but
-%not very sensitive \citep{}. Regardless of that fact,
+%not very sensitive \cite{}. Regardless of that fact,
 %this is one of the only sources for information on patient diagnoses
@@
 
-\subsubsection{Product Descriptions and Categorizations}
+\subsection{Product Descriptions and Catalogs}
 
-Amazon.com, an online retail store, organizes its catalog of products in a
-mulitply-rooted hierarchy and provides textual product descriptions 
+Many web-retails store and organize their catalog of products in a
+mulitply-rooted hierarchy in addition to providing textual product descriptions 
 for most products. Products can be discovered by users
@@
 
-In this experiment, we obtained Amazon.com product categorization data from 
-the Stanford Network Analysis Platform (SNAP) dataset~\citep{SNAP}. 
+Amazon.com is one such retailer.  Their product categorization data is available as part of the 
+ Stanford Network Analysis Platform (SNAP) dataset~\cite{SNAP}.    A representative sub-tree of the amazon.com DVD produce categorization tree is shown in Figure ADLER\ref{ADLER}.  
 Product descriptions were obtained separately from the
-Amazon.com website directly. We limited our dataset to the collection of DVDs
-in the product catalog.
-
+Amazon.com website directly. Once such description is
+\begin{quote}
+{ADLER DVD description }
+\end{quote}
+We study the collection of DVDs
+in the product catalog specifically.
 %We were able to deduce the structure of the
@@
 %for the DVD, {}``The Time Machine.''
-
-Our dataset contains 15,130 product descriptions for training and 1,000
-for testing. The product descriptions are shorter than the discharge summaries
-(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique categories.
+The resulting dataset contains 15,130 product descriptions for training and 1,000
+for testing. The product descriptions consist of
+91.89 terms on average (std dev=53.08). Overall, there are 2,691 unique categories.
 Products are assigned on average 9.01 categories (std dev=4.91). The vocabulary
@@
 
+HSLDA is used here to understand and model the relationship between the product text description and the products' positioning in the product hierarchy.
+We show how to automatically situate a product in a hierarchal product catalog.
+
 \subsection{Comparison Models}
 
-We evaluated HSLDA along with two closely related models against
-the two datasets. The comparison models included sLDA with independent
-regressors (hierarchical constraints on labels ignored) and HSLDA fit by first
-performing LDA then fitting tree-conditional regressions. These models were
-chosen to  highlight several aspects of HSLDA including performance in the
+We compare HSLDA to two closely related models. The comparison models are SLDA with independent
+regressors (hierarchical constraints on labels ignored,  i.e. the regression is not conditional) and HSLDA fit by first
+performing LDA then fitting tree-conditional regressions (rather than jointly inferring the topics and the regression coefficients). These models were
+chosen because they are the strongest available competitors and because they  highlight several pedagogical aspects of HSLDA including performance in the
 absence  of hierarchical constraints, the effect of the combined inference, and
@@
 
-sLDA with independent regressors is the most salient comparison model
-for our work. The distinguishing factor between HSLDA and sLDA is the
-additional structure imposed on the label space, a distinction that we
+SLDA with independent regressors is the most salient comparison model
+for our work. The distinguishing factor between HSLDA and SLDA is the
+additional structure imposed on the label space, a distinction that in developing HDSLA we
 hypothesized would result in a difference in predictive performance.
 
-There are two components to HSLDA, LDA and a hierarchically constrained
-response. The second comparison model is HSLDA fit by performing LDA first
+ The second comparison model, HSLDA fit by performing LDA first
 followed by performing inference over the hierarchically constrained label
-space. In this comparison model, the separate inference processes do not allow
-the responses to influence the low dimensional structure inferred by LDA.
-Combined inference has been shown to improve performance in sLDA
-\citep{BleiMcAuliffe2008}. This comparison model examines not the structuring
-of the label space, but the benefit of combined inference over both the
+space, does not allow
+the responses to influence the topics inferred by LDA.
+Combined inference has been shown to improve performance in SLDA
+\cite{BleiMcAuliffe2008}. This comparison model doesn't examine the value of utilizing the structured nature 
+of the label space, instead it highlights the benefit of combined inference over both the
 documents and the label space.
@@
 negative mean implements a prior belief that missing labels are likely to be
-negative. Thus, we evaluated model performance for all three models with a
+negative. Thus, we show model performance for all three models with a
 range of values for $\mu$, the mean prior parameter for regression coefficients 
@@
 
-The number of topics for all models was set to 50, the prior distributions of
+The number of topics for all models was set to $K=50$, the prior distributions of
 $p\left(\alpha\right)$, $p\left(\alpha^{\prime}\right)$, and
-$p\left(\gamma\right)$ were gamma distributed with a shape parameter of 1 and a
-scale parameters of 1000.
+$p\left(\gamma\right)$ were all chosen to be gamma with a shape parameter of 1 and a
+scale parameter of 1000.
 
@@
 %\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][Clinical data performance.]{\label{fig:1a}\includegraphics[width=.49\textwidth]{figs/clin_pred_varying_mu}}
+\subfigure[][Clinical data performance.]{\label{fig:1a}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/clin_pred_varying_mu}}
 %\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[][Retail product performance.]{\label{fig:1b}\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
+\subfigure[][Retail product performance.]{\label{fig:1b}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/amazon_pred_varying_mu}}
 %\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}
-\caption{ROC curves for out-of-sample label prediction varying $\mu$, the prior mean of the regression parameters. 
-In both figures, solid is HSLDA, dashed are independent regressors + sLDA (hierarchical 
+\caption{ROC curves for HSLDA out-of-sample label prediction varying $\mu$, the prior mean of the regression parameters. 
+In both figures, solid is HSLDA, dashed are independent regressors + SLDA (hierarchical 
 constraints on labels ignored), and dotted is HSLDA fit by running LDA first then running 
@@
 
-We evaluated our model, HSLDA, against the comparison models with a focus on
+We are particularly interested in
 predictive performance on held-out data. Prediction performance was measured
@@
 
-The gold standard for comparison was derived from the testing set in each
-dataset. To make the comparison as fair as possible among models, ancestors of
+In each case the gold standard for testing was derived from the test data. To make the comparison as antagonistic to HSLDA as possible (relative to the other models), ancestors of
 observed nodes in the label hierarchy were ignored, observed nodes were
-considered positive and descendents of observed nodes were considered to be
-negative. Note that this is different from our treatment of the observations during inference.
-Since the sLDA model does not enforce the hierarchical
-constraints, we establish a more equal footing by
-considering only the observed labels as being positive, despite the fact
-that, following the hierarchical constraints, ancestors must also be positive.
-Such a gold standard will likely inflate the number of false positives because
+considered positive and descendants of observed nodes were assumed to be
+negative. Note that this is different from our treatment of the observations during inference where we marginalize over possible settings of unobserved labels.
+For instance, as the SLDA model does not enforce hierarchical
+label constraints, when we consider only observed nodes we penalize HSLDA.  This is because
+ the is-a hierarchical constraints say that the ancestors of positively labeled nodes must also be positive which the SLDA model cannot guarantee.
+Another antagonism of this gold standard is that it is likely to inflate the number of false positives because
 the labels applied to any particular document are usually not as complete as
-they could be.  ICD-9 codes, for instance, lack sensitivity and their use as a
+they could be.  ICD-9 codes, for instance, are known to lack sensitivity and their use as a
 gold standard could lead to correctly positive predictions being labeled as
-false positives~\citep{Birmetal2005}.
+false positives~\cite{Birmetal2005}.
 However, given that the label space is often large (as in our examples) it is a
-moderate assumption that erroneous false positives should not skew results
+reasonable assumption that erroneous false positives should not skew results
 significantly.
 
-Predictive performance in HSLDA is evaluated by $p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)$ 
-for each test document, $\hat{d}$. For efficiency, the expectation of this
-probability distribution was estimated in the following way. Expectations 
+Predictive performance in HSLDA is evaluated by computing
+ \[p\left(y_{l,\hat{d}}\mid w_{1:N_{\hat{d}},\hat{d}}, w_{1:N_d,1:D},  y_{l\in\mathcal{L},1:D}\right)\] 
+for each test document $\hat{d}$ for each observed label $y_{l,\hat{d}}$ (given the test document words). For efficiency, the expectation of this
+probability distribution was approximated in the following way. Expectations 
 of $\mathbf{\bar{z}}_{\hat{d}}$ and $\boldsymbol{\eta}_l$ were estimated with samples
-from the posterior. Using these expectations, we performed Gibbs sampling over
+from the posterior. Fixing these expectations, we performed Gibbs sampling over
 the hierarchy to acquire predictive samples for the documents in the test set.
@@
 %tbp] %  figure placement: here, top, bottom, or page
-\centering \includegraphics[scale=0.4]{figs/ROC_comparison_leafs} \caption{ROC curve for out-of-sample ICD-9 code prediction varying auxiliary variable threshold.
+\centering \includegraphics[scale=0.4]{Chapters/chapter1/figs/ROC_comparison_leafs} \caption{ROC curve for out-of-sample ICD-9 code prediction varying auxiliary variable threshold.
 $\mu = -1.0$ for all three models in this figure.}
@@
 the full HSLDA model had a 
-true positive rate of 0.57 and a false positive rate of 0.13, the sLDA model
+true positive rate of 0.57 and a false positive rate of 0.13, the SLDA model
 had a true positive rate of 0.42 and a false positive rate of 0.07, and the
@@
 
+Example topics (as word lists) learned for the discharge data are given below.  These word lists are computed by ADLER
+
+\begin{extract}
+ADLER
+\end{extract}
+
+ADLER could you write a sentence about how these are cleaner than the SLDA or LDA topics?
+
 Figure~\ref{fig:1b} demonstrates the performance of the model on the retail product data as an ROC curve also varying $\mu$. For
@@
 performance: the full HSLDA model had a true positive rate of 0.85 
-and a false positive rate of 0.30, the sLDA model had a true positive
+and a false positive rate of 0.30, the SLDA model had a true positive
 rate of 0.78 and a false positive rate of 0.14, and the HSLDA model where
@@
 and a false positive rate of 0.16. These results follow a similar pattern to the clinical data. These points are highlighted in Figure~\ref{fig:1b}.
+
+Example topics (as word lists) learned for the amazon.com data are given below.  T
+
+\begin{extract}
+ADLER
+\end{extract}
 
@@
 increased, the models predict in a less sensitive and more specific manner,
-creating the points in the lower left hand corner of the ROC curve. HSLDA with full joint inference outperforms sLDA with
+creating the points in the lower left hand corner of the ROC curve. HSLDA with full joint inference outperforms SLDA with
 independent regressors as well as HSLDA with separately trained

### 2012-02-24 Krantz-v3/Chapters/chapter1/inference.tex
Added booke chapter.

\label{sec:inference}
Our inference goal  is to obtain a representation of the posterior distribution
of the latent variables in the model.  The posterior distribution we seek does not have a simple analytic form from which exact samples can be drawn.
This is usually the case for posterior distributions of non-trivial
probabilistic models and suggests a approximating the posterior distribution by sampling.

In this section we derive the conditional distributions required to sample from the HSLDA posterior distribution using Markov chain Monte Carlo.  A very brief review of Markov chain Monte Carlo can be found in Section \ref{sec:mcmc} at the end of this chapter.

The HSLDA sampler, like the collapsed Gibbs samplers for LDA \cite{Griffiths04}, is itself a collapsed Gibbs sampler in which all of the latent variables that can be analytically marginalized are.  Among others, the parameters $\boldsymbol{\phi}_{1:K}$
and $\boldsymbol{\theta}_{1:D}$ are analytically marginalized prior to deriving the following conditional distributions for sampling.  
  It will often be the
case that the set of labels $\mathcal{L}$ is not fully observed for
every document. We will define $\mathcal{L}_{d}$ to be the subset
of labels which have been observed for document $d$. In an is-a hierarchical regression  it is straightforward
to marginalize the variables $a_{l^{\prime},d}$ and $y_{l^{\prime},d}$
for $l^{\prime}\in\mathcal{L}\backslash\mathcal{L}_{d}$ simply by ignoring them. 
%We can also integrate out the parameters $\boldsymbol{\phi}_{1:K}$
%and $\boldsymbol{\theta}_{1:D}$ as in \cite{Griffiths04}.
 The remaining latent variables (those that are not collapsed out) are $\mathbf{z}=\{z_{1:N_{d},d}\}_{d=1,\ldots,D},\boldsymbol\eta=\{\boldsymbol\eta_{l}\}_{l\in\mathcal{L}},\mathbf{a}=\{a_{l^{\prime},d}\}_{l^{\prime}\in\mathcal{L}_{d},d=1,\ldots,D},\boldsymbol\beta,\alpha,\alpha^{\prime}$
and $\gamma$.

%We will appeal to one of the common methods
%for approximating posterior distributions in the face of intractable
%normalization factors: Markov chain Monte Carlo (MCMC) sampling. 
%Since
% the conditional distributions
%for all variables are relatively simple (enumeration and, in the case of the profit regression coefficients, conjugate) we will use Gibbs sampling. %In particular, we will derive a collapsed Gibbs sampler.
%Given an observation of a set of observed labels and a document, the posterior distribution for the latent variables is
%\begin{equation}
%p\left(\theta,z_{1:N},\phi_{1:K},\boldsymbol\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\boldsymbol\beta,\alpha,\alpha',\gamma\mid w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)\label{eq:Posterior}\end{equation}
%\[
%=\frac{p\left(\theta,z_{1:N},\phi_{1:K},\boldsymbol\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\boldsymbol\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}{\int_{\theta,\phi,a,\boldsymbol\eta,\alpha,\alpha',\boldsymbol\beta,\gamma}\sum_{z}p\left(\theta,z_{1:N},\phi_{1:K},\boldsymbol\eta_{i_{l,c}\in\mathcal{I}},a_{i_{l,c}\in\mathcal{I}},\boldsymbol\beta,\alpha,\alpha',\gamma,w_{1:N},y_{i_{l,c}\in\mathcal{I}};\sigma,\lambda\right)}\]
%
%

\subsection{Gibbs Sampler}
 Let $\mathbf{a}$ be the set of all auxiliary variables, $\mathbf{w}$  the set of all words, $\boldsymbol\eta$  the set of all regression coefficients, and  $\mathbf{z}\backslash z_{n,d}$  the set $\mathbf{z}$ with element $z_{n,d}$ removed.  %The conditional posterior distribution of the latent topic indicators is
 We use the notation $\mathbf{z}_{-(n,d)}$ to
denote $\mathbf{z}_d\backslash z_{n,d}$. %, marginalizing
%$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
%in LDA models see \cite{Griffiths04}.

%\subsection{Sampling latent topic indicators}
%$p(z_{n,d}\mid\mathbf{z}_{-(n,d)},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma)$}

First we consider the conditional distribution of $z_{n,d}$ (the assignment variable
for each word $n=1,\ldots,N_{d}$ in documents $d=1,\ldots,D$). 
%The
%conditional distribution does not include $\theta_{1:D}$ and $\phi_{1:K}$
%because they have been integrated out as in the collapsed Gibbs sampler
%\cite{Griffiths04}. 
%The conditional distribution of $z_{n,d}$ is
%proportional to the joint distribution of its markov blanket. \begin{equation}
%p\left(z_{d,n}\mid\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma\right)\propto p\left(z_{d,n},\mathbf{z_{-\left(d,n\right)}},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma\right)\end{equation}
Following the factorization of the model (refer again to Figure~\ref{fig:graphical_model}), we can write
\begin{eqnarray*}
\lefteqn{p\left(z_{n,d}\mid\mathbf{z}_d\backslash z_{n,d},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma\right)}\\
&\propto&\prod_{l\in\mathcal{L}_{d}}p\left(a_{l,d}\mid\mathbf{z},\boldsymbol\eta_{l}\right)p\left(z_{n,d}\ |\ \mathbf{z}_d\backslash z_{n,d},\mathbf{a},\mathbf{w},\alpha,\boldsymbol\beta,\gamma\right).\end{eqnarray*}
 The product is only over the subset of labels $\mathcal{L}_{d}$
which have been observed for document $d$. By isolating terms that
depend on $z_{n,d}$ and absorbing all other terms into a normalizing
constant as in \cite{Griffiths04} we find 
\begin{eqnarray}
\lefteqn{p\left(z_{n,d}=k\mid\mathbf{z}\backslash z_{n,d},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma\right)\propto} \label{eq:z-likelihood} \\
 & \hspace{2cm}\left(c_{\left(\cdot\right),d}^{k,-\left(n,d\right)}+\alpha\boldsymbol\beta_{k}\right)\frac{c_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}+\gamma}{\left(c_{\left(\cdot\right),\left(\cdot\right)}^{k,-\left(n,d\right)}+V\gamma\right)}\prod_{l\in\mathcal{L}_{d}}\exp\left\{ -\frac{\left(\bar{\mathbf{z}}_{d}^{T}\boldsymbol\eta_{l}-a_{l,d}\right)^{2}}{2}\right\}\nonumber\end{eqnarray}
 where $c_{v,d}^{k,-\left(n,d\right)}$ is the number
of words of type $v$ in document $d$ assigned to topic $k$ omitting
the $n$th word of document $d$. The subscript $(\cdot)$'s
indicate to sum over the range of the replaced variable, i.e.~$ {c_{w_{n,d},\left(\cdot\right)}^{k,-\left(n,d\right)}} = \sum_d {c_{w_{n,d},d}^{k,-\left(n,d\right)}}$.  Here $\mathcal{L}_{d}$ is the set of labels which are observed for document $d$.  We sample from \eqref{eq:z-likelihood}  by enumeration and implicit normalization.
%
%Given Equation \ref{eq:z-likelihood}, $p\left(z_{d,n}\mid\mathbf{z}_{-\left(d,n\right)},\mathbf{a},\mathbf{w},\mathbf{\boldsymbol\eta},\alpha,\boldsymbol\beta,\gamma\right)$
%can be sampled through enumeration.

The conditional posterior distribution of the regression coefficients is given by 
\begin{equation}
p(\boldsymbol\eta_{l}\mid\mathbf{z},\mathbf{a},\sigma) = \mathcal{N}(\hat{\boldsymbol\mu}_{l},\hat{\mathbf{\Sigma}})\label{eqn:regression_param_conditional}
\end{equation}
$\boldsymbol\eta_{l}$ for $l\in\mathcal{L}$. Given that $\boldsymbol\eta_{l}$
and $a_{l,d}$ are distributed normally, the posterior distribution
of $\boldsymbol\eta_{l}$ is normally distributed with mean $\hat{\boldsymbol\mu}_{l}$
and covariance $\hat{\mathbf{\Sigma}}.$% such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
where
\begin{equation*}
\hat{\boldsymbol\mu}_{l}  =  \hat{\mathbf{\Sigma}}\left(\mathbf{1}\frac{\mu}{\sigma}+\bar{\mathbf{Z}}^{T}\mathbf{a}_{l}\right) \qquad \hat{\mathbf{\Sigma}}^{-1}  =  \mathbf{I}\sigma^{-1}+\bar{\mathbf{Z}}^{T}\bar{\mathbf{Z}}
.\end{equation*}
Here $\bar{\mathbf{Z}}$ is a $D\times K$ matrix
such that row $d$ of $\mathbf{\bar{Z}}$ is $\bar{\mathbf{z}}_{d}$, and $\mathbf{a}_{l}=[a_{l,1},a_{l,2},\ldots,a_{l,D}]^{T}$.  The simplicity of this conditional distribution follows from the choice of probit regression  \cite{Albert_Chib_1993}; the specific form of the update is a standard result from Bayesian normal  linear regression \cite{gelmanbda04}. 
%p\left(\boldsymbol\eta_{i_{l,c}}\mid\mathbf{z}_{1:D},\mathbf{a};\sigma\right)=\mathcal{N}\left(\boldsymbol\eta_{i_{l,c}}\mid\hat{\mu}_{i},\hat{\Sigma}_{i}\right)\end{equation}
%\[
%\[
%\subsection{$p\left(a_{l,d}\mid\mathbf{z},\mathbf{Y},\mathbf{\boldsymbol\eta}\right)$}
%and \textmd{$p\left(y_{m,i}\mid\mathbf{a}\right)$}}
%The auxiliary variables $a_{l,d}$ must be sampled for documents $d=1,\ldots,D$
%and $l\in\mathcal{L}_{d}$. The conditional posterior distribution
%of 
It also is a standard probit regression result that the conditional posterior
distribution of $a_{l,d}$ is a truncated normal
distribution~\cite{Albert_Chib_1993}. 
\begin{eqnarray*}
\lefteqn{p\left(a_{l,d}\mid\mathbf{z},\mathbf{Y},\mathbf{\boldsymbol\eta}\right)}\\&\propto&
\begin{cases}
\exp\left\{ -\frac{1}{2}\left(a_{l,d}-\boldsymbol\eta_{l}^{T}\mathbf{\bar{z}}_{d}\right)\right\} \mathbb{I}\left(a_{l,d}y_{l,d}>0\right)\mathbb{I}(a_{l,d}<0) , & y_{\mathrm{pa}(l),d}=-1\\
\exp\left\{ -\frac{1}{2}\left(a_{l,d}-\boldsymbol\eta_{l}^{T}\mathbf{\bar{z}}_{d}\right)\right\} \mathbb{I}\left(a_{l,d}y_{l,d}>0\right) , & y_{\mathrm{pa}(l),d}=1\end{cases}\label{eqn:a_l_d}\end{eqnarray*}

%\begin{equation}
%p\left(a_{l,d,}\mid\mathbf{z},\mathbf{Y},\mathbf{\boldsymbol\eta}\right)\propto\frac{1}{\sqrt{2\pi}}\exp\left\{ -\frac{1}{2}\left(a_{l,d}-\boldsymbol\eta_{l}^{T}\mathbf{\bar{z}}_{d}\right)\right\} \mathbb{I}\left(a_{l,d}y_{l,d}>0\right).\label{eqn:a_l_d}\end{equation}
% This conditional distribution can be sampled using an inverse CDF
%method. %However, if $y_{d,i_{l,c}}$ is unobserved then $a_{d,i_{l,c}}$must
%be sampled jointly with $y_{d,i_{l,c}}$ to ensure that the Markov
%chain is ergodic. Suppose that $a_{d,i_{l,c}}$ is sampled to have
%a negative value and $y_{d,i_{l,c}}$ is apporopriately sampled at
%-1. Although there exist valid states where $a_{d,i_{l,c}}>0$ and
%$y_{d,i_{l,c}}=1$, they will never be reached by such a Markov chain
%since $p\left(a_{d,i_{l,c}}<0\mid y_{d,i_{l,c}}=-1\right)=1$ and
%$p\left(y_{d,i_{l,c}}=-1\mid a_{d,i_{l,c}}<0\right)=1$. Therefore,
%to ensure ergodicity, $a_{d,i_{l,c}}$and $y_{d,i_{l,c}}$ must be
%sampled from the joint distribution as shown in Equation \ref{eq:Probit-Joint}.
%\begin{equation}
%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\boldsymbol\eta}\right)\propto p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)p\left(a_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y},\mathbf{\boldsymbol\eta}\right)\end{equation}
%\begin{equation}
%p\left(y_{i_{l,c}}\mid\mathbf{a},\mathbf{y}_{-\left(l,c\right)}\right)=\delta\left(sign\left(a_{d,i_{l,c}}\right)=y_{i_{l,c}}\right)p\left(y_{i_{l,c}}\mid y_{parents_{l,c}}\right)\prod_{i_{\hat{l},\hat{c}}\in children_{l,c}}p\left(y_{i_{\hat{l},\hat{c}}}\mid y_{i_{l,c}}\right)\end{equation}
%\begin{equation}
%p\left(y_{i_{l,c}}=-1\mid y_{parent{}_{l,c}}\right)=\begin{cases}
%1, & y_{parent_{l,c}}=-1\\
%0.5, & y_{parent_{l,c}}=1\end{cases}\end{equation}
%\[
%p\left(a_{d,i_{l,c}},y_{d,i_{l,c}}\mid\mathbf{z},\mathbf{Y}_{-\left(d,i_{l,c}\right)},\mathbf{\boldsymbol\eta}\right)\]
%\begin{equation}
%=\begin{cases}
%\mathcal{N}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\boldsymbol\eta_{i_{l,c}},1\right)p\left(y_{d,i_{l,c}}\mid a_{d,i_{l,c}}\right), & y_{parent_{l,c}}=1,\forall y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}},y_{i_{\hat{l},\hat{c}}}=-1\\
%trunc\mathcal{N}^{-}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\boldsymbol\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=-1\right), & y_{parent_{l,c}}=-1\\
%trunc\mathcal{N}^{+}\left(a_{d,i_{l,c}}\mid\bar{z}^{T}\boldsymbol\eta_{i_{l,c}},1\right)\delta\left(y_{d,i_{l,c}}=1\right), & \exists y_{i_{\hat{l},\hat{c}}}\in y_{children_{l,c}}\setminus y_{i_{\hat{l},\hat{c}}}=1\\
%0 & otherwise\end{cases}\end{equation}
%where $\mathbf{Y}_{-\left(d,i_{l,c}\right)}$ denotes all of the response
%variables excluding the response variable being sampled.
%
%
%
%\subsection{$p\left(\boldsymbol\beta\mid\mathbf{z},\alpha^{\prime},\alpha\right)$}

%Note that care must be taken to initialize the Gibbs sampler in a valid state.

HSLDA employs a hierarchical Dirichlet prior over topic assignments (i.e.,~$\boldsymbol\beta$ is estimated from data rather than fixed a priori).  This has been shown to improve the quality and stability of inferred topics \cite{WallachMiMc2009}. 
%This prior
%shares many features with the hierarchical Dirichlet process and inference
%over this distribution proceeds in a very similar fashion.
%
Sampling $\boldsymbol\beta$, the vector of global topic proportions,   can be done using the ``direct assignment''
method of \cite{TehJorBea2006} \begin{equation}
\boldsymbol\beta\mid\mathbf{z},\alpha^{\prime},\alpha\sim{\rm Dir}\left(m_{\left(\cdot\right),1}+\alpha^{\prime},m_{\left(\cdot\right),2}+\alpha^{\prime},\ldots,m_{\left(\cdot\right),K}+\alpha^{\prime}.\right)\end{equation}
Here $m_{d,k}$ are auxiliary variables that are required to sample the posterior distribution of $\boldsymbol\beta$.  Their conditional posterior distribution is sampled according to
 \begin{equation}
p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\boldsymbol\beta\right)=\frac{\Gamma\left(\alpha\boldsymbol\beta_{k}\right)}{\Gamma\left(\alpha\boldsymbol\beta_{k}+c^k_{(\cdot),d}\right)}s\left(c^k_{(\cdot),d},m\right)\left(\alpha\boldsymbol\beta_{k}\right)^{m}\end{equation}
where $s\left(n,m\right)$ represents stirling numbers of the first
kind.

%\subsection{$p\left(\alpha\right)$, $p\left(\alpha'\right)$, $p\left(\gamma\right)$}

The hyperparameters $\alpha$, $\alpha^{\prime}$, and $\gamma$ are
%given broad ${\rm Gamma}(1,1000)$ prior distributions and 
sampled
using Metropolis-Hastings.

### 2012-02-24 Krantz-v3/Chapters/chapter1/introduction.tex
Added booke chapter.

--- 2011-06-02 paper/introduction.tex
@@
-There exist many sources of unstructured data that have been partially or
-completely categorized by human editors.  In this paper we focus on
-unstructured text data that has been, at least in part, manually
-categorized.  Examples include but are not limited to webpages and curated
-hierarchical directories of the same \citep{DMOZ}, product descriptions and
-catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
+%This chapter covers one way to utilize supervision in the form of
+%hierarchical and (often) multiple labelings in a similar manner. 
+Consider web
+retail data. Web retailers often have both a browse-able product hierarchy and
+free-text descriptions for all products they sell. The situation of each
+product in a product hierarchy (often multiply situated) constitutes a
+multiple, hierarchical labeling of the free-text product descriptions.
+One might assume that the free-text descriptions of all of the products in a particular node in the product hierarchy are related.
+Basketballs are probably described using language that is similar to that used to describe other basketballs, other balls, and more general sporting goods.
+In both the lay and technical senses, similar products should have product descriptions that share topics.
+What is more, if topic proportions are indicative of the text describing products that are grouped together, it should be possible to use those proportions to decide (via classification) whether or not a particular product should be situated at a particular node in the product hierarchy.
+Conversely, that certain groups of products are known to be clustered together should inform the kinds of topics that are inferred from the product descriptions.
+
+%hierarchical labels should, at least in theory, provide
+%better supervision than the simpler unstructured labels previously considered.
+%Results from applying our model to both medical record and web retail data
+%suggests that this is likely the case. In particular, we observe gains in our
+%primary goal of out-of-sample label prediction that result specifically from
+%leveraging hierarchical supervision.
+
+There are many kinds of problems that have the same characteristics as this: any data  that consists of free text that has been partially or
+completely categorized by human editors for instance; more fully
+% In this paper we focus on
+any bag-of-words data that has been, at least in part, 
+categorized. 
+ Examples include but are not limited to webpages and curated
+hierarchical directories of the same \cite{DMOZ}, product descriptions and
+catalogs, (e.g.~\cite{AMAZON} as available from \cite{SNAP})
 and patient records %hospital treatment transcripts
 and diagnosis codes assigned to them for bookkeeping and insurance purposes.
-In this work we show how to combine 
-these two sources of information using a single model that allows one to
+The model we cover in this chapter shows one way to combine 
+these two sources of information into a single model that allows one to
 categorize new text documents automatically, suggest labels that might be
 inaccurate, compute improved similarities between documents for information
-retrieval purposes, and more. The models and techniques that we develop in
-this paper are applicable to other data as well, namely, any unstructured
-representations of data that have been hierarchically classified (e.g.,~image
-catalogs with bag-of-feature image representations).
+retrieval purposes, and more. %The models and techniques that we develop in
+%this paper are applicable to other data as well, namely, any unstructured
+%representations of data that have been hierarchically classified (e.g.,~image
+%catalogs with bag-of-feature representations).
 
-There are several challenges entailed in incorporating a hierarchy of labels
-into the model. Given a large set of potential labels (often thousands), each
-instance has only a small number of labels associated to it. There are no
-naturally occurring negative labeling in the data, and the absence of a label
-cannot always be interpreted as a negative labeling.
+%There are several challenges entailed in incorporating a hierarchy of labels
+%into the model. Among them, given a large set of potential labels (often thousands), each
+%instance has only a small number of labels associated to it. Furthermore, there are no
+%naturally occurring negative labeling in the data, and the absence of a label
+%cannot always be interpreted as a negative labeling.
 
-Our work operates within the framework of topic modeling. Our approach learns
-topic models of the underlying data and labeling strategies in a joint model,
-while leveraging the hierarchical structure of the labels. For the sake of
-simplicity, we focus on is-a hierarchies, but the model can be
-applied to other hierarchies. We extend supervised latent Dirichlet
-allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
-supervision. We propose an efficient way to incorporate hierarchical
-information into the model. We hypothesize that the context of labels within
-the hierarchy provides valuable information about labeling.
+%Our work operates within the framework of topic modeling. Our approach learns
+%topic models of the underlying data and labeling strategies in a joint model,
+%while leveraging the hierarchical structure of the labels. 
+In this chapter we focus on ``is-a'' hierarchically structured label spaces, but label spaces that are structured differently may be accommodated with relatively minor modifications. 
+%
+%We extend supervised latent Dirichlet
+%allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
+%supervision. We propose an efficient way to incorporate hierarchical
+%information into the model. We hypothesize that the context of labels within
+%the hierarchy provides valuable information about labeling. 
+%
+We show results from modeling real-world datasets in the clinical and web
+retail domains. These results provide evidence that the two views (text and labels) mutually benefit  multi-label
+classification.%In fact, it will be shown that the hierarchical supervised model outperforms a
+That is, modeling the joint  is better than learning 
+topic models and hierarchical classifiers
+independently.
 
-We demonstrate our model on large, real-world datasets in the clinical and web
-retail domains. We observe that hierarchical information is valuable when
-incorporated into the learning and improves our primary goal of multi-label
-classification. Our results show that a joint, hierarchical model outperforms a
-classification with unstructured labels as well as a disjoint model, where the
-topic modeling and the hierarchical classification are carried out
-independently of each other.
-
-The remainder of this paper is as follows. Section~\ref{sec:model}
+The remainder of this chapter is arranged as follows. Section~\ref{sec:model}
 introduces hierarchically supervised LDA (HSLDA), while
@@
 %the final record, and a high error rate is found even in the principal
-%diagnoses \citep{Surjan1999}.
+%diagnoses \cite{Surjan1999}.
 
@@
 %largely due to the provisions of HITECH under the American Recovery
-%and Reinvestment Act \citep{Blumenthal2009}, there has been little
+%and Reinvestment Act \cite{Blumenthal2009}, there has been little
 %movement forward in automating medical coding.

### 2012-02-24 Krantz-v3/Chapters/chapter1/model.tex
Added booke chapter.

--- 2011-10-27 paper/model.tex
@@
 HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We
-will refer to individual groups of bag-of-word data as documents.  Let $w_{n,d}
+call individual groups of bag-of-word data documents (although they could be any data represented by bags of features).  Let $w_{n,d}
 \in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
@@
 In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.
-In the applications we consider, only positive labels are observed.
+In the applications we demonstrate, only positive labels are observed.
+This may not be true of all applications; however, positive-only label imbalance is a common problem.
+How we solve this problem will be discussed later.
 
@@
 %tbp] %  figure placement: here, top, bottom, or page
- \centering \includegraphics[scale=0.4]{Graphical_Model-final} \caption{HSLDA graphical model}
+ \centering \includegraphics[scale=0.4]{Chapters/chapter1/Graphical_Model-final} \caption{Hierarchically supervised latent Dirichlet allocation (HSLDA) graphical model.}
 
@@
 
-%In HSLDA, the bag-of-word document data is modeled using the LDA
-%mixed-membership mixture model with global topic estimation.
-In HSLDA, documents are modeled using the LDA mixed-membership mixture model
-with global topic estimation. Label responses are generated using a conditional
+In HSLDA, the bag-of-word document data is modeled using LDA with
+ full, hierarchical topic estimation (i.e. with the global topic estimated too rather than assumed).
+%In HSLDA, documents are modeled using the LDA mixed-membership mixture model
+%with global topic estimation. 
+Label responses are modeled using a conditional
 hierarchy of probit regressors. The HSLDA graphical model is given in
-Figure~\ref{fig:graphical_model}. In the model, $K$ is the number of LDA
+Figure~\ref{fig:graphical_model}.
+
+ In the corresponding mathematical generative description of HSLDA to follow, $K$ is the number of LDA
 ``topics'' (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$
@@
 function that takes the value $1$ if its argument is true and $0$ otherwise.
+
+\subsection{Ancestral sampling for HSLDA}
+
 The following procedure describes how to generate from the HSLDA generative
@@
 %the generative process is as follows: 
-
+\begin{shortbox}
+\Boxhead{HSLDA Ancestral Sampler}
 \begin{enumerate}
@@
 \begin{itemize}
-\item Draw a label application coefficient $\boldsymbol\eta_{l}\mid\mu,\sigma\sim\mathcal{N}_{K}(\mu \mathbf{1}_K,\sigma \mathbf{I}_{K})$  
+\item Draw a label weight vector $\boldsymbol\eta_{l}\mid\mu,\sigma\sim\mathcal{N}_{K}(\mu \mathbf{1}_K,\sigma \mathbf{I}_{K})$  
 \end{itemize}
@@
 \begin{itemize}
-\item Draw $a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\eta_{l},y_{\mathrm{pa}(l),d}\sim\begin{cases}
-\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1), & y_{\mathrm{pa}(l),d}=1\\
-\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l),d}=-1\end{cases}$ %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
+\item Draw 
+\begin{eqnarray}\lefteqn{a_{l,d}\mid \bar{\mathbf{z}}_d,\boldsymbol\eta_{l},y_{\mathrm{pa}(l),d}}\nonumber \\&\sim&\begin{cases}
+\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1), & y_{\mathrm{pa}(l),d}=1 \\
+\mathcal{N}(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l},1)\mathbb{I}(a_{l,d}<0), & y_{\mathrm{pa}(l),d}=-1 \end{cases}\label{zero_0}\end{eqnarray} %\item Draw $a_{l, d} \ | \ z_{1:N_d,d}, \boldsymbol\beta_l \sim \mathcal{N} \left(\bar z_d^{T} \boldsymbol\beta_{l},1\right)$, where $\bar z_d=N_d^{-1}\sum_{n=1}^{N_d}z_{n,d}$
 
-\item Apply label $l$ to document $d$ according to $a_{l,d}$ \[\hspace{-2cm}
+\item Apply label $l$ to document $d$ according to $a_{l,d}$ \begin{eqnarray}\hspace{-2.2cm}
 y_{l,d}\mid a_{l,d}=\begin{cases}
-\ \ \ 1 & \text{if \ensuremath{a_{l,d}>0}} \\% and \ensuremath{y_{{\rm \mathrm{pa}}(l),d}=1}}\\
--1 & \text{otherwise}\end{cases}\]
+\;\;\, 1 & \text{if \ensuremath{a_{l,d}>0}}  \\% and \ensuremath{y_{{\rm \mathrm{pa}}(l),d}=1}}\\
+ -1 & \text{otherwise} \end{cases}\label{zero_1}\end{eqnarray}
 
@@
 \end{enumerate}
+\end{shortbox}
 
@@
 
-The second half of step 4 is a substantial part of our contribution to the
-general class of supervised LDA models.
-Here, each document is labeled generatively using a hierarchy of conditionally dependent probit regressors \cite{gelmanbda04}.
+The second half of step 4 is what is referred to as supervision in the supervised LDA literature.
+This is where the hierarchical classification of the bag-of-words data takes place and the is-a label constraints are enforced.
 For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.
-Note that label $y_{l,d}$ can only be applied to document $d$ if its parent label $\mathrm{pa}(l)$ is also applied (these expressions are specific to is-a constraints but can be modified to accommodate different constraints).
+Note that $y_{l,d}$ can only be applied to document $d$  (set to 1) if its parent label $\mathrm{pa}(l)$ was also applied (these expressions are specific to is-a constraints but could be modified to accommodate different constraints between labels).
 The regression coefficients $\boldsymbol\eta_l$ are independent a priori, however, the hierarchical coupling in this model induces a posteriori dependence.
-The net effect of this is that label predictors deeper in the label hierarchy are able to focus on finding specific, conditional labeling features.
-We believe this to be a significant source of the empirical label prediction improvement we observe experimentally.
-We test this hypothesis in Section~\ref{sec:experiments}.
+The net effect of this conditional hierarchy of profit regressors is that label predictors deeper in the label hierarchy are able to focus on finding features that distinguish the members of that set, conditioned on the fact that the members are in the same set.
+It should be clear that one can trivially restrict this hierarchy to a depth one hierarchy.
+Also, one can nearly as easily make the conditional classification at each node multi-class rather than single-class if more than one label at each node is required.
+In many cases, however, a binary indicator along with a deeper or more complex tree is sufficient.
+%We believe this to be a significant source of the empirical label prediction improvement we observe experimentally.  We test this hypothesis in Section~\ref{sec:experiments}.
 
@@
 This simplifies posterior inference substantially.
+A review of probit regression can be found near the end of this chapter in Section \ref{sec:probit_review}.
 
@@
 
-In the common case where no negative labels are observed (like the example applications we consider in Section~\ref{sec:experiments}), the model must be explicitly biased towards generating data that has negative labels in order to keep it from learning to assign all labels to all documents.
+In the common case where no negative labels are observed (like the example applications we consider in Section~\ref{sec:experiments}), the model must be explicitly biased towards generating  negative labels in order to keep it from learning to assign positive labels to all documents.
 This is a common problem in modeling unbalanced data.
-To see how this model can be biased in this way we draw the reader's attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.
+To see how this model can achieve this we draw the reader's attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.
 Because $\bar{\mathbf{z}}_d$ is always positive, setting $\mu$ to a negative value results in a bias towards negative labelings, i.e.~for large negative values of $\mu$, all labels become a priori more likely to be negative ($y_{l,d}=-1$).
-We explore the ability of $\mu$ to bias out-of-sample label prediction performance in Section~\ref{sec:experiments}.
+We explore the effect of $\mu$ on out-of-sample label prediction performance in Section~\ref{sec:experiments}.
+In a very real way $\mu$ is a knob that can be adjusted both before inference to induce a broad array of out-of-sample performance characteristics that vary along classical axes like specificity, recall, and accuracy.
+A similar but less principled solution can be effected by changing the decision boundary from $0$ in \eqref{zero_0} and \eqref{zero_1}.
+This technique can be used to vary out-of-sample label bias after learning.
 

### 2012-02-24 Krantz-v3/Chapters/chapter1/probit.tex
Added booke chapter.

\begin{figure}[htbp]
\begin{center}
\includegraphics{Chapters/chapter1/probit_no_aux_vars}
\caption{Probit model, no auxiliary variables}
\label{fig:probit_no_aux_vars}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics{Chapters/chapter1/cdf}
\caption{CDF of $N(0,1)$}
\label{fig:cdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics{Chapters/chapter1/probit}
\caption{Probit model}
\label{fig:probit}
\end{center}
\end{figure}

For reasons that are somewhat obscure to me, statisticians tend to use probit regression for binary classification whereas machine learners tend to use logistic regression.  In a recent paper, my collaborators and I found it useful for computational purposes to use probit regression.  One can find many good primers on probit regression around the web, but, as we all know, there is almost always space for another.

Figure \ref{fig:cdf} is the cumulative distribution function (cdf) of a $N(0,1)$ distribution.  The ``probit'' function is the inverse of the normal cdf.    The normal cdf function, denoted $\Phi(x;\mu,\sigma^2)$ with $\mu$ the mean, $\sigma^2$ the variance and $x$ the argument, is actually much more relevant in this context.

The range of the normal cdf is $(0,1)$ which means that it can be interpreted as a probability.  For instance, one can construct a generalized linear model (a ``probit regression model'') of the form 
\begin{equation}
P(y_i = 1) = \Phi(x_i^T\beta; 0, \sigma^2). \label{eqn:probit}
\end{equation}  Depending on convention (i.e. binary $y_i$ represented as $\{1,0\}$ or $\{1,-1\}$) then the probability of $y_i$ being labeled the opposite way is $P(y_i = -1)$ or $P(y_i = 0) = 1 - P(y_i=1).$  Here $x_i$ is a vector of covariates, $\beta$ is a vector of weights, and $y_i$ is a single, binary valued response.  As always, the close relationship between regression and classification are in full display here : probit regression is a ``generalized linear {\em regression} model'' and a ``binary classifier.''

Figure \ref{fig:probit_no_aux_vars} shows the graphical model for probit regression (minus any priors on the regression weights $\beta$).    In this model we would like to use labeled training data, $\{x_i, y_i\}_{i=1}^N$ to ``learn'' the value of $\beta$ and then to use this value to predict the value of $y_{N+1} | x_{N+1}, \beta.$  We will be being Bayesian about our inference so here, what we really mean is, we will average over the posterior distribution of $\beta$ when making predictions.  This means that we want to draw samples from the posterior distribution of $\beta | \{x_i, y_i\}_{i=1}^N$.  This brings us to Figure \ref{fig:probit} which introduces a set of auxiliary variables $\{u_i\}_{i=1}^N$.  In this note we will demonstrate that the model in  Figure \ref{fig:probit} is the same as the model in Figure \ref{fig:probit_no_aux_vars}  when the $u_i$'s are marginalized out and will suggest that inference in the former is computationally easier.

First -- what is an auxiliary variable?  It is a variable introduced into a model in order to make inference easier but whose existence does not change the distribution of interest.  Auxiliary variables for slice sampling are one particularly clever use of auxiliary variables.  The auxiliary variable trick in probit regression is another.

For the purposes of exposition, let's forget about the $i$ index for a second and focus on a single instance $y$, $x$, and $u$.  The argument we make will hold for all by simply reintroducing subscripts.

To start, let's write down the joint distribution of these quantities (according to the graphical model that includes auxiliary variables).

\begin{equation}
P(y,x,u) = P(y|u)P(u|x,\beta) \label{eqn:joint}
\end{equation}

Clearly, straight away, one can see precisely why this auxiliary variable scheme works.  By the law of total probability we have
\begin{equation}
P(y,x) = \int P(y,x,u)du= \int P(y|u)P(u|x,\beta)du.  \label{eqn:joint_marginalized}
\end{equation}
If by MCMC we generate $S$ samples $\{u^{(s)},y^{(s)},x^{(s)}\}_{s=1}^S \sim P(y,x,u)$ we know that marginalizing $u$ out (i.e.~disregarding its value) we get samples $\{y^{(s)},x^{(s)}\}_{s=1}^S \sim P(y,x).$  %Equation  \ref{eqn:joint_marginalized} 
%\begin{eqnarray}
%y_i &\sim& \Phi(x_i^T\beta; 0, 1)
%\end{eqnarray}

We haven't specified the most important part of the auxiliary variable sampling scheme yet, namely, what $P(y|u)$ and what $P(u|x,\beta)$ are.   Let's try $y = \mbox{sign}(u)$ and $u \sim N(x^T\beta,\sigma^2).$  These choices are nice in a particular way.  First let's verify that the marginalization of $u$ out of this model results in the model specification in Equation \ref{eqn:probit}.

\begin{eqnarray*}
P(y=1|x,\beta) &=&  \int P(y=1|u)P(u|x,\beta)du\\
&=& \int \mathbb{I}(u>0)N(u;x^T\beta,\sigma^2) du \\
&=& \int_0^\infty N(u;x^T\beta,\sigma^2) du \\
&=& 1 - \Phi(0; x^T\beta,\sigma^2) \\ 
&=& \Phi(x^T\beta; 0, \sigma^2)
\end{eqnarray*}
where the last line comes from the fact that for symmetric distributions like the normal distribution, 
$\Phi(x^T\beta; 0, \sigma^2) = 1 - \Phi(-x^T\beta; 0, \sigma^2)$ and the mean of a normal cdf can be translated arbitrarily, i.e. $\Phi(-x^T\beta; 0, \sigma^2) = \Phi(0;x^T\beta, \sigma^2)$ (which comes from adding the offset $x^T\beta$ to the cdf argument and mean).

OK.  So, now, we have established the fact that for a particular sort of auxiliary variable choice, we get the same probit model as we wanted.  Why is this choice nice?

Well, it comes down to sampling $\beta$ and $u$ and $y$.  Generally, sampling $\beta$ in the model without auxiliary variables will require hybrid Monte Carlo (HMC) or Metropolis Hastings of some sort.  Gibbs sampling often comes with substantial benefits.  By making this choice of auxiliary variable the conditional distribution of $u_i$ given everything else is proportional to a truncated Normal distribution, a distribution that is, by nature of its commonness, relatively straightforward to sample from.  The big benefit, though, acrues from the posterior form for sampling $\beta$.  With the $u$'s ``observed'' (as they would be in a Gibbs sampler), the posterior distribution of $\beta$ (for typical choices of prior) is precisely the same as that for linear regression, perhaps the most well studied model in statistics.  Sampling $\beta$ from its posterior distribution typically is quite simple; certainly more so that sampling $\beta$ without the $u$ auxiliary variables.

### 2012-02-24 Krantz-v3/Chapters/chapter1/related_work.tex
Added booke chapter.

This chapter covers classifying or, equivalently, labeling bag-of-words data.  We show how to construct classifiers based on features derived from mixed membership models.  We specifically develop a solution to the concrete problem of hierarchically classifying text documents.  The model and techniques developed in this chapter can be used to solve any labeling or hierarchical classification problem involving data represented as bags-of-words.  If, for instance,  an image is represented by counts of features then it is said to be in bag-of-words representation.   So, for instance, hierarchical classification of images might be tackled using the techniques developed in this chapter, although, such a particular application will not be explicitly considered.

We should agree that classification and regression are largely the same.  Namely, classification is regression with discrete output.  To solve classification problems we must learn from training instances, as in regression, a function $f : \mathcal{X} \rightarrow \mathcal{L}$ which maps from an input space $\mathcal{X}$ to an output (label) space $\mathcal{L}$.   As an example consider the problem of identifying the authors of a set of documents for which the true author is unknown but is one of a set of possible authors.   For such a problem the input space is the richly structured space of possible texts and the output space is the large but largely unstructured list of possible authors.  Training data would consist of text author pairs for texts known to have been authored by authors in the list of possible authors.

To solve such a problem one would need to choose a set of features to extract from the underlying documents. Then, using those features and the training data, a classification function would be learned.  In this case one might choose a multi-class logistic or probit regression function and learn parameters for it from the labeled training instances.   In this chapter we will focus on models that use mixed-membership representations of bag-of-word data as the input features to joint models that can be used, via conditioning, for classification or regression.

The models that we propose in this chapter generalize the author classification example in that the label spaces we will consider are also richly structured (in addition to the input space) and the regression function we learn maps from from the input space to labels in a structured, typically hierarchical label space.  In the same way that mixed membership models are surprisingly effective, the resulting hierarchical classification models seem to be too.

\subsection{Background}

Mixed membership models, particularly latent Dirichlet allocation (LDA) \cite{Blei2003}, have been reviewed in other chapters.   The key property we exploit for purposes of classification is that LDA % is a generative
%probabilistic model of corpora consisting of documents that are bag-of-words.  In particular LDA 
provides a way to extract a latent, low-dimensional representation of each bag-of-words observation consisting of mixing proportions over topic vectors.  Each topic is a frequency histogram of terms.  Each document is a mixture of topics.  This mixture of topics vector is the feature vector we exploit for purposes of document classification.

%Also known as topic models, these models
%infer the latent structure, or topics, of documents in a corpus. Each
%document is represented as a collection of words, generated from a
%set of topic assignments (one for each word), where each topic assignment
%is drawn from a distribution over topics \cite{Blei2003}.

The idea for using this representation for classification was first considered in a paper on so-calling ``supervised'' latent Dirichlet allocation (SLDA).  SLDA built on LDA by incorporating
``supervision'' in the form of an observed exponential family response 
variable per document.  Another way of saying this is that SLDA was the first document classifier for documents in the bag-of-words representation that exploited the LDA topic feature representation.  
%A subtle point should be noted here. 
 SLDA and the subject of this chapter, a generalization of it called hierarchically supervised LDA (HSLDA) \cite{Perotte-2012-NIPS} are not, strictly speaking, regression models.  This is because they model the {\em joint} distribution of labels and bag-of-words data rather than the conditional distribution of the labels given the input.   Having the joint distribution allows us to condition on an input to produce a label for all test instances.  Additionally, during training of the model, knowing the labels effects the learning of the low dimensional representation of the input -- a characteristic discussed later in this chapter.
% This is quite different (and better performing in practice) than LASSO
%(L1 regularized least squares regression) and LDA followed by least
%squares regression \cite{BleiMcAuliffe2008}.
%By performing joint inference, SLDA and HSLDA
%infer topics such that the model predicts the response variable while
%improving word likelihood.

%In this chapter we extend supervised latent Dirichlet allocation (SLDA)
%\cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision. SLDA
%is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per
%document ``supervision,'' often taking the form of a single numerical or
%categorical label. It has been demonstrated that the signal provided by such
%supervision can result in better, task-specific document models and can also
%lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.

%It also has been demonstrated that SLDA has been shown to outperform both LASSO
%(L1 regularized least squares regression) and LDA followed by least squares
%regression~\cite{BleiMcAuliffe2008}.

The problem of applying structured labels to bag-of-words data can be tackled in a number of different ways.  For instance,  SLDA can be used to solve this kind of problem directly; however, doing so requires ignoring the hierarchical
dependencies amongst the labels. In Section~\ref{sec:experiments} we contrast
HSLDA with SLDA applied in this way. 
Other models that incorporate LDA and supervision and so could be used to solve this problem include
LabeledLDA~\cite{Ramage2009} and DiscLDA~\cite{DiscLDA}.  Various applications of these models to 
computer vision and document networks have been explored~\cite{wangbleifeifei08,RelationalLDA}.
None of these models, however, leverage dependency structure in the label space.
%While there has been much work in multi-label classification of text and
%text modeling in general, we focus here on topic modeling approaches.
%Latent Dirichlet allocation (LDA) is a generative probabilistic model which
%represents documents as a mixed-membership bag of word. Each document is
%represented as a collection of words, generated from a set of topic assignments
%(one for each word), where each topic assignment is drawn from a distribution
%over topics~\cite{Blei2003}. SLDA is latent Dirichlet allocation (LDA)
%\cite{Blei2003} augmented with per-document labeling, often taking the form of
%a single numerical or categorical label. Examples of labels include ratings
%associated with online reviews, grades for essays, and the number of times a
%webpage is linked. This approach has been shown to outperform both LASSO
%(L1 regularized least squares regression) and LDA followed by least
%squares regression~\cite{BleiMcAuliffe2008}.

In other non-LDA-based related work, researchers have classified documents into a hierarchy (a closely related task) using naive Bayes classifiers
and support vector machines. Most of this work has been demonstrated on relatively
small datasets, small label spaces, and has focused on single label classification without
a model of documents such as LDA~\cite{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.  The HSLDA model we cover here is not limited in this way.

%One set of models that are particularly relevant to HSLDA are Chang
%and Blei's hierarchical models for document networks (Relational Topic
%Models). In that family of models, they encountered a similar scenario
%where the lack of a link did not truly indicate absence. In hierarchically
%labeled data, negative labels are uncommon and the lack of a label
%in the hierarchy is not equivalent to a negative label. Therefore,
%as in the work of Chang and Blei, we employ regularization to account
%for the lack of negative labels. This will be discussed further in

% Noemie: check out what I wrote here, because I am not really sure this is
% correct. 
%There have been several models that incorporate both latent models of text and
%some form of
%supervision~\cite{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}. One set 
%of models that are particularly relevant to HSLDA are Chang and Blei's
%hierarchical models for document networks (Relational Topic Models). In that
%family of models, they encountered a similar scenario where an unselected label 
%does not always indicate absence. In hierarchical labels, this phenomenon is
%even more pervasive -- there are no explicit negative labels, but it is also
%unclear how to treat the parents of selected labels. Like in the work of Chang
%and Blei, we employ regularization to account for the lack of negative
%labeling. In our experiments, we look at the impact of assigning positive and
%negative instances to the ancestors of selected labels.

### 2012-02-24 Krantz-v3/Chapters/symbollist.tex
Added booke chapter.

\thispagestyle{empty}
\begin{symbollist}{000000}
\symbolentry{$\alpha$}{To solve the generator maintenance scheduling, in the  past, several mathematical techniques have  been applied.}
\symbolentry{$\sigma^2$}{These include integer programming, integer linear programming, dynamic programming, branch and bound etc.}
\symbolentry{$\sum$}{Several heuristic search algorithms have also been developed. In recent years expert systems,}
\symbolentry{$abc$}{fuzzy approaches, simulated annealing and genetic algorithms have also been tested.}
\symbolentry{$\theta\sqrt{abc}$}{This paper presents a survey of the literature}
\symbolentry{$\zeta$}{ over the past fifteen years in the generator}
\symbolentry{$\partial$}{maintenance scheduling. The objective is to}
\symbolentry{sdf}{present a clear picture of the available recent literature}
\symbolentry{ewq}{of the problem, the constraints and the other aspects of}
\symbolentry{bvcn}{the generator maintenance schedule.}
\end{symbollist}

### 2012-02-27 Krantz-v3/Chapters/chapter1/experiments.tex
Added clinical document and product examples

@@
 \begin{quote}
-{ADLER Dischage summary text description snippet }
+{History of Present Illness: Mrs.
+Carmen Sandiego is a 62-year-old female with a past medical history significant for diabetes, hypertension, hyperlipidemia, afib, status post MI in 5/2010 and cholecystectomy in 3/2009.
+The patient presented to the ED on 7/11/2011 with a right sided partial facial hemiparesis along with mild left arm weakness.
+The patient was admitted to the Neurology service and underwent a workup for stroke given her history of MI and many cardiovascular risk factors ...}
 \end{quote} 
@@
 \begin{quote}
-{ADLER DVD description }
+{Winner of five Academy Awards, including Best Picture and Best Director, The Deer Hunter 
+is simultaneously an audacious directorial conceit and one of the greatest films ever made 
+about friendship and the personal impact of war. Like Apocalypse Now, it's hardly a conventional 
+battle film--the soldier's experience was handled with greater authenticity in Platoon--but its 
+depiction of war on an intimate scale packs a devastatingly dramatic punch ... }
 \end{quote}

### 2012-02-28 Krantz-v3/Chapters/chapter1/experiments.tex
touched all kinds of good stuff... ready for adler's final contributions.

@@
 %\subfloat[][\label{fig:1a}]{\includegraphics[width=.49\textwidth]{figs/amazon_pred_varying_mu}}
-\subfigure[][Clinical data performance.]{\label{fig:1a}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/clin_pred_varying_mu}}
+\subfigure[Clinical data performance.]{\label{fig:1a}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/clin_pred_varying_mu}}
 %\subfloat[][\label{fig:1c}]{\includegraphics[width=.49\textwidth]{figs/sens_comparison_leafs}}
-\subfigure[][Retail product performance.]{\label{fig:1b}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/amazon_pred_varying_mu}}
+\subfigure[Retail product performance.]{\label{fig:1b}\includegraphics[width=.48\textwidth]{Chapters/chapter1/figs/amazon_pred_varying_mu}}
 %\subfigure[][]{\label{fig:clinical_roc}\includegraphics[scale=0.4]{figs/ROC_comparison_leafs}}

### 2012-02-28 Krantz-v3/Chapters/chapter1/mcmc.tex
touched all kinds of good stuff... ready for adler's final contributions.

Markov chain Monte Carlo (MCMC) is the name given to a particular technique for generating samples from distributions of interest.  In Bayesian analyses this distribution is often the posterior distribution of the latent quantities given the data and fixed parameters.  The technique is general, elegant, and simple.  The basic idea is that by using Markov chain transition kernel templates, it is possible to construct and simulate a Markov chain whose equilibrium distribution is the distribution of interest.   This means that  the sequence of Markov chain simulation outputs can be used as an empirical stand-in approximation to the distribution of interest.  What is more, general MCMC techniques don't rely upon having a normalized probability distribution as a starting point, merely a function that is at all points proportional to the distribution of interest.   This is particularly convenient in the Bayesian case where the posterior distribution is of interest and does not have an analytic form.  In this case, the posterior distribution is simply proportion to the joint distribution with the observed quantities fixed to their observed values.

Let $\mathbf{z}$ be a set of latent variables.  Let $\mathbf{x}$ be a set of observed quantities.  We could denote by $\theta$ any set of fixed parameters and include it on the conditioning side of every expression to follow in this section but, since it would appear in every expression we'll omit it for purposes of notational clarity.  Our goal is to generate samples distributed according to $P(\mathbf{z} | \mathbf{x})$ (one can think about sampling from a posterior distribution here, however, MCMC applies more generally).  We have a model whose joint distribution is specified and can be evaluated readily for all settings of $\mathbf{z}$ and $\mathbf{x}$.  By Bayes rule we know that

\[P(\mathbf{z} | \mathbf{x} = (x_1, \ldots, x_N) ) \propto P(\mathbf{z}, \mathbf{x}= (x_1, \ldots, x_N)).\]

We denote by $\tilde{P}(\mathbf{z} | \mathbf{x} = (x_1, \ldots, x_N) )$ the normalized version of $P(\mathbf{z} | \mathbf{x} = (x_1, \ldots, x_N) )$.  Assume that we can come up with a function $P(\mathbf{z}^* | \mathbf{z})$ that has the following property (called )

\[P(\mathbf{z*} | \mathbf{x} = (x_1, \ldots, x_N) ) = \sum_ \mathbf{z} P(\mathbf{z}^* | \mathbf{z}) P(\mathbf{z} | \mathbf{x} = (x_1, \ldots, x_N) )\]

### 2012-02-28 Krantz-v3/Chapters/chapter1/probit.tex
touched all kinds of good stuff... ready for adler's final contributions.

@@
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics{Chapters/chapter1/probit_no_aux_vars}
-\caption{Probit model, no auxiliary variables}
-\label{fig:probit_no_aux_vars}
-\end{center}
-\end{figure}
+%\begin{figure}[htbp]
+%\begin{center}
+%\includegraphics{Chapters/chapter1/probit_no_aux_vars}
+%\caption{Probit model, no auxiliary variables}
+%\label{fig:probit_no_aux_vars}
+%\end{center}
+%\end{figure}
+%
+%%
+%%\begin{figure}[htbp]
+%%\begin{center}
+%%\includegraphics{Chapters/chapter1/cdf}
+%%\caption{CDF of $N(0,1)$}
+%%\label{fig:cdf}
+%%\end{center}
+%%\end{figure}
+%
+%\begin{figure}[htbp]
+%\begin{center}
+%\includegraphics{Chapters/chapter1/probit}
+%\caption{Probit model}
+%\label{fig:probit}
+%\end{center}
+%\end{figure}
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics{Chapters/chapter1/cdf}
-\caption{CDF of $N(0,1)$}
-\label{fig:cdf}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics{Chapters/chapter1/probit}
-\caption{Probit model}
-\label{fig:probit}
-\end{center}
-\end{figure}
-
-For reasons that are somewhat obscure to me, statisticians tend to use probit regression for binary classification whereas machine learners tend to use logistic regression.
+%For reasons that are somewhat obscure to me, statisticians tend to use probit regression for binary classification whereas machine learners tend to use logistic regression.
 In a recent paper, my collaborators and I found it useful for computational purposes to use probit regression.
 One can find many good primers on probit regression around the web, but, as we all know, there is almost always space for another.
+%
+%Figure \ref{fig:cdf} is the cumulative distribution function (cdf) of a $N(0,1)$ distribution. 
+%
+For reasons that are somewhat obscure, statisticians tend to use probit regression for binary classification whereas machine learners tend to use logistic regression.
+ The ``probit'' function is the inverse of the normal cumulative distribution function (cdf).
+We denote the normal {\em cdf} function $\Phi(x;\mu,\sigma^2)$ with $\mu$ the mean, $\sigma^2$ the variance and $x$ the argument.
 
-Figure \ref{fig:cdf} is the cumulative distribution function (cdf) of a $N(0,1)$ distribution.
-The ``probit'' function is the inverse of the normal cdf.
-The normal cdf function, denoted $\Phi(x;\mu,\sigma^2)$ with $\mu$ the mean, $\sigma^2$ the variance and $x$ the argument, is actually much more relevant in this context.
-
-The range of the normal cdf is $(0,1)$ which means that it can be interpreted as a probability.  For instance, one can construct a generalized linear model (a ``probit regression model'') of the form 
+The range of the normal cdf is $(0,1)$ which means that it can be interpreted as a probability.
+For instance, one can construct a generalized linear classification model (a ``probit regression model'') of the form
 \begin{equation}
@@
 binary $y_i$ represented as $\{1,0\}$ or $\{1,-1\}$) then the probability of $y_i$ being labeled the opposite way is $P(y_i = -1)$ or $P(y_i = 0) = 1 - P(y_i=1).$  Here $x_i$ is a vector of covariates, $\beta$ is a vector of weights, and $y_i$ is a single, binary valued response.
-As always, the close relationship between regression and classification are in full display here : probit regression is a ``generalized linear {\em regression} model'' and a ``binary classifier.''
+As always, the close relationship between regression and classification are in full display here : probit regression is a ``generalized linear {\em regression} model'' as well as a ``binary classifier.''
 
-Figure \ref{fig:probit_no_aux_vars} shows the graphical model for probit regression (minus any priors on the regression weights $\beta$).
-In this model we would like to use labeled training data, $\{x_i, y_i\}_{i=1}^N$ to ``learn'' the value of $\beta$ and then to use this value to predict the value of $y_{N+1} | x_{N+1}, \beta.$  We will be being Bayesian about our inference so here, what we really mean is, we will average over the posterior distribution of $\beta$ when making predictions.
+%Figure \ref{fig:probit_no_aux_vars} shows the graphical model for probit regression (minus any priors on the regression weights $\beta$).   
+ In this model we would like to use labeled training data, $\{x_i, y_i\}_{i=1}^N$ to ``learn'' the value of $\beta$ and then to use this value to predict the value of $y_{N+1} | x_{N+1}, \beta.$  If we are being Bayesian about our inference here what we really mean is that we will average over the posterior distribution of $\beta$ when making predictions.
 This means that we want to draw samples from the posterior distribution of $\beta | \{x_i, y_i\}_{i=1}^N$.
-This brings us to Figure \ref{fig:probit} which introduces a set of auxiliary variables $\{u_i\}_{i=1}^N$.
-In this note we will demonstrate that the model in  Figure \ref{fig:probit} is the same as the model in Figure \ref{fig:probit_no_aux_vars}  when the $u_i$'s are marginalized out and will suggest that inference in the former is computationally easier.
+To to this efficiently one can introduce a set of auxiliary variables $\{u_i\}_{i=1}^N$.
+ %In this note we will demonstrate that the model in  Figure \ref{fig:probit} is the same as the model in Figure \ref{fig:probit_no_aux_vars}  when the $u_i$'s are marginalized out and will suggest that inference in the former is computationally easier.
 
-First -- what is an auxiliary variable?
-It is a variable introduced into a model in order to make inference easier but whose existence does not change the distribution of interest.
+By auxiliary variable we mean that such variables will be used as an intermediary for purposes of efficiency but will otherwise be uninteresting.
+They are variables introduced into a model in order to make inference easier but whose existence does not change the distribution of interest.
 Auxiliary variables for slice sampling are one particularly clever use of auxiliary variables.
@@
 
-For the purposes of exposition, let's forget about the $i$ index for a second and focus on a single instance $y$, $x$, and $u$.
-The argument we make will hold for all by simply reintroducing subscripts.
+For the purposes of exposition, forget about the $i$ index for a bit and focus on a single instance $y$, $x$, and $u$.  The argument we make will hold for all by simply reintroducing subscripts.
 
-To start, let's write down the joint distribution of these quantities (according to the graphical model that includes auxiliary variables).
+To start, let's propose a factorized joint distribution for these quantities
 
@@
 
-Clearly, straight away, one can see precisely why this auxiliary variable scheme works.  By the law of total probability we have
+Straight away, one can see why this auxiliary variable scheme works.  By the law of total probability we have
 \begin{equation}
@@
 \end{equation}
-If by MCMC we generate $S$ samples $\{u^{(s)},y^{(s)},x^{(s)}\}_{s=1}^S \sim P(y,x,u)$ we know that marginalizing $u$ out (i.e.~disregarding its value) we get samples $\{y^{(s)},x^{(s)}\}_{s=1}^S \sim P(y,x).$  %Equation  \ref{eqn:joint_marginalized}
+So, if by some means we generate $S$ samples $\{u^{(s)},y^{(s)},x^{(s)}\}_{s=1}^S \sim P(y,x,u)$ we know that marginalizing $u$ out (i.e.~disregarding its value) we get samples $\{y^{(s)},x^{(s)}\}_{s=1}^S \sim P(y,x).$  %Equation  \ref{eqn:joint_marginalized}
 %\begin{eqnarray}
@@
 
-OK.  So, now, we have established the fact that for a particular sort of auxiliary variable choice, we get the same probit model as we wanted.  Why is this choice nice?
+Having established the fact that for a particular sort of auxiliary variable choice, we get the same probit model as we wanted, why is this choice nice?
 
@@
 With the $u$'s ``observed'' (as they would be in a Gibbs sampler), the posterior distribution of $\beta$ (for typical choices of prior) is precisely the same as that for linear regression, perhaps the most well studied model in statistics.
-Sampling $\beta$ from its posterior distribution typically is quite simple; certainly more so that sampling $\beta$ without the $u$ auxiliary variables.
+In that case, sampling $\beta$ from its posterior distribution is quite simple usually; certainly more so that sampling $\beta$ without the $u$ auxiliary variables.
+
+The extension to the multivariate HSLDA setting is straightforward and follows this line of reasoning precisely.
+An extended discussion of the techniques suggested here and the multivariate generalization can be found in \cite{gelmanbda04}.

### 2012-02-28 Krantz-v3/Chapters/chapter1/related_work.tex
touched all kinds of good stuff... ready for adler's final contributions.

@@
 The model and techniques developed in this chapter can be used to solve any labeling or hierarchical classification problem involving data represented as bags-of-words.
-If, for instance,  an image is represented by counts of features then it is said to be in bag-of-words representation.
-So, for instance, hierarchical classification of images might be tackled using the techniques developed in this chapter, although, such a particular application will not be explicitly considered.
+For instance,  if an image is represented by counts of features then it can be described as being in a bag-of-words representation.
+So, hierarchical classification of images might be tackled using the techniques developed in this chapter, although, such a particular application will not be explicitly considered.
 
@@
 To solve classification problems we must learn from training instances, as in regression, a function $f : \mathcal{X} \rightarrow \mathcal{L}$ which maps from an input space $\mathcal{X}$ to an output (label) space $\mathcal{L}$.
-As an example consider the problem of identifying the authors of a set of documents for which the true author is unknown but is one of a set of possible authors.
+For example, consider the problem of identifying the authors of a set of documents for which the true author is unknown but is one of a set of possible authors.
 For such a problem the input space is the richly structured space of possible texts and the output space is the large but largely unstructured list of possible authors.

### 2012-02-29 Krantz-v3/Chapters/chapter1/experiments.tex
Added figures and captions to the book chapter and added tables of word lists

@@
 {}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
-pneumonia,'' where the former is a type of the latter.  A representative sub-tree of the ICD-9 code tree is shown in Figure ADLER\ref{ADLER}. It is worth noting that
+pneumonia,'' where the former is a type of the latter.  A representative sub-tree of the ICD-9 code tree is shown in Figure \ref{fig:diagnosis_tree}. It is worth noting that
 the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
@@
 mistakes~\cite{FarzandipourEtAl10}.
+
+\begin{figure}[t]
+%tbp] %  figure placement: here, top, bottom, or page
+\centering \includegraphics[scale=0.4]{Chapters/chapter1/figures/hi_quality_ICD9_tree_cropped} \caption{An illustration of a portion of the ICD9 hierarchy.}
+\label{fig:diagnosis_tree}
+\end{figure}
 
@@
 
+\begin{figure}[t]
+%tbp] %  figure placement: here, top, bottom, or page
+\centering \includegraphics[scale=0.4]{Chapters/chapter1/figures/hi_quality_product_tree_cropped} \caption{An illustration of a portion of the Amazon product hierarchy.}
+\label{fig:product_tree}
+\end{figure}
+
 Amazon.com is one such retailer.  Their product categorization data is available as part of the 
- Stanford Network Analysis Platform (SNAP) dataset~\cite{SNAP}.    A representative sub-tree of the amazon.com DVD produce categorization tree is shown in Figure ADLER\ref{ADLER}.  
+ Stanford Network Analysis Platform (SNAP) dataset~\cite{SNAP}.    A representative sub-tree of the amazon.com DVD produce categorization tree is shown in Figure \ref{fig:product_tree}.  
 Product descriptions were obtained separately from the
@@
 
-Example topics (as word lists) learned for the discharge data are given below.  These word lists are computed by ADLER
+Example topics (as word lists) learned for the discharge data are given below.  These word lists are computed by sorting
+terms in decreasing order based on their probability under a given topic.
 
 \begin{extract}
-ADLER
+\begin{tabular}{|c|c|}
+\hline
+\textbf{Topic 1} & \textbf{Topic 2} \\ \hline
+MASS & WOUND \\
+\hline
+CANCER & FOOT \\
+\hline
+RIGHT & CELLULITIS \\
+\hline
+BREAST & ULCER \\
+\hline
+CHEMOTHERAPY & LEFT \\
+\hline
+METASTATIC & ERYTHEMA \\
+\hline
+LEFT & PAIN \\
+\hline
+LYMPH & SWELLING \\
+\hline
+TUMOR & SKIN \\
+\hline
+BIOPSY & RIGHT \\
+\hline
+CARCINOMA & ABSCESS \\
+\hline
+LUNG & LEG \\
+\hline
+CHEMO & OSTEOMYELITIS \\
+\hline
+ADENOCARCINOMA & TOE \\
+\hline
+NODE & DRAINAGE \\
+\hline
+\end{tabular}
 \end{extract}
-
-ADLER could you write a sentence about how these are cleaner than the SLDA or LDA topics?
 
@@
 
-Example topics (as word lists) learned for the amazon.com data are given below.  T
+Example topics (as word lists) learned for the amazon.com data are given below.  These word lists were also computed by sorting
+terms in decreasing order based on their probability under a given topic.
 
 \begin{extract}
-ADLER
+\begin{tabular}{|c|c|}
+\hline
+\textbf{Topic 1} & \textbf{Topic 2} \\ \hline
+SERIES & BASEBALL \\
+\hline
+EPISODES & TEAM \\
+\hline
+SHOW & GAME \\
+\hline
+SEASON & PLAYERS \\
+\hline
+EPISODE & BASKETBALL \\
+\hline
+FIRST & SPORT \\
+\hline
+TELEVISION & SPORTS \\
+\hline
+SET & NEW \\
+\hline
+TIME & PLAYER \\
+\hline
+TWO & SEASON \\
+\hline
+SECOND & LEAGUE \\
+\hline
+ONE & FOOTBALL \\
+\hline
+CHARACTERS & STARS \\
+\hline
+DISC & FANS \\
+\hline
+GUEST & FIELD \\
+\hline
+\end{tabular}
 \end{extract}

### 2012-03-01 Krantz-v3/Chapters/chapter1/experiments.tex
fixed a few typos

@@
 Combined inference has been shown to improve performance in SLDA
-\cite{BleiMcAuliffe2008}. This comparison model doesn't examine the value of utilizing the structured nature 
+\cite{BleiMcAuliffe2008}. This comparison model does not examine the value of utilizing the structured nature 
 of the label space, instead it highlights the benefit of combined inference over both the

### 2012-03-01 Krantz-v3/Chapters/chapter1/introduction.tex
touched all kinds of good stuff... ready for adler's final contributions.
fixed a few typos

@@
 retail domains. These results provide evidence that the two views (text and labels) mutually benefit  multi-label
-classification.%In fact, it will be shown that the hierarchical supervised model outperforms a
+classification. %In fact, it will be shown that the hierarchical supervised model outperforms a
 That is, modeling the joint  is better than learning 
@@
 Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
-Section~\ref{sec:related_work} reviews related work, and
+and
 Section~\ref{sec:experiments} shows results from applying HSLDA to health care

### 2012-03-01 Krantz-v3/Chapters/chapter1/related_work.tex
fixed a few typos

@@
 
-The models that we propose in this chapter generalize the author classification example in that the label spaces we will consider are also richly structured (in addition to the input space) and the regression function we learn maps from from the input space to labels in a structured, typically hierarchical label space.
+The models that we propose in this chapter generalize the author classification example in that the label spaces we will consider are also richly structured (in addition to the input space) and the regression function we learn maps from the input space to labels in a structured, typically hierarchical label space.
 In the same way that mixed membership models are surprisingly effective, the resulting hierarchical classification models seem to be too.

### 2012-03-01 Krantz-v3/Chapters/chapter1/experiments.tex
INITIAL SUBMISSION VERSION

@@
 It consists of 6,000 discharge summaries and
-their associated ICD-9 codes (7,298 distinct codes overall), representing all
+their associated ICD-9 codes (7,298 distinct codes overall), representing a portion of
 the discharges from the hospital in 2009. All included discharge summaries had associated ICD-9 Codes.
@@
 Amazon.com is one such retailer.  Their product categorization data is available as part of the 
- Stanford Network Analysis Platform (SNAP) dataset~\cite{SNAP}.    A representative sub-tree of the amazon.com DVD produce categorization tree is shown in Figure \ref{fig:product_tree}.  
+ Stanford Network Analysis Platform (SNAP) dataset~\cite{SNAP}.    A representative sub-tree of the amazon.com DVD product category tree is shown in Figure \ref{fig:product_tree}.  
 Product descriptions were obtained separately from the
@@
 
-In each case the gold standard for testing was derived from the test data. To make the comparison as antagonistic to HSLDA as possible (relative to the other models), ancestors of
+In each case the gold standard for testing was derived from the test data. To make the comparison as antagonistic to HSLDA as possible (relative to the other models), in evaluation only, ancestors of
 observed nodes in the label hierarchy were ignored, observed nodes were
@@
 HSLDA model where LDA and the regressions were fit separately had a true
-positive rate of 0.39 and a false positive rate of 0.08. These points are highlighted in Figure~\ref{fig:1a}.
+positive rate of 0.39 and a false positive rate of 0.08.
+These points are highlighted in Figure~\ref{fig:1a}.
+Note that the figure is somewhat misleading because for any one value of $\mu$ HSLDA outperforms the comparison models by a relatively large margin.
 
@@
 \hline
-\textbf{Topic 1} & \textbf{Topic 2} \\ \hline
+\textbf{Topic 1} & \textbf{Topic 2} \\ \hline \hline
 MASS & WOUND \\
@@
 
+These topics closely correspond to common clinical concepts, namely, cancers of the thorax and wounds common to diabetics suffering from poor peripheral circulation.
+Evaluations of the subject coherence of these topics relative to baselines are ongoing, but early results suggest positive findings similar to those reported for other supervised LDA models.
+
 Figure~\ref{fig:1b} demonstrates the performance of the model on the retail product data as an ROC curve also varying $\mu$. For
@@
 \hline
-\textbf{Topic 1} & \textbf{Topic 2} \\ \hline
+\textbf{Topic 1} & \textbf{Topic 2} \\ \hline \hline
 SERIES & BASEBALL \\

### 2012-03-01 Krantz-v3/Chapters/chapter1/inference.tex
touched all kinds of good stuff... ready for adler's final contributions.
fixed a few typos
INITIAL SUBMISSION VERSION

@@
 This is usually the case for posterior distributions of non-trivial
-probabilistic models and suggests a approximating the posterior distribution by sampling.
+probabilistic models and suggests approximating the posterior distribution by sampling.
 
 In this section we derive the conditional distributions required to sample from the HSLDA posterior distribution using Markov chain Monte Carlo.
-A very brief review of Markov chain Monte Carlo can be found in Section \ref{sec:mcmc} at the end of this chapter.
-
+%A very brief review of Markov chain Monte Carlo can be found in Section \ref{sec:mcmc} at the end of this chapter.
+%
 The HSLDA sampler, like the collapsed Gibbs samplers for LDA \cite{Griffiths04}, is itself a collapsed Gibbs sampler in which all of the latent variables that can be analytically marginalized are.
@@
 %The conditional posterior distribution of the latent topic indicators is
- We use the notation $\mathbf{z}_{-(n,d)}$ to
-denote $\mathbf{z}_d\backslash z_{n,d}$. %, marginalizing
+ %We use the notation $\mathbf{z}_{-(n,d)}$ to
+%denote $\mathbf{z}_d\backslash z_{n,d}$. %, marginalizing
 %$\theta_{1:D}$ and $\phi_{1:K}$. For details regarding collapsing
@@
 Here $\mathcal{L}_{d}$ is the set of labels which are observed for document $d$.
-We sample from \eqref{eq:z-likelihood}  by enumeration and implicit normalization.
+We sample from \eqref{eq:z-likelihood}  by first enumerating $z_{n,d}$ and then normalizing.
 %

### 2012-03-01 Krantz-v3/Chapters/chapter1/probit.tex
INITIAL SUBMISSION VERSION

@@
 binary $y_i$ represented as $\{1,0\}$ or $\{1,-1\}$) then the probability of $y_i$ being labeled the opposite way is $P(y_i = -1)$ or $P(y_i = 0) = 1 - P(y_i=1).$  Here $x_i$ is a vector of covariates, $\beta$ is a vector of weights, and $y_i$ is a single, binary valued response.
-As always, the close relationship between regression and classification are in full display here : probit regression is a ``generalized linear {\em regression} model'' as well as a ``binary classifier.''
+The close relationship between regression and classification are in full display here : probit regression is a ``generalized linear {\em regression} model'' as well as a ``binary classifier.''
 
 %Figure \ref{fig:probit_no_aux_vars} shows the graphical model for probit regression (minus any priors on the regression weights $\beta$).   
- In this model we would like to use labeled training data, $\{x_i, y_i\}_{i=1}^N$ to ``learn'' the value of $\beta$ and then to use this value to predict the value of $y_{N+1} | x_{N+1}, \beta.$  If we are being Bayesian about our inference here what we really mean is that we will average over the posterior distribution of $\beta$ when making predictions.
+ In this model we would like to use labeled training data, $\{x_i, y_i\}_{i=1}^N$ to ``learn'' the value of $\beta$ and then to use this value to predict the value of $y_{N+1} | x_{N+1}, \beta.$  Being Bayesian about inference means that we will average over the posterior distribution of $\beta$ when making predictions.
 This means that we want to draw samples from the posterior distribution of $\beta | \{x_i, y_i\}_{i=1}^N$.

### 2012-03-01 Krantz-v3/Chapters/chapter1/related_work.tex
INITIAL SUBMISSION VERSION

@@
-This chapter covers classifying or, equivalently, labeling bag-of-words data.
-We show how to construct classifiers based on features derived from mixed membership models.
+This chapter covers classifying and labeling bag-of-words data.  %We show how to construct classifiers based on features derived from mixed membership models.  
 We specifically develop a solution to the concrete problem of hierarchically classifying text documents.
@@
 
-We should agree that classification and regression are largely the same.
-Namely, classification is regression with discrete output.
+%We should agree that classification and regression are largely the same.  Namely, classification is regression with discrete output.  
 To solve classification problems we must learn from training instances, as in regression, a function $f : \mathcal{X} \rightarrow \mathcal{L}$ which maps from an input space $\mathcal{X}$ to an output (label) space $\mathcal{L}$.
@@
 For such a problem the input space is the richly structured space of possible texts and the output space is the large but largely unstructured list of possible authors.
-Training data would consist of text author pairs for texts known to have been authored by authors in the list of possible authors.
+Training data would consist of text/author pairs for texts known to have been authored by authors in the list of possible authors.
 

### 2012-11-16 Krantz-v3/Chapters/chapter1/HSLDA.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

--- 2012-02-24 Krantz-v3/Chapters/chapter1/model.tex
@@
+% !TEX root =  ../../Krantz_style.tex
+
 \label{sec:model} 
@@
 
-HSLDA is a model for hierarchically, multiply-labeled, bag-of-word data.  We
-call individual groups of bag-of-word data documents (although they could be any data represented by bags of features).  Let $w_{n,d}
+This model (HSLDA) is designed to fit hierarchically, multiply-labeled, bag-of-word data.  We
+call groups of bag-of-words data documents (unordered words in text documents, bag of visual feature representations of images, etc.).  Let $w_{n,d}
 \in \Sigma$ be the $n$th observation in the $d$th document.  Let $\mathbf{w}_d
@@
 $d$.  Let there be $D$ such documents and let the size of the vocabulary be
-$V=|\Sigma|$.  Let the set of labels be $\mathcal{L}=\left\{
-  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $. Each label
+$V=|\Sigma|$.
+
+Let the set of labels be $\mathcal{L}=\left\{
+  l_{1},l_{2},\ldots,l_{\left|\mathcal{L}\right|}\right\} $.
+A label in HSLDA corresponds to a node in the graphical model in Figure~\ref{fig:graphical_model}.
+A label can either be observed or unobserved.
+Documents can be multiply labeled, meaning that subsets of label nodes in the graphical model in Fig.~\ref{fig:graphical_model} can have observed values.
+Each label
 $l \in \mathcal{L}$, except the root, has a parent $\mathrm{pa}(l) \in \mathcal{L}$
@@
 Such a label hierarchy forms a multiply rooted tree.
+Confused readers may wish to consult Figure \ref{fig:diagnosis_tree} or Figure \ref{fig:product_tree} each of which is a  label-tree graphical model.
+
+ In a label forest a node may be observed (the label was ``applied'' and, for instance, was observed to have value 1), unobserved and unknown, or unobserved and constrained to be either -1 or 1 by the structure of the label space.
 Without loss of generality we will consider a tree with a single root $r\in\mathcal{L}$.
-Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label is applied to document $d$ or not.
-In most cases $y_{i,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.
+Each document has a variable $y_{l,d} \in \{-1,1\}$ for every label which indicates whether the label applies to document $d$ or not.
+In most cases $y_{l,d}$ will be unobserved, in some cases we will be able to fix its value because of  constraints on the label hierarchy, and in the relatively minor remainder its value will be observed.
 In the applications we demonstrate, only positive labels are observed.
@@
 
-The constraints imposed by an is-a label hierarchy are that if the $l$th label is applied to document $d$, i.e.,~$y_{l,d} = 1$, then all labels in the label hierarchy up to the root are also applied to document $d$, i.e.,~$y_{\mathrm{pa}(l),d} = 1, y_{\mathrm{pa}(\mathrm{pa}(l)),d} = 1, \ldots, y_{r,d}=1.$  Conversely, if a label $l'$ is marked as not applying to a document then no descendant of that label may be applied to the same.
+The constraints imposed by an is-a label hierarchy are that if the $l$th label is applied to document $d$, i.e.,~$y_{l,d} = 1$, then all labels in the label hierarchy up to the root are also applied to document $d$, i.e.,~$y_{\mathrm{pa}(l),d} = 1, y_{\mathrm{pa}(\mathrm{pa}(l)),d} = 1, \ldots, y_{r,d}=1.$  Conversely, if a label $l'$ is marked as not applying to a document (i.e.~$y_{l',d}=-1$) then no descendant label of that label can take value 1.
 We assume that at least one label is applied to every document.
@@
 Label responses are modeled using a conditional
-hierarchy of probit regressors. The HSLDA graphical model is given in
+hierarchy of probit regressors and will be discussed next. The full HSLDA graphical model is given in
 Figure~\ref{fig:graphical_model}.
 
- In the corresponding mathematical generative description of HSLDA to follow, $K$ is the number of LDA
+\subsection{Generative Model}
+
+In the following box  the HSLDA generative model  is given for the ``is-a hierarchy'' set of label constraints.   In the box and what follows in this chapter, $K$ is the number of LDA
 ``topics'' (distributions over the elements of $\Sigma$), $\boldsymbol\phi_k$
@@
 function that takes the value $1$ if its argument is true and $0$ otherwise.
-
-\subsection{Ancestral sampling for HSLDA}
-
-The following procedure describes how to generate from the HSLDA generative
-model.
 
@@
 \begin{shortbox}
-\Boxhead{HSLDA Ancestral Sampler}
+\Boxhead{HSLDA Generative Model}
 \begin{enumerate}
@@
 \end{enumerate}
+\label{box:generative_model}
 \end{shortbox}
@@
 For every label $l \in \mathcal{L}$, both the empirical topic distribution for document $d$ and whether or not its parent label was applied (i.e.~$\mathbb{I}(y_{\mathrm{pa}(l),d}=1)$) are used to determine whether or not  label $l$ is to be applied to  document $d$ as well.
+\eqref{zero_0} and \eqref{zero_1} comprise a probit regression model in an auxiliary variable formulation (see Section \ref{sec:probit_review}).
+Note that in the case that the parent label is applied, i.e.~$y_{\mathrm{pa}(l),d}=1,$ the child label $y_{l,d}$ is applied with probability $P(\bar{\mathbf{z}}^{T}_d\boldsymbol\eta_{l}>0).$  This is a conditional probit regression model for classification where $\boldsymbol\eta_l$ are the class-conditional regression parameters.
+The auxiliary variables $a_{l,d}$ make inference tractable but are not fundamental to the model -- only the labels and regression parameters are actually of interest.
+
 Note that $y_{l,d}$ can only be applied to document $d$  (set to 1) if its parent label $\mathrm{pa}(l)$ was also applied (these expressions are specific to is-a constraints but could be modified to accommodate different constraints between labels).
-The regression coefficients $\boldsymbol\eta_l$ are independent a priori, however, the hierarchical coupling in this model induces a posteriori dependence.
-The net effect of this conditional hierarchy of profit regressors is that label predictors deeper in the label hierarchy are able to focus on finding features that distinguish the members of that set, conditioned on the fact that the members are in the same set.
-It should be clear that one can trivially restrict this hierarchy to a depth one hierarchy.
+Note that multiple labels can be applied to the same document.
+The regression coefficients $\boldsymbol\eta_l$ which generate the labels are independent a priori, however, the hierarchical coupling in this model and conditional label dependency structure induces a posteriori dependence.
+The net effect of this conditional hierarchy of profit regressors is that child label predictors deeper in the label hierarchy are able to focus on finding features that distinguish label paths in the tree, conditioned on the fact that all the children of any particular node are by design members of some more general parent set.
+It should be clear that one can trivially restrict this hierarchy to a depth one hierarchy recovering SLDA with probit link and univariate categorical labels.
 Also, one can nearly as easily make the conditional classification at each node multi-class rather than single-class if more than one label at each node is required.
@@
 
-Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations.
+Note that the choice of variables $a_{l,d}$ and how they are distributed were  driven at least in part by posterior inference efficiency considerations (see Appendix \ref{sec:probit_review}).
 In particular, choosing probit-style auxiliary variable distributions for the $a_{l,d}$'s  yields conditional posterior distributions for both the auxiliary variables \eqref{eqn:a_l_d} and the regression coefficients \eqref{eqn:regression_param_conditional} which are analytic.
@@
 
-In the common case where no negative labels are observed (like the example applications we consider in Section~\ref{sec:experiments}), the model must be explicitly biased towards generating  negative labels in order to keep it from learning to assign positive labels to all documents.
-This is a common problem in modeling unbalanced data.
+\subsection{Dealing with Label Imbalance}
+
+In the common case where no negative labels are observed (like the example applications we consider in Section~\ref{sec:example_applications}), the model must be explicitly biased towards generating  negative labels in order to keep it from learning to only assign positive labels to all documents.
+This is a common problem in modeling with unbalanced labels.
 To see how this model can achieve this we draw the reader's attention to the $\mu$ parameter and, to a lesser extent, the $\sigma$ parameter above.
 Because $\bar{\mathbf{z}}_d$ is always positive, setting $\mu$ to a negative value results in a bias towards negative labelings, i.e.~for large negative values of $\mu$, all labels become a priori more likely to be negative ($y_{l,d}=-1$).
-We explore the effect of $\mu$ on out-of-sample label prediction performance in Section~\ref{sec:experiments}.
+We explore the effect of $\mu$ on out-of-sample label prediction performance in Section~\ref{sec:example_applications}.
 In a very real way $\mu$ is a knob that can be adjusted both before inference to induce a broad array of out-of-sample performance characteristics that vary along classical axes like specificity, recall, and accuracy.

### 2012-11-16 Krantz-v3/Chapters/chapter1/discussion.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

@@
+% !TEX root =  ../../Krantz_style.tex
+
+\section{Related Work}
+\label{sec:related_work}
+
+HSLDA does not, of course, stand alone.  Models for structured labeling of bag of words data can be designed in a number of different ways.
+
+As shown  in Section~\ref{sec:example_applications},  SLDA can be used to solve this kind of problem directly; however, doing so requires ignoring the hierarchical
+dependencies amongst the labels.  Other models that incorporate LDA and supervision and so also could be used to solve this problem include
+LabeledLDA~\cite{Ramage2009} and DiscLDA~\cite{DiscLDA}.  Various applications of these models to 
+computer vision and document networks have been explored~\cite{wangbleifeifei08,RelationalLDA}.
+None of these models, however, leverage dependency structure in the label space.
+
+In other non-LDA-based related work, researchers have classified documents into hierarchies (a closely related task) using naive Bayes classifiers
+and support vector machines. Most of this work has been demonstrated on relatively
+small datasets, small label spaces, and has focused on single label classification without
+a model of documents such as LDA~\cite{mccallum99building,Dumais2000HCW,Kollerilprints291,Chakrabarti1998SFS}.  
+\section{Discussion}
+\label{sec:discussion}
+
 %Despite the proliferation of LDA related models in recent years, one could
@@
 can be expressed as label prediction problems for bag-of-word data. A
-surprisingly large amount of that kind of data possess structured labels,
+surprisingly large amount data possess structured labels,
 either hierarchically constrained or otherwise.  HSLDA directly addresses
@@
 
+There are many kinds of problems that have the same characteristics as this: any data  that consists of free text that has been partially or
+completely categorized by human editors for instance; more fully
+% In this paper we focus on
+any bag-of-words data that has been, at least in part, 
+categorized. 
+ Examples include but are not limited to webpages and curated
+hierarchical directories of the same \cite{DMOZ}, product descriptions and
+catalogs, (e.g.~\cite{AMAZON} as available from \cite{SNAP})
+and patient records %hospital treatment transcripts
+and diagnosis codes assigned to them for bookkeeping and insurance purposes.
+The model we cover in this chapter shows one way to combine 
+these two sources of information into a single model that allows one to
+categorize new text documents automatically, suggest labels that might be
+inaccurate, compute improved similarities between documents for information
+retrieval purposes, and more. %The models and techniques that we develop in
+%this paper are applicable to other data as well, namely, any unstructured
+%representations of data that have been hierarchically classified (e.g.,~image
+%catalogs with bag-of-feature representations).
+
+%There are several challenges entailed in incorporating a hierarchy of labels
+%into the model. Among them, given a large set of potential labels (often thousands), each
+%instance has only a small number of labels associated to it. Furthermore, there are no
+%naturally occurring negative labeling in the data, and the absence of a label
+%cannot always be interpreted as a negative labeling.
+
+%Our work operates within the framework of topic modeling. Our approach learns
+%topic models of the underlying data and labeling strategies in a joint model,
+%while leveraging the hierarchical structure of the labels. 
+%In this chapter we focus on ``is-a'' hierarchically structured label spaces, but label spaces that are structured differently may be accommodated with relatively minor modifications.
+
 Extensions to this work include unbounded topic cardinality variants and

### 2012-11-16 Krantz-v3/Chapters/chapter1/example_applications.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

--- 2012-03-01 Krantz-v3/Chapters/chapter1/experiments.tex
@@
-Having described HSLDA and given details on how to do inference in it, we turn to demonstrations of using HSLDA to solve problems from  two different domains: predicting medical
-diagnosis codes from hospital discharge summaries and predicting product
-categories from Amazon.com product descriptions.
+% !TEX root =  ../../Krantz_style.tex
+
+%Having described HSLDA and given details on how to do inference in it, we turn to demonstrations of using HSLDA to solve problems from  two different domains: predicting medical
+%diagnosis codes from hospital discharge summaries and predicting product
+%categories from Amazon.com product descriptions.
 
@@
 regressors (hierarchical constraints on labels ignored,  i.e. the regression is not conditional) and HSLDA fit by first
-performing LDA then fitting tree-conditional regressions (rather than jointly inferring the topics and the regression coefficients). These models were
+performing LDA then fitting probit regressors that respect the conditional label hierarchy (rather than jointly inferring the topics and the regression coefficients). These models were
 chosen because they are the strongest available competitors and because they  highlight several pedagogical aspects of HSLDA including performance in the
@@
 scale parameter of 1000.
+Different values of $K$ corresponding to different numbers of topics were explored, however, the results that we show in the following are not substantially changed in character.
+As is usual in mixed-membership models, there is an ideal number of topics that should be used for out-of-sample prediction tasks, however, a full model-selection search varying topic cardinality was not performed for these datasets.
 
@@
 correct labels at a cost of an increase in the number of false positives
-relative to the comparison models.
+relative to the comparison models.  However, as shown in Figure~\ref{fig:1a}  HSLDA outperforms or no worse than the comparison models across across the full range of specificities.
 

### 2012-11-16 Krantz-v3/Chapters/chapter1/inference.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

@@
 Our inference goal  is to obtain a representation of the posterior distribution
-of the latent variables in the model.  The posterior distribution we seek does not have a simple analytic form from which exact samples can be drawn.
+of the latent variables in the model.
+This posterior distribution can then be used for predictive inference of labels for held-out documents among other things.
+Unfortunately the posterior distribution we seek does not have a simple analytic form from which exact samples can be drawn.
 This is usually the case for posterior distributions of non-trivial
@@
 The HSLDA sampler, like the collapsed Gibbs samplers for LDA \cite{Griffiths04}, is itself a collapsed Gibbs sampler in which all of the latent variables that can be analytically marginalized are.
-Among others, the parameters $\boldsymbol{\phi}_{1:K}$
-and $\boldsymbol{\theta}_{1:D}$ are analytically marginalized prior to deriving the following conditional distributions for sampling.  
-  It will often be the
-case that the set of labels $\mathcal{L}$ is not fully observed for
-every document. We will define $\mathcal{L}_{d}$ to be the subset
-of labels which have been observed for document $d$. In an is-a hierarchical regression  it is straightforward
-to marginalize the variables $a_{l^{\prime},d}$ and $y_{l^{\prime},d}$
-for $l^{\prime}\in\mathcal{L}\backslash\mathcal{L}_{d}$ simply by ignoring them. 
+Among others, the topic distributions $\boldsymbol{\phi}_{1:K}$
+and document specific topic assignment distributions $\boldsymbol{\theta}_{1:D}$ are analytically marginalized prior to deriving the following conditional distributions for sampling.
+
+  It will usually be the
+case that values $y_{l,d}$ will not be known for all labels $l \in \mathcal{L}$ in the space of possible labels.
+Values for $y_{l,d}$ that are enforced by label constraints and observed labels are set to their constrained values prior to inference and treated as observed.
+We will define $\mathcal{L}_{d}$ to be the subset
+of labels which have been observed (or observed via filling in from constraints) for document $d$.   Marginalizing the probit regression auxiliary variables $a_{l^{\prime},d}$ and $y_{l^{\prime},d}$
+for $l^{\prime}\in\mathcal{L}\backslash\mathcal{L}_{d}$ is simple in the is-a hierarchy case because they can simply be ignored. 
 %We can also integrate out the parameters $\boldsymbol{\phi}_{1:K}$
 %and $\boldsymbol{\theta}_{1:D}$ as in \cite{Griffiths04}.
- The remaining latent variables (those that are not collapsed out) are $\mathbf{z}=\{z_{1:N_{d},d}\}_{d=1,\ldots,D},\boldsymbol\eta=\{\boldsymbol\eta_{l}\}_{l\in\mathcal{L}},\mathbf{a}=\{a_{l^{\prime},d}\}_{l^{\prime}\in\mathcal{L}_{d},d=1,\ldots,D},\boldsymbol\beta,\alpha,\alpha^{\prime}$
+ The remaining latent variables (those that are not collapsed out) are the topic indicators $\mathbf{z}=\{z_{1:N_{d},d}\}_{d=1,\ldots,D},$ the probit regression parameters $\boldsymbol\eta=\{\boldsymbol\eta_{l}\}_{l\in\mathcal{L}},$ the auxiliary variables $\mathbf{a}=\{a_{l^{\prime},d}\}_{l^{\prime}\in\mathcal{L}_{d},d=1,\ldots,D},$ the global topic proportions $\boldsymbol\beta$, and the concentration parameters $\alpha,\alpha^{\prime}$
 and $\gamma$.
@@
 \subsection{Gibbs Sampler}
- Let $\mathbf{a}$ be the set of all auxiliary variables, $\mathbf{w}$  the set of all words, $\boldsymbol\eta$  the set of all regression coefficients, and  $\mathbf{z}\backslash z_{n,d}$  the set $\mathbf{z}$ with element $z_{n,d}$ removed.
+ Let $\mathbf{a}$ be the set of all probit regression auxiliary variables, $\mathbf{w}$  the set of all words, $\boldsymbol\eta$  the set of all regression coefficients, and  $\mathbf{z}\backslash z_{n,d}$  the set $\mathbf{z}$ with element $z_{n,d}$ removed.
 %The conditional posterior distribution of the latent topic indicators is
@@
 distribution of $a_{l,d}$ is a truncated normal
-distribution~\cite{Albert_Chib_1993}. 
+distribution~\cite{Albert_Chib_1993} (see also Section \ref{sec:probit_review}). 
 \begin{eqnarray*}
@@
 \boldsymbol\beta\mid\mathbf{z},\alpha^{\prime},\alpha\sim{\rm Dir}\left(m_{\left(\cdot\right),1}+\alpha^{\prime},m_{\left(\cdot\right),2}+\alpha^{\prime},\ldots,m_{\left(\cdot\right),K}+\alpha^{\prime}.\right)\end{equation}
-Here $m_{d,k}$ are auxiliary variables that are required to sample the posterior distribution of $\boldsymbol\beta$.  Their conditional posterior distribution is sampled according to
+Here $m_{d,k}$ are additional auxiliary variables that are introduced by the direct assignment method to sample the posterior distribution of $\boldsymbol\beta$.
+Their conditional posterior distribution is sampled according to
  \begin{equation}
 p\left(m_{d,k}=m\mid\mathbf{z},\mathbf{m}_{-\left(d,k\right)},\boldsymbol\beta\right)=\frac{\Gamma\left(\alpha\boldsymbol\beta_{k}\right)}{\Gamma\left(\alpha\boldsymbol\beta_{k}+c^k_{(\cdot),d}\right)}s\left(c^k_{(\cdot),d},m\right)\left(\alpha\boldsymbol\beta_{k}\right)^{m}\end{equation}
-where $s\left(n,m\right)$ represents stirling numbers of the first
+where $s\left(n,m\right)$ denotes Stirling numbers of the first
 kind.
-
 %\subsection{$p\left(\alpha\right)$, $p\left(\alpha'\right)$, $p\left(\gamma\right)$}
-
 The hyperparameters $\alpha$, $\alpha^{\prime}$, and $\gamma$ are
@@
 sampled
-using Metropolis-Hastings.
+using Metropolis-Hastings. \newline
+
+It remains now to show that HSLDA works.  To do so we demonstrate results from modeling real-world datasets in the clinical and web
+retail domains. These results provide evidence that the two views (text and labels) mutually benefit  multi-label
+classification. %In fact, it will be shown that the hierarchical supervised model outperforms a
+That is, modeling the joint  is better than learning 
+topic models and hierarchical classifiers
+independently.

### 2012-11-16 Krantz-v3/Chapters/chapter1/introduction.tex
First full edit pass in prep. for resubmission.  Handing to Adler for a review.

--- 2011-09-08 NYAS_abstract/introduction.tex
@@
-There exist many sources of unstructured data that have been partially or
-completely categorized by human editors.  In this paper, we focus on
-unstructured text data that has been, at least in part, manually
-categorized.  Examples include but are not limited to webpages and curated
-hierarchical directories of the same \citep{DMOZ}, product descriptions and
-catalogs, %(e.g.~\citep{AMAZON} as available from \citep{SNAP})
-and patient records %hospital treatment transcripts
-and diagnosis codes assigned to them for bookkeeping and insurance purposes.
-In this work we show how to combine 
-these two sources of information using a single model that allows one to
-categorize new text documents automatically, suggest labels that might be
-inaccurate, compute improved similarities between documents for information
-retrieval purposes, and more. The models and techniques that we develop in
-this paper are applicable to other data as well, namely, any unstructured
-representations of data that have been hierarchically classified (e.g.,~image
-catalogs with bag-of-feature image representations).
+% !TEX root =  ../../Krantz_style.tex
 
-There are several challenges entailed in incorporating a hierarchy of labels
-into the model. Among them, given a large set of potential labels (often thousands), each
-instance has only a small number of labels associated to it. Furthermore, there are no
-naturally occurring negative labeling in the data, and the absence of a label
-cannot always be interpreted as a negative labeling.
+Documents frequently come with additional information like labels or popularity ratings.
+Contemporary examples include the product ratings that  accompany product descriptions, the number of ``likes'' that webpages have attracted, grades associated with assigned essays, and so forth.
 
-Our work operates within the framework of topic modeling. Our approach learns
-topic models of the underlying data and labeling strategies in a joint model,
-while leveraging the hierarchical structure of the labels. For the sake of
-simplicity, we focus on is-a hierarchies, but the model can be
-applied to other structured label spaces. We extend supervised latent Dirichlet
-allocation (sLDA)~\cite{BleiMcAuliffe2008} to take advantage of hierarchical
-supervision. We propose an efficient way to incorporate hierarchical
-information into the model. We hypothesize that the context of labels within
-the hierarchy provides valuable information about labeling.
+This chapter covers one way to jointly model documents and the labels applied to them.
+In particular we focus on our own work on modeling documents with more complicated labels that themselves possess some kind of structural organization.
+Consider typical product catalogs.
+They usually contain text  descriptions of products that have been organized into hierarchical product directories.
+The situation of a product into such a hierarchy (the path or paths in the product hierarchy that lead to it) can be thought of as a structured label.
+Jointly modeling the document and such a label is useful for automatically labeling new documents (corresponding in this example to automatically situating a new product in the product directory) and more.
 
-We demonstrate our model on large, real-world datasets in the clinical and web
-retail domains. We observe that hierarchical information is valuable when
-incorporated into the learning and improves our primary goal of multi-label
-classification. Our results show that a joint, hierarchical model outperforms a
-classification with unstructured labels as well as a disjoint model, where the
-topic model and the hierarchical classification are inferred
-independently of each other.
+Collections of hierarchically labelled  documents abound, text and otherwise.
+We will consider hierarchically labelled  patient clinical records in later sections.
+Applications like situating web-pages in hierarchical link directories are left for others to explore.
 
-The remainder of this paper is as follows. Section~\ref{sec:model}
-introduces hierarchically supervised LDA (HSLDA), while
-Section~\ref{sec:inference} details a sampling approach to inference in HSLDA. 
-Section~\ref{sec:related_work} reviews related work, and
-Section~\ref{sec:experiments} shows results from applying HSLDA to health care
-and web retail data.
+%
+%This chapter covers classifying and labeling bag-of-words data.  %We show how to construct classifiers based on features derived from mixed membership models.  
+%We specifically develop a solution to the concrete problem of hierarchically classifying text documents.
+The model and techniques developed in this chapter can be used to solve any labeling or hierarchical classification problem involving data represented as bags-of-words.
+For instance,  if an image is represented by counts of features then it can be described as being in a bag-of-words representation.
+So, hierarchical classification of images might be tackled using the techniques developed in this chapter, although, such a particular application will not be explicitly considered.
+%
+%
+%
+%
+%%We should agree that classification and regression are largely the same.  Namely, classification is regression with discrete output.  
+%To solve classification problems we must learn from training instances, as in regression, a function $f : \mathcal{X} \rightarrow \mathcal{L}$ which maps from an input space $\mathcal{X}$ to an output (label) space $\mathcal{L}$.
+For example, consider the problem of identifying the authors of a set of documents for which the true author is unknown but is one of a set of possible authors.
+For such a problem the input space is the richly structured space of possible texts and the output space is the large but largely unstructured list of possible authors.
+Training data would consist of text/author pairs for texts known to have been authored by authors in the list of possible authors.
+%
+Because text documents are notorious difficult to directly model we take an approach common to other chapters in this book.
+We use a mixed-membership model of the text document to represent the document as bag-of-words drawn from a document specific mixture of topic distributions.
+The modeling choices we make in relating this representation to structured labels follows, as too its relationship to prior art.
 
-%In this paper we describe the use of a topic model based on supervised
-%latent Dirichlet allocation (sLDA) to identify topics within narrative
-%discharge summaries and to automate the assignment of diagnostic codes,
-%specifically International Classification of Disease 9th Revision,
-%Clinical Modification (ICD-9-CM) codes. There are a number of advantages
-%to this approach. First, manually coding diagnoses is a time-consuming
-%and notoriously unreliable process. Many diagnoses are omitted in
-%the final record, and a high error rate is found even in the principal
-%diagnoses \citep{Surjan1999}.
+%solve our problem one would need to choose a set of features to extract from the underlying documents.
+In this chapter we will focus on models that use mixed-membership representations of bag-of-word data as the input features to joint models that can be used, via conditioning, for classification or regression.
 
-%Our main contribution is to show how to utilize supervision in the form of
-%hierarchical and (often) multiple labelings in a similar manner. Consider web
-%retail data. Web retailers often have both a browse-able product hierarchy and
-%free-text descriptions for all products they sell. The situation of each
-%product in a product hierarchy (often multiply situated) constitutes a
-%multiple, hierarchical labeling of the free-text product descriptions. We
-%hypothesize that such hierarchical labels should, at least in theory, provide
-%better supervision than the simpler unstructured labels previously considered.
-%Results from applying our model to both medical record and web retail data
-%suggests that this is likely the case. In particular, we observe gains in our
-%primary goal of out-of-sample label prediction that result specifically from
-%leveraging hierarchical supervision.
+%The models that we propose in this chapter generalize the author classification example in that the label spaces we will consider are also richly structured (in addition to the input space) and the regression function we learn maps from the input space to labels in a structured, typically hierarchical label space.
+In the same way that mixed membership models are surprisingly effective, the resulting hierarchical classification models seem to be too.
 
-%\begin{itemize}
-%\item Benefits of combining human categorization information into ``topic models''
-%\item LDA solved free text
-%\item supervised LDA improves LDA (extra info) and allows new inference (predict links, etc.)
-%\item amazon, freshdirect, netflix, dmoz, pandora (music genome)
-%\end{itemize}
+\subsection{Background}
 
-% Informatics journal paper
+Mixed membership models, including the model upon which we build, latent Dirichlet allocation (LDA) \cite{Blei2003}, have been reviewed in other chapters.
+The key property we exploit for purposes of classification is that LDA % is a generative
+%probabilistic model of corpora consisting of documents that are bag-of-words.  In particular LDA 
+provides a way to extract a latent, low-dimensional representation of text and other documents consisting of the  frequency of word assignments to the topics that are assumed to have generated them.
+A topic is a distribution over words.
+Each document is a bag of words drawn from a document specific mixture of topics.
+%This mixture of topics vector is the feature vector we exploit for purposes of document classification.
+
+%Also known as topic models, these models
+%infer the latent structure, or topics, of documents in a corpus. Each
+%document is represented as a collection of words, generated from a
+%set of topic assignments (one for each word), where each topic assignment
+%is drawn from a distribution over topics \cite{Blei2003}.
+
+Building a joint model of documents and labels using this representation is not new.  It was first introduced by Blei and McAuliffe in a paper on 
+ ``supervised'' latent Dirichlet allocation (SLDA) \cite{BleiMcAuliffe2008}.  SLDA built on LDA by incorporating
+``supervision'' in the form of an observed exponential family response 
+variable per document.
+
+\subsubsection{Latent Dirichlet Allocation}
+
+To explain both SLDA and to set the stage for our work it helps to introduce our notation for LDA.
+Assume that there are $K$ topics.
+Let $\boldsymbol\phi_{k}\sim{\rm Dir}_{V}(\gamma\mathbf{1}_V)$ be ``topic''  $k$, i.e.~a distribution over a finite set of words.
+Here $V$ simply labels the variables to indicate that they have to do with the vocabulary.
+The vector $\mathbf{1}_V$ consists of all ones and has length equal to the size of the vocabulary.
+The distribution ${\rm Dir}$ is the Dirichlet distribution.
+The constant $\gamma$ controls the smoothness of the inferred topics.
+Larger values lead to smoother topic estimates.
+Let $\boldsymbol\beta\mid\alpha'\sim{\rm Dir}_{K}\left(\alpha^{\prime}\mathbf{1}_K\right)$  be a ``global'' distribution over topics where $K$ indicates that the distribution and variable sizes are equal to the number of topics $K$ and $\alpha^{\prime}$ controls the relative proportion of topics globally, large $\alpha^\prime$ leading to all topics being roughly responsible for the same number of words.
+Intuitively $\boldsymbol\beta$ is something like the average topic proportion independent of any particular document.
+Per document $d$ topic distributions $\boldsymbol\theta_d\mid\boldsymbol\beta,\alpha\sim{\rm Dir}_{K}\left(\alpha\boldsymbol\beta\right)$ are modeled as being deviations from the global distribution over topics where larger values of $\alpha$ result in all documents' topic distributions being more similar.
+We will use  $z_{n,d}\mid\boldsymbol\theta_d\sim{\rm Multinomial}(\boldsymbol\theta_d)$ to indicate which topic generated the $n$th word of document $d$.
+Drawing $n$th word in document $d$ from the  indicated topic $w_{n,d}\mid z_{n,d},\boldsymbol\phi_{1:K}\sim{\rm Multinomial}(\boldsymbol\phi_{z_{n,d}})$ completes our notation of standard LDA.
+
+\subsubsection{Supervised Latent Dirichlet Allocation}
+
+Supervised LDA adds another per document observation, a label $y_d$.
+It is modeled as being generated by a generalized linear model (GLM) $y_d | \bar{{\mathbf z}}_d, \boldsymbol\eta,\delta \sim \mathrm{GLM}(\bar{\mathbf{z}}_d,\boldsymbol\eta,\delta)$.
+Brushing aside exponential family and GLM link function generalities, what SLDA does is regress labels against the empirical distributions of assigned topic indicator variables $\bar{\mathbf{z}}_d = \{\bar{z}_{1,d}, \ldots, \bar{z}_{K,d}\}$ where $\bar{z}_{k,d} = \frac{\sum_n \mathbb{I}(z_{n,d}=k)}{\sum_n \sum_j \mathbb{I}(z_{n,d}=j)}$ is the fraction of words assigned to topic $k$ ($\mathbb{I}(\cdot)$ is the indicator function that returns one if its argument is true).
+If the document labels are real valued then one example choice for the regression relationship would be $y_d | {\bar{\mathbf z}}_d, \boldsymbol\eta \sim \mathcal{N}(\bar{\mathbf{z}}_d^T\boldsymbol\eta,\delta).$  Using an generalized linear model in the exponential family to parameterize this regression relationship allows for a wide variety of distributions over different kinds of label spaces to be represented in the same mathematical formalism.
+A variational expectation maximization algorithm was proposed in \cite{BleiMcAuliffe2008} to learn model parameters.
+Experimental results in \cite{BleiMcAuliffe2008}  showed both excellent out of sample label prediction and improved topics.
+Topic improvement was measured by using the empirical topic proportions from SLDA  as features for external, discriminative approaches to label prediction.
+Regression models built on SLDA topics outperformed the same built on LDA derived topics.
+
+In one sense SLDA is more general than the model we present in this chapter, namely, the labels need not be categorical valued.
+%In our $y_d | {\bar{\mathbf z}}_d, \boldsymbol\eta \sim \mathcal{N}(\bar{\mathbf{z}}_d^T\boldsymbol\eta,\delta)$ example the labels are real-valued for instance.
+The main subject of this chapter, a generalization of SLDA called hierarchically supervised LDA (HSLDA) \cite{Perotte-2012-NIPS} does not deal with real-valued labels, however, it is more general than SLDA in the case of categorical labels.
+The exponential family/GLM regression framework can theoretically account for multivariate labels and potentially even structured categorical labels.
+HSLDA is, however,  a specific, practical way to model with structured categorical labels.
+Because we focus on hierarchically structured categorical labels we refer to our model as a mixed membership hierarchical classification model.
+
 %
-% Despite the growing emphasis on meaningful
-%use of technology in medicine, many aspects of medical record-keeping
-%remain a manual process. Diagnostic coding for billing and insurance
-%purposes is often handled by professional medical coders who must
-%explore a patient's extensive clinical record before assigning the
-%proper codes. So while electronic health records (EHRs) should be
-%adopted by most medical institutions within the next several years,
-%largely due to the provisions of HITECH under the American Recovery
-%and Reinvestment Act \citep{Blumenthal2009}, there has been little
-%movement forward in automating medical coding.
+%Another way of saying this is that SLDA was the first document classifier for documents in the bag-of-words representation that exploited the LDA topic feature representation.  
+%%A subtle point should be noted here. 
+% SLDA and the subject of this chapter, a generalization of it called hierarchically supervised LDA (HSLDA) \cite{Perotte-2012-NIPS} are not, strictly speaking, regression models.
+This is because they model the {\em joint} distribution of labels and bag-of-words data rather than the conditional distribution of the labels given the input.
+Having the joint distribution allows us to condition on an input to produce a label for all test instances.
+Additionally, during training of the model, knowing the labels effects the learning of the low dimensional representation of the input -- a characteristic discussed later in this chapter.
+% This is quite different (and better performing in practice) than LASSO
+%(L1 regularized least squares regression) and LDA followed by least
+%squares regression \cite{BleiMcAuliffe2008}.
+%By performing joint inference, SLDA and HSLDA
+%infer topics such that the model predicts the response variable while
+%improving word likelihood.
 
-%An automated process would ideally produce a more complete and accurate
-%diagnosis lists. Also, this model will reveal information about the
-%medical records themselves. For example, we may gain an understanding
-%of what a specific code actually means in terms of clinical narratives.
-%Similarly, viewing the distribution of topics over discharge summaries
-%may reveal information about the latent structure of clinician documentation.
-%Lastly, the sLDA model would provide a novel approach to dealing with
-%the problem of high dimensionality when representing narrative text
-%in a vector space specifically by reducing dimensions from an entire
-%vocabulary of potentially tens of thousands of words to a set of several
-%dozen topics.
+%In this chapter we extend supervised latent Dirichlet allocation (SLDA)
+%\cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision. SLDA
+%is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per
+%document ``supervision,'' often taking the form of a single numerical or
+%categorical label. It has been demonstrated that the signal provided by such
+%supervision can result in better, task-specific document models and can also
+%lead to good label prediction for out-of-sample data \cite{BleiMcAuliffe2008}.
 
-%In this work we extend supervised latent Dirichlet allocation (sLDA)
-%\cite{BleiMcAuliffe2008} to take advantage of hierarchical supervision.  sLDA
-%is latent Dirichlet allocation (LDA) \cite{Blei2003} augmented with per
-%document ``supervision'';  often taking the form of a single numerical or
-%categorical ``label.''  More generally this supervision is just extra per
-%document data;  for instance its quality or relevance (e.g.~online review
-%scores), marks given to written work (e.g.~essay grades), or the number of
-%times a web page is linked.  These labels are usually generatively modeled as
-%having been conditionally drawn from some distribution that depends on the
-%document-specific topic mixture.  It has been demonstrated that the signal
-%provided by such supervision can result in better, task-specific document
-%models and can also lead to good label prediction for out-of-sample data
-%\cite{BleiMcAuliffe2008}.
+%It also has been demonstrated that SLDA has been shown to outperform both LASSO
+%(L1 regularized least squares regression) and LDA followed by least squares
+%regression~\cite{BleiMcAuliffe2008}.
+
+%While there has been much work in multi-label classification of text and
+%text modeling in general, we focus here on topic modeling approaches.
+%Latent Dirichlet allocation (LDA) is a generative probabilistic model which
+%represents documents as a mixed-membership bag of word. Each document is
+%represented as a collection of words, generated from a set of topic assignments
+%(one for each word), where each topic assignment is drawn from a distribution
+%over topics~\cite{Blei2003}. SLDA is latent Dirichlet allocation (LDA)
+%\cite{Blei2003} augmented with per-document labeling, often taking the form of
+%a single numerical or categorical label. Examples of labels include ratings
+%associated with online reviews, grades for essays, and the number of times a
+%webpage is linked. This approach has been shown to outperform both LASSO
+%(L1 regularized least squares regression) and LDA followed by least
+%squares regression~\cite{BleiMcAuliffe2008}.
+
+%One set of models that are particularly relevant to HSLDA are Chang
+%and Blei's hierarchical models for document networks (Relational Topic
+%Models). In that family of models, they encountered a similar scenario
+%where the lack of a link did not truly indicate absence. In hierarchically
+%labeled data, negative labels are uncommon and the lack of a label
+%in the hierarchy is not equivalent to a negative label. Therefore,
+%as in the work of Chang and Blei, we employ regularization to account
+%for the lack of negative labels. This will be discussed further in
+
+% Noemie: check out what I wrote here, because I am not really sure this is
+% correct. 
+%There have been several models that incorporate both latent models of text and
+%some form of
+%supervision~\cite{Ramage2009,DiscLDA,wangbleifeifei08,RelationalLDA}. One set 
+%of models that are particularly relevant to HSLDA are Chang and Blei's
+%hierarchical models for document networks (Relational Topic Models). In that
+%family of models, they encountered a similar scenario where an unselected label 
+%does not always indicate absence. In hierarchical labels, this phenomenon is
+%even more pervasive -- there are no explicit negative labels, but it is also
+%unclear how to treat the parents of selected labels. Like in the work of Chang
+%and Blei, we employ regularization to account for the lack of negative
+%labeling. In our experiments, we look at the impact of assigning positive and
+%negative instances to the ancestors of selected labels.

### 2012-12-18 Krantz-v3/Chapters/chapter1/example_applications.tex
Auto-Backup Commit of All Projects

@@
 
-Many web-retails store and organize their catalog of products in a
+Many web-retailers store and organize their catalog of products in a
 mulitply-rooted hierarchy in addition to providing textual product descriptions 

### 2012-12-18 Krantz-v3/Chapters/chapter1/inference.tex
Auto-Backup Commit of All Projects

@@
 of $\boldsymbol\eta_{l}$ is normally distributed with mean $\hat{\boldsymbol\mu}_{l}$
-and covariance $\hat{\mathbf{\Sigma}}.$% such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
+and covariance $\hat{\mathbf{\Sigma}},$% such that % (probably not the right place for this) We evaluated the model over various values of $\sigma$ where $\sigma=\left\{ 0.01,0.1,0.25,1,2\right\} $.
 where

### 2012-12-18 Krantz-v3/Chapters/chapter1/introduction.tex
Auto-Backup Commit of All Projects

@@
 Brushing aside exponential family and GLM link function generalities, what SLDA does is regress labels against the empirical distributions of assigned topic indicator variables $\bar{\mathbf{z}}_d = \{\bar{z}_{1,d}, \ldots, \bar{z}_{K,d}\}$ where $\bar{z}_{k,d} = \frac{\sum_n \mathbb{I}(z_{n,d}=k)}{\sum_n \sum_j \mathbb{I}(z_{n,d}=j)}$ is the fraction of words assigned to topic $k$ ($\mathbb{I}(\cdot)$ is the indicator function that returns one if its argument is true).
-If the document labels are real valued then one example choice for the regression relationship would be $y_d | {\bar{\mathbf z}}_d, \boldsymbol\eta \sim \mathcal{N}(\bar{\mathbf{z}}_d^T\boldsymbol\eta,\delta).$  Using an generalized linear model in the exponential family to parameterize this regression relationship allows for a wide variety of distributions over different kinds of label spaces to be represented in the same mathematical formalism.
+If the document labels are real valued then one example choice for the regression relationship would be $y_d | {\bar{\mathbf z}}_d, \boldsymbol\eta \sim \mathcal{N}(\bar{\mathbf{z}}_d^T\boldsymbol\eta,\delta).$  Using a generalized linear model in the exponential family to parameterize this regression relationship allows for a wide variety of distributions over different kinds of label spaces to be represented in the same mathematical formalism.
 A variational expectation maximization algorithm was proposed in \cite{BleiMcAuliffe2008} to learn model parameters.

### 2013-08-16 Krantz-v3/Chapters/chapter1/HSLDA.tex
Responses to 2nd reviewer request

@@
 In HSLDA, the bag-of-word document data is modeled using LDA with
- full, hierarchical topic estimation (i.e. with the global topic estimated too rather than assumed).
+ full, hierarchical topic estimation (i.e. global topic proportions are also estimated).
 %In HSLDA, documents are modeled using the LDA mixed-membership mixture model
@@
 
-Here $\bar{\mathbf{z}}_d^T = [\bar{z}_{1}, \ldots, \bar{z}_k, \ldots, \bar{z}_K]$ is the empirical topic distribution for document $d$, in which each entry is the percentage of the words in that document that come from topic $k$, $\bar{z}_{k}=N_{d}^{-1}\sum_{n=1}^{N_d}\mathbb{I}(z_{n,d}=k).$
+Here $\bar{\mathbf{z}}_d^T = [\bar{z}_{1}, \ldots, \bar{z}_k, \ldots, \bar{z}_K]$ is the empirical topic distribution for document $d$, in which each entry is the percentage of the words in that document that come from topic $k$, $\bar{z}_{k}=N_{d}^{-1}\sum_{n=1}^{N_d}\mathbb{I}(z_{n,d}=k).$  As in \cite{BleiMcAuliffe2008}, the response variables are directly dependent on $\bar{\mathbf{z}}_d^T$ because this directly couples the topic assignments used to explain the words and the topic assignments used to explain the responses.
 

### 2013-08-16 Krantz-v3/Chapters/chapter1/abstract.tex
Responses to 2nd reviewer request

--- 2011-06-02 paper/abstract.tex
@@
-We introduce hierarchically supervised latent Dirchlet allocation (HSLDA), a
-model for hierarchically and multiply labeled bag-of-word data.  Examples of
-such data include web pages and their placement in directories, product
+Placing documents within a hierarchical structure is a common task and can be viewed as a multi-label classification with hierarchical structure in the label space.
+Examples of such data include web pages and their placement in directories, product
 descriptions and associated categories from product hierarchies, and free-text
-clinical records and their assigned diagnosis codes. Out-of-sample label
+clinical records and their assigned diagnosis codes. We present a
+model for hierarchically and multiply labeled bag-of-word data called hierarchically supervised latent Dirichlet allocation (HSLDA).  Out-of-sample label
 prediction is the primary goal of this work, but improved lower-dimensional
